### 2019-01-22 kgswc19.tex
Initial source

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepackage{tikz}
\usepackage{enumitem}
\usepackage{gensymb}
\usepackage{amsmath}
\usepgfplotslibrary{clickable}

\usepackage{listings}
\lstset{language=SQL,morekeywords={PREFIX,java,rdf,rdfs,url}}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Change estimation of SPARQL queries in Wikidata\thanks{Supported by CONICYT PFCHA/ DOCTORADO BECAS CHILE/2017 - 72180000 and Millennium Institute for Foundational Research on Data.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Alberto Moya\orcidID{0000-1111-2222-3333} \and
Aidan Hogan\orcidID{1111-2222-3333-4444} \and
Jesus P\'erez\orcidID{2222--3333-4444-5555} \and
Grethel Coello\orcidID{2222--3333-4444-5555}}
%
\authorrunning{A. Moya et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Computer Science Department, University of Chile, Chile\\
\email{\{amoya,ahogan,jperez,gcoello\}@dcc.uhile.cl}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Linked Open Datasets contain descriptions that change over time. The resources described by these datasets are continuously created, moved, deleted, linked, and unlinked. Applications that leverage Linked Data must be aware of these change dynamics to deliver accurate services. In this work, we tackle the problem of estimating future changes in query results based on query syntax features. We propose a model that exploits the characteristics of RDF to capture the dynamics of the predicates. Our Dynamics Model complements syntax features extracted from queries to assess if query results will change in the near future using classifiers and long future using Linear Regression. To evaluate our proposal, we select Wikidata as our dataset and we build a curated collection of queries. Our results show that the accuracy of our proposal approaches that obtained using information on the historical evolution of the query results but at a lower cost.

\keywords{Change estimation \and SPARQL query \and Link Data dynamics.}
\end{abstract}
%
%
	%
\section{Introduction}
\label{sec:intro}
%
The Linked Data (LD) principles\footnote{https://www.w3.org/DesignIssues/LinkedData.html} provide a flexible publishing paradigm to integrate and interlink data on the Web. Following the LD principles, a lot of organizations are publishing valuable information online using Resource Description Framework (RDF) and other LD technologies. The standardized approach to query these kinds of data is SPARQL, the recommendation of the W3C\footnote{https://www.w3.org/} to query RDF. There are many public SPARQL endpoints that allow any data consumer to query RDF datasets on the Web~\cite{VandenbusscheUM17}. Since the LD principles have been created, until now, the LOD cloud\footnote{https://lod-cloud.net/} has grown significantly~\cite{SchmachtenbergBP14}.

Many applications that consume Linked Data face challenges related to changes in the underlying data, whose key is keeping data fresh. Since datasets change over time, long-running applications that cache and repeatedly use query results obtained from a SPARQL endpoint may resubmit the queries regularly to ensure up-to-dateness. As a result, applications often get the same results in several query executions, resulting in an unnecessary computation, and other times, they give the user stale results when the backend data change between re-executions.

By caching the answers of queries and using the cached replies whenever possible, client-side caches reduce the network traffic between clients and servers, the load on servers, and the average user-perceived latency of query processing. However, for caches to be useful, cache consistency must be maintained, that is, cached copies should be updated when the data that gave rise to it change.
Ensuring a strong consistency in which no stale copy of the modified document will ever be returned to the user could be expensive. Thus, several weak consistency approaches have been proposed that try to keep the local data of the applications updated at lower cost~\cite{KnuthHS16,DividinoGS15,NishiokaS17}.

Some works have proposed to learn from historical entity evolution, but a query could involve many entities. Considering the historical evolution of all entities relating to a query is expensive because it is necessary to have the previous complete versions of data and to obtain the results of the execution of the query in each version. In addition, several recent works have explored the dynamics of the query results with the intention of finding patterns that allow for characterizing, recognizing, and predicting changes~\cite{KaferAUOH13,NishiokaS16,GonzalezH18}. Recent works have found some general correlations with datasets dynamics, such as, the domains, predicates, and schema~\cite{KaferAUOH13,NishiokaS17,UmbrichHHPD10}. Based on dynamic knowledge, hybrid approaches have been developed to return fresher results and at the same time speed up queries, but freshness estimates are a major and still current challenge, because of the difficulty of predicting the underlying dynamics on the growing Web of Data.

In this work, we tackle the problem of estimating future changes in query results based on data statistics and methods from Machine Learning. We devise a solution where we do not have complete information from the past, but a reduced statistical model that summarizes the historical dynamics based on predicates. Subsequently, we develop classifiers for assessing whether or not query results will change in the future. Additionally, we propose to use a set of features extracted from the query syntax.

We apply this framework to compute a predictive Dynamics Model for 23 versions of the Wikidata knowledge graph~\cite{VrandecicK14}, evaluating its suitability for predicting future changes in the query results. We select Wikidata as: (1) it provides a history of weekly versions that we can use to evaluate predictions and (2) it is edited by thousands of users, which means that significant changes are observed weekly. In addition, we build a dataset of real queries on Wikidata to monitor and to check the validity of our proposed approach.

In summary, the \textbf{contributions} of this paper are: (i) we build a dataset of real queries for Wikidata, (ii) we provide an analysis of the evolution of the results of those queries over 23 weekly versions of Wikidata, (iii) we propose a classifier that judges whether or not the results of a query will change using a predicate-based Dynamics Model and features from the syntax of the query, and (iv) we show that the proposed features and classifier work well, and our results are close to the more expensive previous approaches that assume full knowledge of historical data.

The rest of this document is organized as follows: In Section~\ref{sec:related}, we review related works. Section~\ref{sec:preliminar} provides the formal definition of the problem. Section~\ref{sec:approach} describes our proposal. In Section~\ref{sec:data}, we explain our datasets. Section~\ref{sec:eval} presents the design and results of the experiments. We conclude and review future directions in Section~\ref{sec:conclusion}.

%
\section{Related Work}
\label{sec:related}
%

Many applications can benefit from change prediction in query results. Both, Smart Caching (to save computational resources) and Hybrid Architectures (to deliver fresh results in less time) require synchronization techniques. Several techniques have been developed following strong consistency, based on HTTP metadata~\cite{fielding2rfc} and notifications~\cite{PassantM10,Fitzpatrick10,TrampFEA10}. Other works seek weak consistency based on probabilities, such as Empirical Distributions~\cite{NeumaierU16}, Poisson Processes~\cite{UmbrichHHPD10}, and Markov Chains~\cite{UmbrichMP15}, as well as other works based on heuristics~\cite{AliciAOCU12,UmbrichMP15,KnuthHS16}, metrics~\cite{DividinoGS15,KnuthHS16,AkhtarAL17}, and Machine Learning~\cite{NishiokaS17,GonzalezH18}. In the case of metadata, some works~\cite{UmbrichHHPD10,DividinoKG14,Kjernsmo15,NeumaierU16} show that these data are often not provided or are wrong. In the case of notifications, several tools have been developed, such as rsine~\cite{MaderMS14} and Boca~RDF~\cite{MissierACDG07}, but these mechanisms must keep a subscriber list and open connection states that can also cause scalability problems. The rest of the approaches achieve interesting results, but the main deficiency in these approaches is that their quality depends on having a long and complete history of updates. In the case of ML-based approaches, they constitute a general framework where important improvements can be achieved by incorporating relevant features to the models. In our approach, we propose an ML-based solution that dispenses with the expensive history of updates by a lighter model with statistics from this history and syntactic query features.

Some works have tried to understand the dynamics of the LOD datasets~\cite{UmbrichHHPD10,UmbrichKL10,KaferAUOH13,DividinoSGG13,NishiokaS16}. Some features found to describe the dynamics of a resource are Pay-Level Domain (PLD)~\cite{NishiokaS16,NishiokaS17}, predicates~\cite{KaferAUOH13,NishiokaS17}, and Characteristic Sets (CS)~\cite{NishiokaS16,GonzalezH18} that are correlated with the dynamics of several datasets studied. This knowledge can be used to estimates the freshness of cached data. Umbrich et al.~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12} identified two variants to obtain knowledge about freshness for different query patterns based on predicates. In more recent work Dehghanzadeh et al.~\cite{DehghanzadehPKUHD14} proposed a solution to estimate the freshness using cardinality estimation techniques based on predicates and CS. However, to the best of our knowledge, there is no proposal to estimate the freshness of cached query results using Machine Learning models based on syntactic features from queries and reduced historical statistics. In addition, the previous approaches usually impose restrictions on queries, such as not considering complex queries, e.g., with filters. Also, we do not find works that report the quality of the proposed models on complex real queries.

%
\section{Preliminaries: RDF and SPARQL}
\label{sec:preliminar}

According to the LD principles, the data published on the Web must be described using the Resource Description Framework (RDF). RDF is a conceptual model based on directed graphs that serves to provide descriptive information about the resources on the Web. The RDF nodes correspond to resources, literals (strings), or blank nodes, which model variables in the graph. Resources are identified by an Internationalized Resource Identifier (IRI). The nodes are linked through labeled and directed arcs, which represent semantic relationships.

\begin{definition}[RDF Triple and Dataset]
	Let I, L, and B be disjoint sets of IRIs, literals, and blank nodes respectively. A tuple $ (s, p, o) \in (I \cup  B)  \times (I)  \times  (I \cup  B \cup  L)$  is called an RDF $ triple $, where $ s $ is called subject, $ p $ is called predicate, and $ o $ is called object. An RDF $dataset$ is a set of RDF $triples$.
\end{definition}

SPARQL is the recommended query language to retrieve and manipulate data stored in the RDF format. In this work, we focus on SPARQL SELECT queries, which return the set of variables and their mapping results. Here we define a SPARQL 1.0 query.

\begin{definition}[SPARQL expression]
	Let V be a set of variables disjoint from $U \cup B \cup L$. A SPARQL $expression$ is built recursively as follows. (1) A triple $pattern t \in  (U \cup V ) \times (U \cup V ) \times (L \cup U \cup V )$ is an $expression$. (2) If $Q_1$ and $Q_2$ are $expressions$ and R is a filter condition, then $Q_1$ FILTER R, $Q_1$ UNION $Q_2$, $Q_1$ OPTIONAL $Q_2$, $Q_1$ AND $Q_2$ are $expressions$.
\end{definition}

A triple $ pattern $ in the second position can contain an expression called \textit{Property Paths} since SPARQL~1.1 to allow for navigational querying over RDF graphs. Other features at SPARQL~1.1 will be seen in experiments.

\begin{definition}[Property Paths]
	\label{eq:pp}
	Property Paths expressions are defined by the grammar:\\

	\begin{tabular}{ll}    \hline
		Rules &  \\    \hline
		e $\rightarrow$ p & a predicate \\

		e $\rightarrow$ \^{}e    &  inverse path \\

		e $\rightarrow$ e $/$ e    &  a path of $ e_1 $ followed by $ e_2 $\\

		e $\rightarrow$ e $\mid$ e    &  a path of $ e_1 $ or $ e_2 $ \\

		e $\rightarrow$ e*    &  a path of zero or more $ e $ \\

		e $\rightarrow$ e+    &  a path of one or more $ e $ \\

		e $\rightarrow$ e?    &  a path of zero or one $ e $ \\

		e $\rightarrow$ !p    &  any predicate not $ p $ \\

		e $\rightarrow$ !( p1,...,pk, \^{}p1,...,\^{}pk)    &  any (inverse) predicate not listed \\

		e $\rightarrow$ ( e )    &  brackets used for grouping \\    \hline
	\end{tabular}

	where $p, p1,...,pk$ are IRIs in $I$.
\end{definition}

\section{Our Approach}
\label{sec:approach}

Next, we formalize the problem addressed in this work and present our proposed solution.

\subsection{Problem Definition and Proposed Solution}

For this work, we consider a dynamic RDF dataset, denoted by $\mathcal{D}$. In addition, we assume that we have consecutive sequence of $m$ past versions $\mathcal{D} = (d_1, d_2, ..., d_m)$ constituting different revisions fetched at regular intervals. We also consider a SPARQL SELECT query $q$, of which we would like to know if its results are going to be valid for a long time. For that, we have the results of $q$ on each of the $m$ past versions of $\mathcal{D}$ and we need to know when we should update it. Then, we call $q(d_i)$ the result of executing the query $q$ on the $d_i$ version of $\mathcal{D}$.\\

In addition, we consider delta ($\delta$) a function to compute the binary dissimilarity between two results for a query $q$ with the following expression:

\begin{equation}
\label{eq:delta}
\delta(q(d_i), q(d_j)) = \begin{cases}
0 & q(d_i) = q(d_j) \\
1 & q(d_i) \neq q(d_j) 
\end{cases}
\end{equation}

\begin{example}
	\label{ex:dataset}    
	Consider the following examples of seven versions of a dynamic RDF dataset that contains a subject and four properties. Also, consider the queries~\ref{lst:sparql0}, \ref{lst:sparql1} and \ref{lst:sparql2}.
\end{example}

\setlength{\tabcolsep}{6pt}
\begin{tabular}{l l l}    \hline	
	\multicolumn{3}{c}{versions $d_1$ and $d_2$} \\    \hline
	:Atlantic\_Ocean  & :instance\_of  & ocean          \\
	:Atlantic\_Ocean  & :temperature   & 40$\degree$F          \\
	:Atlantic\_Ocean  & :area          & $10^7\pm10^3$ $km^2$ \\
	:Atlantic\_Ocean  & :image         & AOcean1.png    \\    \hline
	\multicolumn{3}{c}{versions $d_3$, $d_4$ and $d_5$} \\    \hline
	:Atlantic\_Ocean  & :instance\_of  & ocean          \\
	:Atlantic\_Ocean  & :temperature   & 50$\degree$F             \\
	:Atlantic\_Ocean  & :area          & $10^7\pm10^3$ $km^2$ \\
	:Atlantic\_Ocean  & :image         & AOcean1.png    \\    \hline
	\multicolumn{3}{c}{versions $d_6$ and $d_7$} \\    \hline
	:Atlantic\_Ocean  & :instance\_of  & ocean          \\
	:Atlantic\_Ocean  & :temperature   & 60$\degree$F    \\
	:Atlantic\_Ocean  & :area          & $10^7\pm10^3$ $km^2$ \\
	:Atlantic\_Ocean  & :image         & AOcean1.png    \\
	:Atlantic\_Ocean  & :image         & AOcean2.png    \\    \hline
\end{tabular}\\

\begin{lstlisting}[captionpos=b, caption=SPARQL query about ocean temperatures., label=lst:sparql0,
basicstyle=\ttfamily,frame=single]
SELECT ?ocean ?temp
WHERE {
  ?ocean :instance_of :ocean .
  ?ocean :temperature ?temp
}
\end{lstlisting}

\begin{lstlisting}[captionpos=b, caption=SPARQL query about ocean areas., label=lst:sparql1,
basicstyle=\ttfamily,frame=single]
SELECT ?ocean ?area
WHERE {
  ?ocean :instance_of :ocean .
  ?ocean :area ?area
}
\end{lstlisting}

\begin{lstlisting}[captionpos=b, caption=SPARQL query about ocean images., label=lst:sparql2,
basicstyle=\ttfamily,frame=single]
SELECT ?ocean ?image
WHERE {
?ocean :instance_of :ocean .
?ocean :image ?image
}
\end{lstlisting}

$\delta(q_1(d_1), q_1(d_3)) = 1$ and $\delta(q_2(d_2), q_2(d_4)) = 0$

Based on the $\delta$ function (see Equation~\ref{eq:delta}), the problem of Time-to-Live (TTL) estimation, denoted by $TTL(q,m)$ is defined by the size of the largest interval from $m$ for which the query result does not change (see Equation~\ref{eq:ttl}).

\begin{equation}
\label{eq:ttl}
TTL(q,m)=min\{t \in \mathbf{Z}^+ \mid \delta(q(d_m), q(d_{m+t})) \neq 0 \}
\end{equation}

As a relaxed version of this problem, there is the problem of One-Step-Change (OSC) estimation to predict the change in the next interval. (see Equation~\ref{eq:osc})

\begin{equation}
\label{eq:osc}
OSC(q,m) = \delta(q(d_m), q(d_{m+1}))
\end{equation}

Considering the dataset from Example~\ref{ex:dataset} and queries~\ref{lst:sparql0}, \ref{lst:sparql1} and \ref{lst:sparql2}, we show the change history of the query $q_1$ (see Figure~\ref{fig:problem}). We mark with a discontinuous line the OSC and with a continuous line the TTL at version $ d_3 $. The OSC is equal to zero, which means that the results will not change in the next version $d_4$, and the TTL equals three.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.7\linewidth]{img/problem.pdf}
	\caption{Example change history of query $q_1$.}
	\label{fig:problem}
\end{figure}

We propose a scheme that automatically identifies whether a SPARQL query will change in the next version (OSC) or when it will change (TTL), depending on the syntax of the query and statistics about the dynamics of predicates. Figure~\ref{fig:schema} shows the components of our approach, which receives as input a SPARQL query. The Feature Extractor analyzes the query and extracts information about the syntax of the query. The Dynamics Model considers some of these features to provide a dynamics indicator of the query. This indicator, together with the previous features extracted, is passed through a classifier that finally predicts the value for the problems posed in Equation~\ref{eq:osc} and Equation~\ref{eq:ttl}.

\begin{figure*}[!ht]
	\centering
	\includegraphics[width=0.7\linewidth]{img/schema.pdf}
	\caption{ Proposed architecture. It receives a SPARQL query as input; uses a pre-calculated Dynamics Model to estimate a query dynamic feature, which is used with other features to predict the OSC and TTL values. }
	\label{fig:schema}
\end{figure*}

\subsection{Feature Extractor}

In order to predict when the results of a query will change, we propose a framework based on Machine Learning using features extracted from the queries. Our features are based on the syntactic analysis of the queries; then some features are delivered to the Dynamics Model, which provides an indicator based on the probability of change for the constant predicates within the triple patterns of the query.

The extracted features try to capture the basic characteristics of a query. We try to discover the degree of correlation between the propensity to change, with respect to numerical values, such as the number of variables, the number of triple patterns, the number of predicate or entity constants, or the number of projected variables. We also believe that some qualitative variables may be related to the change, such as the existence of some solution-modifiers.

We also consider features that describe the complexity of the queries, such as the presence of subqueries, the existence of recursive expressions in the Property Paths of the triple patterns or the existence of any type of negation.

Figure~\ref{fig:query} shows the complete list of evaluated features and shows the nominals. In the case of negation, we consider the statements MINUS, FILTER NOT EXISTS, and FILTER NOT BOUND and by the recursive path we understand those Property Paths that have patterns of the form $e*$ or $e+$ (see Definition~\ref{eq:pp}).

\begin{lstlisting}[captionpos=b, caption=Query example, label=fig:query,
basicstyle=\ttfamily,frame=single]
SELECT ?item
WHERE {
  ?item :instance_of :human .
  ?item :gender :female .
    { ?item :place_of_birth :Wales }
  UNION
    { ?item :place_of_birth ?pob .
      ?pob :located_in* :Wales }
  OPTIONAL { ?sitelink schema:about ?item .
             ?sitelink schema:inLanguage "cy" }
  FILTER (!BOUND(?sitelink))
}
LIMIT 100
\end{lstlisting}

\begin{center}
	\begin{tabular}{|l|ll|}    \hline
		1  & \# triples        & 7                             \\
		2  & \# vars           & 3                             \\
		3  & \# projected vars & 1                             \\
		4  & \# predicates     & 6                             \\
		5  & FILTER            & TRUE                          \\
		6  & LIMIT             & TRUE                          \\
		7  & UNION             & TRUE                          \\
		8  & SUBQUERY          & FALSE                         \\
		9  & *PATH             & TRUE                          \\
		10 & NEGATION          & TRUE                          \\
		11 & predicates        & {[}$p_1$,...,$p_k${]}    \\    \hline
	\end{tabular}
\end{center}

\subsection{RDF Dynamics Model}

We propose a model to estimate the dynamics of the results of a SPARQL query on an RDF dataset. The intuition behind our model is to capture how many triples change for a predicate in a time interval. For this, in the following, we define \textit{Predicate Multiplicity of an RDF dataset}.

\begin{definition}[Predicate Multiplicity of an RDF Dataset]
	Given an RDF dataset $d$ and a predicate $p$, the multiplicity of $p$ in $d$, denoted by $M(p,d)$, is:
	\begin{equation}
	\label{eq:pm}
	M(p,d) = |\{(x,y,z) \in d : y = p \}|
	\end{equation}
\end{definition}

\begin{example}
	\label{ex:pm}    
	Considering Example~\ref{ex:dataset}, the Predicate Multiplicity of predicate \textit{image} in $d_6$ dataset is:    $M(image,d_6) = 2$
\end{example}

Then, we must capture the change between two versions of an RDF dataset at the triples level, which represents triples added and deleted between the versions. For this purpose, we define a \textit{Difference Set} as an RDF dataset denoted by $\Delta(d_1, d_2)$ applying low-level change detection techniques.

\begin{definition}[Difference Set]
	Let $d_1$ and $d_2$ be two RDF datasets; a Difference Set denoted by $\Delta(d_1, d_2)$ is:
	\begin{equation}
	\label{eq:ds}
	\Delta(d_1, d_2) = \{t \mid t \in d_2 \wedge t \notin d_1\} \cup \{t \mid t \in d_1 \wedge t \notin d_2\}
	\end{equation}
\end{definition}

\begin{example}
	\label{ex:ds}
	We show the Difference Set between versions $d_1$ and $d_6$:\\
	\begin{center}    
		\begin{tabular}{lll}    \hline
			\multicolumn{3}{c}{$\Delta(d_1, d_6)$} \\    \hline
			-    :Atlantic\_Ocean  & :temperature   & 40$\degree$F          \\
			+    :Atlantic\_Ocean  & :temperature   & 60$\degree$F          \\
			+    :Atlantic\_Ocean  & :image         & AOcean2.png              \\    \hline
		\end{tabular}
	\end{center}

	We can now denote the Predicate Multiplicity of a Difference Set of $d_1$ and $d_2$ by $M(p, \Delta(d_1, d_2))$, e.g., $M(image, \Delta(d_1, d_6)) = 1$.
\end{example}

Following the intuition of our model, we now look for the multiplicity of a predicate $p$ of a dynamic dataset $\mathcal{D}$ in a time interval $[n,m]$.

\begin{definition}[Aggregated Predicate Multiplicity of a dynamic dataset]
	Let $\mathcal{D} = (d_1, d_2, ..., d_m)$ be a dynamic dataset, and $p$ a predicate. The Aggregated Predicate Multiplicity of $\mathcal{D}$ in a time interval $[n,m]$, denoted by $AM(p, \mathcal{D},[n,m])$ is:
	\begin{equation}
	\label{eq:apm}
	AM(p, \mathcal{D},[n,m]) = \sum_{i=n}^{m-1}M(p, \Delta(d_i, d_{i+1}))
	\end{equation}
\end{definition}

\begin{example}
	\label{ex:apm}
	$AM(temperature, \mathcal{D},[1, 7]) =$

	$ \sum_{i=1}^{6}M(temperature, \Delta(d_i, d_{i+1})) = 4$

	\begin{center}    
		\begin{tabular}{lll}    \hline
			\multicolumn{3}{c}{$\sum_{i=1}^{6}\Delta(d_i, d_{i+1})$} \\    \hline
			-    :Atlantic\_Ocean  & :temperature   & 40$\degree$F          \\
			+    :Atlantic\_Ocean  & :temperature   & 50$\degree$F          \\            
			-    :Atlantic\_Ocean  & :temperature   & 50$\degree$F          \\
			+    :Atlantic\_Ocean  & :temperature   & 60$\degree$F          \\
			+    :Atlantic\_Ocean  & :image         & AOcean2.png              \\    \hline
		\end{tabular}
	\end{center}
\end{example}

We can now denote the Aggregated Multiplicity of a dynamic dataset in a time interval $[n,m]$ by:

\begin{equation}
\label{eq:am}
AM(\mathcal{D},[n,m]) = \sum_{p \in P}AM(p, \mathcal{D},[n,m])
\end{equation}\\
Where $ P $ denotes the set of all predicates that have changed in the interval $[n,m]$.

\begin{example}
	\label{ex:am}
	$AM(\mathcal{D},[1,7]) = AM(temperature, \mathcal{D},[1,7])$

	$+ AM(image, \mathcal{D},[1,7]) = 5$
\end{example}

Finally, the \textit{Dynamics} of a predicate is given by the \textit{Aggregated Predicate Multiplicity of a dynamic dataset} $\mathcal{D}$ divided by \textit{Aggregated Multiplicity of a dynamic dataset}.

\begin{equation}
\label{eq:dyn}
Dyn(p, \mathcal{D},[n,m]) = \frac{AM(p, \mathcal{D},[n,m])}{AM(\mathcal{D},[n,m])}
\end{equation}

\begin{example}
	\label{ex:dyn}
	Considering our example:

	$ Dyn(temperature, \mathcal{D},[1,7]) = \frac{4}{5} = 0.8$

	$ Dyn(image, \mathcal{D},[1,7]) = \frac{1}{5} = 0.2$

	$ Dyn(area, \mathcal{D},[1,7]) = \frac{0}{5} = 0$
\end{example}

Considering our proposed model, the Dynamics Model component receives the set of predicates extracted from the queries, calculates the $Dyn(p, \mathcal{D},[n,m])$ for each predicate $p$ and provides the value of a Transformation Function on these indicators. The idea behind the Transformation Function is to estimate the dynamics of the query based on the Dynamics of the predicates, e.g., Average and Maximum.

\section{Datasets}
\label{sec:data}

We conducted an empirical study to evaluate the effectiveness of our proposal to predict change in SPARQL query results. For this, we need an RDF dataset, with a lot of information, preferably, cross-domain, and that is constantly updated since we assume that the information of the dataset is correct, complete, but dynamic. We also need a set of queries appropriate to the chosen dataset, preferably a heterogeneous set defined by real users and covering a large part of the dataset.

\subsection{RDF Dataset}

We use Wikidata as one of the largest cross-domain knowledge graphs. There are several different kinds of data dumps available in JSON, XML, and RDF formats. We use 23 snapshots from 18/04/2017 to 27/09/2017, which are captured weekly. We worked with the truthy version in the RDF N-triples format, which represents statements that have the best non-deprecated rank for a given property.

The first version has 1,102,242,331 triples, 57,499,191 unique IRIs, 3,276 unique predicates, 933,588,860 literals, and 7,449 blank nodes, while the final version has 1,924,967,162 triples (+74\%), 80,637,164 unique IRIs (+40\%), 3,660 unique predicates (+11\%), 1,676,930,476 literals (+79\%) and 8,909 blank nodes (+19\%). Wikidata is maintained collaboratively by volunteer editors. Figure~\ref{graph:datasize} shows the growth maintained over the period of some of these indicators. The gap in week 12 is due to the fact that we were not able to obtain the data for that week. Even though the information is relatively incremental, there are deletions. Figure~\ref{graph:delta} shows the ratio of triples eliminated and added each week.

\begin{figure}[h]
	\begin{minipage}[b]{0.40\linewidth}
		\centering
		\begin{tikzpicture}[thick, scale=0.5]
		\begin{axis}[
		%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
		xlabel={Version},
		ymin=0,
		xtick={0,5,10,15,20,25},
		legend pos=north west,
		ymajorgrids=true,
		grid style=dashed,
		]

		\addplot[
		color=blue,
		mark=square,
		]
		coordinates {
			(1,1102242331)(2,1138975810)(3,1141303957)(4,1170283653)(5,1176093821)(6,1184177503)(7,1190456714)(8,1211813217)(9,1239145151)(10,1271353260)(11,1293099057)
		};

		\addplot[
		color=blue,
		mark=square,
		]
		coordinates {
			(13,1348622441)(14,1418460320)(15,1432143355)(16,1470298276)(17,1506140622)(18,1542278452)(19,1593334993)(20,1613832004)(21,1684774955)(22,1771601730)(23,1843086909)(24,1924967162)
		};

		%\legend{# entities}
		%\legend{# predicates}
		%\legend{# literals}
		%\legend{# blank nodes}

		\end{axis}
		\end{tikzpicture}
		\caption{Evolution of the number of triples on Wikidata.}
		\label{graph:datasize}
	\end{minipage}
	\hspace{0.5cm}
	\begin{minipage}[b]{0.40\linewidth}
		\centering
		\begin{tikzpicture}[thick, scale=0.5]
		\begin{axis}[
		x tick label style={
			/pgf/number format/1000 sep=},
		%ylabel=\#Triples,
		enlargelimits=0.05,
		legend style={at={(0.5,-0.1)},
			anchor=north,legend columns=-1},
		ybar interval=0.7,
		]
		\addplot 
		coordinates {(1,40137188)(2,4563382)(3,29302968)(4,8906402)(5,11455047)(6,11435682)(7,27229516)(8,26968577)(9,38952400)(10,20283316)(11,0)(12,0)(13,70073220)(14,17685830)(15,45405782)(16,41199256)(17,42633484)(18,56683899)(19,25309217)(20,25030143)(21,22548371)(22,23330356)(23,37009794)};
		\addplot 
		coordinates {(1,3403743)(2,2235268)(3,1116111)(4,3096270)(5,3371448)(6,5156519)(7,5873153)(8,1336452)(9,6744349)(10,999601)(11,0)(12,0)(13,1123311)(14,4002828)(15,7250907)(16,5356971)(17,6495717)(18,5627376)(19,4812243)(20,790590)(21,681118)(22,1194899)(23,998105)};
		\legend{Added, Removed}
		\end{axis}
		\end{tikzpicture}
		\caption{Triples added and removed in Wikidata.}
		\label{graph:delta}
	\end{minipage}
\end{figure}

\subsection{Queries Dataset}

Given our dynamic, cross-domain, large dataset, with historical information about its evolution, we now need a collection of queries for the dataset.

As opposed to RDF data, which can be easily obtained in the form of dumps, at the time of conducting this research, we did not find any structured dataset with queries following the Wikidata data model. Besides, Wikidata query logs are inaccessible, which would be interesting to understand the current usage of these data.

We found a set of 388 queries of examples about Wikidata\footnote{https://www.wikidata.org/wiki/Wikidata:SPARQL\_query\_service/queries/examples}, distributed in 18 categories. We consider these queries to evaluate our proposal. We crawled the sample queries from Wikidata in March 2018. This set is not a query log, but a wiki page containing selected queries sent by users. These queries are intended to show the behavior of the system or highlight the characteristics of the Wikidata dataset and, therefore, should reflect real-world queries.

Finally, the queries had to be curated manually, because they had syntax errors; others were not considered, because they asked information external to Wikidata, such as DBpedia\footnote{http://dbpedia.org/}, bigdata\footnote{http://www.bigdata.com/}; others asked for qualifiers that are not in the truthy version. We also eliminated the queries that returned bindings to blank nodes to facilitate the comparison between the results (they were only 10).

We consider query results as changed if it is not the same results returned for this query in the previous evaluation. Therefore, we need deterministic queries, that is, we remove those that could change their result even when the data is the same, such as, queries that contain the keyword SAMPLE and LIMIT without a total order. In our case, we eliminate the queries with the keyword SAMPLE and add ORDER BY to all the queries with all the projected variables. The latter change also facilitates the detection of changes in the results by iterating over both results sets in their sorted order.

Many queries use the custom Label service of Wikidata\footnote{https://www.mediawiki.org/wiki/Wikidata\_Query\_Service/User\_Manual\#Label\_service}. The service is very helpful when you want to retrieve labels, as it reduces the complexity of SPARQL queries that would otherwise be needed to achieve the same effect. This service can contain one or more language codes, separated by commas. Languages are preferred in the order in which they are specified. If no label is available in any of the specified languages, the Q-id of the entity (without any prefix) is its label. For our purposes, we need to consult only the information of the truthy version of Wikidata. Therefore, these queries were modified by the following procedure:

\begin{itemize}
	\item If an unbound variable in SELECT is named $?NAMELabel$, then we generated the label (rdfs:label) for the entity in variable $ ?NAME $ for each specified languages.
	\item If an unbound variable in SELECT is named $ ?NAMEAltLabel $, then we generated the alias (skos:altLabel) for the entity in variable $ ?NAME $ for each specified languages.
	\item If an unbound variable in SELECT is named $ ?NAMEDescription $, then we generated the description (schema:description) for the entity in variable $ ?NAME $ for each specified languages.
\end{itemize}
Then, each label, alia, and description are delivered according to the language order.

Figures~\ref{fig:languageB} and \ref{fig:languageA} shows an example of several transformations of a query.

\begin{lstlisting}[captionpos=b, caption=Query transformation. Before., label=fig:languageB,
basicstyle=\ttfamily,frame=single]
#Cats
SELECT ?item ?itemLabel
WHERE {
  ?item wdt:P31 wd:Q146 .
  SERVICE wikibase:label
    { bd:serviceParam wikibase:language "en,es" }
}
\end{lstlisting}

\begin{lstlisting}[captionpos=b, caption=Query transformation. After., label=fig:languageA,
basicstyle=\ttfamily,frame=single]
#Cats
SELECT ?item ?itemLabel
WHERE {
  ?item wdt:P31 wd:Q146 .
  OPTIONAL {?item rdfs:label ?en.FILTER(LANG(?en)="en")}
  OPTIONAL {?item rdfs:label ?es.FILTER(LANG(?es)="es")}
  BIND(str(COALESCE(?en, ?es, strafter(str(?item),
    "http://www.wikidata.org/entity/"))) AS ?itemLabel)
}
ORDER BY ?item ?itemLabel
\end{lstlisting}

In the end, we have 221 curated queries. An analysis of the basic use of SPARQL characteristics was carried out counting the keywords in the queries (see Table~\ref{tab:Keyword}). The first block in Table~\ref{tab:Keyword} describes the type of queries where we see that the set only has SELECT queries.
The second block contains solution modifiers. We notice that all (100\%) of the queries use ORDER BY which is due to the determination process used. LIMIT is used in (18.10\%) and DISTINCT is used more widely (29.41\%).
The third block has keywords associated with the SPARQL algebra operators that occur in the body. We see that FILTER (92.31\%) and OPTIONAL (82.35\%) are quite common, which is due to the language service elimination process used and to a lesser extent UNION (6.79\%) and NOT EXISTS (5.43\%) with some MINUS (1.81\%) and EXISTS (0.45\%).
The fourth block has aggregation operators. We see the highest use of GROUP BY (23.08\%), which is combined with most aggregation operators. Within the operators, the use of COUNT (19.91\%) stands out and the rest are used scarcely (<1.81\%).

\begin{table}[h]
	\centering
	\caption{Keyword count in queries.}
	\label{tab:Keyword}
	\scalebox{0.9}{
		\begin{tabular}{lrr} \hline
			Element       & Absolute & Relative        \\    \hline
			SELECT        & 221      & 100.00\%      \\
			ASK           & 0        & 0.00\%        \\
			DESCRIBE      & 0        & 0.00\%        \\
			CONSTRUCT     & 0        & 0.00\%        \\    \hline
			DISTINCT      & 65       & 29.41\%       \\
			LIMIT         & 40       & 18.10\%       \\
			OFFSET        & 0        & 0.00\%        \\
			ORDER BY      & 221      & 100.00\%      \\    \hline
			FILTER        & 204      & 92.31\%       \\
			UNION         & 15       & 6.79\%        \\
			OPTIONAL      & 182      & 82.35\%       \\
			NOT EXISTS    & 12       & 5.43\%        \\
			MINUS         & 4        & 1.81\%        \\
			EXISTS        & 1        & 0.45\%        \\    \hline
			COUNT         & 44       & 19.91\%       \\
			MAX           & 2        & 0.90\%        \\
			MIN           & 1        & 0.45\%        \\
			AVG           & 1        & 0.45\%        \\
			SUM           & 2        & 0.90\%        \\
			GROUP\_CONCAT & 4        & 1.81\%        \\
			GROUP BY      & 51       & 23.08\%       \\
			HAVING        & 3        & 1.36\%        \\    \hline
	\end{tabular}}
\end{table}

Once we have shown a basic analysis about the presence of keywords in the queries, we analyze the structure of the triple patterns. This is important, because our notion of freshness focuses on predicates. This restricts our approach to triple patterns with a constant in the predicate position. Therefore, our first objective is to verify the frequency of such patterns in the queries. Table~\ref{tab:Pattern} shows that V C V (71.73\%) is the most used, V C C (18.05\%) is also very common and in total 90.35\% of the triple patterns have a constant in the predicate position, which helps validate our approach of using a model based on the dynamics of predicates.

\begin{table}[h]
	\centering
	\caption{Triple patterns (C=Constant, V=Variable).}
	\label{tab:Pattern}
	\scalebox{0.9}{
		\begin{tabular}{lrr}    \hline
			Pattern & Absolute & Relative \\    \hline
			V C V   & 751      & 71.73\%    \\
			V C C   & 189      & 18.05\%    \\
			V V V   & 9        & 0.86\%     \\
			C C V   & 6        & 0.57\%     \\
			V V C   & 5        & 0.48\%     \\
			C V V   & 3        & 0.29\%     \\
			C V C   & 1        & 0.10\%      \\
			C C C   & 0        & 0.00\%        \\    \hline
			V P C   & 69       & 6.59\%     \\
			V P V   & 13       & 1.24\%     \\
			C P V   & 1        & 0.10\%      \\
			C P C   & 0        & 0.00\%        \\    \hline
	\end{tabular}}
\end{table}

Table~\ref{tab:Pattern} also shows that 7.93\% of triple patterns contain a Property Path in the predicate position. Then we set out to analyze these triples more deeply. The Property Paths are expressions that are formed recursively by applying a set of rules. Table~\ref{tab:Path} shows the frequency of occurrence of these patterns in our queries. It shows that $ P \rightarrow P * $ (74.70\%) and $ P \rightarrow P + $ (9.64\%) are often applied, which corresponds to our intuition to have a feature that indicates if there are triple patterns offering arbitrary length paths (*PATH). We should also notice that there is no negation at the level of the path. Our feature extractor is able to extract all constants of predicates within paths.

\begin{table}[h]
	\centering
	\caption{Structure of navigational Property Paths.}
	\label{tab:Path}
	\scalebox{0.9}{
		\begin{tabular}{lrr}    \hline
			Rules                & Absolute & Relative \\    \hline
			$P \rightarrow P* $   & 62       & 74.70\%     \\
			$P \rightarrow P / P$  & 56       & 67.47\%    \\
			$P \rightarrow P+ $   & 8        & 9.64\%     \\
			$P \rightarrow P \mid P$  & 6        & 7.23\%     \\
			$P \rightarrow P? $   & 6        & 7.23\%     \\
			$P \rightarrow !(P)$ & 0        & 0.00\%        \\
			$P \rightarrow C $    & 83       & 100.00\%      \\    \hline
	\end{tabular}}
\end{table}

Finally, we evaluate the dynamics of the results of the queries considering Equation~\ref{eq:delta} between query results of the same query over consecutive versions of Wikidata.

Figure~\ref{graph:changeV} shows the sum of the $ \delta $ function for all the queries considering the consecutive results of each query for 23 weeks. We identify that approximately half of the results of the queries change every week.

Figure~\ref{graph:changeH} shows for each query how many changes registered over 23 weeks. We identify queries that change every week, others that never changed and a fairly uniform distribution of changes in the period.

\begin{figure}[h]
	\begin{minipage}[b]{0.40\linewidth}
		\centering
		\begin{tikzpicture}[thick, scale=0.5]
		\begin{axis}[
		%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
		xlabel={Version},
		ylabel={Changes},
		ymin=0, ymax=221,
		xtick={0,5,10,15,20,25},
		legend pos=north west,
		ymajorgrids=true,
		grid style=dashed,
		]

		\addplot[
		color=blue,
		mark=square,
		]
		coordinates {
			(1,115)(2,112)(3,118)(4,116)(5,121)(6,111)(7,127)(8,109)(9,118)(10,115)
		};

		\addplot[
		color=blue,
		mark=square,
		]
		coordinates {
			(12,105)(13,113)(14,118)(15,117)(16,111)(17,118)(18,112)(19,116)(20,106)(21,106)(22,129)
		};

		%\legend{# entities}
		%\legend{# predicates}
		%\legend{# literals}
		%\legend{# blank nodes}

		\end{axis}
		\end{tikzpicture}
		\caption{Query results changes by time.}
		\label{graph:changeV}
	\end{minipage}
	\hspace{0.7cm}
	\begin{minipage}[b]{0.40\linewidth}
		\centering
		\begin{tikzpicture}[thick, scale=0.5]
		\begin{axis}[
		%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
		xlabel={Queries},
		ymin=0,
		legend pos=north west,
		ymajorgrids=true,
		grid style=dashed,
		]

		\addplot[
		color=blue,
		mark=square,
		]
		coordinates {
			(1,21)(2,21)(3,21)(4,21)(5,21)(6,21)(7,21)(8,21)(9,21)(10,21)(11,21)(12,21)(13,21)(14,21)(15,21)(16,21)(17,21)(18,21)(19,21)(20,21)(21,21)(22,21)(23,21)(24,21)(25,21)(26,21)(27,21)(28,21)(29,21)(30,21)(31,21)(32,21)(33,21)(34,21)(35,21)(36,21)(37,21)(38,21)(39,21)(40,21)(41,21)(42,21)(43,21)(44,21)(45,20)(46,20)(47,20)(48,20)(49,20)(50,20)(51,20)(52,19)(53,19)(54,19)(55,19)(56,19)(57,19)(58,19)(59,19)(60,18)(61,18)(62,18)(63,18)(64,18)(65,18)(66,18)(67,17)(68,17)(69,17)(70,17)(71,17)(72,17)(73,17)(74,16)(75,16)(76,16)(77,16)(78,16)(79,16)(80,15)(81,15)(82,15)(83,15)(84,14)(85,14)(86,14)(87,14)(88,14)(89,14)(90,14)(91,14)(92,14)(93,13)(94,13)(95,13)(96,13)(97,13)(98,13)(99,12)(100,12)(101,12)(102,12)(103,12)(104,12)(105,12)(106,12)(107,12)(108,12)(109,11)(110,11)(111,11)(112,11)(113,10)(114,10)(115,10)(116,10)(117,10)(118,9)(119,9)(120,9)(121,9)(122,9)(123,9)(124,9)(125,9)(126,8)(127,8)(128,8)(129,8)(130,8)(131,8)(132,8)(133,8)(134,8)(135,8)(136,7)(137,7)(138,7)(139,7)(140,7)(141,7)(142,7)(143,6)(144,6)(145,6)(146,6)(147,6)(148,6)(149,6)(150,6)(151,5)(152,5)(153,5)(154,4)(155,4)(156,4)(157,4)(158,4)(159,4)(160,4)(161,4)(162,4)(163,4)(164,4)(165,4)(166,3)(167,3)(168,3)(169,3)(170,3)(171,3)(172,2)(173,2)(174,2)(175,2)(176,2)(177,2)(178,2)(179,2)(180,2)(181,2)(182,2)(183,2)(184,2)(185,2)(186,2)(187,2)(188,1)(189,1)(190,1)(191,1)(192,1)(193,1)(194,1)(195,1)(196,1)(197,1)(198,1)(199,1)(200,1)(201,1)(202,1)(203,1)(204,0)(205,0)(206,0)(207,0)(208,0)(209,0)(210,0)(211,0)(212,0)(213,0)(214,0)(215,0)(216,0)(217,0)(218,0)(219,0)(220,0)(221,0)
		};

		%\legend{# entities}
		%\legend{# predicates}
		%\legend{# literals}
		%\legend{# blank nodes}

		\end{axis}
		\end{tikzpicture}
		\caption{Query results changes by queries.}
		\label{graph:changeH}
	\end{minipage}
\end{figure}

Figure~\ref{graph:deltas} shows an analysis of the query result monotonicity in the evaluated period. Red dots mean only removed results, greens added, and blues added and removed. It can be clearly seen that query results are not monotonic.

\begin{figure}[h]	
	\begin{minipage}[b]{0.40\linewidth}
		\centering	
		\begin{tikzpicture}[thick, scale=0.5]			
		\begin{axis}[
		%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
		xlabel={Version},
		ylabel={Changes},
		ymin=0,
		xtick={0,5,10,15,20,25},
		legend pos=north west,
		ymajorgrids=true,
		grid style=dashed,
		]

		\addplot[
		color=green,
		mark=square,
		]
		coordinates {
			(1,14)(2,38)(3,27)(4,24)(5,32)(6,23)(7,36)(8,21)(9,24)(10,23)
		};

		\addplot[
		color=green,
		mark=square,
		]
		coordinates {
			(12,26)(13,26)(14,27)(15,29)(16,23)(17,25)(18,20)(19,24)(20,13)(21,18)(22,31)
		};

		\addplot[
		color=red,
		mark=square,
		]
		coordinates {
			(1,8)(2,5)(3,5)(4,8)(5,6)(6,7)(7,5)(8,9)(9,9)(10,6)
		};

		\addplot[
		color=red,
		mark=square,
		]
		coordinates {
			(12,3)(13,7)(14,7)(15,8)(16,7)(17,6)(18,5)(19,6)(20,7)(21,6)(22,10)
		};

		\addplot[
		color=blue,
		mark=square,
		]
		coordinates {
			(1,93)(2,69)(3,86)(4,84)(5,83)(6,81)(7,86)(8,79)(9,85)(10,86)
		};

		\addplot[
		color=blue,
		mark=square,
		]
		coordinates {
			(12,76)(13,80)(14,84)(15,80)(16,81)(17,87)(18,87)(19,86)(20,86)(21,82)(22,88)
		};

		%\legend{# entities}
		%\legend{# predicates}
		%\legend{# literals}
		%\legend{# blank nodes}			
		\end{axis}
		\end{tikzpicture}				
		\label{graph:deltaV}
	\end{minipage}
	\hspace{0.7cm}
	\begin{minipage}[b]{0.40\linewidth}
		\centering
		\begin{tikzpicture}[thick, scale=0.5]
		\begin{axis}[
		%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
		xlabel={Queries},
		ymin=0,
		legend pos=north west,
		ymajorgrids=true,
		grid style=dashed,
		]

		\addplot[
		color=red,
		mark=square,
		]
		coordinates {
			(1,8)(2,6)(3,6)(4,5)(5,5)(6,5)(7,4)(8,4)(9,4)(10,4)(11,4)(12,4)(13,4)(14,4)(15,3)(16,3)(17,3)(18,3)(19,3)(20,2)(21,2)(22,2)(23,2)(24,2)(25,2)(26,2)(27,2)(28,2)(29,2)(30,2)(31,1)(32,1)(33,1)(34,1)(35,1)(36,1)(37,1)(38,1)(39,1)(40,1)(41,1)(42,1)(43,1)(44,1)(45,1)(46,1)(47,1)(48,1)(49,1)(50,1)(51,1)(52,1)(53,1)(54,1)(55,1)(56,1)(57,1)(58,1)(59,1)(60,1)(61,1)(62,1)(63,1)(64,1)(65,1)(66,1)(67,1)(68,1)(69,1)(70,1)(71,1)(72,1)(73,1)(74,0)(75,0)(76,0)(77,0)(78,0)(79,0)(80,0)(81,0)(82,0)(83,0)(84,0)(85,0)(86,0)(87,0)(88,0)(89,0)(90,0)(91,0)(92,0)(93,0)(94,0)(95,0)(96,0)(97,0)(98,0)(99,0)(100,0)(101,0)(102,0)(103,0)(104,0)(105,0)(106,0)(107,0)(108,0)(109,0)(110,0)(111,0)(112,0)(113,0)(114,0)(115,0)(116,0)(117,0)(118,0)(119,0)(120,0)(121,0)(122,0)(123,0)(124,0)(125,0)(126,0)(127,0)(128,0)(129,0)(130,0)(131,0)(132,0)(133,0)(134,0)(135,0)(136,0)(137,0)(138,0)(139,0)(140,0)(141,0)(142,0)(143,0)(144,0)(145,0)(146,0)(147,0)(148,0)(149,0)(150,0)(151,0)(152,0)(153,0)(154,0)(155,0)(156,0)(157,0)(158,0)(159,0)(160,0)(161,0)(162,0)(163,0)(164,0)(165,0)(166,0)(167,0)(168,0)(169,0)(170,0)(171,0)(172,0)(173,0)(174,0)(175,0)(176,0)(177,0)(178,0)(179,0)(180,0)(181,0)(182,0)(183,0)(184,0)(185,0)(186,0)(187,0)(188,0)(189,0)(190,0)(191,0)(192,0)(193,0)(194,0)(195,0)(196,0)(197,0)(198,0)(199,0)(200,0)(201,0)(202,0)(203,0)(204,0)(205,0)(206,0)(207,0)(208,0)(209,0)(210,0)(211,0)(212,0)(213,0)(214,0)(215,0)(216,0)(217,0)(218,0)(219,0)(220,0)(221,0)
		};

		\addplot[
		color=green,
		mark=square,
		]
		coordinates {
			(1,13)(2,13)(3,13)(4,12)(5,11)(6,11)(7,10)(8,10)(9,10)(10,10)(11,9)(12,9)(13,9)(14,8)(15,8)(16,8)(17,8)(18,8)(19,8)(20,7)(21,7)(22,7)(23,7)(24,7)(25,7)(26,7)(27,7)(28,7)(29,7)(30,7)(31,6)(32,6)(33,6)(34,6)(35,6)(36,6)(37,6)(38,6)(39,6)(40,6)(41,6)(42,5)(43,5)(44,5)(45,5)(46,5)(47,5)(48,5)(49,5)(50,4)(51,4)(52,4)(53,4)(54,4)(55,4)(56,4)(57,4)(58,4)(59,4)(60,4)(61,4)(62,4)(63,3)(64,3)(65,3)(66,3)(67,3)(68,3)(69,3)(70,3)(71,3)(72,3)(73,3)(74,3)(75,3)(76,3)(77,3)(78,3)(79,2)(80,2)(81,2)(82,2)(83,2)(84,2)(85,2)(86,2)(87,2)(88,2)(89,2)(90,2)(91,2)(92,2)(93,2)(94,2)(95,2)(96,2)(97,2)(98,2)(99,2)(100,2)(101,2)(102,2)(103,2)(104,1)(105,1)(106,1)(107,1)(108,1)(109,1)(110,1)(111,1)(112,1)(113,1)(114,1)(115,1)(116,1)(117,1)(118,1)(119,1)(120,1)(121,1)(122,1)(123,1)(124,1)(125,1)(126,1)(127,1)(128,1)(129,1)(130,1)(131,0)(132,0)(133,0)(134,0)(135,0)(136,0)(137,0)(138,0)(139,0)(140,0)(141,0)(142,0)(143,0)(144,0)(145,0)(146,0)(147,0)(148,0)(149,0)(150,0)(151,0)(152,0)(153,0)(154,0)(155,0)(156,0)(157,0)(158,0)(159,0)(160,0)(161,0)(162,0)(163,0)(164,0)(165,0)(166,0)(167,0)(168,0)(169,0)(170,0)(171,0)(172,0)(173,0)(174,0)(175,0)(176,0)(177,0)(178,0)(179,0)(180,0)(181,0)(182,0)(183,0)(184,0)(185,0)(186,0)(187,0)(188,0)(189,0)(190,0)(191,0)(192,0)(193,0)(194,0)(195,0)(196,0)(197,0)(198,0)(199,0)(200,0)(201,0)(202,0)(203,0)(204,0)(205,0)(206,0)(207,0)(208,0)(209,0)(210,0)(211,0)(212,0)(213,0)(214,0)(215,0)(216,0)(217,0)(218,0)(219,0)(220,0)(221,0)
		};

		\addplot[
		color=blue,
		mark=square,
		]
		coordinates {
			(1,22)(2,22)(3,22)(4,22)(5,22)(6,22)(7,22)(8,22)(9,22)(10,22)(11,22)(12,22)(13,22)(14,22)(15,22)(16,22)(17,22)(18,22)(19,22)(20,22)(21,22)(22,22)(23,22)(24,22)(25,22)(26,22)(27,22)(28,22)(29,22)(30,22)(31,22)(32,21)(33,21)(34,21)(35,21)(36,21)(37,21)(38,21)(39,21)(40,21)(41,20)(42,20)(43,20)(44,20)(45,20)(46,19)(47,19)(48,19)(49,19)(50,18)(51,18)(52,18)(53,18)(54,16)(55,16)(56,15)(57,15)(58,15)(59,15)(60,15)(61,15)(62,15)(63,14)(64,14)(65,13)(66,13)(67,12)(68,11)(69,11)(70,11)(71,11)(72,11)(73,11)(74,11)(75,11)(76,11)(77,10)(78,10)(79,10)(80,10)(81,9)(82,9)(83,9)(84,9)(85,8)(86,8)(87,8)(88,8)(89,7)(90,7)(91,7)(92,7)(93,7)(94,7)(95,7)(96,7)(97,7)(98,7)(99,7)(100,7)(101,6)(102,6)(103,6)(104,6)(105,6)(106,6)(107,6)(108,6)(109,6)(110,6)(111,5)(112,5)(113,5)(114,5)(115,5)(116,5)(117,5)(118,5)(119,4)(120,4)(121,4)(122,4)(123,4)(124,4)(125,4)(126,4)(127,4)(128,4)(129,4)(130,4)(131,4)(132,4)(133,4)(134,3)(135,3)(136,3)(137,3)(138,3)(139,3)(140,3)(141,2)(142,2)(143,2)(144,2)(145,2)(146,2)(147,2)(148,2)(149,2)(150,2)(151,2)(152,2)(153,2)(154,2)(155,1)(156,1)(157,1)(158,1)(159,1)(160,1)(161,1)(162,1)(163,1)(164,1)(165,1)(166,1)(167,1)(168,1)(169,1)(170,1)(171,1)(172,1)(173,1)(174,1)(175,1)(176,1)(177,1)(178,1)(179,1)(180,0)(181,0)(182,0)(183,0)(184,0)(185,0)(186,0)(187,0)(188,0)(189,0)(190,0)(191,0)(192,0)(193,0)(194,0)(195,0)(196,0)(197,0)(198,0)(199,0)(200,0)(201,0)(202,0)(203,0)(204,0)(205,0)(206,0)(207,0)(208,0)(209,0)(210,0)(211,0)(212,0)(213,0)(214,0)(215,0)(216,0)(217,0)(218,0)(219,0)(220,0)(221,0)
		};
		%\legend{# entities}
		%\legend{# predicates}
		%\legend{# literals}
		%\legend{# blank nodes}

		\end{axis}
		\end{tikzpicture}
		\label{graph:deltaH}
	\end{minipage}
	\caption{Types of changes by version and query. Only addition of results (green), only elimination of results (red), or both (blue).}
	\label{graph:deltas}
\end{figure}

\begin{comment}
\tikzset{mark options={mark size=0.5}}
\begin{figure}[h]
\centering
\begin{tikzpicture}
\begin{axis}[%
clickable coords={(xy): \thisrow{label}},%
xlabel={Time},
scatter/classes={%
c={blue},%+-3
b={red},%-2
a={green}}]%+1
\addplot[scatter,only marks,%
scatter src=explicit symbolic]%
table[meta=label] {treschange.dat};
\end{axis}
\end{tikzpicture}
\caption{Query results monotony.}
\label{graph:changeT}
\end{figure}
\end{comment}

\section{Evaluation}
\label{sec:eval}

We evaluate the performance of our proposal to predict changes in the results of our queries. For this, we use the Wikidata dataset and our query dataset explained previously.
To build the Difference Sets ($\Delta$) for our model we used an approach based on a merge-sort to scan over the snapshots as follows:

\begin{enumerate}
	\item Sort all statements by their syntactic order (subject-predicate-object).
	\item Perform a pairwise comparison of the statements by scanning two snapshots in linear time.
	\item Trigger a detection of the change as soon as the order of the statements differs between two snapshots.
\end{enumerate}

Then we configure the parameters of our model. For the time interval, we consider three versions of the dataset equivalent to the last three weeks. As the Transformation Function, we evaluate three basic functions: Maximum, Minimum, and Average. For example, Maximum will take the dynamics of the most dynamic property (used as a predicate in the query) as a feature. 
In addition, given that existing approaches are based on the historical results of the queries, we evaluate the substitution of our model for a historical change indicator that counts how many times the results have changed in the last three weeks.

\subsection{Measures}

Depending on whether the query results really changed, we consider four possible outcomes of this decision:

\begin{enumerate}[label=\roman*]
	\item \textbf{True positive (tp):} Changed results are detected. Our proposal is correct in its prediction.
	\item \textbf{False positive (fp):} No changed results are not detected. Our proposal is incorrect in its prediction.
	\item \textbf{True negative (tn):} No changed results are detected. Our proposal is correct in its prediction.
	\item \textbf{False negative (fn):} Changed results are not detected. Our proposal is incorrect in its prediction.

\end{enumerate}

We measured precision (see Equation~\ref{eq:precision}), recall (see Equation~\ref{eq:recall}) and f-measure ($f_1$) (see Equation~\ref{eq:f1}).

\begin{equation}
\label{eq:precision}
precision = \frac{tp}{tp + fp}
\end{equation}

\begin{equation}
\label{eq:recall}
recall = \frac{tp}{tp + fn}
\end{equation}

\begin{equation}
\label{eq:f1}
f_1 = 2 \times \frac{precision \times recall}{precision + recall}
\end{equation}

\subsection{Results}

First, we consider the Pearson correlation between the characteristics. We found only a couple of highly correlated variables (\# triples and \# vars: 0.943). The influence of the use of various Transformation Functions in our model was also evaluated. In this case, Average was the Transformation Function that obtained the best result followed by Maximum.

We tested a list of classifiers, including Decision Tree, Nearest Neighbors, and Naive Bayes among others listed in Table~\ref{tab:classifier}. Our pilot experiment shows that Nearest Neighbors, Gradient Boosting Classifier, and Random Forest work better than the others. We evaluated the quality of our proposal by forming a set of classifiers to predict the change of 221 queries in 15 weeks, which is equivalent to 3315 objects to classify, described in terms of our proposed characteristics. We split the data 80\% for training and 20\% for tests. To avoid overfitting, we use cross-validation in the training data.

Table~\ref{tab:classifier} shows the quality values of the classifiers evaluated in terms of $F_1$. In the state of the art, most of the works use information of the history of the results of the queries, which requires storing previous complete past versions to evaluate the queries. To have a measure of the positioning of our proposal, we evaluate the incorporation of a new feature that counts how many times the results of the query change on average in the last three versions. Then, we did two executions: one with our syntactic characteristics and this new indicator, which we call the past, and a second with our syntactic characteristics and information about dynamic predicates from our model. Table~\ref{tab:classifier} shows that our algorithm achieved less quality. Despite obtaining lower results, the quality of our proposal is not far from that achieved by observing the results of queries in the past, but certainly, our approach is less expensive since it does not require knowledge of past query results.

Subsequently, we individually evaluate the classification performance using different groups of characteristics to determine the influence of each characteristic. Table~\ref{tab:features} shows the quality when each feature is not used: a lower value indicates that the characteristic has a large positive influence. For this, we use the Nearest Neighbors as classifier and Average as the Transformation Function. We can notice a slight fall in the feature Dyn and Negation.

Finally, we show in Figure~\ref{graph:Modelsize} the number of triples in Wikidata versions vs the number of triples in the Difference Sets, in which our Dynamics Model are based. It is clearly visible that our model requires much less storage space.

\begin{table*}[h]
	\centering
	\caption{Tested classifiers.}
	\label{tab:classifier}
	\begin{tabular}{|l|lll|lll|}    \hline
		& \multicolumn{3}{|l|}{Past Results}           & \multicolumn{3}{|l|}{Dynamics Model}        \\    \hline
		Classifier                   & Precision & Recall & F1 & Precision & Recall & F1 \\    \hline
		Nearest Neighbors            & 0.827869 & 0.776923 & 0.801587 & 0.742188 & 0.730769 & 0.736434 \\
		Gradient Boosting Classifier & 0.848233 & 0.784615 & 0.815185 & 0.734127 & 0.711538 & 0.722656 \\
		Random Forest                & 0.818548 & 0.780769 & 0.799213 & 0.725146 & 0.715385 & 0.720232 \\
		Decision Tree                & 0.830544 & 0.763462 & 0.795591 & 0.700382 & 0.705769 & 0.703065 \\
		Linear SVM                   & 0.865342 & 0.753846 & 0.805755 & 0.735808 & 0.648077 & 0.689162 \\
		Neural Net                   & 0.836364 & 0.796154 & 0.815764 & 0.693989 & 0.488462 & 0.573363 \\
		Logistic Regression          & 0.836032 & 0.794231 & 0.814596 & 0.616253 & 0.525000    & 0.566978 \\
		Naive Bayes                  & 0.792100   & 0.732692 & 0.761239 & 0.690323 & 0.411538 & 0.515663 \\    \hline
	\end{tabular}
\end{table*}

\begin{table*}[]
	\centering
	\caption{Feature analysis using Nearest Neighbors.}
	\label{tab:features}
	\begin{tabular}{|l|rrr|}        \hline
		Feature           & Precision & Recall & F1 \\    \hline
		\# triples        & 0.741414 & 0.705769 & 0.723152  \\
		\# vars           & 0.738878 & 0.714615 & 0.726544 \\
		\# predicates    & 0.732558 & 0.726923 & 0.729730 \\
		\# projected vars & 0.726744 & 0.721154 & 0.723938 \\
		FILTER            & 0.738878 & 0.714615 & 0.726544 \\
		LIMIT             & 0.732558 & 0.726923 & 0.729730 \\
		UNION             & 0.733640  & 0.730769 & 0.732201 \\
		SUBQUERY         & 0.732558 & 0.736923 & 0.734734  \\
		*PATH             & 0.732558 & 0.726923 & 0.729730 \\
		NEGATION          & 0.713462 & 0.713462 & 0.713462  \\
		DYN               & 0.706046 & 0.711538 & 0.708781    \\ \hline
	\end{tabular}
\end{table*}

\begin{figure}[h]
	\centering
	\begin{tikzpicture}[thick, scale=0.5]
	\begin{axis}[
	%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
	xlabel={Version},
	ymin=0,
	xtick={0,5,10,15,20,25},
	legend pos=north west,
	ymajorgrids=true,
	grid style=dashed,
	]

	\addplot[
	color=red,
	mark=square,
	]
	coordinates {
		(1,1102242331)(2,1138975810)(3,1141303957)(4,1170283653)(5,1176093821)(6,1184177503)(7,1190456714)(8,1211813217)(9,1239145151)(10,1271353260)(11,1293099057)
	};

	\addplot[
	color=red,
	mark=square,
	]
	coordinates {
		(13,1348622441)(14,1418460320)(15,1432143355)(16,1470298276)(17,1506140622)(18,1542278452)(19,1593334993)(20,1613832004)(21,1684774955)(22,1771601730)(23,1843086909)(24,1924967162)
	};

	\addplot[
	color=blue,
	mark=square,
	]
	coordinates {
		(2,43540931)(3,6798650)(4,30419079)(5,12002672)(6,14826495)(7,16592201)(8,33102669)(9,28305029)(10,45696749)(11,21282917)
	};

	\addplot[
	color=blue,
	mark=square,
	]
	coordinates {
		(13,71196531)(14,21688658)(15,52656689)(16,46556227)(17,49129201)(18,62311275)(19,30121460)(20,25820733)(21,23229489)(22,24525255)(23,38007899)
	};

	%\legend{# entities}
	%\legend{# predicates}
	%\legend{# literals}
	%\legend{# blank nodes}

	\end{axis}
	\end{tikzpicture}
	\caption{Number of triples in Wikidata vs number of triples in Difference Sets.}
	\label{graph:Modelsize}
\end{figure}

\section{Conclusion and Future Work}
\label{sec:conclusion}

In this paper we propose classifiers that predict whether or not the results of a query will change in the short time interval; for this, we propose a model based on the dynamics of the predicates in the datasets and a set of characteristics extracted from the syntax of the queries. The proposed framework is flexible for incorporating new features extracted from the queries, from the current results, or from the model based on the multiplicity of predicates.

We select the datasets to evaluate our proposal considering desirable characteristics, such as availability, size, dynamics, and integrity. Choosing Wikidata, we present statistics to understand such characteristics. In the case of the queries dataset, we did not find any existing dataset of queries that complied with the requirements, so we curated a query dataset for Wikidata based on query examples proposed by users.

Finally, we evaluate our proposal, making a weekly prediction of change in the results of the queries and evaluating the success rates of the different classifiers each week based on precision, recall, and F-measure.
We measure the effectiveness of the characteristic based on our model of predicate dynamics versus a characterization of dynamics based on the historical results. The latter is a common factor in the previous proposals, but it requires a high cost. Our results show that our approach using predicate dynamics is competitive for prediction while being less expensive to support. An analysis of the sensitivity of the model to the different characteristics was also carried out, finding that our characteristics based on the dynamics of the predicates have the greatest influence.

Although the results obtained are encouraging, they are still preliminary. We consider for future work evaluating the quality of our proposal over a longer period. In addition, we believe that more complex query characteristics can improve the results, such as considering operator semantics; for example, to see if negation is affecting predicates or if the predicates are affected by the existence of filters. On the other hand, it would be worth expanding the set of queries in the test dataset, in order to increase the size, complexity, and coverage. Recently a query dataset for Wikidata was published\footnote{https://lists.w3.org/Archives/Public/semantic-web/2018Aug/0023.html}. It would be good to evaluate this query dataset, as well as to evaluate the sensitivity of our proposal to consider longer intervals and different Transformation Functions.

Finally, we would like to evaluate the quality of the characteristics and the proposed model to estimate the current Time-to-Live of the queries, rather than a binary One-Step-Change classification.

\bibliographystyle{splncs04}
\bibliography{ref}

\end{document}

### 2019-01-22 kgswc19.tex
Remove some autors (Grethel y Jesus) temporary

@@
 \author{Alberto Moya\orcidID{0000-1111-2222-3333} \and
-Aidan Hogan\orcidID{1111-2222-3333-4444} \and
-Jesus P\'erez\orcidID{2222--3333-4444-5555} \and
-Grethel Coello\orcidID{2222--3333-4444-5555}}
+Aidan Hogan\orcidID{1111-2222-3333-4444}}% \and
+%Jesus P\'erez\orcidID{2222--3333-4444-5555} \and
+%Grethel Coello\orcidID{2222--3333-4444-5555}}
 %
@@
 Finally, we evaluate our proposal, making a weekly prediction of change in the results of the queries and evaluating the success rates of the different classifiers each week based on precision, recall, and F-measure.
+We also apply linear regression techniques to predict, specifically, when the results of a query will change.
 We measure the effectiveness of the characteristic based on our model of predicate dynamics versus a characterization of dynamics based on the historical results.

### 2019-01-23 kgswc19.tex
reducing ...

@@
 \institute{Computer Science Department, University of Chile, Chile\\
-\email{\{amoya,ahogan,jperez,gcoello\}@dcc.uhile.cl}}
+\email{\{amoya,ahogan\}@dcc.uhile.cl}}
 %
@@
 We propose a model that exploits the characteristics of RDF to capture the dynamics of the predicates.
-Our Dynamics Model complements syntax features extracted from queries to assess if query results will change in the near future using classifiers and long future using Linear Regression.
+Our Dynamics Model complements syntax features extracted from queries to assess if query results will change in the near future using classifiers and long future using Multiple Linear Regression techniques.
 To evaluate our proposal, we select Wikidata as our dataset and we build a curated collection of queries.
@@
 
-We use Wikidata as one of the largest cross-domain knowledge graphs.
-There are several different kinds of data dumps available in JSON, XML, and RDF formats.
-We use 23 snapshots from 18/04/2017 to 27/09/2017, which are captured weekly.
-We worked with the truthy version in the RDF N-triples format, which represents statements that have the best non-deprecated rank for a given property.
-
+We use 23 Wikidata snapshots from 18/04/2017 to 27/09/2017, which are captured weekly in truthy version, which represents statements that have the best non-deprecated rank for a given property. 
 The first version has 1,102,242,331 triples, 57,499,191 unique IRIs, 3,276 unique predicates, 933,588,860 literals, and 7,449 blank nodes, while the final version has 1,924,967,162 triples (+74\%), 80,637,164 unique IRIs (+40\%), 3,660 unique predicates (+11\%), 1,676,930,476 literals (+79\%) and 8,909 blank nodes (+19\%).
-Wikidata is maintained collaboratively by volunteer editors.
 Figure~\ref{graph:datasize} shows the growth maintained over the period of some of these indicators.
@@
 
-Given our dynamic, cross-domain, large dataset, with historical information about its evolution, we now need a collection of queries for the dataset.
-
-As opposed to RDF data, which can be easily obtained in the form of dumps, at the time of conducting this research, we did not find any structured dataset with queries following the Wikidata data model.
-Besides, Wikidata query logs are inaccessible, which would be interesting to understand the current usage of these data.
-
-We found a set of 388 queries of examples about Wikidata\footnote{https://www.wikidata.org/wiki/Wikidata:SPARQL\_query\_service/queries/examples}, distributed in 18 categories.
-We consider these queries to evaluate our proposal.
-We crawled the sample queries from Wikidata in March 2018.
-This set is not a query log, but a wiki page containing selected queries sent by users.
+As opposed to RDF data, which can be easily obtained in the form of dumps, at the time of conducting this research, we did not find any structured dataset with queries following the Wikidata data model and Wikidata query logs where inaccessible.
+We considered then 388 queries of examples from Wikidata page\footnote{https://www.wikidata.org/wiki/Wikidata:SPARQL\_query\_service/queries/examples}.
 These queries are intended to show the behavior of the system or highlight the characteristics of the Wikidata dataset and, therefore, should reflect real-world queries.
 
-Finally, the queries had to be curated manually, because they had syntax errors; others were not considered, because they asked information external to Wikidata, such as DBpedia\footnote{http://dbpedia.org/}, bigdata\footnote{http://www.bigdata.com/}; others asked for qualifiers that are not in the truthy version.
+Then, the queries had to be curated manually, because they had syntax errors; others asked information external to Wikidata or asked for qualifiers that are not present in the truthy version.
 We also eliminated the queries that returned bindings to blank nodes to facilitate the comparison between the results (they were only 10).
-
-We consider query results as changed if it is not the same results returned for this query in the previous evaluation.
-Therefore, we need deterministic queries, that is, we remove those that could change their result even when the data is the same, such as, queries that contain the keyword SAMPLE and LIMIT without a total order.
+Therefore, we need deterministic queries, that is, we remove those that could change their result even when the data is the same.
 In our case, we eliminate the queries with the keyword SAMPLE and add ORDER BY to all the queries with all the projected variables.
 The latter change also facilitates the detection of changes in the results by iterating over both results sets in their sorted order.
-
-Many queries use the custom Label service of Wikidata\footnote{https://www.mediawiki.org/wiki/Wikidata\_Query\_Service/User\_Manual\#Label\_service}.
-The service is very helpful when you want to retrieve labels, as it reduces the complexity of SPARQL queries that would otherwise be needed to achieve the same effect.
-This service can contain one or more language codes, separated by commas.
-Languages are preferred in the order in which they are specified.
-If no label is available in any of the specified languages, the Q-id of the entity (without any prefix) is its label.
-For our purposes, we need to consult only the information of the truthy version of Wikidata.
-Therefore, these queries were modified by the following procedure:
-
-\begin{itemize}
-	\item If an unbound variable in SELECT is named $?NAMELabel$, then we generated the label (rdfs:label) for the entity in variable $ ?NAME $ for each specified languages.
-	\item If an unbound variable in SELECT is named $ ?NAMEAltLabel $, then we generated the alias (skos:altLabel) for the entity in variable $ ?NAME $ for each specified languages.
-	\item If an unbound variable in SELECT is named $ ?NAMEDescription $, then we generated the description (schema:description) for the entity in variable $ ?NAME $ for each specified languages.
-\end{itemize}
-Then, each label, alia, and description are delivered according to the language order.
+Also, we remove the use of custom Label service of Wikidata\footnote{https://www.mediawiki.org/wiki/Wikidata\_Query\_Service/User\_Manual\#Label\_service}.
+We could not query online services.
+Therefore, these queries were modified.
 
@@
 
-In the end, we have 221 curated queries.
-An analysis of the basic use of SPARQL characteristics was carried out counting the keywords in the queries (see Table~\ref{tab:Keyword}).
-The first block in Table~\ref{tab:Keyword} describes the type of queries where we see that the set only has SELECT queries.
-The second block contains solution modifiers.
-We notice that all (100\%) of the queries use ORDER BY which is due to the determination process used.
-LIMIT is used in (18.10\%) and DISTINCT is used more widely (29.41\%).
-The third block has keywords associated with the SPARQL algebra operators that occur in the body.
-We see that FILTER (92.31\%) and OPTIONAL (82.35\%) are quite common, which is due to the language service elimination process used and to a lesser extent UNION (6.79\%) and NOT EXISTS (5.43\%) with some MINUS (1.81\%) and EXISTS (0.45\%).
-The fourth block has aggregation operators.
-We see the highest use of GROUP BY (23.08\%), which is combined with most aggregation operators.
-Within the operators, the use of COUNT (19.91\%) stands out and the rest are used scarcely (<1.81\%).
+In the end, we have 221 curated queries. An analysis of the basic use of SPARQL features was carried out counting the keywords in the queries (see Table~\ref{tab:Keyword}).
 
@@
 This restricts our approach to triple patterns with a constant in the predicate position.
-Therefore, our first objective is to verify the frequency of such patterns in the queries.
-Table~\ref{tab:Pattern} shows that V C V (71.73\%) is the most used, V C C (18.05\%) is also very common and in total 90.35\% of the triple patterns have a constant in the predicate position, which helps validate our approach of using a model based on the dynamics of predicates.
+Table~\ref{tab:Pattern} shows that in total 90.35\% of the triple patterns have a constant in the predicate position, which validate our approach of using a model based on the dynamics of predicates.
 
@@
 Then we set out to analyze these triples more deeply.
-The Property Paths are expressions that are formed recursively by applying a set of rules.
 Table~\ref{tab:Path} shows the frequency of occurrence of these patterns in our queries.

### 2019-01-23 kgswc19.tex
revised 80%

@@
 %
-The Linked Data (LD) principles\footnote{https://www.w3.org/DesignIssues/LinkedData.html} provide a flexible publishing paradigm to integrate and interlink data on the Web.
-Following the LD principles, a lot of organizations are publishing valuable information online using Resource Description Framework (RDF) and other LD technologies.
-The standardized approach to query these kinds of data is SPARQL, the recommendation of the W3C\footnote{https://www.w3.org/} to query RDF.
-There are many public SPARQL endpoints that allow any data consumer to query RDF datasets on the Web~\cite{VandenbusscheUM17}.
-Since the LD principles have been created, until now, the LOD cloud\footnote{https://lod-cloud.net/} has grown significantly~\cite{SchmachtenbergBP14}.
-
-Many applications that consume Linked Data face challenges related to changes in the underlying data, whose key is keeping data fresh.
+Many applications that consume Linked Data (LD) face challenges related to changes in the underlying data, whose key is keeping data fresh.
 Since datasets change over time, long-running applications that cache and repeatedly use query results obtained from a SPARQL endpoint may resubmit the queries regularly to ensure up-to-dateness.
@@
 By caching the answers of queries and using the cached replies whenever possible, client-side caches reduce the network traffic between clients and servers, the load on servers, and the average user-perceived latency of query processing.
-However, for caches to be useful, cache consistency must be maintained, that is, cached copies should be updated when the data that gave rise to it change.
+However, for caches to be useful, cached copies should be updated when the data that gave rise to it change.
 Ensuring a strong consistency in which no stale copy of the modified document will ever be returned to the user could be expensive.
@@
 Considering the historical evolution of all entities relating to a query is expensive because it is necessary to have the previous complete versions of data and to obtain the results of the execution of the query in each version.
-In addition, several recent works have explored the dynamics of the query results with the intention of finding patterns that allow for characterizing, recognizing, and predicting changes~\cite{KaferAUOH13,NishiokaS16,GonzalezH18}.
-Recent works have found some general correlations with datasets dynamics, such as, the domains, predicates, and schema~\cite{KaferAUOH13,NishiokaS17,UmbrichHHPD10}.
-Based on dynamic knowledge, hybrid approaches have been developed to return fresher results and at the same time speed up queries, but freshness estimates are a major and still current challenge, because of the difficulty of predicting the underlying dynamics on the growing Web of Data.
+In addition, recent works have explored the dynamics of the LD with the intention of finding patterns that allow for characterizing, recognizing, and predicting changes~\cite{KaferAUOH13,NishiokaS16,GonzalezH18} finding some general correlations, such as, the domains, predicates, and schema~\cite{KaferAUOH13,NishiokaS17,UmbrichHHPD10}.
+Based on dynamic knowledge, hybrid approaches have been developed to return fresher query results and at the same time speed up queries, but freshness estimates are a major and still current challenge, because of the difficulty of predicting the underlying dynamics on the growing Web of Data.
 
@@
 We devise a solution where we do not have complete information from the past, but a reduced statistical model that summarizes the historical dynamics based on predicates.
-Subsequently, we develop classifiers for assessing whether or not query results will change in the future.
-Additionally, we propose to use a set of features extracted from the query syntax.
+We compute a predictive Dynamics Model for 23 versions of the Wikidata knowledge graph~\cite{VrandecicK14}, evaluating its suitability for predicting future changes in the query results.
+We select Wikidata as: (1) it provides a history of weekly versions that we can use to evaluate predictions and (2) it is edited by thousands of users, which means that significant changes are observed weekly.
 
-We apply this framework to compute a predictive Dynamics Model for 23 versions of the Wikidata knowledge graph~\cite{VrandecicK14}, evaluating its suitability for predicting future changes in the query results.
-We select Wikidata as: (1) it provides a history of weekly versions that we can use to evaluate predictions and (2) it is edited by thousands of users, which means that significant changes are observed weekly.
-In addition, we build a dataset of real queries on Wikidata to monitor and to check the validity of our proposed approach.
-
-In summary, the \textbf{contributions} of this paper are: (i) we build a dataset of real queries for Wikidata, (ii) we provide an analysis of the evolution of the results of those queries over 23 weekly versions of Wikidata, (iii) we propose a classifier that judges whether or not the results of a query will change using a predicate-based Dynamics Model and features from the syntax of the query, and (iv) we show that the proposed features and classifier work well, and our results are close to the more expensive previous approaches that assume full knowledge of historical data.
+In summary, the \textbf{contributions} of this paper are: (i) we build a dataset of real queries for Wikidata, (ii) we provide an analysis of the evolution of the results of those queries over 23 weekly versions of Wikidata, (iii) we evaluated several classifiers to judges whether or not the results of a query will change regression techniques to determine specifically when a query results will change using a predicate-based Dynamics Model and features from the syntax of the query, and (iv) we show that the proposed features and classifier work well, and our results are close to the more expensive previous approaches that assume full knowledge of historical data.
 
@@
 Many applications can benefit from change prediction in query results.
-Both, Smart Caching (to save computational resources) and Hybrid Architectures (to deliver fresh results in less time) require synchronization techniques.
+Both, Smart Caching to save computational resources and Hybrid Architectures to deliver fresh results in less time.
 Several techniques have been developed following strong consistency, based on HTTP metadata~\cite{fielding2rfc} and notifications~\cite{PassantM10,Fitzpatrick10,TrampFEA10}.
 Other works seek weak consistency based on probabilities, such as Empirical Distributions~\cite{NeumaierU16}, Poisson Processes~\cite{UmbrichHHPD10}, and Markov Chains~\cite{UmbrichMP15}, as well as other works based on heuristics~\cite{AliciAOCU12,UmbrichMP15,KnuthHS16}, metrics~\cite{DividinoGS15,KnuthHS16,AkhtarAL17}, and Machine Learning~\cite{NishiokaS17,GonzalezH18}.
-In the case of metadata, some works~\cite{UmbrichHHPD10,DividinoKG14,Kjernsmo15,NeumaierU16} show that these data are often not provided or are wrong.
-In the case of notifications, several tools have been developed, such as rsine~\cite{MaderMS14} and Boca~RDF~\cite{MissierACDG07}, but these mechanisms must keep a subscriber list and open connection states that can also cause scalability problems.
+In the case of metadata, many works~\cite{UmbrichHHPD10,DividinoKG14,Kjernsmo15,NeumaierU16} have show that these data are often not provided or are wrong.
+In the case of notifications, several tools have been developed, such as rsine~\cite{MaderMS14} and Boca~RDF~\cite{MissierACDG07}, but these mechanisms must keep a subscriber list and it cause scalability problems.
 The rest of the approaches achieve interesting results, but the main deficiency in these approaches is that their quality depends on having a long and complete history of updates.
@@
 Some works have tried to understand the dynamics of the LOD datasets~\cite{UmbrichHHPD10,UmbrichKL10,KaferAUOH13,DividinoSGG13,NishiokaS16}.
-Some features found to describe the dynamics of a resource are Pay-Level Domain (PLD)~\cite{NishiokaS16,NishiokaS17}, predicates~\cite{KaferAUOH13,NishiokaS17}, and Characteristic Sets (CS)~\cite{NishiokaS16,GonzalezH18} that are correlated with the dynamics of several datasets studied.
-This knowledge can be used to estimates the freshness of cached data.
-Umbrich et al.~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12} identified two variants to obtain knowledge about freshness for different query patterns based on predicates.
+Some features found to describe the dynamics of a resource are Pay-Level Domain (PLD)~\cite{NishiokaS16,NishiokaS17}, predicates~\cite{KaferAUOH13,NishiokaS17}, and Characteristic Sets (CS)~\cite{NishiokaS16,GonzalezH18}.
+This knowledge could be used to estimates the freshness of cached data.
+Umbrich et al.~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12} identified variants to obtain knowledge about freshness for different query patterns based on predicates.
 In more recent work Dehghanzadeh et al.~\cite{DehghanzadehPKUHD14} proposed a solution to estimate the freshness using cardinality estimation techniques based on predicates and CS.
 However, to the best of our knowledge, there is no proposal to estimate the freshness of cached query results using Machine Learning models based on syntactic features from queries and reduced historical statistics.
-In addition, the previous approaches usually impose restrictions on queries, such as not considering complex queries, e.g., with filters.
-Also, we do not find works that report the quality of the proposed models on complex real queries.
 
@@
 
-In addition, we consider delta ($\delta$) a function to compute the binary dissimilarity between two results for a query $q$ with the following expression:
+Let delta ($\delta$) be a function to compute the binary dissimilarity between two results for a query $q$ with the following expression:
 
@@
 	\label{ex:dataset}    
-	Consider the following examples of seven versions of a dynamic RDF dataset that contains a subject and four properties.
-Also, consider the queries~\ref{lst:sparql0}, \ref{lst:sparql1} and \ref{lst:sparql2}.
+	Consider the following example of seven versions of a RDF dataset. Also, consider the queries~\ref{lst:sparql0}, \ref{lst:sparql1} and \ref{lst:sparql2}.
 \end{example}
@@
 
-In order to predict when the results of a query will change, we propose a framework based on Machine Learning using features extracted from the queries.
 Our features are based on the syntactic analysis of the queries; then some features are delivered to the Dynamics Model, which provides an indicator based on the probability of change for the constant predicates within the triple patterns of the query.
 
-The extracted features try to capture the basic characteristics of a query.
-We try to discover the degree of correlation between the propensity to change, with respect to numerical values, such as the number of variables, the number of triple patterns, the number of predicate or entity constants, or the number of projected variables.
+We try to discover the degree of correlation between the propensity to change, with respect to numerical values, such as the number of variables, the number of triple patterns or number of predicate.
 We also believe that some qualitative variables may be related to the change, such as the existence of some solution-modifiers.
-
 We also consider features that describe the complexity of the queries, such as the presence of subqueries, the existence of recursive expressions in the Property Paths of the triple patterns or the existence of any type of negation.
@@
 Considering our proposed model, the Dynamics Model component receives the set of predicates extracted from the queries, calculates the $Dyn(p, \mathcal{D},[n,m])$ for each predicate $p$ and provides the value of a Transformation Function on these indicators.
-The idea behind the Transformation Function is to estimate the dynamics of the query based on the Dynamics of the predicates, e.g., Average and Maximum.
+The intention of the Transformation Function is to estimate the dynamics of the query based on the Dynamics of the predicates.
 
@@
 We considered then 388 queries of examples from Wikidata page\footnote{https://www.wikidata.org/wiki/Wikidata:SPARQL\_query\_service/queries/examples}.
-These queries are intended to show the behavior of the system or highlight the characteristics of the Wikidata dataset and, therefore, should reflect real-world queries.
 
 Then, the queries had to be curated manually, because they had syntax errors; others asked information external to Wikidata or asked for qualifiers that are not present in the truthy version.
-We also eliminated the queries that returned bindings to blank nodes to facilitate the comparison between the results (they were only 10).
-Therefore, we need deterministic queries, that is, we remove those that could change their result even when the data is the same.
+We also eliminated the queries that returned bindings to blank nodes to facilitate the comparison between the results.
+Therefore, we need deterministic queries.
 In our case, we eliminate the queries with the keyword SAMPLE and add ORDER BY to all the queries with all the projected variables.
-The latter change also facilitates the detection of changes in the results by iterating over both results sets in their sorted order.
-Also, we remove the use of custom Label service of Wikidata\footnote{https://www.mediawiki.org/wiki/Wikidata\_Query\_Service/User\_Manual\#Label\_service}.
-We could not query online services.
-Therefore, these queries were modified.
+The latter change also facilitates the detection of changes in the results.
+Also, we removed the use of custom Label service of Wikidata\footnote{https://www.mediawiki.org/wiki/Wikidata\_Query\_Service/User\_Manual\#Label\_service}.
 
@@
 
-In the end, we have 221 curated queries. An analysis of the basic use of SPARQL features was carried out counting the keywords in the queries (see Table~\ref{tab:Keyword}).
-
-\begin{table}[h]
-	\centering
-	\caption{Keyword count in queries.}
-	\label{tab:Keyword}
-	\scalebox{0.9}{
-		\begin{tabular}{lrr} \hline
-			Element       & Absolute & Relative        \\    \hline
-			SELECT        & 221      & 100.00\%      \\
-			ASK           & 0        & 0.00\%        \\
-			DESCRIBE      & 0        & 0.00\%        \\
-			CONSTRUCT     & 0        & 0.00\%        \\    \hline
-			DISTINCT      & 65       & 29.41\%       \\
-			LIMIT         & 40       & 18.10\%       \\
-			OFFSET        & 0        & 0.00\%        \\
-			ORDER BY      & 221      & 100.00\%      \\    \hline
-			FILTER        & 204      & 92.31\%       \\
-			UNION         & 15       & 6.79\%        \\
-			OPTIONAL      & 182      & 82.35\%       \\
-			NOT EXISTS    & 12       & 5.43\%        \\
-			MINUS         & 4        & 1.81\%        \\
-			EXISTS        & 1        & 0.45\%        \\    \hline
-			COUNT         & 44       & 19.91\%       \\
-			MAX           & 2        & 0.90\%        \\
-			MIN           & 1        & 0.45\%        \\
-			AVG           & 1        & 0.45\%        \\
-			SUM           & 2        & 0.90\%        \\
-			GROUP\_CONCAT & 4        & 1.81\%        \\
-			GROUP BY      & 51       & 23.08\%       \\
-			HAVING        & 3        & 1.36\%        \\    \hline
-	\end{tabular}}
-\end{table}
-
-Once we have shown a basic analysis about the presence of keywords in the queries, we analyze the structure of the triple patterns.
+In the end, we have 221 curated queries.
+Then, we analyze the structure of the triple patterns of the queries.
 This is important, because our notion of freshness focuses on predicates.
-This restricts our approach to triple patterns with a constant in the predicate position.
+This restricts our approach to queries with triple patterns with a constant in the predicate position.
 Table~\ref{tab:Pattern} shows that in total 90.35\% of the triple patterns have a constant in the predicate position, which validate our approach of using a model based on the dynamics of predicates.
@@
 \begin{table}[h]
-	\centering
+	\begin{minipage}[b]{0.45\linewidth}
 		\caption{Triple patterns (C=Constant, V=Variable).}
-	\label{tab:Pattern}
-	\scalebox{0.9}{
 		\begin{tabular}{lrr}    \hline
@@
 			C P C   & 0        & 0.00\%        \\    \hline
-	\end{tabular}}
-\end{table}
-
-Table~\ref{tab:Pattern} also shows that 7.93\% of triple patterns contain a Property Path in the predicate position.
-Then we set out to analyze these triples more deeply.
-Table~\ref{tab:Path} shows the frequency of occurrence of these patterns in our queries.
-It shows that $ P \rightarrow P * $ (74.70\%) and $ P \rightarrow P + $ (9.64\%) are often applied, which corresponds to our intuition to have a feature that indicates if there are triple patterns offering arbitrary length paths (*PATH).
-We should also notice that there is no negation at the level of the path.
-Our feature extractor is able to extract all constants of predicates within paths.
-
-\begin{table}[h]
-	\centering
+		\end{tabular}
+		\label{tab:Pattern}
+	\end{minipage}
+	\hspace{1cm}
+	\begin{minipage}[b]{0.45\linewidth}
 		\caption{Structure of navigational Property Paths.}
 		\label{tab:Path}
-	\scalebox{0.9}{
 		\begin{tabular}{lrr}    \hline
@@
 			$P \rightarrow C $    & 83       & 100.00\%      \\    \hline
-	\end{tabular}}
+		\end{tabular}
+	\end{minipage}
 \end{table}
+
+Table~\ref{tab:Pattern} also shows that 7.93\% of triple patterns contain a Property Path in the predicate position.
+Then we set out to analyze these triples more deeply.
+Table~\ref{tab:Path} shows the frequency of occurrence of these patterns in our queries.
+It shows that $ P \rightarrow P * $ (74.70\%) and $ P \rightarrow P + $ (9.64\%) are often applied, which corresponds to our intuition to have a feature that indicates if there are triple patterns offering arbitrary length paths (*PATH).
+We should also notice that there is no negation at the level of the path.
+Our feature extractor is able to extract all constants of predicates within paths.
 
@@
 
-Figure~\ref{graph:deltas} shows an analysis of the query result monotonicity in the evaluated period.
-Red dots mean only removed results, greens added, and blues added and removed.
-It can be clearly seen that query results are not monotonic.
-
-\begin{figure}[h]	
-	\begin{minipage}[b]{0.40\linewidth}
-		\centering	
-		\begin{tikzpicture}[thick, scale=0.5]			
-		\begin{axis}[
-		%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
-		xlabel={Version},
-		ylabel={Changes},
-		ymin=0,
-		xtick={0,5,10,15,20,25},
-		legend pos=north west,
-		ymajorgrids=true,
-		grid style=dashed,
-		]
-
-		\addplot[
-		color=green,
-		mark=square,
-		]
-		coordinates {
-			(1,14)(2,38)(3,27)(4,24)(5,32)(6,23)(7,36)(8,21)(9,24)(10,23)
-		};
-
-		\addplot[
-		color=green,
-		mark=square,
-		]
-		coordinates {
-			(12,26)(13,26)(14,27)(15,29)(16,23)(17,25)(18,20)(19,24)(20,13)(21,18)(22,31)
-		};
-
-		\addplot[
-		color=red,
-		mark=square,
-		]
-		coordinates {
-			(1,8)(2,5)(3,5)(4,8)(5,6)(6,7)(7,5)(8,9)(9,9)(10,6)
-		};
-
-		\addplot[
-		color=red,
-		mark=square,
-		]
-		coordinates {
-			(12,3)(13,7)(14,7)(15,8)(16,7)(17,6)(18,5)(19,6)(20,7)(21,6)(22,10)
-		};
-
-		\addplot[
-		color=blue,
-		mark=square,
-		]
-		coordinates {
-			(1,93)(2,69)(3,86)(4,84)(5,83)(6,81)(7,86)(8,79)(9,85)(10,86)
-		};
-
-		\addplot[
-		color=blue,
-		mark=square,
-		]
-		coordinates {
-			(12,76)(13,80)(14,84)(15,80)(16,81)(17,87)(18,87)(19,86)(20,86)(21,82)(22,88)
-		};
-
-		%\legend{# entities}
-		%\legend{# predicates}
-		%\legend{# literals}
-		%\legend{# blank nodes}			
-		\end{axis}
-		\end{tikzpicture}				
-		\label{graph:deltaV}
-	\end{minipage}
-	\hspace{0.7cm}
-	\begin{minipage}[b]{0.40\linewidth}
-		\centering
-		\begin{tikzpicture}[thick, scale=0.5]
-		\begin{axis}[
-		%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
-		xlabel={Queries},
-		ymin=0,
-		legend pos=north west,
-		ymajorgrids=true,
-		grid style=dashed,
-		]
-
-		\addplot[
-		color=red,
-		mark=square,
-		]
-		coordinates {
-			(1,8)(2,6)(3,6)(4,5)(5,5)(6,5)(7,4)(8,4)(9,4)(10,4)(11,4)(12,4)(13,4)(14,4)(15,3)(16,3)(17,3)(18,3)(19,3)(20,2)(21,2)(22,2)(23,2)(24,2)(25,2)(26,2)(27,2)(28,2)(29,2)(30,2)(31,1)(32,1)(33,1)(34,1)(35,1)(36,1)(37,1)(38,1)(39,1)(40,1)(41,1)(42,1)(43,1)(44,1)(45,1)(46,1)(47,1)(48,1)(49,1)(50,1)(51,1)(52,1)(53,1)(54,1)(55,1)(56,1)(57,1)(58,1)(59,1)(60,1)(61,1)(62,1)(63,1)(64,1)(65,1)(66,1)(67,1)(68,1)(69,1)(70,1)(71,1)(72,1)(73,1)(74,0)(75,0)(76,0)(77,0)(78,0)(79,0)(80,0)(81,0)(82,0)(83,0)(84,0)(85,0)(86,0)(87,0)(88,0)(89,0)(90,0)(91,0)(92,0)(93,0)(94,0)(95,0)(96,0)(97,0)(98,0)(99,0)(100,0)(101,0)(102,0)(103,0)(104,0)(105,0)(106,0)(107,0)(108,0)(109,0)(110,0)(111,0)(112,0)(113,0)(114,0)(115,0)(116,0)(117,0)(118,0)(119,0)(120,0)(121,0)(122,0)(123,0)(124,0)(125,0)(126,0)(127,0)(128,0)(129,0)(130,0)(131,0)(132,0)(133,0)(134,0)(135,0)(136,0)(137,0)(138,0)(139,0)(140,0)(141,0)(142,0)(143,0)(144,0)(145,0)(146,0)(147,0)(148,0)(149,0)(150,0)(151,0)(152,0)(153,0)(154,0)(155,0)(156,0)(157,0)(158,0)(159,0)(160,0)(161,0)(162,0)(163,0)(164,0)(165,0)(166,0)(167,0)(168,0)(169,0)(170,0)(171,0)(172,0)(173,0)(174,0)(175,0)(176,0)(177,0)(178,0)(179,0)(180,0)(181,0)(182,0)(183,0)(184,0)(185,0)(186,0)(187,0)(188,0)(189,0)(190,0)(191,0)(192,0)(193,0)(194,0)(195,0)(196,0)(197,0)(198,0)(199,0)(200,0)(201,0)(202,0)(203,0)(204,0)(205,0)(206,0)(207,0)(208,0)(209,0)(210,0)(211,0)(212,0)(213,0)(214,0)(215,0)(216,0)(217,0)(218,0)(219,0)(220,0)(221,0)
-		};
-
-		\addplot[
-		color=green,
-		mark=square,
-		]
-		coordinates {
-			(1,13)(2,13)(3,13)(4,12)(5,11)(6,11)(7,10)(8,10)(9,10)(10,10)(11,9)(12,9)(13,9)(14,8)(15,8)(16,8)(17,8)(18,8)(19,8)(20,7)(21,7)(22,7)(23,7)(24,7)(25,7)(26,7)(27,7)(28,7)(29,7)(30,7)(31,6)(32,6)(33,6)(34,6)(35,6)(36,6)(37,6)(38,6)(39,6)(40,6)(41,6)(42,5)(43,5)(44,5)(45,5)(46,5)(47,5)(48,5)(49,5)(50,4)(51,4)(52,4)(53,4)(54,4)(55,4)(56,4)(57,4)(58,4)(59,4)(60,4)(61,4)(62,4)(63,3)(64,3)(65,3)(66,3)(67,3)(68,3)(69,3)(70,3)(71,3)(72,3)(73,3)(74,3)(75,3)(76,3)(77,3)(78,3)(79,2)(80,2)(81,2)(82,2)(83,2)(84,2)(85,2)(86,2)(87,2)(88,2)(89,2)(90,2)(91,2)(92,2)(93,2)(94,2)(95,2)(96,2)(97,2)(98,2)(99,2)(100,2)(101,2)(102,2)(103,2)(104,1)(105,1)(106,1)(107,1)(108,1)(109,1)(110,1)(111,1)(112,1)(113,1)(114,1)(115,1)(116,1)(117,1)(118,1)(119,1)(120,1)(121,1)(122,1)(123,1)(124,1)(125,1)(126,1)(127,1)(128,1)(129,1)(130,1)(131,0)(132,0)(133,0)(134,0)(135,0)(136,0)(137,0)(138,0)(139,0)(140,0)(141,0)(142,0)(143,0)(144,0)(145,0)(146,0)(147,0)(148,0)(149,0)(150,0)(151,0)(152,0)(153,0)(154,0)(155,0)(156,0)(157,0)(158,0)(159,0)(160,0)(161,0)(162,0)(163,0)(164,0)(165,0)(166,0)(167,0)(168,0)(169,0)(170,0)(171,0)(172,0)(173,0)(174,0)(175,0)(176,0)(177,0)(178,0)(179,0)(180,0)(181,0)(182,0)(183,0)(184,0)(185,0)(186,0)(187,0)(188,0)(189,0)(190,0)(191,0)(192,0)(193,0)(194,0)(195,0)(196,0)(197,0)(198,0)(199,0)(200,0)(201,0)(202,0)(203,0)(204,0)(205,0)(206,0)(207,0)(208,0)(209,0)(210,0)(211,0)(212,0)(213,0)(214,0)(215,0)(216,0)(217,0)(218,0)(219,0)(220,0)(221,0)
-		};
-
-		\addplot[
-		color=blue,
-		mark=square,
-		]
-		coordinates {
-			(1,22)(2,22)(3,22)(4,22)(5,22)(6,22)(7,22)(8,22)(9,22)(10,22)(11,22)(12,22)(13,22)(14,22)(15,22)(16,22)(17,22)(18,22)(19,22)(20,22)(21,22)(22,22)(23,22)(24,22)(25,22)(26,22)(27,22)(28,22)(29,22)(30,22)(31,22)(32,21)(33,21)(34,21)(35,21)(36,21)(37,21)(38,21)(39,21)(40,21)(41,20)(42,20)(43,20)(44,20)(45,20)(46,19)(47,19)(48,19)(49,19)(50,18)(51,18)(52,18)(53,18)(54,16)(55,16)(56,15)(57,15)(58,15)(59,15)(60,15)(61,15)(62,15)(63,14)(64,14)(65,13)(66,13)(67,12)(68,11)(69,11)(70,11)(71,11)(72,11)(73,11)(74,11)(75,11)(76,11)(77,10)(78,10)(79,10)(80,10)(81,9)(82,9)(83,9)(84,9)(85,8)(86,8)(87,8)(88,8)(89,7)(90,7)(91,7)(92,7)(93,7)(94,7)(95,7)(96,7)(97,7)(98,7)(99,7)(100,7)(101,6)(102,6)(103,6)(104,6)(105,6)(106,6)(107,6)(108,6)(109,6)(110,6)(111,5)(112,5)(113,5)(114,5)(115,5)(116,5)(117,5)(118,5)(119,4)(120,4)(121,4)(122,4)(123,4)(124,4)(125,4)(126,4)(127,4)(128,4)(129,4)(130,4)(131,4)(132,4)(133,4)(134,3)(135,3)(136,3)(137,3)(138,3)(139,3)(140,3)(141,2)(142,2)(143,2)(144,2)(145,2)(146,2)(147,2)(148,2)(149,2)(150,2)(151,2)(152,2)(153,2)(154,2)(155,1)(156,1)(157,1)(158,1)(159,1)(160,1)(161,1)(162,1)(163,1)(164,1)(165,1)(166,1)(167,1)(168,1)(169,1)(170,1)(171,1)(172,1)(173,1)(174,1)(175,1)(176,1)(177,1)(178,1)(179,1)(180,0)(181,0)(182,0)(183,0)(184,0)(185,0)(186,0)(187,0)(188,0)(189,0)(190,0)(191,0)(192,0)(193,0)(194,0)(195,0)(196,0)(197,0)(198,0)(199,0)(200,0)(201,0)(202,0)(203,0)(204,0)(205,0)(206,0)(207,0)(208,0)(209,0)(210,0)(211,0)(212,0)(213,0)(214,0)(215,0)(216,0)(217,0)(218,0)(219,0)(220,0)(221,0)
-		};
-		%\legend{# entities}
-		%\legend{# predicates}
-		%\legend{# literals}
-		%\legend{# blank nodes}
-
-		\end{axis}
-		\end{tikzpicture}
-		\label{graph:deltaH}
-	\end{minipage}
-	\caption{Types of changes by version and query. Only addition of results (green), only elimination of results (red), or both (blue).}
-	\label{graph:deltas}
-\end{figure}
-
-\begin{comment}
-\tikzset{mark options={mark size=0.5}}
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}
-\begin{axis}[%
-clickable coords={(xy): \thisrow{label}},%
-xlabel={Time},
-scatter/classes={%
-c={blue},%+-3
-b={red},%-2
-a={green}}]%+1
-\addplot[scatter,only marks,%
-scatter src=explicit symbolic]%
-table[meta=label] {treschange.dat};
-\end{axis}
-\end{tikzpicture}
-\caption{Query results monotony.}
-\label{graph:changeT}
-\end{figure}
-\end{comment}
-
 \section{Evaluation}
@@
 
-\subsection{Measures}
-
-Depending on whether the query results really changed, we consider four possible outcomes of this decision:
-
-\begin{enumerate}[label=\roman*]
-	\item \textbf{True positive (tp):} Changed results are detected. Our proposal is correct in its prediction.
-	\item \textbf{False positive (fp):} No changed results are not detected. Our proposal is incorrect in its prediction.
-	\item \textbf{True negative (tn):} No changed results are detected. Our proposal is correct in its prediction.
-	\item \textbf{False negative (fn):} Changed results are not detected. Our proposal is incorrect in its prediction.
-
-\end{enumerate}
-
-We measured precision (see Equation~\ref{eq:precision}), recall (see Equation~\ref{eq:recall}) and f-measure ($f_1$) (see Equation~\ref{eq:f1}).
-
-\begin{equation}
-\label{eq:precision}
-precision = \frac{tp}{tp + fp}
-\end{equation}
-
-\begin{equation}
-\label{eq:recall}
-recall = \frac{tp}{tp + fn}
-\end{equation}
-
-\begin{equation}
-\label{eq:f1}
-f_1 = 2 \times \frac{precision \times recall}{precision + recall}
-\end{equation}
+In addition, we train a Machine Learning model to determine the life spam of a query results.
+To train the model we consider the changes observed in the monitored interval.
+For each update of the results of a query, we have created an example of training with the previously presented features.
+The target TTL value that we learn is the time interval during which the results of the query remain fresh from the current update time.
 

### 2019-01-23 kgswc19.tex
removed pages from references added TTL to the schema picture source updated

@@
 %
-\author{Alberto Moya\orcidID{0000-1111-2222-3333} \and
+\author{Alberto Moya\orcidID{0000-0002-7003-5087} \and
 Aidan Hogan\orcidID{1111-2222-3333-4444}}% \and
@@
 
-In summary, the \textbf{contributions} of this paper are: (i) we build a dataset of real queries for Wikidata, (ii) we provide an analysis of the evolution of the results of those queries over 23 weekly versions of Wikidata, (iii) we evaluated several classifiers to judges whether or not the results of a query will change regression techniques to determine specifically when a query results will change using a predicate-based Dynamics Model and features from the syntax of the query, and (iv) we show that the proposed features and classifier work well, and our results are close to the more expensive previous approaches that assume full knowledge of historical data.
+In summary, the \textbf{contributions} of this paper are: (i) we build a dataset of real queries for Wikidata, (ii) we provide an analysis of the evolution of the results of those queries over 23 weekly versions of Wikidata, (iii) we evaluated several classifiers to judges whether or not the results of a query will change and regression techniques to determine specifically when a query results will change using a predicate-based Dynamics Model and features from the syntax of the query, and (iv) we show that the proposed features and classifier work well, and our results are close to the more expensive previous approaches that assume full knowledge of historical data.
 
@@
 We tested a list of classifiers, including Decision Tree, Nearest Neighbors, and Naive Bayes among others listed in Table~\ref{tab:classifier}.
-Our pilot experiment shows that Nearest Neighbors, Gradient Boosting Classifier, and Random Forest work better than the others.
-We evaluated the quality of our proposal by forming a set of classifiers to predict the change of 221 queries in 15 weeks, which is equivalent to 3315 objects to classify, described in terms of our proposed characteristics.
+Our pilot experiment shows that Nearest Neighbors and Decision Tree work better than the others.
+We evaluated the quality of our proposal by forming a set of classifiers to predict the change of 221 queries in 15 weeks, which is equivalent to 3315 objects to classify, described in terms of our proposed features.
 We split the data 80\% for training and 20\% for tests.
@@
 Table~\ref{tab:classifier} shows the quality values of the classifiers evaluated in terms of $F_1$.
-In the state of the art, most of the works use information of the history of the results of the queries, which requires storing previous complete past versions to evaluate the queries.
-To have a measure of the positioning of our proposal, we evaluate the incorporation of a new feature that counts how many times the results of the query change on average in the last three versions.
-Then, we did two executions: one with our syntactic characteristics and this new indicator, which we call the past, and a second with our syntactic characteristics and information about dynamic predicates from our model.
+To have a measure of the positioning of our proposal, we added a new feature that counts how many times the results of the query change on average in the last three versions.
+Then, we did two executions with our syntactic features: one using this new indicator, and the second with our dynamic model.
 Table~\ref{tab:classifier} shows that our algorithm achieved less quality.
@@
 
-Subsequently, we individually evaluate the classification performance using different groups of characteristics to determine the influence of each characteristic.
-Table~\ref{tab:features} shows the quality when each feature is not used: a lower value indicates that the characteristic has a large positive influence.
-For this, we use the Nearest Neighbors as classifier and Average as the Transformation Function.
-We can notice a slight fall in the feature Dyn and Negation.
+Finally, to predict the time that the results of a query will remain invariant (TTL), a continuous value (number of weeks) will be determined.
+For that we employed techniques of multiple linear regression using the sklearn and tensorflow library.
+The latter allows us to see how the periods of estimation and the convergence analysis will behave.
 
-Finally, we show in Figure~\ref{graph:Modelsize} the number of triples in Wikidata versions vs the number of triples in the Difference Sets, in which our Dynamics Model are based.
-It is clearly visible that our model requires much less storage space.
+We were able to predict the time of life (in relation to the number of invariant weeks) with an average square error of 0.43 and a variance score of 0.01.
+These results allow to predict future changes in query results with good accuracy using linear regression techniques.
 
@@
 	\label{tab:classifier}
-	\begin{tabular}{|l|lll|lll|}    \hline
-		& \multicolumn{3}{|l|}{Past Results}           & \multicolumn{3}{|l|}{Dynamics Model}        \\    \hline
-		Classifier                   & Precision & Recall & F1 & Precision & Recall & F1 \\    \hline
-		Nearest Neighbors            & 0.827869 & 0.776923 & 0.801587 & 0.742188 & 0.730769 & 0.736434 \\
-		Gradient Boosting Classifier & 0.848233 & 0.784615 & 0.815185 & 0.734127 & 0.711538 & 0.722656 \\
-		Random Forest                & 0.818548 & 0.780769 & 0.799213 & 0.725146 & 0.715385 & 0.720232 \\
-		Decision Tree                & 0.830544 & 0.763462 & 0.795591 & 0.700382 & 0.705769 & 0.703065 \\
-		Linear SVM                   & 0.865342 & 0.753846 & 0.805755 & 0.735808 & 0.648077 & 0.689162 \\
-		Neural Net                   & 0.836364 & 0.796154 & 0.815764 & 0.693989 & 0.488462 & 0.573363 \\
-		Logistic Regression          & 0.836032 & 0.794231 & 0.814596 & 0.616253 & 0.525000    & 0.566978 \\
-		Naive Bayes                  & 0.792100   & 0.732692 & 0.761239 & 0.690323 & 0.411538 & 0.515663 \\    \hline
+	\begin{tabular}{|l|r|r|}    \hline		
+		Classifier                   & Past Results & Dynamics Model \\    \hline
+		Base Dummy 					 & 0.497869 & 0.486923 \\
+		Nearest Neighbors            & 0.701587 & 0.635923 \\
+		Decision Tree                & 0.695591 & 0.651462 \\
+		Linear SVM                   & 0.705755 & 0.593846 \\
+		Naive Bayes                  & 0.661239 & 0.587692 \\    \hline
 	\end{tabular}
 \end{table*}
+\begin{comment}
 
-\begin{table*}[]
+\begin{figure}
 	\centering
-	\caption{Feature analysis using Nearest Neighbors.}
-	\label{tab:features}
-	\begin{tabular}{|l|rrr|}        \hline
-		Feature           & Precision & Recall & F1 \\    \hline
-		\# triples        & 0.741414 & 0.705769 & 0.723152  \\
-		\# vars           & 0.738878 & 0.714615 & 0.726544 \\
-		\# predicates    & 0.732558 & 0.726923 & 0.729730 \\
-		\# projected vars & 0.726744 & 0.721154 & 0.723938 \\
-		FILTER            & 0.738878 & 0.714615 & 0.726544 \\
-		LIMIT             & 0.732558 & 0.726923 & 0.729730 \\
-		UNION             & 0.733640  & 0.730769 & 0.732201 \\
-		SUBQUERY         & 0.732558 & 0.736923 & 0.734734  \\
-		*PATH             & 0.732558 & 0.726923 & 0.729730 \\
-		NEGATION          & 0.713462 & 0.713462 & 0.713462  \\
-		DYN               & 0.706046 & 0.711538 & 0.708781    \\ \hline
-	\end{tabular}
-\end{table*}
+	\begin{tikzpicture}[thick, scale=1]
+	\begin{axis}
 
-\begin{figure}[h]
-	\centering
-	\begin{tikzpicture}[thick, scale=0.5]
-	\begin{axis}[
-	%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
-	xlabel={Version},
-	ymin=0,
-	xtick={0,5,10,15,20,25},
-	legend pos=north west,
-	ymajorgrids=true,
-	grid style=dashed,
-	]
-
-	\addplot[
-	color=red,
-	mark=square,
-	]
-	coordinates {
-		(1,1102242331)(2,1138975810)(3,1141303957)(4,1170283653)(5,1176093821)(6,1184177503)(7,1190456714)(8,1211813217)(9,1239145151)(10,1271353260)(11,1293099057)
-	};
-
-	\addplot[
-	color=red,
-	mark=square,
-	]
-	coordinates {
-		(13,1348622441)(14,1418460320)(15,1432143355)(16,1470298276)(17,1506140622)(18,1542278452)(19,1593334993)(20,1613832004)(21,1684774955)(22,1771601730)(23,1843086909)(24,1924967162)
-	};
-
-	\addplot[
-	color=blue,
-	mark=square,
-	]
-	coordinates {
-		(2,43540931)(3,6798650)(4,30419079)(5,12002672)(6,14826495)(7,16592201)(8,33102669)(9,28305029)(10,45696749)(11,21282917)
-	};
-
-	\addplot[
-	color=blue,
-	mark=square,
-	]
-	coordinates {
-		(13,71196531)(14,21688658)(15,52656689)(16,46556227)(17,49129201)(18,62311275)(19,30121460)(20,25820733)(21,23229489)(22,24525255)(23,38007899)
-	};
-
-	%\legend{# entities}
-	%\legend{# predicates}
-	%\legend{# literals}
-	%\legend{# blank nodes}
-
+	\addplot 
+	coordinates {("Base Dummy",40137188)("Nearest Neighbors",4563382)("Decision Tree",29302968)("Linear SVM",8906402)("Naive Bayes",11455047)};
+	\addplot 
+	coordinates {(Base Dummy,40137188)(Nearest Neighbors,4563382)(Decision Tree,29302968)(Linear SVM,8906402)(Naive Bayes,11455047)};
+	\legend{Added, Removed}
 	\end{axis}
 	\end{tikzpicture}
-	\caption{Number of triples in Wikidata vs number of triples in Difference Sets.}
-	\label{graph:Modelsize}
+	\caption{Triples added and removed in Wikidata.}
+	\label{graph:delta}
 \end{figure}
+\end{comment}
 
@@
 
-In this paper we propose classifiers that predict whether or not the results of a query will change in the short time interval; for this, we propose a model based on the dynamics of the predicates in the datasets and a set of characteristics extracted from the syntax of the queries.
+In this paper we propose classifiers that predict whether or not the results of a query will change in the short time interval; for this, we propose a model based on the dynamics of the predicates in the datasets and a set of features extracted from the syntax of the queries.
 The proposed framework is flexible for incorporating new features extracted from the queries, from the current results, or from the model based on the multiplicity of predicates.
@@
 
-Finally, we evaluate our proposal, making a weekly prediction of change in the results of the queries and evaluating the success rates of the different classifiers each week based on precision, recall, and F-measure.
+Finally, we evaluate our proposal, making a weekly prediction of change in the results of the queries and evaluating the success rates of the different classifiers each week based on F-measure.
 We also apply linear regression techniques to predict, specifically, when the results of a query will change.
@@
 Our results show that our approach using predicate dynamics is competitive for prediction while being less expensive to support.
-An analysis of the sensitivity of the model to the different characteristics was also carried out, finding that our characteristics based on the dynamics of the predicates have the greatest influence.
 
@@
 We consider for future work evaluating the quality of our proposal over a longer period.
-In addition, we believe that more complex query characteristics can improve the results, such as considering operator semantics; for example, to see if negation is affecting predicates or if the predicates are affected by the existence of filters.
+In addition, we believe that more complex query features can improve the results, such as considering operator semantics; for example, to see if negation is affecting predicates or if the predicates are affected by the existence of filters.
 On the other hand, it would be worth expanding the set of queries in the test dataset, in order to increase the size, complexity, and coverage.
 Recently a query dataset for Wikidata was published\footnote{https://lists.w3.org/Archives/Public/semantic-web/2018Aug/0023.html}.
-It would be good to evaluate this query dataset, as well as to evaluate the sensitivity of our proposal to consider longer intervals and different Transformation Functions.
-
-Finally, we would like to evaluate the quality of the characteristics and the proposed model to estimate the current Time-to-Live of the queries, rather than a binary One-Step-Change classification.
+It would be good to evaluate this query dataset, as well as consider a probabilistic approach based on query algebra (rather than syntax).
 

### 2019-01-25 kgswc19.tex
Aidan revisions to title and intro

@@
 
+\newcommand{\ah}[1]{{\color{blue}\textsc{ah:} #1}}
+
 \usepackage{listings}
@@
 %
-\title{Change estimation of SPARQL queries in Wikidata\thanks{Supported by CONICYT PFCHA/ DOCTORADO BECAS CHILE/2017 - 72180000 and Millennium Institute for Foundational Research on Data.}}
+\title{Estimating the dynamics of SPARQL query results over Wikidata\thanks{Supported by CONICYT PFCHA/ DOCTORADO BECAS CHILE/2017 - 72180000 and Millennium Institute for Foundational Research on Data.}}
 %
@@
 %
-\author{Alberto Moya\orcidID{0000-0002-7003-5087} \and
-Aidan Hogan\orcidID{1111-2222-3333-4444}}% \and
-%Jesus P\'erez\orcidID{2222--3333-4444-5555} \and
-%Grethel Coello\orcidID{2222--3333-4444-5555}}
+\author{Alberto Moya \and
+Aidan Hogan\orcidID{0000-0001-9482-1982}}
 %
-\authorrunning{A. Moya et al.}
+\authorrunning{A. Moya and A. Hogan}
 % First names are abbreviated in the running head.
@@
 %
-\institute{Computer Science Department, University of Chile, Chile\\
+\institute{IMFD; DCC, University of Chile\\
 \email{\{amoya,ahogan\}@dcc.uhile.cl}}
@@
 \begin{abstract}
-Linked Open Datasets contain descriptions that change over time.
-The resources described by these datasets are continuously created, moved, deleted, linked, and unlinked.
-Applications that leverage Linked Data must be aware of these change dynamics to deliver accurate services.
-In this work, we tackle the problem of estimating future changes in query results based on query syntax features.
-We propose a model that exploits the characteristics of RDF to capture the dynamics of the predicates.
-Our Dynamics Model complements syntax features extracted from queries to assess if query results will change in the near future using classifiers and long future using Multiple Linear Regression techniques.
-To evaluate our proposal, we select Wikidata as our dataset and we build a curated collection of queries.
-Our results show that the accuracy of our proposal approaches that obtained using information on the historical evolution of the query results but at a lower cost.
+In this work, we tackle the problem of estimating future changes in SPARQL query results over dynamic datasets; such work has applications for caching and remote synchronization.
+Our proposal is based on syntactic query features and (optionally) knowledge of the dynamics of predicates in the underlying dataset; the resulting model can be used to predict (1) whether or not the results for a query will change in the near future using classifiers, and (2) when they might change (if at all) using linear regression.
+To evaluate our proposal, we select Wikidata as our dataset and build a curated collection of real-world SPARQL queries.
+Our results show that the accuracy of the predictions of our proposed model is competitive with that obtained using knowledge of the complete historical evolution of the query results but at a lower cost.
 
-\keywords{Change estimation \and SPARQL query \and Link Data dynamics.}
+\keywords{SPARQL \and Linked Data \and Dynamics \and Wikidata}
 \end{abstract}
@@
 %
-Many applications that consume Linked Data (LD) face challenges related to changes in the underlying data, whose key is keeping data fresh.
-Since datasets change over time, long-running applications that cache and repeatedly use query results obtained from a SPARQL endpoint may resubmit the queries regularly to ensure up-to-dateness.
-As a result, applications often get the same results in several query executions, resulting in an unnecessary computation, and other times, they give the user stale results when the backend data change between re-executions.
+Many applications that consume Linked Data (LD) face challenges related to remote changes in the underlying data.
+By caching remote data and using the cache whenever possible, client-side caches reduce the network traffic between clients and servers, the load on servers, and the average user-perceived latency of responses.
+However, since datasets change over time, for caches to be useful, cached copies should be updated when the underlying data that they reflect change; predicting such remote changes, however, is a challenging problem, particularly when data are accessed through a SPARQL endpoint.
 
-By caching the answers of queries and using the cached replies whenever possible, client-side caches reduce the network traffic between clients and servers, the load on servers, and the average user-perceived latency of query processing.
-However, for caches to be useful, cached copies should be updated when the data that gave rise to it change.
-Ensuring a strong consistency in which no stale copy of the modified document will ever be returned to the user could be expensive.
-Thus, several weak consistency approaches have been proposed that try to keep the local data of the applications updated at lower cost~\cite{KnuthHS16,DividinoGS15,NishiokaS17}.
+Since datasets change over time, long-running applications that cache and repeatedly use query results obtained from an external SPARQL endpoint may resubmit the queries regularly to ensure up-to-dateness.
+As a result, without further information as to the dynamics of a particular SPARQL query, applications face the choice of either performing frequent query executions that may be redundant and repeatedly return the same results (aiming for stronger consistency at the cost of more frequent updates), or performing infrequent query executions that may lead to stale data being persisted in the application when the underlying sources change (accepting weaker consistency at the cost of less fresh data).
+Given the costs on clients and servers of repetitive requests served over the Web  and the potential efficiency gains offered by local caches, several weak consistency approaches have been proposed that try to keep the local data of the applications updated at lower cost by predicting changes~\cite{KnuthHS16,DividinoGS15,NishiokaS17}.
 
-Some works have proposed to learn from historical entity evolution, but a query could involve many entities.
-Considering the historical evolution of all entities relating to a query is expensive because it is necessary to have the previous complete versions of data and to obtain the results of the execution of the query in each version.
-In addition, recent works have explored the dynamics of the LD with the intention of finding patterns that allow for characterizing, recognizing, and predicting changes~\cite{KaferAUOH13,NishiokaS16,GonzalezH18} finding some general correlations, such as, the domains, predicates, and schema~\cite{KaferAUOH13,NishiokaS17,UmbrichHHPD10}.
-Based on dynamic knowledge, hybrid approaches have been developed to return fresher query results and at the same time speed up queries, but freshness estimates are a major and still current challenge, because of the difficulty of predicting the underlying dynamics on the growing Web of Data.
+Some works \ah{Which?
+Cite at the end of the sentence} have proposed to predict dynamics based on the historical evolution of entities, but following such an approach for SPARQL queries is expensive because (1) a SPARQL query may involve potentially many entities; and (2) it is necessary to have the previous complete versions of data to analyse the entities relevant to a query.
+Recent works have explored the dynamics of Linked Data with the intention of finding patterns that allow for characterizing, recognizing, and predicting changes based on analysis of the domains, predicates, and schema~\cite{UmbrichHHPD10,KaferAUOH13,NishiokaS16,NishiokaS17,GonzalezH18}.
+Among these works, we can find hybrid approaches that have been developed to return fresher query results with faster response times by decomposing a query into dynamic sub-queries answered remotely, and static sub-queries answer over local caches~\cite{UmbrichKHP12}, but such an approach only focuses on estimating the dynamics of individual triple patterns, and does not consider, for example, query operators.
+More generally, coping with changes in remote data still presents a major challenge for applications leveraging dynamic Linked Data.
 
-In this work, we tackle the problem of estimating future changes in query results based on data statistics and methods from Machine Learning.
-We devise a solution where we do not have complete information from the past, but a reduced statistical model that summarizes the historical dynamics based on predicates.
-We compute a predictive Dynamics Model for 23 versions of the Wikidata knowledge graph~\cite{VrandecicK14}, evaluating its suitability for predicting future changes in the query results.
-We select Wikidata as: (1) it provides a history of weekly versions that we can use to evaluate predictions and (2) it is edited by thousands of users, which means that significant changes are observed weekly.
+In this work, we tackle the problem of estimating future changes in query results based on query features, high-level statistics about the dynamics of predicates, as well as classification and regression methods from Machine Learning.
+We thus propose a solution where we do not have complete historical information about the results of a query, but rather a much more lightweight statistical model that saves on space and obviates the need to compute query results for multiple past versions.
+In order to assess the predictions made by our proposal, we perform experiments over 23 weekly versions of the Wikidata knowledge graph~\cite{VrandecicK14}, evaluating various configurations in terms of the accuracy for predicting future changes in the query results; we select Wikidata as: (1) it provides a history of weekly versions that we can use to evaluate predictions; (2) it is edited by thousands of users, which means that significant changes are observed weekly; (3) it provides a large-scale diverse dataset; (4) it receives millions of queries on a daily basis~\cite{MalyshevKGGB18}; and (5) a diverse set of exemplar SPARQL queries has been proposed by the community that we can use for our evaluation.
 
-In summary, the \textbf{contributions} of this paper are: (i) we build a dataset of real queries for Wikidata, (ii) we provide an analysis of the evolution of the results of those queries over 23 weekly versions of Wikidata, (iii) we evaluated several classifiers to judges whether or not the results of a query will change and regression techniques to determine specifically when a query results will change using a predicate-based Dynamics Model and features from the syntax of the query, and (iv) we show that the proposed features and classifier work well, and our results are close to the more expensive previous approaches that assume full knowledge of historical data.
+The contributions of this paper are thus as follows: (i) we propose a supervised method that uses static query features and suitable classifiers to predict whether or not results will change in a given point in the near future; (ii) we propose to augment this model with lightweight statistics of the dynamics of predicates; (iii) we further experiment with linear regression models to predict when results will change; (iv) we prepare a benchmark for predicting the dynamics of query results based on 23 weekly versions of Wikidata and a set of \ah{XXX} real-world SPARQL queries; (v) we use this benchmark to compare the results obtained by our lightweight models versus an ``ideal model'' that knows of changes in the query's results for previous versions.
+Our results show that our proposed models are competitive with the more expensive ideal model.
 
 The rest of this document is organized as follows: In Section~\ref{sec:related}, we review related works.
-Section~\ref{sec:preliminar} provides the formal definition of the problem.
-Section~\ref{sec:approach} describes our proposal.
-In Section~\ref{sec:data}, we explain our datasets.
+Section~\ref{sec:preliminar} provides a formal definition of the problem.
+Section~\ref{sec:approach} describes our proposed models.
+In Section~\ref{sec:data}, we describe our evaluation data and queries based on Wikidata.
 Section~\ref{sec:eval} presents the design and results of the experiments.

### 2019-01-26 kgswc19.tex
edits up to section 4

@@
 
+\newcommand{\Iris}{\ensuremath{\mathbf{I}}}
+\newcommand{\Lits}{\ensuremath{\mathbf{L}}}
+\newcommand{\Bnodes}{\ensuremath{\mathbf{B}}}
+\newcommand{\Vars}{\ensuremath{\mathbf{V}}}
+
+\newcommand{\da}{\ensuremath{:\nolinebreak\mkern-1.2mu\nolinebreak=}}
+
 \newcommand{\ah}[1]{{\color{blue}\textsc{ah:} #1}}
+
+\usepackage{xcolor}
+\usepackage{hyperref}
+\definecolor{dark-blue}{rgb}{0.0,0.0,0.2}
+\definecolor{dark-green}{rgb}{0.0,0.2,0.0}
+\definecolor{dark-red}{rgb}{0.2,0.0,0.0}
+\hypersetup{
+    colorlinks, linkcolor={dark-red},
+    citecolor={dark-green}, urlcolor={dark-blue},
+    pdftitle={Estimating the dynamics of SPARQL query results over Wikidata},    % title
+    pdfauthor={Alberto Moya, Aidan Hogan},     % author
+    pdfsubject={KGSWC 2019},   % subject of the document
+    pdfkeywords={SPARQL;} {Linked Data;} {Dynamics;} {Wikidata;}, % list of keywords
+}
 
@@
 %
-\title{Estimating the dynamics of SPARQL query results over Wikidata\thanks{Supported by CONICYT PFCHA/ DOCTORADO BECAS CHILE/2017 - 72180000 and Millennium Institute for Foundational Research on Data.}}
+\title{Estimating the dynamics of\\SPARQL query results over Wikidata\thanks{This work was supported by CONICYT PFCHA, Doctorado Becas Chile/2017 -- 72180000, the Millennium Institute for Foundational Research on Data (IMFD) and FONDECYT Grant No.
+1181896.}}
 %
@@
 Many applications that consume Linked Data (LD) face challenges related to remote changes in the underlying data.
-By caching remote data and using the cache whenever possible, client-side caches reduce the network traffic between clients and servers, the load on servers, and the average user-perceived latency of responses.
-However, since datasets change over time, for caches to be useful, cached copies should be updated when the underlying data that they reflect change; predicting such remote changes, however, is a challenging problem, particularly when data are accessed through a SPARQL endpoint.
+Client-side caches can reduce the network traffic between clients and servers, the load on servers, and the average latency of responses.
+However, since datasets change over time, for caches to be useful, they should be updated when the underlying data that they reflect change; predicting such remote changes, however, is a challenging problem, particularly when data are accessed as the results of queries to a SPARQL endpoint.
 
 Since datasets change over time, long-running applications that cache and repeatedly use query results obtained from an external SPARQL endpoint may resubmit the queries regularly to ensure up-to-dateness.
-As a result, without further information as to the dynamics of a particular SPARQL query, applications face the choice of either performing frequent query executions that may be redundant and repeatedly return the same results (aiming for stronger consistency at the cost of more frequent updates), or performing infrequent query executions that may lead to stale data being persisted in the application when the underlying sources change (accepting weaker consistency at the cost of less fresh data).
+As a result, without further information as to the dynamics of a particular SPARQL query, applications face the choice of either performing frequent query executions that may be redundant and repeatedly return the same results (aiming for stronger consistency at the cost of more frequent updates), or performing infrequent query executions that may lead to stale data being persisted in the application when the underlying sources change (accepting weaker consistency to improve efficiency/scalability).
 Given the costs on clients and servers of repetitive requests served over the Web  and the potential efficiency gains offered by local caches, several weak consistency approaches have been proposed that try to keep the local data of the applications updated at lower cost by predicting changes~\cite{KnuthHS16,DividinoGS15,NishiokaS17}.
@@
 
-The contributions of this paper are thus as follows: (i) we propose a supervised method that uses static query features and suitable classifiers to predict whether or not results will change in a given point in the near future; (ii) we propose to augment this model with lightweight statistics of the dynamics of predicates; (iii) we further experiment with linear regression models to predict when results will change; (iv) we prepare a benchmark for predicting the dynamics of query results based on 23 weekly versions of Wikidata and a set of \ah{XXX} real-world SPARQL queries; (v) we use this benchmark to compare the results obtained by our lightweight models versus an ``ideal model'' that knows of changes in the query's results for previous versions.
-Our results show that our proposed models are competitive with the more expensive ideal model.
+The contributions of this paper are thus as follows: (i) we propose a supervised method that uses static query features and suitable classifiers to predict whether or not results will change in a given point in the near future; (ii) we propose to augment this model with lightweight statistics of the dynamics of predicates; (iii) we further experiment with linear regression models to predict when results will change; (iv) we prepare a benchmark for predicting the dynamics of query results based on 23 weekly versions of Wikidata and a set of 221 real-world SPARQL queries; (v) we use this benchmark to compare the results obtained by our lightweight models versus an ``clairvoyant model''~\cite{KnuthHS16} that knows of changes in the query's results for previous versions.
+Our results show that our proposed models are competitive with the costly clairvoyant model.
 
@@
 
-Many applications can benefit from change prediction in query results.
-Both, Smart Caching to save computational resources and Hybrid Architectures to deliver fresh results in less time.
-Several techniques have been developed following strong consistency, based on HTTP metadata~\cite{fielding2rfc} and notifications~\cite{PassantM10,Fitzpatrick10,TrampFEA10}.
-Other works seek weak consistency based on probabilities, such as Empirical Distributions~\cite{NeumaierU16}, Poisson Processes~\cite{UmbrichHHPD10}, and Markov Chains~\cite{UmbrichMP15}, as well as other works based on heuristics~\cite{AliciAOCU12,UmbrichMP15,KnuthHS16}, metrics~\cite{DividinoGS15,KnuthHS16,AkhtarAL17}, and Machine Learning~\cite{NishiokaS17,GonzalezH18}.
-In the case of metadata, many works~\cite{UmbrichHHPD10,DividinoKG14,Kjernsmo15,NeumaierU16} have show that these data are often not provided or are wrong.
-In the case of notifications, several tools have been developed, such as rsine~\cite{MaderMS14} and Boca~RDF~\cite{MissierACDG07}, but these mechanisms must keep a subscriber list and it cause scalability problems.
-The rest of the approaches achieve interesting results, but the main deficiency in these approaches is that their quality depends on having a long and complete history of updates.
-In the case of ML-based approaches, they constitute a general framework where important improvements can be achieved by incorporating relevant features to the models.
-In our approach, we propose an ML-based solution that dispenses with the expensive history of updates by a lighter model with statistics from this history and syntactic query features.
+A variety of works have addressed the issue of modeling and consuming dynamic Linked Data from a broad range of perspectives~\cite{PassantM10,TrampFEA10,UmbrichHHPD10,UmbrichKL10,KaferAUOH13,DividinoSGG13,DividinoKG14,Kjernsmo15,NishiokaS16,GonzalezH18}.
+One of the major challenges considered is that of keeping cached copies of remote dynamic data -- cached for reasons of efficiency and scalability -- up-to-date on the consumer side, which we refer to as the synchronization problem.
 
-Some works have tried to understand the dynamics of the LOD datasets~\cite{UmbrichHHPD10,UmbrichKL10,KaferAUOH13,DividinoSGG13,NishiokaS16}.
-Some features found to describe the dynamics of a resource are Pay-Level Domain (PLD)~\cite{NishiokaS16,NishiokaS17}, predicates~\cite{KaferAUOH13,NishiokaS17}, and Characteristic Sets (CS)~\cite{NishiokaS16,GonzalezH18}.
-This knowledge could be used to estimates the freshness of cached data.
-Umbrich et al.~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12} identified variants to obtain knowledge about freshness for different query patterns based on predicates.
-In more recent work Dehghanzadeh et al.~\cite{DehghanzadehPKUHD14} proposed a solution to estimate the freshness using cardinality estimation techniques based on predicates and CS.
-However, to the best of our knowledge, there is no proposal to estimate the freshness of cached query results using Machine Learning models based on syntactic features from queries and reduced historical statistics.
+Some works have addressed the synchronization problem on the publisher side, proposing notification mechanisms that keep registered consumers informed about relevant changes to the data~\cite{MissierACDG07,PassantM10,TrampFEA10,MaderMS14,webSub18}; although such approaches may facilitate strong consistency -- meaning that consumers are kept up-to-date with the remote data on the publisher side -- they centralize the burden  of synchronization on the publisher, potentially leading to scalability issues.
+
+On the other hand, a variety of works have looked at building models of remote data that can help to predict which data are most dynamic, and which are most static, indicating which subsets of the data may need be refreshed from the remote source more often~\cite{PassantM10,TrampFEA10,UmbrichHHPD10,UmbrichKL10,KaferAUOH13,DividinoSGG13,NishiokaS16,GonzalezH18}; such works consider changes in RDF datasets at differing levels of granularity, including documents~\cite{KaferAUOH13}, domains~\cite{KaferAUOH13,NishiokaS16,NishiokaS17}, predicates~\cite{KaferAUOH13,NishiokaS17}, characteristic sets\footnote{A characteristic set is the set of predicate terms used to describe a given subject~\cite{NeumannM11}.}~\cite{NishiokaS16,GonzalezH18}, etc.
+Features at different levels of granularity can be fed into different predictive models based on Poisson Processes~\cite{UmbrichHHPD10}, Markov Chains~\cite{UmbrichMP15}, Empirical Distributions~\cite{NeumaierU16}, Machine Learning classification and regression~\cite{NishiokaS17,GonzalezH18}, as well as a variety of other heuristics~\cite{AliciAOCU12,UmbrichMP15,KnuthHS16} and metrics~\cite{DividinoGS15,KnuthHS16,AkhtarAL17}.
+Such approaches obviate the need for a subscription/notification mechanism.
+On the other hand, to ensure strong consistency in the presence of highly dynamic data, consumers may need to conservatively send a great many refresh requests to the server, which may be even more costly than a subscription/notification mechanism; hence such approaches are better suited for scenarios where weak consistency is more acceptable.
+
+% queries: PassantM10, SPARQL push, 
+%
+
+Specifically regarding the dynamics of SPARQL query results, Passant and Mendes~\cite{PassantM10} proposed \textsc{sparqlPuSH} as a notification framework based  on \textsc{PubSubHubBub} (recently standardized as \textsc{WebSub}~\cite{webSub18}) aiming for strong consistency.
+On the other hand, aiming for weak consistency, Umbrich et al.~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12} proposed various methods to obtain knowledge about dynamics for different query patterns, mainly based on predicates.
+Dehghanzadeh et al.~\cite{DehghanzadehPKUHD14} later proposed a method to estimate the freshness using cardinality estimation techniques based on predicates and characteristic sets.
+Combining the notion of subscription-based notifications and predicting dynamics, Knuth et al.~\cite{KnuthHS16} propose a middleware to which consumers may subscribe and that periodically ranks and schedules refreshes for queries according to policies that take into account how likely the results are to be stale, how long ago the query was last refreshed, how many results previously changed, how long the query takes to run, etc.
+
+Our proposal aims to model the dynamics of SPARQL query results in order to predict (1) whether or not the results will have changed in a fixed point in the near future and,  (2) when the results will change.
+Our work thus complements existing works aiming for weak consistency, but rather investigates the potential of using static features of a query as a lightweight alternative with the intuition that, for example, queries with more triple patterns or certain query operators are more likely to have dynamic results.
+For the purposes of comparison, we complement these static features with further knowledge of the dynamics of predicates~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12,DehghanzadehPKUHD14} and knowledge of how results for the input query have changed over past versions~\cite{KnuthHS16}, as considered in previous works.
+We feed these data into Machine Learning classification and regression models to predict future changes (hence, for the purposes of supervised learning, we require knowledge of the changes of some queries over time, though not necessarily for the input query).
+The benefits of such an approach can be seen in a number of ways: (i) as complementing existing approaches; (ii) as an alternative to predicate-based statistics, which requires analysis of prior versions of the underlying data; (iii) as an alternative to historical results, which requires having a record of (or computing at runtime) the results of the query over several historical versions.
+We further remark that, to the best of our knowledge, no existing work aims to predict the dynamics of SPARQL query results using Machine Learning models.
 
@@
 
-According to the LD principles, the data published on the Web must be described using the Resource Description Framework (RDF).
-RDF is a conceptual model based on directed graphs that serves to provide descriptive information about the resources on the Web.
-The RDF nodes correspond to resources, literals (strings), or blank nodes, which model variables in the graph.
-Resources are identified by an Internationalized Resource Identifier (IRI).
-The nodes are linked through labeled and directed arcs, which represent semantic relationships.
-
-\begin{definition}[RDF Triple and Dataset]
-	Let I, L, and B be disjoint sets of IRIs, literals, and blank nodes respectively.
-A tuple $ (s, p, o) \in (I \cup  B)  \times (I)  \times  (I \cup  B \cup  L)$  is called an RDF $ triple $, where $ s $ is called subject, $ p $ is called predicate, and $ o $ is called object.
-An RDF $dataset$ is a set of RDF $triples$.
-\end{definition}
+RDF is a conceptual data model based on directed graphs that can be used to describe resources on the Web.
+\emph{RDF terms} are elements of the set $\Iris \cup \Bnodes \cup \Lits$ composed of IRIs $\Iris$, literals $\Lits$, and blank nodes $\Bnodes$.
+A tuple $ (s, p, o) \in (\Iris \cup  \Bnodes)  \times (\Iris)  \times  (\Iris \cup \Bnodes \cup  \Lits)$  is called an \emph{RDF triple}, where $s$ is called \emph{subject}, $p$ is called \emph{predicate}, and $o$ is called \emph{object}.
+An \emph{RDF graph} is a set of RDF triples.
 
 SPARQL is the recommended query language to retrieve and manipulate data stored in the RDF format.
-In this work, we focus on SPARQL SELECT queries, which return the set of variables and their mapping results.
-Here we define a SPARQL 1.0 query.
+In this work, we focus on SPARQL \texttt{SELECT} queries, where we will first define a SPARQL 1.0 query.
+Let $\Vars$ be a set of variables disjoint from the set of RDF terms.
+A \emph{SPARQL expression} is built recursively as follows.
+(1) A \emph{triple pattern} $t \in  (\Iris \cup \Bnodes \cup \Lits \cup \Vars) \times (\Iris \cup \Vars) \times (\Iris \cup \Lits \cup \Bnodes \cup \Vars)$ is an \emph{expression}.
+(2) If $Q_1$ and $Q_2$ are expressions and $R$ is a filter condition, then $Q_1$ \texttt{FILTER} $R$, $Q_1$ \texttt{UNION} $Q_2$, $Q_1$ \texttt{OPTIONAL} $Q_2$, $Q_1$ \texttt{AND} $Q_2$ are $expressions$.
+Finally, if $Q$ is an expression, $V$ a list of variables and $\Delta$ a boolean value, $\texttt{SELECT}_V^\Delta Q$ is a SPARQL \texttt{SELECT} query, where $V$ denotes the projected variables, and $\Delta$ the \texttt{DISTINCT} option that when true, removes duplicate results.
+The semantics of a SPARQL \texttt{SELECT} query $Q$ is defined in terms of its \emph{evaluation} over an RDF graph $G$, denoted $Q(G)$, giving a set of partial mappings from projected variables to the set $\Iris \cup \Lits \cup \Bnodes$; we refer to P{\'{e}}rez et al.~\cite{PerezAG09} for definitions.
+(According to the standard, a SPARQL query is evaluated over a set of \emph{named graphs}; however, since the Wikidata query service indexes data in one RDF graph, to keep discussion concise, we consider RDF graphs, though our methods can be generalized in a straighforward manner to RDF datasets/named graphs.)
 
-\begin{definition}[SPARQL expression]
-	Let V be a set of variables disjoint from $U \cup B \cup L$.
-A SPARQL $expression$ is built recursively as follows.
-(1) A triple $pattern t \in  (U \cup V ) \times (U \cup V ) \times (L \cup U \cup V )$ is an $expression$.
-(2) If $Q_1$ and $Q_2$ are $expressions$ and R is a filter condition, then $Q_1$ FILTER R, $Q_1$ UNION $Q_2$, $Q_1$ OPTIONAL $Q_2$, $Q_1$ AND $Q_2$ are $expressions$.
-\end{definition}
+Our method supports SPARQL 1.1 \texttt{SELECT} queries, which allow a variety of additional features.
+One key feature in this extension is that of \textit{property paths}, which allows for matching arbitrary length paths in an RDF graph, potentially returning or matching the endpoints of the path.
+An IRI $p$ is a \emph{path expression}; if $e$, $e_1$ and $e_2$ are path expressions, then $\texttt{\^{}}e$ (inverse of $e$), $e_1 \texttt{/} e_2$ ($e_1$ followed by $e_2$), $e_1 \texttt{|} e_2$ ($e_1$ or $e_2$), $e\texttt{*}$ (zero or more $e$), $e\texttt{+}$ (one or more $e$), $e\texttt{?}$ (zero or one $e$), and $\texttt{(}$e$\texttt{)}$ (parentheses used for precedence) are also \textit{path expressions}; finally, if $p$, $p_1$, $\ldots$ $p_n$ are IRIs, then $\texttt{!}p$, $\texttt{!}\texttt{(}p_1 \texttt{|} \ldots \texttt{|} p_n\texttt{)}$ and $\texttt{!}\texttt{(}p_1 \texttt{|} \ldots \texttt{|} p_k \texttt{|}  \texttt{{\^{}}}p_{k+1} \texttt{|} \ldots \texttt{|} \texttt{{\^{}}}p_n\texttt{)}$ for $k+1 \leq n$ (negated property sets) are path expressions.
+Thereafter, a \textit{path pattern} $(s,e,o)$ where $e$ is a path expression is an expression.
+Other features supported in SPARQL~1.1 include sub-queries, negation, aggregation, value binding, and so forth; for brevity, we do not introduce definitions for all such features.
 
-A triple $ pattern $ in the second position can contain an expression called \textit{Property Paths} since SPARQL~1.1 to allow for navigational querying over RDF graphs.
-Other features at SPARQL~1.1 will be seen in experiments.
+%\begin{table}
+%\centering
+%	\begin{tabular}{ll}    \hline
+%		Rules &  \\    \hline
+%		e $\rightarrow$ p & a predicate \\
+%		
+%		e $\rightarrow$ \^{}e    &  inverse path \\
+%		
+%		e $\rightarrow$ e $/$ e    &  a path of $ e_1 $ followed by $ e_2 $\\
+%		
+%		e $\rightarrow$ e $\mid$ e    &  a path of $ e_1 $ or $ e_2 $ \\
+%		
+%		e $\rightarrow$ e*    &  a path of zero or more $ e $ \\
+%		
+%		e $\rightarrow$ e+    &  a path of one or more $ e $ \\
+%		
+%		e $\rightarrow$ e?    &  a path of zero or one $ e $ \\
+%		
+%		e $\rightarrow$ !p    &  any predicate not $ p $ \\
+%		
+%		e $\rightarrow$ !( p1,...,pk, \^{}p1,...,\^{}pk)    &  any (inverse) predicate not listed \\
+%		
+%		e $\rightarrow$ ( e )    &  brackets used for grouping \\    \hline
+%	\end{tabular}
+%	\caption{Property path syntax}
+%\end{table}
+%
+%\begin{definition}[Property Paths]
+%	\label{eq:pp}
+%	Property Paths expressions are defined by the grammar:\\
+%	
+%	
+%	
+%	where $p, p1,...,pk$ are IRIs in $I$.
+%\end{definition}
 
-\begin{definition}[Property Paths]
-	\label{eq:pp}
-	Property Paths expressions are defined by the grammar:\\
-
-	\begin{tabular}{ll}    \hline
-		Rules &  \\    \hline
-		e $\rightarrow$ p & a predicate \\
-
-		e $\rightarrow$ \^{}e    &  inverse path \\
-
-		e $\rightarrow$ e $/$ e    &  a path of $ e_1 $ followed by $ e_2 $\\
-
-		e $\rightarrow$ e $\mid$ e    &  a path of $ e_1 $ or $ e_2 $ \\
-
-		e $\rightarrow$ e*    &  a path of zero or more $ e $ \\
-
-		e $\rightarrow$ e+    &  a path of one or more $ e $ \\
-
-		e $\rightarrow$ e?    &  a path of zero or one $ e $ \\
-
-		e $\rightarrow$ !p    &  any predicate not $ p $ \\
-
-		e $\rightarrow$ !( p1,...,pk, \^{}p1,...,\^{}pk)    &  any (inverse) predicate not listed \\
-
-		e $\rightarrow$ ( e )    &  brackets used for grouping \\    \hline
-	\end{tabular}
-
-	where $p, p1,...,pk$ are IRIs in $I$.
-\end{definition}
-
-\section{Our Approach}
+\section{Predicting the Dynamics of Query Results}
 \label{sec:approach}
 
-Next, we formalize the problem addressed in this work and present our proposed solution.
+We consider a dynamic RDF graph to be a sequence of $n$ versions of an RDF graph denoted $\mathcal{G} \da ( G_1, \ldots, G_n )$; in practice, we assume these versions to have regular intervals (e.g., hourly, daily, weekly, etc.).
+Further given a SPARQL (1.1) \texttt{SELECT} query $Q$, we would like to have knowledge of the dynamics of its results.
+The first such problem we consider -- which we call One Shot Change (OSC) -- accepts $\mathcal{G}$, $Q$ and a positive integer $k$ as input and outputs a boolean value predicting whether or not the results will change from $G_n$ to some future version $G_{n+k}$ (i.e., $Q(G_n) \neq Q(G_{n+k})$?).
+The second (more difficult) problem -- which we call Time-To-Live (TTL) -- accepts only $\mathcal{G}$ and $Q$ and outputs a positive integer $k$ as a prediction for the lowest such value where $Q(G_n) \neq Q(G_{n+k})$.
 
-\subsection{Problem Definition and Proposed Solution}
+\ah{How do we define quality under bag semantics, without \texttt{DISTINCT}? Do we consider equality of unique results, or multiplicity of results?}
 
-For this work, we consider a dynamic RDF dataset, denoted by $\mathcal{D}$.
-In addition, we assume that we have consecutive sequence of $m$ past versions $\mathcal{D} = (d_1, d_2, ..., d_m)$ constituting different revisions fetched at regular intervals.
-We also consider a SPARQL SELECT query $q$, of which we would like to know if its results are going to be valid for a long time.
-For that, we have the results of $q$ on each of the $m$ past versions of $\mathcal{D}$ and we need to know when we should update it.
-Then, we call $q(d_i)$ the result of executing the query $q$ on the $d_i$ version of $\mathcal{D}$.\\
-
-Let delta ($\delta$) be a function to compute the binary dissimilarity between two results for a query $q$ with the following expression:
-
-\begin{equation}
-\label{eq:delta}
-\delta(q(d_i), q(d_j)) = \begin{cases}
-0 & q(d_i) = q(d_j) \\
-1 & q(d_i) \neq q(d_j) 
-\end{cases}
-\end{equation}
-
-\begin{example}
-	\label{ex:dataset}    
-	Consider the following example of seven versions of a RDF dataset. Also, consider the queries~\ref{lst:sparql0}, \ref{lst:sparql1} and \ref{lst:sparql2}.
-\end{example}
-
-\setlength{\tabcolsep}{6pt}
-\begin{tabular}{l l l}    \hline	
-	\multicolumn{3}{c}{versions $d_1$ and $d_2$} \\    \hline
-	:Atlantic\_Ocean  & :instance\_of  & ocean          \\
-	:Atlantic\_Ocean  & :temperature   & 40$\degree$F          \\
-	:Atlantic\_Ocean  & :area          & $10^7\pm10^3$ $km^2$ \\
-	:Atlantic\_Ocean  & :image         & AOcean1.png    \\    \hline
-	\multicolumn{3}{c}{versions $d_3$, $d_4$ and $d_5$} \\    \hline
-	:Atlantic\_Ocean  & :instance\_of  & ocean          \\
-	:Atlantic\_Ocean  & :temperature   & 50$\degree$F             \\
-	:Atlantic\_Ocean  & :area          & $10^7\pm10^3$ $km^2$ \\
-	:Atlantic\_Ocean  & :image         & AOcean1.png    \\    \hline
-	\multicolumn{3}{c}{versions $d_6$ and $d_7$} \\    \hline
-	:Atlantic\_Ocean  & :instance\_of  & ocean          \\
-	:Atlantic\_Ocean  & :temperature   & 60$\degree$F    \\
-	:Atlantic\_Ocean  & :area          & $10^7\pm10^3$ $km^2$ \\
-	:Atlantic\_Ocean  & :image         & AOcean1.png    \\
-	:Atlantic\_Ocean  & :image         & AOcean2.png    \\    \hline
-\end{tabular}\\
-
-\begin{lstlisting}[captionpos=b, caption=SPARQL query about ocean temperatures., label=lst:sparql0,
-basicstyle=\ttfamily,frame=single]
-SELECT ?ocean ?temp
-WHERE {
-  ?ocean :instance_of :ocean .
-  ?ocean :temperature ?temp
-}
-\end{lstlisting}
-
-\begin{lstlisting}[captionpos=b, caption=SPARQL query about ocean areas., label=lst:sparql1,
-basicstyle=\ttfamily,frame=single]
-SELECT ?ocean ?area
-WHERE {
-  ?ocean :instance_of :ocean .
-  ?ocean :area ?area
-}
-\end{lstlisting}
-
-\begin{lstlisting}[captionpos=b, caption=SPARQL query about ocean images., label=lst:sparql2,
-basicstyle=\ttfamily,frame=single]
-SELECT ?ocean ?image
-WHERE {
-?ocean :instance_of :ocean .
-?ocean :image ?image
-}
-\end{lstlisting}
-
-$\delta(q_1(d_1), q_1(d_3)) = 1$ and $\delta(q_2(d_2), q_2(d_4)) = 0$
-
-Based on the $\delta$ function (see Equation~\ref{eq:delta}), the problem of Time-to-Live (TTL) estimation, denoted by $TTL(q,m)$ is defined by the size of the largest interval from $m$ for which the query result does not change (see Equation~\ref{eq:ttl}).
-
-\begin{equation}
-\label{eq:ttl}
-TTL(q,m)=min\{t \in \mathbf{Z}^+ \mid \delta(q(d_m), q(d_{m+t})) \neq 0 \}
-\end{equation}
-
-As a relaxed version of this problem, there is the problem of One-Step-Change (OSC) estimation to predict the change in the next interval. (see Equation~\ref{eq:osc})
-
-\begin{equation}
-\label{eq:osc}
-OSC(q,m) = \delta(q(d_m), q(d_{m+1}))
-\end{equation}
-
-Considering the dataset from Example~\ref{ex:dataset} and queries~\ref{lst:sparql0}, \ref{lst:sparql1} and \ref{lst:sparql2}, we show the change history of the query $q_1$ (see Figure~\ref{fig:problem}).
+%Let delta ($\delta$) be a function to compute the binary dissimilarity between two results for a query $q$ with the following expression:
+%
+%\begin{equation}
+%\label{eq:delta}
+%\delta(q(d_i), q(d_j)) = \begin{cases}
+%0 & q(d_i) = q(d_j) \\
+%1 & q(d_i) \neq q(d_j) 
+%\end{cases}
+%\end{equation}
+%
+%\begin{example}
+%	\label{ex:dataset}    
+%	Consider the following example of seven versions of a RDF dataset. Also, consider the queries~\ref{lst:sparql0}, \ref{lst:sparql1} and \ref{lst:sparql2}.
+%\end{example}
+%
+%\setlength{\tabcolsep}{6pt}
+%\begin{tabular}{l l l}    \hline	
+%	\multicolumn{3}{c}{versions $d_1$ and $d_2$} \\    \hline
+%	:Atlantic\_Ocean  & :instance\_of  & ocean          \\
+%	:Atlantic\_Ocean  & :temperature   & 40$\degree$F          \\
+%	:Atlantic\_Ocean  & :area          & $10^7\pm10^3$ $km^2$ \\
+%	:Atlantic\_Ocean  & :image         & AOcean1.png    \\    \hline
+%	\multicolumn{3}{c}{versions $d_3$, $d_4$ and $d_5$} \\    \hline
+%	:Atlantic\_Ocean  & :instance\_of  & ocean          \\
+%	:Atlantic\_Ocean  & :temperature   & 50$\degree$F             \\
+%	:Atlantic\_Ocean  & :area          & $10^7\pm10^3$ $km^2$ \\
+%	:Atlantic\_Ocean  & :image         & AOcean1.png    \\    \hline
+%	\multicolumn{3}{c}{versions $d_6$ and $d_7$} \\    \hline
+%	:Atlantic\_Ocean  & :instance\_of  & ocean          \\
+%	:Atlantic\_Ocean  & :temperature   & 60$\degree$F    \\
+%	:Atlantic\_Ocean  & :area          & $10^7\pm10^3$ $km^2$ \\
+%	:Atlantic\_Ocean  & :image         & AOcean1.png    \\
+%	:Atlantic\_Ocean  & :image         & AOcean2.png    \\    \hline
+%\end{tabular}\\
+%
+%\begin{lstlisting}[captionpos=b, caption=SPARQL query about ocean temperatures., label=lst:sparql0,
+%basicstyle=\ttfamily,frame=single]
+%SELECT ?ocean ?temp
+%WHERE {
+%  ?ocean :instance_of :ocean .
+%  ?ocean :temperature ?temp
+%}
+%\end{lstlisting}
+%
+%\begin{lstlisting}[captionpos=b, caption=SPARQL query about ocean areas., label=lst:sparql1,
+%basicstyle=\ttfamily,frame=single]
+%SELECT ?ocean ?area
+%WHERE {
+%  ?ocean :instance_of :ocean .
+%  ?ocean :area ?area
+%}
+%\end{lstlisting}
+%
+%\begin{lstlisting}[captionpos=b, caption=SPARQL query about ocean images., label=lst:sparql2,
+%basicstyle=\ttfamily,frame=single]
+%SELECT ?ocean ?image
+%WHERE {
+%?ocean :instance_of :ocean .
+%?ocean :image ?image
+%}
+%\end{lstlisting}
+%
+%$\delta(q_1(d_1), q_1(d_3)) = 1$ and $\delta(q_2(d_2), q_2(d_4)) = 0$
+%
+%Based on the $\delta$ function (see Equation~\ref{eq:delta}), the problem of Time-to-Live (TTL) estimation, denoted by $TTL(q,m)$ is defined by the size of the largest interval from $m$ for which the query result does not change (see Equation~\ref{eq:ttl}).
+%
+%\begin{equation}
+%\label{eq:ttl}
+%TTL(q,m)=min\{t \in \mathbf{Z}^+ \mid \delta(q(d_m), q(d_{m+t})) \neq 0 \}
+%\end{equation}
+%
+%As a relaxed version of this problem, there is the problem of One-Step-Change (OSC) estimation to predict the change in the next interval. (see Equation~\ref{eq:osc})
+%
+%\begin{equation}
+%\label{eq:osc}
+%OSC(q,m) = \delta(q(d_m), q(d_{m+1}))
+%\end{equation}
+%
+%Considering the dataset from Example~\ref{ex:dataset} and queries~\ref{lst:sparql0}, \ref{lst:sparql1} and \ref{lst:sparql2}, we show the change history of the query $q_1$ (see Figure~\ref{fig:problem}).
 We mark with a discontinuous line the OSC and with a continuous line the TTL at version $ d_3 $.
 The OSC is equal to zero, which means that the results will not change in the next version $d_4$, and the TTL equals three.
+%
+%\begin{figure}[h]
+%	\centering
+%	\includegraphics[width=0.7\linewidth]{img/problem.pdf}
+%	\caption{Example change history of query $q_1$.}
+%	\label{fig:problem}
+%\end{figure}
 
-\begin{figure}[h]
+\paragraph{Architecture:} We propose a method to predict answers for the OSC and TTL problems.
+Our method primarily considers static features of the query: these include statistics such as the number of triple patterns, as well as the presence of query operators, such as \texttt{UNION}, \texttt{OPTIONAL}, specific path expressions, etc.; our model can also consider statistics about predicates in the dynamic RDF graph, as well as in which past versions of the dynamic graph the results for the query changed.
+Figure~\ref{fig:schema} shows the components of our approach, which receives as input a SPARQL query (the architecture itself already encodes information about the dynamic graph and the value of $k$ for OSC, which we set in experiments to 1, predicting changes in the next version).
+The Feature Extractor decomposes the query into a set of numeric features.
+The Dynamics Model accepts some features of the query, as well as information from the dynamic graph, in order to compose further features.
+The final features are passed to a classifier in the case of the OSC problem, and to a regression model in the case of the TTL problem, to generate the final output prediction.
+We now describe the Feature Extractor and Dynamics Model in more detail.
+
+\begin{figure*}[!t]
 	\centering
-	\includegraphics[width=0.7\linewidth]{img/problem.pdf}
-	\caption{Example change history of query $q_1$.}
-	\label{fig:problem}
-\end{figure}
-
-We propose a scheme that automatically identifies whether a SPARQL query will change in the next version (OSC) or when it will change (TTL), depending on the syntax of the query and statistics about the dynamics of predicates.
-Figure~\ref{fig:schema} shows the components of our approach, which receives as input a SPARQL query.
-The Feature Extractor analyzes the query and extracts information about the syntax of the query.
-The Dynamics Model considers some of these features to provide a dynamics indicator of the query.
-This indicator, together with the previous features extracted, is passed through a classifier that finally predicts the value for the problems posed in Equation~\ref{eq:osc} and Equation~\ref{eq:ttl}.
-
-\begin{figure*}[!ht]
-	\centering
-	\includegraphics[width=0.7\linewidth]{img/schema.pdf}
-	\caption{ Proposed architecture.
-It receives a SPARQL query as input; uses a pre-calculated Dynamics Model to estimate a query dynamic feature, which is used with other features to predict the OSC and TTL values.
-}
+	\includegraphics[width=0.9\linewidth]{img/schema.pdf}
+	\caption{Proposed architecture with Feature Extractor and Dynamics Model \ah{What is $g$ here and why is it output from the Dynamics Model (rather than input)?}}
 	\label{fig:schema}
@@
 
-\subsection{Feature Extractor}
-
-Our features are based on the syntactic analysis of the queries; then some features are delivered to the Dynamics Model, which provides an indicator based on the probability of change for the constant predicates within the triple patterns of the query.
+\paragraph{Feature Extractor:} Our features are based on the syntactic analysis of the queries; then some features are delivered to the Dynamics Model, which provides an indicator based on the probability of change for the constant predicates within the triple patterns of the query.
 

### 2019-01-26 kgswc19.tex
Some further edits in Section 4

@@
 \usepgfplotslibrary{clickable}
+
+\usepackage{textcomp}
+\usepackage{amssymb}
+\usepackage{xspace}
+
+\newcommand{\tickYes}{\checkmark}
+\newcommand{\tickNo}{\textsl{\textsf{X}}}
+\newcommand{\ntextnumero}{{\fontfamily{txr}\selectfont \textnumero}\xspace}
 
@@
 }
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% SPARQL Listing style
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage{listings}
+
+\colorlet{punct}{red!60!black}
+\definecolor{background}{HTML}{EEEEEE}
+\definecolor{delim}{RGB}{120,20,40}
+\definecolor{keyw}{RGB}{0,0,192}
+\colorlet{numb}{magenta!60!black}
+
+\lstdefinelanguage{sparql}{
+	sensitive=false,
+	extendedchars=true,
+	literate={á}{{\'a}}1 {é}{{\'e}}1 {í}{{\'{\i}}}1 {ó}{{\'o}}1 {ú}{{\'u}}1
+	{Á}{{\'A}}1 {É}{{\'E}}1 {Í}{{\'I}}1 {Ó}{{\'O}}1 {Ú}{{\'U}}1
+	{ü}{{\"u}}1 {Ü}{{\"U}}1 {ñ}{{\~n}}1 {Ñ}{{\~N}}1 {¿}{{?``}}1 {¡}{{!``}}1
+	{<}{{{\color{delim}<}}}{1}
+	{>}{{{\color{delim}>}}}{1}
+	{?}{{{\color{delim}?}}}{1}
+	{*}{{{\color{delim}*}}}{1}
+	{+}{{{\color{delim}+}}}{1}
+	{/}{{{\color{delim}/}}}{1}
+	{,}{{{\color{punct}{,}}}}{1}
+	{;}{{{\color{punct}{;}}}}{1}
+	{.}{{{\color{punct}{.}}}}{1}
+	{:}{{{\color{punct}{:}}}}{1}
+	{\{}{{{\color{delim}{\{}}}}{1} {\}}{{{\color{delim}{\}}}}}{1},
+	morekeywords={ask,select,from,where,order,by,distinct,limit,offset,optional,union,filter,prefix,bound,desc,regex,str,group,not,exists,minus,service,certain,maybe}
+}
+
+\lstdefinestyle{sparqld}{
+	basicstyle=\scriptsize\ttfamily,
+	identifierstyle=\color{black},
+	keywordstyle=\color{keyw}\bfseries,
+	ndkeywordstyle=\color{greenCode}\bfseries,
+	stringstyle=\color{ocherCode}\ttfamily,
+	commentstyle=\color{darkgray}\ttfamily,
+	language={sparql},
+	tabsize=2,
+	showtabs=false,
+	showspaces=false,
+	showstringspaces=false,
+	extendedchars=true,
+	escapechar=`,
+	frame={single},
+	breaklines=true,
+	basewidth=0.5em,
+	moredelim=[is][\color{magenta}]{~}{~},
+	moredelim=**[is][\color{gray}]{£}{£},
+	moredelim=**[is][\color{blue!50!black}]{$}{$},
+	moredelim=**[is][\color{green!50!black}]{¬}{¬},
+	xleftmargin=2ex,
+	xrightmargin=1ex,
+	aboveskip=1.5ex,
+	belowskip=1.5ex
+}
+
+\makeatletter
+\newcommand{\sqbox}{%
+	\collectbox{%
+		\@tempdima=\dimexpr\width-\totalheight\relax
+		\ifdim\@tempdima<\z@
+		\fbox{\hbox{\hspace{-.5\@tempdima}\BOXCONTENT\hspace{-.5\@tempdima}}}%
+		\else
+		\ht\collectedbox=\dimexpr\ht\collectedbox+.5\@tempdima\relax
+		\dp\collectedbox=\dimexpr\dp\collectedbox+.5\@tempdima\relax
+		\fbox{\BOXCONTENT}%
+		\fi
+	}%
+}
+\makeatother
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% /SPARQL Listing style
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\usepackage{floatrow}
+% Table float box with bottom caption, box width adjusted to content
+%\newfloatcommand{capbtabbox}{table}[][\FBwidth]
 
@@
 
-\paragraph{Feature Extractor:} Our features are based on the syntactic analysis of the queries; then some features are delivered to the Dynamics Model, which provides an indicator based on the probability of change for the constant predicates within the triple patterns of the query.
+\paragraph{Feature Extractor:} Our initial set of features are based on analysis of the input SPARQL query.
+Figure~\ref{fig:query} shows the complete list of evaluated features and shows the nominals.
+In the case of negation, we consider the statements MINUS, FILTER NOT EXISTS, and FILTER NOT BOUND and by the recursive path we understand those Property Paths that have patterns of the form $e*$ or $e+$ (see Definition~\ref{eq:pp}).
 
@@
 
-Figure~\ref{fig:query} shows the complete list of evaluated features and shows the nominals.
-In the case of negation, we consider the statements MINUS, FILTER NOT EXISTS, and FILTER NOT BOUND and by the recursive path we understand those Property Paths that have patterns of the form $e*$ or $e+$ (see Definition~\ref{eq:pp}).
-
-\begin{lstlisting}[captionpos=b, caption=Query example, label=fig:query,
-basicstyle=\ttfamily,frame=single]
+\begin{figure}[!t]
+{\centering
+	\begin{minipage}{0.46\textwidth}
+	\begin{lstlisting}[style=sparqld]
 SELECT ?item
@@
       ?pob :located_in* :Wales }
-  OPTIONAL { ?sitelink schema:about ?item .
+  OPTIONAL 
+    { ?sitelink schema:about ?item .
       ?sitelink schema:inLanguage "cy" }
@@
 	\end{lstlisting}
-
-\begin{center}
-	\begin{tabular}{|l|ll|}    \hline
-		1  & \# triples        & 7                             \\
-		2  & \# vars           & 3                             \\
-		3  & \# projected vars & 1                             \\
-		4  & \# predicates     & 6                             \\
-		5  & FILTER            & TRUE                          \\
-		6  & LIMIT             & TRUE                          \\
-		7  & UNION             & TRUE                          \\
-		8  & SUBQUERY          & FALSE                         \\
-		9  & *PATH             & TRUE                          \\
-		10 & NEGATION          & TRUE                          \\
-		11 & predicates        & {[}$p_1$,...,$p_k${]}    \\    \hline
+	\end{minipage}}
+	\hfill
+	\resizebox{0.46\textwidth}{!}{
+	\begin{tabular}{rlr}    \toprule
+		1  & \ntextnumero of triples        & 7			\\
+		2  & \ntextnumero of variables      & 3			\\
+		3  & \ntextnumero of projected vars & 1			\\
+		4  & \ntextnumero of predicates     & 6			\\ \midrule
+		5  & \texttt{FILTER}            	& \tickYes	\\
+		6  & \texttt{LIMIT}             	& \tickYes	\\
+		7  & \texttt{UNION}             	& \tickYes	\\
+		8  & \textit{Sub-query}         	& \tickNo	\\
+		9  & \textit{Recursive path}		& \tickYes	\\
+		10 & \textit{Negation}         		& \tickYes	\\ \midrule
+		11 & \textit{Predicates}			& $\{$ \texttt{:instance\_of}, ... $\}$     \\    \bottomrule
 	\end{tabular}	
-\end{center}
+	}
+\caption{Example query and the features extracted from it \label{fig:query}}	
+\end{figure}
 

### 2019-01-26 kgswc19.tex
Finished pass up to end of Section 4

@@
 
+One topic we do wish to highlight however -- as it relates to the behavior of a query over a dynamic RDF graph -- is that of the \emph{monotonicity} of SPARQL queries~\cite{ArenasP11,ArenasU17}.
+We say that a \texttt{SELECT} query $Q$ is monotone if and only if $G_1 \subseteq G_2$ implies that $Q(G_1) \subseteq Q(G_2)$ for RDF graphs $G_1$ and $G_2$; intuitively, as data are extended, the results of a monotone query can only be extended.
+Monotonic SPARQL features include, for example, joins, unions, paths and filters; on the other hand, non-monotonic SPARQL features include negation and optional~\cite{ArenasP11}.
+
 %\begin{table}
@@
 \paragraph{Feature Extractor:} Our initial set of features are based on analysis of the input SPARQL query.
-Figure~\ref{fig:query} shows the complete list of evaluated features and shows the nominals.
-In the case of negation, we consider the statements MINUS, FILTER NOT EXISTS, and FILTER NOT BOUND and by the recursive path we understand those Property Paths that have patterns of the form $e*$ or $e+$ (see Definition~\ref{eq:pp}).
+Figure~\ref{fig:query} shows the list of evaluated features, as well as -- for the purposes of illustration -- their values for an accompanying example query.
 
-We try to discover the degree of correlation between the propensity to change, with respect to numerical values, such as the number of variables, the number of triple patterns or number of predicate.
-We also believe that some qualitative variables may be related to the change, such as the existence of some solution-modifiers.
-We also consider features that describe the complexity of the queries, such as the presence of subqueries, the existence of recursive expressions in the Property Paths of the triple patterns or the existence of any type of negation.
+The first four features capture statistics about the query, indicating loosely its size and complexity.
+One may consider that the higher these four values are, in general, the more dynamic we can expect the query to be since there are more ``opportunities'' for the query results to be affected by change.
+In some cases, however, the hypothesized correlation is not direct since for example adding more triple patterns may serve to narrow the query down and focus it on a static part of the graph (e.g., when looking for the movies of directors, adding a triple pattern to restrict the results to directors who have died).
+Hence it will be of interest to see how these features affect the predictions made experimentally.
+
+The next six features indicate the query operators used; to reduce the dimensionality of the problem, we capture only a subset of popularly-used features and group related features into one dimension: in the case of \textit{Recursive path}, we group queries that contain as expression of the form $e$\texttt{*} or $e$\texttt{+}.
+On the other hand, in \textit{Negation}, we group non-monotonic features that allow for modeling difference (\texttt{MINUS}, \texttt{NOT EXISTS}, \texttt{OPTIONAL}, \texttt{!BOUND}).
+These features -- though course-grained -- are straightforward to extract from a query, and offer valuable insights into how the query may behave in a dynamic query; for example, the \textit{Negation} feature captures information about the (non-)monotonicity of the query, while we suppose that \textit{Recursive paths}, which may traverse an arbitrary number of triples in the graph, might be sensitive to change.
+Again, such relations are not direct where experiments are required for further insights.
+
+\ah{Is OPTIONAL not captured? Why were these features captured? Why not BIND or OFFSET or GROUP BY?}
+
+The final feature -- a raw set of all predicates used -- is rather a meta-feature that serves as input for the next part of the proposed model.
 
@@
 	\hfill
-	\resizebox{0.46\textwidth}{!}{
+	\resizebox{0.48\textwidth}{!}{
 	\begin{tabular}{rlr}    \toprule
-		1  & \ntextnumero of triples        & 7			\\
+		1  & \ntextnumero of triple patterns        & 7			\\
 		2  & \ntextnumero of variables      & 3			\\
-		3  & \ntextnumero of projected vars & 1			\\
+		3  & \ntextnumero of projected variables & 1			\\
 		4  & \ntextnumero of predicates     & 6			\\ \midrule
@@
 
-\subsection{RDF Dynamics Model}
+\paragraph{Dynamics Model:} The next component in our architecture is the Dynamics Model which (optionally) incorporates information about the query \textit{and} the dynamic graph under consideration.
 
-We propose a model to estimate the dynamics of the results of a SPARQL query on an RDF dataset.
-The intuition behind our model is to capture how many triples change for a predicate in a time interval.
-For this, in the following, we define \textit{Predicate Multiplicity of an RDF dataset}.
+The first feature of the Dynamics Model aims to capture how many triples change for a predicate in a time interval, with the idea that -- as in previous works~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12,DehghanzadehPKUHD14} -- predicates capture rich information about the dynamics of a dataset, where the results of queries with dynamic predicates will be more sensitive to change; furthermore, given that in most datasets the number of predicates is relatively low (in the thousands or tens of thousand), such statistics are quite lightweight to maintain and compute.
 
-\begin{definition}[Predicate Multiplicity of an RDF Dataset]
-	Given an RDF dataset $d$ and a predicate $p$, the multiplicity of $p$ in $d$, denoted by $M(p,d)$, is:
-	\begin{equation}
-	\label{eq:pm}
-	M(p,d) = |\{(x,y,z) \in d : y = p \}|
-	\end{equation}
-\end{definition}
+Formally, given two RDF graphs $G_i$ and $G_j$, we denote by $G_i \oplus G_j$ the set of triples $(G_i \cup G_j) - (G_i \cap G_j)$ where ``$-$'' denotes set difference; in other words, $G_i \oplus G_j$ denotes the triples in one graph or the other but not both (noting that $G_i \oplus G_j = G_j \oplus G_i$).
+Next, given an RDF graph $G$ and an IRI $p$, let $\#(G,p) \da |\{(x,y,z) \in G : y = p \}|$ denote the number of triples in $G$ with predicate $p$.
+Next, given a dynamic RDF graph $\mathcal{G} \da ( G_1, \ldots, G_n )$ and a predicate $p$, we denote by $\Delta(\mathcal{G},p) \da \Sigma_{i=1}^{n-1} \#(G_i \oplus G_{i+1},p)$ the sum of the number of triples with the predicate $p$ that changed between pairs of consecutive versions.
+
+\ah{We don't need to define the interval in the latter function since we could project an interval from $\mathcal{G}$ when needed (more general and readable I think; e.g., $\Delta(\mathcal{G}[i,j],p)$).
+Also I don't get why you divide by Aggregated Predicate Multiplicity; doesn't this just scale all predicates by the same constant for an interval?
+I think it would perhaps make more sense to normalize by $\#(G_i \cup G_{i+1},p)$, or even just go down the route of Jaccard Similarity: $\Delta(\mathcal{G},p) \da \Sigma_{i=1}^{n-1}\frac{\#(G_i \cap G_{i+1},p)}{\#(G_i \cup G_{i+1},p)}$}
+
+%\begin{definition}[Predicate Multiplicity of an RDF Dataset]
+%	Given an RDF dataset $d$ and a predicate $p$, the multiplicity of $p$ in $d$, denoted by $M(p,d)$, is:
+%	\begin{equation}
+%	\label{eq:pm}
+%	M(p,d) = |\{(x,y,z) \in d : y = p \}|
+%	\end{equation}
+%\end{definition}
+%
+%
+%\begin{example}
+%	\label{ex:pm}    
+%	Considering Example~\ref{ex:dataset}, the Predicate Multiplicity of predicate \textit{image} in $d_6$ dataset is:    $M(image,d_6) = 2$
+%\end{example}
+%
+%Then, we must capture the change between two versions of an RDF dataset at the triples level, which represents triples added and deleted between the versions.
+For this purpose, we define a \textit{Difference Set} as an RDF dataset denoted by $\Delta(d_1, d_2)$ applying low-level change detection techniques.
+%
+%\begin{definition}[Difference Set]
+%	Let $d_1$ and $d_2$ be two RDF datasets; a Difference Set denoted by $\Delta(d_1, d_2)$ is:
+%	\begin{equation}
+%	\label{eq:ds}
+%	\Delta(d_1, d_2) = \{t \mid t \in d_2 \wedge t \notin d_1\} \cup \{t \mid t \in d_1 \wedge t \notin d_2\}
+%	\end{equation}
+%\end{definition}
+%
+%\begin{example}
+%	\label{ex:ds}
+%	We show the Difference Set between versions $d_1$ and $d_6$:\\
+%	\begin{center}    
+%		\begin{tabular}{lll}    \hline
+%			\multicolumn{3}{c}{$\Delta(d_1, d_6)$} \\    \hline
+%			-    :Atlantic\_Ocean  & :temperature   & 40$\degree$F          \\
+%			+    :Atlantic\_Ocean  & :temperature   & 60$\degree$F          \\
+%			+    :Atlantic\_Ocean  & :image         & AOcean2.png              \\    \hline
+%		\end{tabular}
+%	\end{center}
+%	
+%	We can now denote the Predicate Multiplicity of a Difference Set of $d_1$ and $d_2$ by $M(p, \Delta(d_1, d_2))$, e.g., $M(image, \Delta(d_1, d_6)) = 1$.
+%\end{example}
+%
+%
+%Following the intuition of our model, we now look for the multiplicity of a predicate $p$ of a dynamic dataset $\mathcal{D}$ in a time interval $[n,m]$.
+%
+%\begin{definition}[Aggregated Predicate Multiplicity of a dynamic dataset]
+%	Let $\mathcal{D} = (d_1, d_2, ..., d_m)$ be a dynamic dataset, and $p$ a predicate.
+The Aggregated Predicate Multiplicity of $\mathcal{D}$ in a time interval $[n,m]$, denoted by $AM(p, \mathcal{D},[n,m])$ is:
+%	\begin{equation}
+%	\label{eq:apm}
+%	AM(p, \mathcal{D},[n,m]) = \sum_{i=n}^{m-1}M(p, \Delta(d_i, d_{i+1}))
+%	\end{equation}
+%\end{definition}
+%
+%\begin{example}
+%	\label{ex:apm}
+%	$AM(temperature, \mathcal{D},[1, 7]) =$
+%	
+%	$ \sum_{i=1}^{6}M(temperature, \Delta(d_i, d_{i+1})) = 4$
+%	
+%	\begin{center}    
+%		\begin{tabular}{lll}    \hline
+%			\multicolumn{3}{c}{$\sum_{i=1}^{6}\Delta(d_i, d_{i+1})$} \\    \hline
+%			-    :Atlantic\_Ocean  & :temperature   & 40$\degree$F          \\
+%			+    :Atlantic\_Ocean  & :temperature   & 50$\degree$F          \\            
+%			-    :Atlantic\_Ocean  & :temperature   & 50$\degree$F          \\
+%			+    :Atlantic\_Ocean  & :temperature   & 60$\degree$F          \\
+%			+    :Atlantic\_Ocean  & :image         & AOcean2.png              \\    \hline
+%		\end{tabular}
+%	\end{center}
+%\end{example}
+%
+%We can now denote the Aggregated Multiplicity of a dynamic dataset in a time interval $[n,m]$ by:
+%
+%\begin{equation}
+%\label{eq:am}
+%AM(\mathcal{D},[n,m]) = \sum_{p \in P}AM(p, \mathcal{D},[n,m])
+%\end{equation}\\
+%Where $ P $ denotes the set of all predicates that have changed in the interval $[n,m]$.
+%
+%\begin{example}
+%	\label{ex:am}
+%	$AM(\mathcal{D},[1,7]) = AM(temperature, \mathcal{D},[1,7])$
+%	
+%	$+ AM(image, \mathcal{D},[1,7]) = 5$
+%\end{example}
+%
+%Finally, the \textit{Dynamics} of a predicate is given by the \textit{Aggregated Predicate Multiplicity of a dynamic dataset} $\mathcal{D}$ divided by \textit{Aggregated Multiplicity of a dynamic dataset}.
+%
+%\begin{equation}
+%\label{eq:dyn}
+%Dyn(p, \mathcal{D},[n,m]) = \frac{AM(p, \mathcal{D},[n,m])}{AM(\mathcal{D},[n,m])}
+%\end{equation}
+%
+%\begin{example}
+%	\label{ex:dyn}
+%	Considering our example:
+%	
+%	$ Dyn(temperature, \mathcal{D},[1,7]) = \frac{4}{5} = 0.8$
+%	
+%	$ Dyn(image, \mathcal{D},[1,7]) = \frac{1}{5} = 0.2$
+%	
+%	$ Dyn(area, \mathcal{D},[1,7]) = \frac{0}{5} = 0$
+%\end{example}
+%
+%Considering our proposed model, the Dynamics Model component receives the set of predicates extracted from the queries, calculates the $Dyn(p, \mathcal{D},[n,m])$ for each predicate $p$ and provides the value of a Transformation Function on these indicators.
+The intention of the Transformation Function is to estimate the dynamics of the query based on the Dynamics of the predicates.
+
+\begin{figure}[!t]
+{\centering
+	\begin{minipage}{0.32\textwidth}
+	\begin{lstlisting}[style=sparqld]
+:LMessi :name "L. Messi" ;
+ :instance_of :Human ;
+ :children 1 ;
+ :played :2014WC .
+	\end{lstlisting}
+	\end{minipage}	
+\hfill
+\begin{minipage}{0.32\textwidth}
+	\begin{lstlisting}[style=sparqld]
+:LMessi :name "L. Messi" ;
+ :children 2 ;
+ :played :2014WC, :2018WC .
+	\end{lstlisting}
+\end{minipage}	
+\hfill
+\begin{minipage}{0.32\textwidth}
+	\begin{lstlisting}[style=sparqld]
+:LMessi :name "L. Messi" ;
+ :instance_of :Human ;
+ :children 3 ;
+ :played :2014WC, :2018WC .
+	\end{lstlisting}
+\end{minipage}	
+}
+
+\caption{Example dynamic RDF graph $\mathcal{G} = (G_1, G_2, G_3)$ (from left to right, resp.) \label{fig:dg}}
+\end{figure}
+
+\begin{figure}[!t]
+{\centering
+\hfill
+	\begin{minipage}{0.35\textwidth}
+	\begin{lstlisting}[style=sparqld]
+:LMessi :instance_of :Human ;
+ :children 1 , 2 ;
+ :played :2014WC .
+	\end{lstlisting}
+	\end{minipage}	
+\hfill
+\begin{minipage}{0.35\textwidth}
+	\begin{lstlisting}[style=sparqld]
+:LMessi :instance_of :Human ;
+ :children 2 , 3 .
+	\end{lstlisting}
+\end{minipage}	
+\hfill
+}
+
+\caption{Changes for $\mathcal{G}$ in Figure~\ref{fig:dgc}, showing $G_1 \oplus G_2$ (left) and $G_2 \oplus G_3$ (right) \label{fig:dgc}}
+\end{figure}
 
 \begin{example}
-	\label{ex:pm}    
-	Considering Example~\ref{ex:dataset}, the Predicate Multiplicity of predicate \textit{image} in $d_6$ dataset is:    $M(image,d_6) = 2$
+Consider the example dynamic RDF graph $\mathcal{G}$ in Figure~\ref{fig:dg} (based on real data from Wikidata, with IRIs modified for the purposes of readability).
+In Figure~\ref{fig:dgc} we show the pairwise changes between each version: $G_1 \oplus G_2$ and $G_3 \oplus G_4$.
+For a predicate $p$, the value $\Delta(\mathcal{G},p)$ is then the sum of triples with predicate $p$ in the graphs $G_1 \oplus G_2$ and $G_2 \oplus G_3$.
+Looking at Figure~\ref{fig:dgc}, for example, \texttt{:name} does not appear ($\Delta(\mathcal{G},\texttt{:name})=0$), \texttt{:children} appears twice in $G_1 \oplus G_2$ and twice in $G_2 \oplus G_3$ ($\Delta(\mathcal{G},\texttt{:children})=4$), and so forth.
 \end{example}
-
-Then, we must capture the change between two versions of an RDF dataset at the triples level, which represents triples added and deleted between the versions.
-For this purpose, we define a \textit{Difference Set} as an RDF dataset denoted by $\Delta(d_1, d_2)$ applying low-level change detection techniques.
-
-\begin{definition}[Difference Set]
-	Let $d_1$ and $d_2$ be two RDF datasets; a Difference Set denoted by $\Delta(d_1, d_2)$ is:
-	\begin{equation}
-	\label{eq:ds}
-	\Delta(d_1, d_2) = \{t \mid t \in d_2 \wedge t \notin d_1\} \cup \{t \mid t \in d_1 \wedge t \notin d_2\}
-	\end{equation}
-\end{definition}
-
-\begin{example}
-	\label{ex:ds}
-	We show the Difference Set between versions $d_1$ and $d_6$:\\
-	\begin{center}    
-		\begin{tabular}{lll}    \hline
-			\multicolumn{3}{c}{$\Delta(d_1, d_6)$} \\    \hline
-			-    :Atlantic\_Ocean  & :temperature   & 40$\degree$F          \\
-			+    :Atlantic\_Ocean  & :temperature   & 60$\degree$F          \\
-			+    :Atlantic\_Ocean  & :image         & AOcean2.png              \\    \hline
-		\end{tabular}
-	\end{center}
-
-	We can now denote the Predicate Multiplicity of a Difference Set of $d_1$ and $d_2$ by $M(p, \Delta(d_1, d_2))$, e.g., $M(image, \Delta(d_1, d_6)) = 1$.
-\end{example}
-
-Following the intuition of our model, we now look for the multiplicity of a predicate $p$ of a dynamic dataset $\mathcal{D}$ in a time interval $[n,m]$.
-
-\begin{definition}[Aggregated Predicate Multiplicity of a dynamic dataset]
-	Let $\mathcal{D} = (d_1, d_2, ..., d_m)$ be a dynamic dataset, and $p$ a predicate.
-The Aggregated Predicate Multiplicity of $\mathcal{D}$ in a time interval $[n,m]$, denoted by $AM(p, \mathcal{D},[n,m])$ is:
-	\begin{equation}
-	\label{eq:apm}
-	AM(p, \mathcal{D},[n,m]) = \sum_{i=n}^{m-1}M(p, \Delta(d_i, d_{i+1}))
-	\end{equation}
-\end{definition}
-
-\begin{example}
-	\label{ex:apm}
-	$AM(temperature, \mathcal{D},[1, 7]) =$
-
-	$ \sum_{i=1}^{6}M(temperature, \Delta(d_i, d_{i+1})) = 4$
-
-	\begin{center}    
-		\begin{tabular}{lll}    \hline
-			\multicolumn{3}{c}{$\sum_{i=1}^{6}\Delta(d_i, d_{i+1})$} \\    \hline
-			-    :Atlantic\_Ocean  & :temperature   & 40$\degree$F          \\
-			+    :Atlantic\_Ocean  & :temperature   & 50$\degree$F          \\            
-			-    :Atlantic\_Ocean  & :temperature   & 50$\degree$F          \\
-			+    :Atlantic\_Ocean  & :temperature   & 60$\degree$F          \\
-			+    :Atlantic\_Ocean  & :image         & AOcean2.png              \\    \hline
-		\end{tabular}
-	\end{center}
-\end{example}
-
-We can now denote the Aggregated Multiplicity of a dynamic dataset in a time interval $[n,m]$ by:
-
-\begin{equation}
-\label{eq:am}
-AM(\mathcal{D},[n,m]) = \sum_{p \in P}AM(p, \mathcal{D},[n,m])
-\end{equation}\\
-Where $ P $ denotes the set of all predicates that have changed in the interval $[n,m]$.
-
-\begin{example}
-	\label{ex:am}
-	$AM(\mathcal{D},[1,7]) = AM(temperature, \mathcal{D},[1,7])$
-
-	$+ AM(image, \mathcal{D},[1,7]) = 5$
-\end{example}
-
-Finally, the \textit{Dynamics} of a predicate is given by the \textit{Aggregated Predicate Multiplicity of a dynamic dataset} $\mathcal{D}$ divided by \textit{Aggregated Multiplicity of a dynamic dataset}.
-
-\begin{equation}
-\label{eq:dyn}
-Dyn(p, \mathcal{D},[n,m]) = \frac{AM(p, \mathcal{D},[n,m])}{AM(\mathcal{D},[n,m])}
-\end{equation}
-
-\begin{example}
-	\label{ex:dyn}
-	Considering our example:
-
-	$ Dyn(temperature, \mathcal{D},[1,7]) = \frac{4}{5} = 0.8$
-
-	$ Dyn(image, \mathcal{D},[1,7]) = \frac{1}{5} = 0.2$
-
-	$ Dyn(area, \mathcal{D},[1,7]) = \frac{0}{5} = 0$
-\end{example}
-
-Considering our proposed model, the Dynamics Model component receives the set of predicates extracted from the queries, calculates the $Dyn(p, \mathcal{D},[n,m])$ for each predicate $p$ and provides the value of a Transformation Function on these indicators.
-The intention of the Transformation Function is to estimate the dynamics of the query based on the Dynamics of the predicates.
 

### 2019-01-26 kgswc19.tex
Finished pass up to end of Section 5

@@
 \definecolor{keyw}{RGB}{0,0,192}
+\definecolor{stri}{RGB}{100,0,130}
 \colorlet{numb}{magenta!60!black}
@@
 	{:}{{{\color{punct}{:}}}}{1}
-	{\{}{{{\color{delim}{\{}}}}{1} {\}}{{{\color{delim}{\}}}}}{1},
-	morekeywords={ask,select,from,where,order,by,distinct,limit,offset,optional,union,filter,prefix,bound,desc,regex,str,group,not,exists,minus,service,certain,maybe}
+	{\{}{{{\color{delim}{\{}}}}{1} {\}}{{{\color{delim}{\}}}}}{1}
+	{(}{{{\color{punct}(}}}{1} {)}{{{\color{punct})}}}{1},
+	morekeywords={ask,select,from,where,order,by,distinct,limit,offset,optional,union,filter,prefix,bound,desc,lang,regex,str,group,not,exists,minus,service,certain,maybe,coalesce,bind,as,strafter}
 }
@@
 	ndkeywordstyle=\color{greenCode}\bfseries,
-	stringstyle=\color{ocherCode}\ttfamily,
-	commentstyle=\color{darkgray}\ttfamily,
+	stringstyle=\color{stri}\ttfamily,
+	commentstyle=\color{gray}\ttfamily,
 	language={sparql},
@@
 	basewidth=0.5em,
+	comment=[l]{\#},
+	morestring=[b]",
 	moredelim=[is][\color{magenta}]{~}{~},
@@
 \paragraph{Architecture:} We propose a method to predict answers for the OSC and TTL problems.
-Our method primarily considers static features of the query: these include statistics such as the number of triple patterns, as well as the presence of query operators, such as \texttt{UNION}, \texttt{OPTIONAL}, specific path expressions, etc.; our model can also consider statistics about predicates in the dynamic RDF graph, as well as in which past versions of the dynamic graph the results for the query changed.
+Our method primarily considers static features of the query: these include statistics such as the number of triple patterns, as well as the presence of query operators, such as \texttt{UNION}, \texttt{FILTER}, specific path expressions, etc.; our model can also consider statistics about predicates in the dynamic RDF graph, as well as in which past versions of the dynamic graph the results for the query changed.
 Figure~\ref{fig:schema} shows the components of our approach, which receives as input a SPARQL query (the architecture itself already encodes information about the dynamic graph and the value of $k$ for OSC, which we set in experiments to 1, predicting changes in the next version).
-The Feature Extractor decomposes the query into a set of numeric features.
+The Feature Extractor decomposes the query into a set of numeric features, described presently.
 The Dynamics Model accepts some features of the query, as well as information from the dynamic graph, in order to compose further features.
@@
 
-The final feature -- a raw set of all predicates used -- is rather a meta-feature that serves as input for the next part of the proposed model.
+The final feature is the set of all predicates used in a triple pattern, including (where applicable) the IRIs used in a path expression; this is rather a meta-feature that serves as input for the next part of the proposed model.
 
@@
 }
-
 \caption{Changes for $\mathcal{G}$ in Figure~\ref{fig:dgc}, showing $G_1 \oplus G_2$ (left) and $G_2 \oplus G_3$ (right) \label{fig:dgc}}
@@
 
+Given a query $Q$ with a set of predicates $\{ p_i, ..., p_n \}$, we may then consider a variety of aggregate functions over $\Delta(\mathcal{G},p_1), ..., \Delta(\mathcal{G},p_n)$, such as $\mathrm{max}$, $\mathrm{mean}$, etc., to compute a final numeric feature for the query, representing a summary of the level of dynamicity of the predicates it contains.
+
 \section{Datasets}
@@
 
-We conducted an empirical study to evaluate the effectiveness of our proposal to predict change in SPARQL query results.
-For this, we need an RDF dataset, with a lot of information, preferably, cross-domain, and that is constantly updated since we assume that the information of the dataset is correct, complete, but dynamic.
-We also need a set of queries appropriate to the chosen dataset, preferably a heterogeneous set defined by real users and covering a large part of the dataset.
-
-\subsection{RDF Dataset}
-
-We use 23 Wikidata snapshots from 18/04/2017 to 27/09/2017, which are captured weekly in truthy version, which represents statements that have the best non-deprecated rank for a given property. 
-The first version has 1,102,242,331 triples, 57,499,191 unique IRIs, 3,276 unique predicates, 933,588,860 literals, and 7,449 blank nodes, while the final version has 1,924,967,162 triples (+74\%), 80,637,164 unique IRIs (+40\%), 3,660 unique predicates (+11\%), 1,676,930,476 literals (+79\%) and 8,909 blank nodes (+19\%).
-Figure~\ref{graph:datasize} shows the growth maintained over the period of some of these indicators.
-The gap in week 12 is due to the fact that we were not able to obtain the data for that week.
-Even though the information is relatively incremental, there are deletions.
-Figure~\ref{graph:delta} shows the ratio of triples eliminated and added each week.
-
-\begin{figure}[h]
-	\begin{minipage}[b]{0.40\linewidth}
+In order to evaluate the effectiveness of our proposal for predicting changes in SPARQL query results, we require a dynamic RDF graph, with access to various historical versions; preferably this graph contains real-world, large-scale, diverse data, and with sufficient changes between versions to provide both positive and negative examples of queries whose results change.
+Furthermore, we require a set of SPARQL queries that can be answered against this dynamic graph; preferably these queries again should be diverse, representative of real-world user queries, of a variety of shapes and sizes, using a variety of query operators, and with a mix of both dynamic and static results over the dynamic graph.
+In particular, we choose Wikidata~\cite{VrandecicK14} for our experiments, which, we shall argue, meets the aforementioned requirements.
+We first give details of the dynamic data we collect from Wikidata; thereafter, we discuss the queries we use in our experiments.
+
+\paragraph{RDF Data:} We use 23 Wikidata snapshots from 18/04/2017 to 27/09/2017, which are captured weekly in the \emph{truthy version} that contains triples (without qualifiers) that have the best non-deprecated rank for a given property.
+The first version has 1,102,242,331 triples and 3,276 unique predicates, while the final version has 1,924,967,162 triples (+74\%) and 3,660 unique predicates (+11\%).
+Figure~\ref{fig:wikidata}, on the left, shows the growth in triples as the weeks progress, while on the right, shows the numbers of triples added and removed week-to-week.
+We see that although many more triples are added than removed, there are some triples removed each week.
+We further note some peaks in triples added in some weeks, which may be due to bulk imports of data.
+The gap in week 12 is due to the fact that we were not able to obtain the data for that week (we thus exclude the triples added/removed from weeks 11 to 12, and 12 to 13).
+
+\ah{Regarding: ``The first version has 1,102,242,331 triples, 57,499,191 unique IRIs, 3,276 unique predicates, 933,588,860 literals, and 7,449 blank nodes, while the final version has 1,924,967,162 triples (+74\%), 80,637,164 unique IRIs (+40\%), 3,660 unique predicates (+11\%), 1,676,930,476 literals (+79\%) and 8,909 blank nodes (+19\%).'' I assume that the literals and blank nodes are not unique?
+Also the IRIs are unique per position?
+(In any case, such data is not 100\% needed; what's more important are the unique triples and unique predicates.)}
+
+\ah{Would it be possible to get a table of the top-10 most and least dynamic predicates by number of triples added/removed over the weeks, and also the number of triples with each such predicate in the first and last week?}
+
+\begin{figure}[t]
+	%\begin{minipage}[b]{0.47\linewidth}
 		\centering
-		\begin{tikzpicture}[thick, scale=0.5]
+		\begin{tikzpicture}[thick,font=\scriptsize]
 		\begin{axis}[
@@
 		xlabel={Version},
+		width=5cm, 
+		height=5cm,
 		ymin=0,
-		xtick={0,5,10,15,20,25},
+		xtick={1,2,...,23},
+		xticklabels={},
+		extra x ticks={1,23},
 		legend pos=north west,
@@
 		\end{tikzpicture}
-		\caption{Evolution of the number of triples on Wikidata.}
-		\label{graph:datasize}
-	\end{minipage}
-	\hspace{0.5cm}
-	\begin{minipage}[b]{0.40\linewidth}
-		\centering
-		\begin{tikzpicture}[thick, scale=0.5]
+%		\caption{Evolution of the number of triples on Wikidata.}
+%		\label{graph:datasize}
+%	\end{minipage}
+	\hspace{0.1cm}
+%	\begin{minipage}[b]{0.47\linewidth}
+%		\centering
+		\begin{tikzpicture}[thick,font=\scriptsize]
 		\begin{axis}[
+		xlabel={Version},
 		x tick label style={
@@
 		%ylabel=\#Triples,
-		enlargelimits=0.05,
-		legend style={at={(0.5,-0.1)},
+		width=9cm, 
+		height=5cm,
+		enlarge x limits=0.04,
+		enlarge y limits=0,
+		ymax=80000000,
+		legend style={at={(0.25,0.95)},
 			anchor=north,legend columns=-1},
-		ybar interval=0.7,
+		ybar=0pt,
+		xtick=data,
+		xtick align=inside,
+		bar width=3pt,
+		ymajorgrids,
+		grid style=dashed,
 		]
@@
 		\end{tikzpicture}
-		\caption{Triples added and removed in Wikidata.}
-		\label{graph:delta}
-	\end{minipage}
+\caption{Evolution of total triples in Wikidata (left) and number of triples added to and removed from the $n$\textsuperscript{th} version for the $n+1$\textsuperscript{th} version (right) \label{fig:wikidata}}
 \end{figure}
 
-\subsection{Queries Dataset}
-
-As opposed to RDF data, which can be easily obtained in the form of dumps, at the time of conducting this research, we did not find any structured dataset with queries following the Wikidata data model and Wikidata query logs where inaccessible.
-We considered then 388 queries of examples from Wikidata page\footnote{https://www.wikidata.org/wiki/Wikidata:SPARQL\_query\_service/queries/examples}.
-
-Then, the queries had to be curated manually, because they had syntax errors; others asked information external to Wikidata or asked for qualifiers that are not present in the truthy version.
-We also eliminated the queries that returned bindings to blank nodes to facilitate the comparison between the results.
-Therefore, we need deterministic queries.
-In our case, we eliminate the queries with the keyword SAMPLE and add ORDER BY to all the queries with all the projected variables.
-The latter change also facilitates the detection of changes in the results.
-Also, we removed the use of custom Label service of Wikidata\footnote{https://www.mediawiki.org/wiki/Wikidata\_Query\_Service/User\_Manual\#Label\_service}.
-
-Figures~\ref{fig:languageB} and \ref{fig:languageA} shows an example of several transformations of a query.
-
-\begin{lstlisting}[captionpos=b, caption=Query transformation. Before., label=fig:languageB,
-basicstyle=\ttfamily,frame=single]
+%\subsection{Queries Dataset}
+
+\paragraph{Queries:} In order to achieve a set of SPARQL queries that are answerable over Wikidata and with which we could run experiments, we took the user-contributed example queries from the Wikidata Query Service\footnote{https://www.wikidata.org/wiki/Wikidata:SPARQL\_query\_service/queries/examples}, consisting of 388 SPARQL 1.1 \texttt{SELECT} queries of varying degrees of complexity, touching upon various domains of data, and with a mix of varying query operators.\footnote{Very recently, a large log of millions of SPARQL queries have been released for Wikidata~\cite{MalyshevKGGB18}; these queries were not available when conducting our experiments, but would be interesting to analyze as part of future work.}
+
+Certain queries had to be removed from the original set of 388: some had syntax errors; others asked for information external to Wikidata (using \texttt{SERVICE}) or asked for qualifiers that are not present in the truthy version.
+We also eliminated a handful of queries that returned bindings to blank nodes to facilitate comparison between the results.
+Furthermore, we needed to transform certain queries for the purposes of our experiments.
+Some queries used features particular to the Wikidata Query Service, such as using a custom label \texttt{SERVICE} call to specify language preferences; we converted these queries to standard SPARQL using \texttt{COALESCE} to capture preferences.
+Furthermore, some queries featured non-deterministic elements -- such as \texttt{LIMIT}/\texttt{OFFSET} without ordering, \texttt{SAMPLE} or temporal functions -- such as \texttt{TODAY()} -- that may lead to changes in results not related to changes in the data; we remove queries with \texttt{SAMPLE} and temporal functions, and add an \texttt{ORDER BY} clause to all queries to ensure determinism with \texttt{LIMIT}/\texttt{OFFSET} and more generally to facilitate quick comparison of results.
+Figure~\ref{fig:transform} shows an example of a Wikidata query with the \texttt{LABEL} service rewritten and a deterministic ordering added.
+As a result of this process of filtering and transformation, we end up with a total of 221 deterministic queries.
+
+\begin{figure}[!t]
+\centering
+{
+	\begin{minipage}{0.8\textwidth}
+	\begin{lstlisting}[style=sparqld]
 #Cats
@@
  ?item wdt:P31 wd:Q146 .
-  SERVICE wikibase:label
-    { bd:serviceParam wikibase:language "en,es" }
+ SERVICE wikibase:label { bd:serviceParam wikibase:language "en,es" }
 }
 	\end{lstlisting}
-
-\begin{lstlisting}[captionpos=b, caption=Query transformation. After., label=fig:languageA,
-basicstyle=\ttfamily,frame=single]
+	\end{minipage}	
+}
+
+{
+\begin{minipage}{0.8\textwidth}
+	\begin{lstlisting}[style=sparqld]
 #Cats
@@
   ?item wdt:P31 wd:Q146 .
-  OPTIONAL {?item rdfs:label ?en.FILTER(LANG(?en)="en")}
-  OPTIONAL {?item rdfs:label ?es.FILTER(LANG(?es)="es")}
-  BIND(str(COALESCE(?en, ?es, strafter(str(?item),
-    "http://www.wikidata.org/entity/"))) AS ?itemLabel)
+  OPTIONAL { ?item rdfs:label ?en . FILTER(lang(?en)="en") }
+  OPTIONAL { ?item rdfs:label ?es . FILTER(lang(?es)="es") }
+  BIND(str(coalesce(?en, ?es, strafter(str(?item),
+    "http://www.wikidata.org/entity/"))) AS ?itemLabel )
 }
@@
 	\end{lstlisting}
-
-In the end, we have 221 curated queries.
-Then, we analyze the structure of the triple patterns of the queries.
-This is important, because our notion of freshness focuses on predicates.
-This restricts our approach to queries with triple patterns with a constant in the predicate position.
-Table~\ref{tab:Pattern} shows that in total 90.35\% of the triple patterns have a constant in the predicate position, which validate our approach of using a model based on the dynamics of predicates.
-
-\begin{table}[h]
-	\begin{minipage}[b]{0.45\linewidth}
-		\caption{Triple patterns (C=Constant, V=Variable).}
-		\begin{tabular}{lrr}    \hline
-			Pattern & Absolute & Relative \\    \hline
+\end{minipage}	
+}
+\caption{Query transformation before (above) and after (below) \label{fig:transform}}
+\end{figure}
+
+In Table~\ref{tab:Pattern}, we provide statistics on the distribution of different types of triple patterns in the 221 queries; of key importance is that 90.35\% of the triple patterns have a constant in the predicate position, meaning that they are compatible with a model based on the dynamics of predicates.
+We also see that 83 triple patterns (7.93\%) feature a path expression, where Table~\ref{tab:Path} provides the distribution of these expressions (remarking that multiple patterns can be used in one expression); of note here is that recursion is commonly found (74.70\% use $e\texttt{*}$ while 9.64\% use $e\texttt{+}$), and that we find no negated property paths.
+
+\begin{table}[t]
+	\begin{minipage}[b]{0.47\linewidth}
+		\centering
+		\caption{Distribution of triple patterns (C: Constant, V: Variable, P: Path). \label{tab:Pattern}}
+		\begin{tabular}{lrr}    \toprule
+			\textbf{Pattern} & \textbf{\ntextnumero} & \textbf{\%} \\    \midrule
 			V C V   & 751      & 71.73\%    \\
@@
 			C V C   & 1        & 0.10\%      \\
-			C C C   & 0        & 0.00\%        \\    \hline
+			C C C   & 0        & 0.00\%        \\    \midrule
 			V P C   & 69       & 6.59\%     \\
@@
 			C P V   & 1        & 0.10\%      \\
-			C P C   & 0        & 0.00\%        \\    \hline
+			C P C   & 0        & 0.00\%        \\    \bottomrule
 		\end{tabular}
-		\label{tab:Pattern}
 	\end{minipage}
-	\hspace{1cm}
-	\begin{minipage}[b]{0.45\linewidth}
-		\caption{Structure of navigational Property Paths.}
-		\label{tab:Path}
-		\begin{tabular}{lrr}    \hline
-			Rules                & Absolute & Relative \\    \hline
-			$P \rightarrow P* $   & 62       & 74.70\%     \\
-			$P \rightarrow P / P$  & 56       & 67.47\%    \\
-			$P \rightarrow P+ $   & 8        & 9.64\%     \\
-			$P \rightarrow P \mid P$  & 6        & 7.23\%     \\
-			$P \rightarrow P? $   & 6        & 7.23\%     \\
-			$P \rightarrow !(P)$ & 0        & 0.00\%        \\
-			$P \rightarrow C $    & 83       & 100.00\%      \\    \hline
+	\hfill
+	\begin{minipage}[b]{0.47\linewidth}
+		\centering
+		\caption{Distribution of paths\label{tab:Path}}
+		\begin{tabular}{lrr}    \toprule
+			\textbf{Pattern}      & \textbf{\ntextnumero} & \textbf{\%} \\    \midrule
+			$e\texttt{*}$   & 62       & 74.70\%     \\
+			$e_1 \texttt{/} e_2$  & 56       & 67.47\%    \\
+			$e\texttt{+} $   & 8        & 9.64\%     \\
+			$e_1 \texttt{|} e_2$  & 6        & 7.23\%     \\
+			$e\texttt{?}$    & 6        & 7.23\%     \\
+			$\texttt{!(}e\texttt{)}$ & 0        & 0.00\%        \\ \midrule
+			$e$ & 83 & 100.00\% \\
+			\bottomrule
+		%	$P \rightarrow C $    & 83       & 100.00\%      \\    \bottomrule
 		\end{tabular}
@@
 
-Table~\ref{tab:Pattern} also shows that 7.93\% of triple patterns contain a Property Path in the predicate position.
-Then we set out to analyze these triples more deeply.
-Table~\ref{tab:Path} shows the frequency of occurrence of these patterns in our queries.
-It shows that $ P \rightarrow P * $ (74.70\%) and $ P \rightarrow P + $ (9.64\%) are often applied, which corresponds to our intuition to have a feature that indicates if there are triple patterns offering arbitrary length paths (*PATH).
-We should also notice that there is no negation at the level of the path.
-Our feature extractor is able to extract all constants of predicates within paths.
-
-Finally, we evaluate the dynamics of the results of the queries considering Equation~\ref{eq:delta} between query results of the same query over consecutive versions of Wikidata.
-
-Figure~\ref{graph:changeV} shows the sum of the $ \delta $ function for all the queries considering the consecutive results of each query for 23 weeks.
-We identify that approximately half of the results of the queries change every week.
-
-Figure~\ref{graph:changeH} shows for each query how many changes registered over 23 weeks.
-We identify queries that change every week, others that never changed and a fairly uniform distribution of changes in the period.
-
-\begin{figure}[h]
-	\begin{minipage}[b]{0.40\linewidth}
+Finally, we look at how the results for these queries change over the 23 weeks.
+Figure~\ref{graph:changeV} shows how many queries had some results change between two consecutive versions, where we see that approximately half of the queries see changes in their results each week.
+But are these always the same queries that change every week?
+Figure~\ref{graph:changeV} shows the number of weeks in which some result changed for each query; the queries are ordered by the number of changes in their results, where we can see that the results of 44 queries (19.90\%) change each time, 18 queries (8\%) never change, and more generally, we note a quite uniform distribution of queries in between.
+More generally speaking, we conclude that our query set has a good balance of queries whose results never change, queries whose results always change, and queries whose results sometimes change.
+
+\begin{figure}[t]
+	\begin{minipage}[b]{0.47\linewidth}
 		\centering
-		\begin{tikzpicture}[thick, scale=0.5]
+		\begin{tikzpicture}[thick,font=\scriptsize]
 		\begin{axis}[
-		%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
 		xlabel={Version},
-		ylabel={Changes},
+		ylabel={Queries changed},
+		ylabel near ticks,
+		width=6.5cm, 
+		height=5cm,
 		ymin=0, ymax=221,
-		xtick={0,5,10,15,20,25},
+		xtick={1,2,...,23},
+		xticklabels={},
+		extra x ticks={1,23},
 		legend pos=north west,
@@
 		\end{tikzpicture}
-		\caption{Query results changes by time.}
+		\caption{Number of queries with some results changed for each version}
 		\label{graph:changeV}
@@
 	\hspace{0.7cm}
-	\begin{minipage}[b]{0.40\linewidth}
+	\begin{minipage}[b]{0.47\linewidth}
 		\centering
-		\begin{tikzpicture}[thick, scale=0.5]
+		\begin{tikzpicture}[thick,font=\scriptsize]
 		\begin{axis}[
-		%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
-		xlabel={Queries},
+		xlabel={Query},
+		ylabel={Versions changed},
+		ylabel near ticks,
+		width=6.5cm, 
+		height=5cm,
 		ymin=0,
@@
 		\end{tikzpicture}
-		\caption{Query results changes by queries.}
+		\caption{Number of versions in which some results changed for each query}
 		\label{graph:changeH}

### 2019-01-27 kgswc19.tex
Finished first pass over the entire paper

@@
 
-\section{Datasets}
+\section{Gold Standard Dataset}
 \label{sec:data}
@@
 The gap in week 12 is due to the fact that we were not able to obtain the data for that week (we thus exclude the triples added/removed from weeks 11 to 12, and 12 to 13).
+The dynamic graph thus considers a total of 32.3 billion triples across 23 versions.
 
@@
 		ymin=0,
-		xtick={1,2,...,23},
+		xtick={1,2,...,24},
 		xticklabels={},
-		extra x ticks={1,23},
+		extra x ticks={1,24},
 		legend pos=north west,
@@
 		color=blue,
-		mark=square,
+		mark=diamond,
 		]
@@
 		\addplot[
-		color=blue,
-		mark=square,
+		color=blue
 		]
@@
 
+\paragraph{Feature extraction:} To compute the query features, we use Apache Jena to parse the query and extract the necessary statistics and determine the presence of the relevant query operators.
+In order to extract statistics on the dynamics of individual predicates in the dynamic graph, a challenge here is scalability, since we work with a total of 32.3 billion triples; hence we (1) sort each version of Wikidata, (2) apply a merge-sort iterator over each pair of consecutive versions ($G_i$ and $G_{i+1}$) to detect triples that changed ($G_i \oplus G_j$), writing a separate file for triples that were deleted ($G_i - G_j$) and added ($G_j - G_i$), (3) from these files, we can then compute and sum the number of triples changed for each predicate between each pair of versions.
+Finally, in order to create the ground truth in terms of which results change between which versions, we index each version in Virtuoso and compute the results for each query against each version and write them to disk; we then compare consecutive pairs of results with a merge-sort.
+
 \section{Evaluation}
@@
 
-We evaluate the performance of our proposal to predict changes in the results of our queries. For this, we use the Wikidata dataset and our query dataset explained previously.
+%We evaluate the performance of our proposal to predict changes in the results of our queries.
+For this, we use the Wikidata dataset and our query dataset explained previously.
 To build the Difference Sets ($\Delta$) for our model we used an approach based on a merge-sort to scan over the snapshots as follows:
-
-\begin{enumerate}
-	\item Sort all statements by their syntactic order (subject-predicate-object).
-	\item Perform a pairwise comparison of the statements by scanning two snapshots in linear time.
-	\item Trigger a detection of the change as soon as the order of the statements differs between two snapshots.
-\end{enumerate}
-
-Then we configure the parameters of our model.
+%
+%\begin{enumerate}
+%	\item Sort all statements by their syntactic order (subject-predicate-object).
+%	\item Perform a pairwise comparison of the statements by scanning two snapshots in linear time.
+%	\item Trigger a detection of the change as soon as the order of the statements differs between two snapshots.
+%\end{enumerate}
+
+Our experiments aim to ascertain the (relative) quality of predictions that can be made in the context of One Shot Change (OSC) and Time To Live (TTL) based on query features, statistics of predicate dynamics, as well as information about the historical evolution of queries.
+\ah{Update with research questions later}
+
+\paragraph{Setting:} Given that our tasks are prediction based, we must thus build a training set and a test set.
+With 23 versions, we must hold out at least one version for testing, leaving a maximum interval of 22 past versions to use for predictions.
+However, the more versions we add to interval of past versions considered, the fewer examples we can generate; for example, if we consider 22 versions for training, we will only have one ground truth label for each query.
+In the end, we opted to consider the 3 previous versions ($G_{i-3}, G_{i-2}, G_{i-1}$) to predict changes for the subsequent version ($G_i$); this allows us to predict from $G_4$ to $G_{23}$ inclusive, providing 20 examples per query and 4420 examples in total.
+We remark that no feature explicitly indicates which specific version numbers are given and which are to be predicted.
+(\ah{Discuss the gap: \textit{We evaluated the quality of our proposal by forming a set of classifiers to predict the change of 221 queries in 15 weeks, which is equivalent to 3315 objects to classify, described in terms of our proposed features.}}) \ah{Could we do experiments with different intervals?}
+
+In the case of predicate dynamics, as discussed in Section~\ref{sec:approach}, there are potentially many predicates per query, each with its own value for $\Delta(\mathcal{G},p)$, where to reduce (and fix) dimensionality, we may apply an aggregate function to choose the min, max or mean value over all predicates.
+In preliminary experiments, we found the mean value to offer the best results, followed by the max value; hence in what follows we adopt the mean value in our experiments.
+
+We tested the Pearson correlation ($\rho$) between the features available in our data.
+We found only one pair of highly correlated variables: \textbf{\textnumero\ of triple patterns} and \textbf{\textnumero\ of variables} ($\rho = 0.943$).
+\ah{Not sure about this part: what does it add?
+What does ``highly correlated mean?'' Also it's surprising that, e.g., \textbf{\textnumero of predicates} is not also highly correlated here.}
+
+For OSC prediction, we selected four classifiers for testing: Decision Trees, Linear SVM, Naive Bayes and Nearest Neighbors.
+We split the data to use 80\% for training and 20\% for tests.
+To avoid overfitting, we use cross-validation in the training data.
+We consider three types of features: static query features (\textsc{q}): predicate dynamics (\textsc{p}) and information on historical results (\texttt{r}).
+\ah{We should be more explicit on precisely what \textsc{r} is?} We may also consider combinations of these features: \textsc{qp}, \textsc{qs}, \textsc{pr} and \textsc{qpr}.
+In our experiments, we compare \textsc{qp} with \textsc{r}.
+\ah{Is this correct?
+Or would it be \textsc{qr}?} \ah{Why only 2?
+Could we not compare all seven?}
+
+%In the case of predicate dynamics, as discussed in Section~\ref{sec:approach}, there are potentially many predicates per query, each with its own value for $\Delta(\mathcal{G},p)$, where to reduce (and fix) dimensionality, we may apply an aggregate function to choose the min, max or mean value over all predicates.
+In preliminary experiments, we found the mean value to offer the best results, followed by the max value; hence in what follows we adopt the mean value in our experiments.
+
+%Then we configure the parameters of our model.
 For the time interval, we consider three versions of the dataset equivalent to the last three weeks.
@@
 For example, Maximum will take the dynamics of the most dynamic property (used as a predicate in the query) as a feature.
-In addition, given that existing approaches are based on the historical results of the queries, we evaluate the substitution of our model for a historical change indicator that counts how many times the results have changed in the last three weeks.
-
-In addition, we train a Machine Learning model to determine the life spam of a query results.
+%In addition, given that existing approaches are based on the historical results of the queries, we evaluate the substitution of our model for a historical change indicator that counts how many times the results have changed in the last three weeks.
+
+In addition, we train a Machine Learning model to determine the TTL of a query's results.
 To train the model we consider the changes observed in the monitored interval.
@@
 The target TTL value that we learn is the time interval during which the results of the query remain fresh from the current update time.
-
-\subsection{Results}
-
-First, we consider the Pearson correlation between the characteristics.
-We found only a couple of highly correlated variables (\# triples and \# vars: 0.943).
-The influence of the use of various Transformation Functions in our model was also evaluated.
-In this case, Average was the Transformation Function that obtained the best result followed by Maximum.
-
-We tested a list of classifiers, including Decision Tree, Nearest Neighbors, and Naive Bayes among others listed in Table~\ref{tab:classifier}.
-Our pilot experiment shows that Nearest Neighbors and Decision Tree work better than the others.
-We evaluated the quality of our proposal by forming a set of classifiers to predict the change of 221 queries in 15 weeks, which is equivalent to 3315 objects to classify, described in terms of our proposed features.
-We split the data 80\% for training and 20\% for tests.
-To avoid overfitting, we use cross-validation in the training data.
-
-Table~\ref{tab:classifier} shows the quality values of the classifiers evaluated in terms of $F_1$.
-To have a measure of the positioning of our proposal, we added a new feature that counts how many times the results of the query change on average in the last three versions.
-Then, we did two executions with our syntactic features: one using this new indicator, and the second with our dynamic model.
-Table~\ref{tab:classifier} shows that our algorithm achieved less quality.
-Despite obtaining lower results, the quality of our proposal is not far from that achieved by observing the results of queries in the past, but certainly, our approach is less expensive since it does not require knowledge of past query results.
-
-Finally, to predict the time that the results of a query will remain invariant (TTL), a continuous value (number of weeks) will be determined.
-For that we employed techniques of multiple linear regression using the sklearn and tensorflow library.
+\ah{This paragraph is a little vague/unclear to me.}
+
+\paragraph{Results:} Table~\ref{tab:classifier} shows the $F_1$-measure for the predictions made; for reference, we include a baseline that randomly guesses yes/no.
+The results show that all configurations clearly outperform the random baseline; however, with the highest result being an $F_1$-measure of 0.706, we can also see that predicting changes in query results is indeed a challenging problem.
+Looking in further detail, as can be expected, prediction performs best when the \textsc{r} feature -- with knowledge of changes in the historical value of results of the input query -- is available.
+However, the \textsc{qp} features -- which do not assume the availability of such information -- are often competitive: more specifically, the best result for \textsc{r} is 0.706 with Linear SVM, while the best results for \textsc{qp} is 0.651 with Decision Trees.
+
+Finally, to predict the TTL of the results of the query -- outputting the number of versions for which the query results can be expected to remain the same -- we employed techniques of multiple linear regression using the \textsc{sklearn} and \textsc{tensorflow} libraries.
 The latter allows us to see how the periods of estimation and the convergence analysis will behave.
-
-We were able to predict the time of life (in relation to the number of invariant weeks) with an average square error of 0.43 and a variance score of 0.01.
+We were able to predict the TTL (in relation to the number of invariant weeks) with an average square error of 0.43 and a variance score of 0.01.
 These results allow to predict future changes in query results with good accuracy using linear regression techniques.
-
-\begin{table*}[h]
+\ah{Could do with a comparison here; the 0.43 result is quite abstract.}
+
+\begin{table*}[t]
 	\centering
-	\caption{Tested classifiers.}
+	\caption{$F_1$-measure for tested classifiers on OSC}
 	\label{tab:classifier}
-	\begin{tabular}{|l|r|r|}    \hline		
-		Classifier                   & Past Results & Dynamics Model \\    \hline
-		Base Dummy 					 & 0.497869 & 0.486923 \\
-		Nearest Neighbors            & 0.701587 & 0.635923 \\
-		Decision Tree                & 0.695591 & 0.651462 \\
-		Linear SVM                   & 0.705755 & 0.593846 \\
-		Naive Bayes                  & 0.661239 & 0.587692 \\    \hline
+	\begin{tabular}{lrr}    \toprule	
+		\textbf{Classifier}                   & \textsc{r} & \textsc{qp} \\    \midrule
+		\textit{Random Baseline}				 & 0.497869 & 0.486923 \\\midrule
+		Decision Trees                & 0.695591 & \underline{0.651462} \\
+		Linear SVM                   & \underline{0.705755} & 0.593846 \\
+		Naive Bayes                  & 0.661239 & 0.587692 \\ 
+		Nearest Neighbors            & 0.701587 & 0.635923 \\   \bottomrule
 	\end{tabular}
@@
 
-In this paper we propose classifiers that predict whether or not the results of a query will change in the short time interval; for this, we propose a model based on the dynamics of the predicates in the datasets and a set of features extracted from the syntax of the queries.
-The proposed framework is flexible for incorporating new features extracted from the queries, from the current results, or from the model based on the multiplicity of predicates.
-
-We select the datasets to evaluate our proposal considering desirable characteristics, such as availability, size, dynamics, and integrity.
-Choosing Wikidata, we present statistics to understand such characteristics.
-In the case of the queries dataset, we did not find any existing dataset of queries that complied with the requirements, so we curated a query dataset for Wikidata based on query examples proposed by users.
-
-Finally, we evaluate our proposal, making a weekly prediction of change in the results of the queries and evaluating the success rates of the different classifiers each week based on F-measure.
-We also apply linear regression techniques to predict, specifically, when the results of a query will change.
-We measure the effectiveness of the characteristic based on our model of predicate dynamics versus a characterization of dynamics based on the historical results.
-The latter is a common factor in the previous proposals, but it requires a high cost.
-Our results show that our approach using predicate dynamics is competitive for prediction while being less expensive to support.
-
-Although the results obtained are encouraging, they are still preliminary.
-We consider for future work evaluating the quality of our proposal over a longer period.
-In addition, we believe that more complex query features can improve the results, such as considering operator semantics; for example, to see if negation is affecting predicates or if the predicates are affected by the existence of filters.
-On the other hand, it would be worth expanding the set of queries in the test dataset, in order to increase the size, complexity, and coverage.
-Recently a query dataset for Wikidata was published\footnote{https://lists.w3.org/Archives/Public/semantic-web/2018Aug/0023.html}.
-It would be good to evaluate this query dataset, as well as consider a probabilistic approach based on query algebra (rather than syntax).
+In this paper we propose methods to predict whether or not (OSC) and/or when (TTL) the results of an input query will change in the future.
+Such methods have applications for synchronization, caching, and more besides.
+More specifically, we propose a framework based on classifiers and regression that accept features extracted from the queries, from past versions of the data, or from the combination of both.
+Considering this framework, we propose that there is a conceptual trade-off between the cost of computing features and their value for prediction: features extracted from queries alone are the most efficient to extract, not requiring historical data, but are quite coarse-grained for prediction; on the other hand, features based on changes in results in previous versions for an (unseen) input query are often the most costly to acquire, but offer fine-grained information for prediction; finally, features based on the dynamics of predicates used in the data offer a balance between the two, allowing to summarize historical data into succinct statistics, while offering more fine-grained information than static query features and less fine-grained information than historical results.
+
+Our goal then is to explore this trade-off experimentally.
+For this, we define some desirable characteristics for the queries and data used in the experimental setting, such as availability, size, dynamics, diversity, etc.
+Settling on Wikidata, we collect 23 versions and curate 221 user-generated queries from which we compute results, statistics and ultimately features that form the gold standard for the tasks.
+Analysing these data and queries, we conclude that they meet the desired characteristics, offering a large-scale, diverse, real-world setting; more importantly, we find that our queries offer an ideal distribution of queries whose results always change, queries whose results never change, and (more importantly) queries whose results change intermittently and for which prediction is thus a non-trivial task.
+We make this gold standard available at \ah{...} -- including queries, results and features -- so that it can be reused in further research.
+
+Finally, we use this gold standard dataset to evaluate the trade-off identified between different types of features for predictions in the OSC and TTL tasks.
+For OSC, we evaluate the success rates of a variety of classifiers based on the $F_1$-measure of the predictions.
+Our results show that although features based on historical changes to query results perform best ($F_1 = 0.706$), considering static query features and predicate dynamics alone is competitive ($F_1 =0.651$) when historical results are not available or are prohibitively costly to compute.
+
+In terms of future work, gold standards based on datasets other than Wikidata could be pursued.
+Also, our gold standard based on Wikidata could be extended to consider more versions spanning a longer period of time and/or more queries; as previously mentioned, a very large query dataset for Wikidata was recently published~\cite{MalyshevKGGB18} from which it should be possible to extend the Wikidata gold standard (the challenge being to keep a balance of static and dynamic queries with diverse features).
+Concerning the prediction tasks themselves, there are a variety of other features and approaches that could be applied; one promising direction may be, for example, to apply a more fine-grained analysis of the query, considering for example estimates of the selectivity of particular triples patterns, and combining this information with statistics on the dynamics of the data and the query operators used; our current gold standard could be used to validate such developments.
+Finally, it would be interesting to investigate the effectiveness of these techniques in practice, developing caching systems, synchronization schedules, and other applications, based on the proposed predictions.
+
+\ah{Some reference can be shortened to just expand the acronym, no more, for example; \textit{Advances in Information Retrieval -- 34th European Conference on IR Research, ECIR, 2012}; can become \textit{European Conference on IR Research, ECIR, 2012}.
+Addresses and dates are not necessary (keep the year of course).
+The workshop name is enough; no need to also give the conference.
+Keeping the publisher for conference is good, but maybe IEEE Computer Society can become IEEE.}
 

### 2019-01-28 kgswc19.tex
Second quick pass on the paper

@@
 \usepgfplotslibrary{clickable}
+\usetikzlibrary{shapes,arrows,positioning,fit,backgrounds,matrix,chains,patterns,trees,calc,arrows.meta}
 
@@
 \begin{abstract}
-In this work, we tackle the problem of estimating future changes in SPARQL query results over dynamic datasets; such work has applications for caching and remote synchronization.
-Our proposal is based on syntactic query features and (optionally) knowledge of the dynamics of predicates in the underlying dataset; the resulting model can be used to predict (1) whether or not the results for a query will change in the near future using classifiers, and (2) when they might change (if at all) using linear regression.
-To evaluate our proposal, we select Wikidata as our dataset and build a curated collection of real-world SPARQL queries.
-Our results show that the accuracy of the predictions of our proposed model is competitive with that obtained using knowledge of the complete historical evolution of the query results but at a lower cost.
+We address the problem of estimating when the results of an input SPARQL query results over dynamic RDF datasets will change.
+We propose a framework that extracts features from the query and/or from past versions of the target dataset.
+The resulting features are then used to predict (1) whether or not the results for a query will change in the near future using classifiers, and (2) when they might change (if at all) using linear regression.
+To evaluate our proposal, we create a gold standard based on 23 versions of Wikidata and a curated collection of 221 SPARQL queries.
+Our results show that the quality of predictions using query features and lightweight statistics of the dynamics of individual predicates is competitive with that obtained using (more costly to derive) knowledge of the complete historical changes in the query results.
 
@@
 Recent works have explored the dynamics of Linked Data with the intention of finding patterns that allow for characterizing, recognizing, and predicting changes based on analysis of the domains, predicates, and schema~\cite{UmbrichHHPD10,KaferAUOH13,NishiokaS16,NishiokaS17,GonzalezH18}.
-Among these works, we can find hybrid approaches that have been developed to return fresher query results with faster response times by decomposing a query into dynamic sub-queries answered remotely, and static sub-queries answer over local caches~\cite{UmbrichKHP12}, but such an approach only focuses on estimating the dynamics of individual triple patterns, and does not consider, for example, query operators.
+Among these works, we can find hybrid approaches that have been developed to return fresher query results with faster response times by decomposing a query into dynamic sub-queries executed remotely, and static sub-queries executed over local caches~\cite{UmbrichKHP12}, but such an approach only focuses on estimating the dynamics of individual triple patterns, and does not consider, for example, query operators.
 More generally, coping with changes in remote data still presents a major challenge for applications leveraging dynamic Linked Data.
 
-In this work, we tackle the problem of estimating future changes in query results based on query features, high-level statistics about the dynamics of predicates, as well as classification and regression methods from Machine Learning.
-We thus propose a solution where we do not have complete historical information about the results of a query, but rather a much more lightweight statistical model that saves on space and obviates the need to compute query results for multiple past versions.
-In order to assess the predictions made by our proposal, we perform experiments over 23 weekly versions of the Wikidata knowledge graph~\cite{VrandecicK14}, evaluating various configurations in terms of the accuracy for predicting future changes in the query results; we select Wikidata as: (1) it provides a history of weekly versions that we can use to evaluate predictions; (2) it is edited by thousands of users, which means that significant changes are observed weekly; (3) it provides a large-scale diverse dataset; (4) it receives millions of queries on a daily basis~\cite{MalyshevKGGB18}; and (5) a diverse set of exemplar SPARQL queries has been proposed by the community that we can use for our evaluation.
-
-The contributions of this paper are thus as follows: (i) we propose a supervised method that uses static query features and suitable classifiers to predict whether or not results will change in a given point in the near future; (ii) we propose to augment this model with lightweight statistics of the dynamics of predicates; (iii) we further experiment with linear regression models to predict when results will change; (iv) we prepare a benchmark for predicting the dynamics of query results based on 23 weekly versions of Wikidata and a set of 221 real-world SPARQL queries; (v) we use this benchmark to compare the results obtained by our lightweight models versus an ``clairvoyant model''~\cite{KnuthHS16} that knows of changes in the query's results for previous versions.
-Our results show that our proposed models are competitive with the costly clairvoyant model.
-
-The rest of this document is organized as follows: In Section~\ref{sec:related}, we review related works.
-Section~\ref{sec:preliminar} provides a formal definition of the problem.
-Section~\ref{sec:approach} describes our proposed models.
-In Section~\ref{sec:data}, we describe our evaluation data and queries based on Wikidata.
+In this work, we address the problem of estimating if and/or when SPARQL query results will change in a future version of a dynamic RDF dataset.
+We propose a general architecture based on Machine Learning that extracts static features from a query, as well as dynamic features from the query and dataset.
+These features are fed into a classifier to predict whether or not the query results will change in a fixed point in the future, or, more ambitiously, can be fed into a regression model to predict when the query results are likely to change.
+With respect to the features used, we propose that there is a trade-off between those that are coarse-grained but easy to compute (e.g., static query features) versus those that are fine-grained but more costly to compute (e.g., historical changes in query results), with further features falling in the middle (e.g., high-level statistics about the dynamics of triples with particular predicates).
+Per this trade-off, which features to use for predicting the dynamics of query results may then depend on the particular application.
+
+In order to better understand this conceptual trade-off -- and more generally, to evaluate the quality of predictions made by our framework -- we create a novel gold standard based on 23 weekly versions of the Wikidata knowledge graph~\cite{VrandecicK14} and 221 user-generated queries; we show this gold standard to have a variety of desirable features, including (most importantly) a balance of queries whose results never change, always change, as well as non-trivial queries whose results intermittently change.
+Using this gold standard, our experiments show that while the most accurate predictions are based on historical changes to the input query's results, predictions based only on query features and predicate dynamics remain competitive (while being much more efficient to maintain/extract).
+
+\paragraph{Contributions:} (1) We propose a general architecture for predicting when/if the results of a SPARQL query will change in a future version of a dynamic RDF dataset.
+(2) We propose a number of features to instantiate this architecture based on analysis of the query, analysis of the dynamics of predicates in the data, and analysis of historical changes in the query results.
+(3) We create a gold standard for these tasks based on 23 weekly versions of Wikidata and a set of 221 real-world SPARQL queries.
+(4) We use this gold standard to compare the predictions obtained using different types of features and classifiers.
+
+\paragraph{Paper structure:} In Section~\ref{sec:related}, we review related works.
+Section~\ref{sec:preliminar} provides a formal statement of the problem.
+Section~\ref{sec:approach} describes our proposed architecture and features.
+In Section~\ref{sec:data}, we describe the creation of our gold standard based on Wikidata.
 Section~\ref{sec:eval} presents the design and results of the experiments.
@@
 Dehghanzadeh et al.~\cite{DehghanzadehPKUHD14} later proposed a method to estimate the freshness using cardinality estimation techniques based on predicates and characteristic sets.
-Combining the notion of subscription-based notifications and predicting dynamics, Knuth et al.~\cite{KnuthHS16} propose a middleware to which consumers may subscribe and that periodically ranks and schedules refreshes for queries according to policies that take into account how likely the results are to be stale, how long ago the query was last refreshed, how many results previously changed, how long the query takes to run, etc.
-
-Our proposal aims to model the dynamics of SPARQL query results in order to predict (1) whether or not the results will have changed in a fixed point in the near future and,  (2) when the results will change.
-Our work thus complements existing works aiming for weak consistency, but rather investigates the potential of using static features of a query as a lightweight alternative with the intuition that, for example, queries with more triple patterns or certain query operators are more likely to have dynamic results.
+Combining the notion of subscription-based notifications and predicting dynamics, Knuth et al.~\cite{KnuthHS16} propose a middleware to which consumers may subscribe that periodically ranks and schedules refreshes for queries according to policies that take into account how likely the results are to be stale, how long ago the query was last refreshed, how many results previously changed, how long the query takes to run, etc.
+
+\paragraph{Novelty:} Given a query and a dynamic RDF dataset, we aim to predict (1) whether or not the query's results will have changed in a fixed point in the near future and/or (2) when the results will change.
+Our work thus complements existing works aiming for weak consistency, but (i) generalizes the problem, proposing a framework that can incorporate statistics on predicate dynamics~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12,DehghanzadehPKUHD14} and historical changes in query results~\cite{KnuthHS16}, as proposed in prior works, as well as novel types of features, (ii) introduces new features based on query operators and statistics; (iii) creates a novel gold standard based on Wikidata and presents comparative results that indicate the relative predictive power inherent in different types of features.
+Furthermore, to the best of our knowledge, no prior work uses Machine Learning models to predict changes in SPARQL query results.
+\ah{Is this last claim true?
+We should try to be as sure as possible.}
+
+%investigates the potential of using static features of a query as a lightweight alternative with the intuition that, for example, queries with more triple patterns or certain query operators are more likely to have dynamic results.
 For the purposes of comparison, we complement these static features with further knowledge of the dynamics of predicates~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12,DehghanzadehPKUHD14} and knowledge of how results for the input query have changed over past versions~\cite{KnuthHS16}, as considered in previous works.
@@
 We further remark that, to the best of our knowledge, no existing work aims to predict the dynamics of SPARQL query results using Machine Learning models.
+\ah{I should make this more concise later.}
 
@@
 The semantics of a SPARQL \texttt{SELECT} query $Q$ is defined in terms of its \emph{evaluation} over an RDF graph $G$, denoted $Q(G)$, giving a set of partial mappings from projected variables to the set $\Iris \cup \Lits \cup \Bnodes$; we refer to P{\'{e}}rez et al.~\cite{PerezAG09} for definitions.
-(According to the standard, a SPARQL query is evaluated over a set of \emph{named graphs}; however, since the Wikidata query service indexes data in one RDF graph, to keep discussion concise, we consider RDF graphs, though our methods can be generalized in a straighforward manner to RDF datasets/named graphs.)
+(According to the standard, a SPARQL query is evaluated over a set of \emph{named graphs}; however, since the Wikidata query service indexes data in one RDF graph, to keep discussion concise, we consider RDF graphs, though our methods can be generalized in a straightforward manner to the named graphs setting.)
 
@@
 One key feature in this extension is that of \textit{property paths}, which allows for matching arbitrary length paths in an RDF graph, potentially returning or matching the endpoints of the path.
-An IRI $p$ is a \emph{path expression}; if $e$, $e_1$ and $e_2$ are path expressions, then $\texttt{\^{}}e$ (inverse of $e$), $e_1 \texttt{/} e_2$ ($e_1$ followed by $e_2$), $e_1 \texttt{|} e_2$ ($e_1$ or $e_2$), $e\texttt{*}$ (zero or more $e$), $e\texttt{+}$ (one or more $e$), $e\texttt{?}$ (zero or one $e$), and $\texttt{(}$e$\texttt{)}$ (parentheses used for precedence) are also \textit{path expressions}; finally, if $p$, $p_1$, $\ldots$ $p_n$ are IRIs, then $\texttt{!}p$, $\texttt{!}\texttt{(}p_1 \texttt{|} \ldots \texttt{|} p_n\texttt{)}$ and $\texttt{!}\texttt{(}p_1 \texttt{|} \ldots \texttt{|} p_k \texttt{|}  \texttt{{\^{}}}p_{k+1} \texttt{|} \ldots \texttt{|} \texttt{{\^{}}}p_n\texttt{)}$ for $k+1 \leq n$ (negated property sets) are path expressions.
+An IRI $p$ is a \emph{path expression}; if $e$, $e_1$ and $e_2$ are path expressions, then $\texttt{\^{}}e$ (inverse of $e$), $e_1 \texttt{/} e_2$ ($e_1$ followed by $e_2$), $e_1 \texttt{|} e_2$ ($e_1$ or $e_2$), $e\texttt{*}$ (zero or more $e$), $e\texttt{+}$ (one or more $e$), $e\texttt{?}$ (zero or one $e$), and $\texttt{(}$e$\texttt{)}$ (parentheses used for precedence) are also \textit{path expressions}; finally, if $p$, $p_1$, $\ldots,$ $p_n$ are IRIs, then $\texttt{!}p$, $\texttt{!}\texttt{(}p_1 \texttt{|} \ldots \texttt{|} p_n\texttt{)}$ and $\texttt{!}\texttt{(}p_1 \texttt{|} \ldots \texttt{|} p_k \texttt{|}  \texttt{{\^{}}}p_{k+1} \texttt{|} \ldots \texttt{|} \texttt{{\^{}}}p_n\texttt{)}$ for $k+1 \leq n$ (negated property sets) are path expressions.
 Thereafter, a \textit{path pattern} $(s,e,o)$ where $e$ is a path expression is an expression.
@@
 The first such problem we consider -- which we call One Shot Change (OSC) -- accepts $\mathcal{G}$, $Q$ and a positive integer $k$ as input and outputs a boolean value predicting whether or not the results will change from $G_n$ to some future version $G_{n+k}$ (i.e., $Q(G_n) \neq Q(G_{n+k})$?).
-The second (more difficult) problem -- which we call Time-To-Live (TTL) -- accepts only $\mathcal{G}$ and $Q$ and outputs a positive integer $k$ as a prediction for the lowest such value where $Q(G_n) \neq Q(G_{n+k})$.
-
-\ah{How do we define quality under bag semantics, without \texttt{DISTINCT}? Do we consider equality of unique results, or multiplicity of results?}
+The second (more difficult) problem -- which we call Time-To-Live (TTL) -- accepts only $\mathcal{G}$ and $Q$ and outputs a positive integer $k$ as a prediction for the lowest such value where $Q(G_n) \neq Q(G_{n+k})$; in the case of bag semantics, we define ``$\neq$'' in terms of bag inequality, meaning that two bags of query results are different if the multiplicity of any result differs.
+Our problem is thus concerned with predicting if/when the results of a query will change, rather than predicting how  -- or to what extent -- they will change.
+
+\ah{Is this how we define quality under bag semantics, without \texttt{DISTINCT}? Do we consider equality of unique results, or multiplicity of results?}
 
@@
 
-\paragraph{Architecture:} We propose a method to predict answers for the OSC and TTL problems.
-Our method primarily considers static features of the query: these include statistics such as the number of triple patterns, as well as the presence of query operators, such as \texttt{UNION}, \texttt{FILTER}, specific path expressions, etc.; our model can also consider statistics about predicates in the dynamic RDF graph, as well as in which past versions of the dynamic graph the results for the query changed.
-Figure~\ref{fig:schema} shows the components of our approach, which receives as input a SPARQL query (the architecture itself already encodes information about the dynamic graph and the value of $k$ for OSC, which we set in experiments to 1, predicting changes in the next version).
-The Feature Extractor decomposes the query into a set of numeric features, described presently.
-The Dynamics Model accepts some features of the query, as well as information from the dynamic graph, in order to compose further features.
+\paragraph{Architecture:} In Figure~\ref{fig:schema}, we provide an overview of our proposed architecture for making predictions with respect to the OSC and TTL tasks.
+A SPARQL query $Q$ and a dynamic RDF graph $\mathcal{G}$ are given as input.
+The first component parses the query and generates static features which include statistics about the number of elements in the query (e.g., number of triple patterns, variables, etc.) as well as the presence of particular query operators (e.g., \texttt{UNION}, \texttt{FILTER}, recursive path expressions, etc.).
+The second component may take as input some features of the query, as well as information from the dynamic graph, and produce a set of further features; this component may compute, for example, statistics about the dynamicity of predicates used in the query, or statistics about changes in the query results over past versions.
 The final features are passed to a classifier in the case of the OSC problem, and to a regression model in the case of the TTL problem, to generate the final output prediction.
-We now describe the Feature Extractor and Dynamics Model in more detail.
-
-\begin{figure*}[!t]
+We will now describe in more detail the query and data features considered in this work.
+
+%\begin{figure*}[!t]
+%	\centering
+%	\includegraphics[width=0.9\linewidth]{img/schema.pdf}
+%	\caption{Proposed architecture with Feature Extractor and Dynamics Model \ah{What is $g$ here and why is it output from the Dynamics Model (rather than input)?}}
+%	\label{fig:schema}
+%\end{figure*}
+
+\begin{figure}[tb]
+\newlength{\hgap}
+\setlength{\hgap}{0.7cm}
+
+\newcommand{\tb}[1]{\begin{tabular}{c} #1 \end{tabular}}
+\newcommand{\hsp}{\vphantom{Ag}}
+
+\tikzset{
+    every node/.style={
+    	font=\sffamily,
+    },
+	f/.style={ 
+		draw,
+		text centered,
+		style={inner sep=0.8},
+		font=\footnotesize,
+		minimum width=1.3em,
+		minimum height=1.2em
+	},
+	i/.style={ 
+		draw,
+		text centered,
+		rounded corners
+	},
+	o/.style={ 
+		draw,
+		text centered,
+		rounded corners
+	},	
+	c/.style={ 
+		draw,
+		top color=white, 
+		bottom color=black!5
+	},	
+	arrout/.style={
+		->,
+		-latex,
+		black!25},
+	arrin/.style={
+        <-,
+        latex-,
+		black!25},	
+}
+
 \centering
-	\includegraphics[width=0.9\linewidth]{img/schema.pdf}
-	\caption{Proposed architecture with Feature Extractor and Dynamics Model \ah{What is $g$ here and why is it output from the Dynamics Model (rather than input)?}}
-	\label{fig:schema}
-\end{figure*}
-
-\paragraph{Feature Extractor:} Our initial set of features are based on analysis of the input SPARQL query.
+\resizebox{0.9\textwidth}{!}{
+\begin{tikzpicture}
+\node[i] (q) {$Q$};
+
+\node[c,right=\hgap of q] (qfe) {\tb{Query \\ Feature \\ Extractor}}
+  edge[arrin] (q);
+
+\node[f,right=1.5\hgap of qfe] (f4) {\tiny ...};
+\node[f,anchor=south] (f3) at (f4.north) {\tiny ...};
+\node[f,anchor=south] (f2) at (f3.north) {\tiny ...};
+\node[f,anchor=south] (f1) at (f2.north) {$f_1$};
+\node[f,anchor=north] (f5) at (f4.south) {\tiny ...};
+\node[f,anchor=north] (f6) at (f5.south) {\tiny ...};
+\node[f,anchor=north] (fn) at (f6.south) {$f_n$};
+
+\draw [arrout] (qfe.east) -- + (0.3,0) |- (f1);
+\draw [arrout] (qfe.east) -- + (0.3,0) |- (f2);
+\draw [arrout] (qfe.east) -- + (0.3,0) |- (f3);
+\draw [arrout] (qfe.east) -- + (0.3,0) |- (f4);
+\draw [arrout] (qfe.east) -- + (0.3,0) |- (f5);
+\draw [arrout] (qfe.east) -- + (0.3,0) |- (f6);
+\draw [arrout] (qfe.east) -- + (0.3,0) |- (fn);
+
+\node[c,right=\hgap of fn] (dfe) {\tb{Data\\Feature\\Extractor}}
+  edge[arrin] (fn);
+
+\draw[arrout] (f6.east) -- (dfe.west|-f6.east);
+
+\node[i,anchor=south] (g) at (q.mid|-dfe.south) {$\mathcal{G}$}
+  edge[arrout] (dfe.west|-g.east);
+
+\node[f,right=\hgap of dfe] (dfk) {$f'_k$}
+  edge[arrin] (dfe);    
+\node[f,anchor=south] (df6) at (dfk.north) {\tiny ...}
+  edge[arrin] (dfe.east|-df6);  
+\node[f,anchor=south] (df5) at (df6.north) {$f_j$}
+  edge[arrin] (f5);    
+\node[f,anchor=south] (df4) at (df5.north) {\tiny ...}
+  edge[arrin] (f4);  
+\node[f,anchor=south] (df3) at (df4.north) {\tiny ...}
+  edge[arrin] (f3);      
+\node[f,anchor=south] (df2) at (df3.north) {\tiny ...}
+  edge[arrin] (f2);   
+\node[f,anchor=south] (df1) at (df2.north) {$f_1$}
+  edge[arrin] (f1);
+
+\node[c,right=1.5\hgap of df2,minimum width=6.5em] (c) {Classification\hsp};
+
+\draw [arrout] (df1.east) -- + (0.3,0) |- (c);
+\draw [arrout] (df2.east) -- + (0.3,0) |- (c);
+\draw [arrout] (df3.east) -- + (0.3,0) |- (c);
+\draw [arrout] (df4.east) -- + (0.3,0) |- (c);
+\draw [arrout] (df5.east) -- + (0.3,0) |- (c);
+\draw [arrout] (df6.east) -- + (0.3,0) |- (c);
+\draw [arrout] (dfk.east) -- + (0.3,0) |- (c);
+
+\node[c,right=1.5\hgap of df6,minimum width=6.5em] (r) {Regression\hsp};
+
+\draw [arrout] (df1.east) -- + (0.3,0) |- (r);
+\draw [arrout] (df2.east) -- + (0.3,0) |- (r);
+\draw [arrout] (df3.east) -- + (0.3,0) |- (r);
+\draw [arrout] (df4.east) -- + (0.3,0) |- (r);
+\draw [arrout] (df5.east) -- + (0.3,0) |- (r);
+\draw [arrout] (df6.east) -- + (0.3,0) |- (r);
+\draw [arrout] (dfk.east) -- + (0.3,0) |- (r);
+
+\node[o,right=\hgap of c] (osc) {\tb{OSC \\ Prediction}}
+  edge[arrin] (c);
+
+\node[o,right=\hgap of r] (ttl) {\tb{TTL \\ Prediction}}
+  edge[arrin] (r);
+
+\end{tikzpicture}
+}
+
+\caption{Proposed architecture for predicting change in query results\label{fig:schema}}
+\end{figure}
+
+\paragraph{Query Features:} Our initial set of features are based on analysis of the input SPARQL query.
 Figure~\ref{fig:query} shows the list of evaluated features, as well as -- for the purposes of illustration -- their values for an accompanying example query.
@@
 One may consider that the higher these four values are, in general, the more dynamic we can expect the query to be since there are more ``opportunities'' for the query results to be affected by change.
-In some cases, however, the hypothesized correlation is not direct since for example adding more triple patterns may serve to narrow the query down and focus it on a static part of the graph (e.g., when looking for the movies of directors, adding a triple pattern to restrict the results to directors who have died).
+In some cases, however, the hypothesized correlation is not direct since, for example, adding more triple patterns may serve to narrow the query down and focus it on a static part of the graph (e.g., when looking for the movies of directors, adding a triple pattern to restrict the results to directors who have died).
 Hence it will be of interest to see how these features affect the predictions made experimentally.
@@
 On the other hand, in \textit{Negation}, we group non-monotonic features that allow for modeling difference (\texttt{MINUS}, \texttt{NOT EXISTS}, \texttt{OPTIONAL}, \texttt{!BOUND}).
-These features -- though course-grained -- are straightforward to extract from a query, and offer valuable insights into how the query may behave in a dynamic query; for example, the \textit{Negation} feature captures information about the (non-)monotonicity of the query, while we suppose that \textit{Recursive paths}, which may traverse an arbitrary number of triples in the graph, might be sensitive to change.
-Again, such relations are not direct where experiments are required for further insights.
+These features -- though course-grained -- are straightforward to extract from a query, and offer valuable insights into how the query may behave in a dynamic query; for example, the \textit{Negation} feature captures information about the (non-)monotonicity of the query, while we suppose that \textit{Recursive paths}, which may traverse an arbitrary number of triples in the graph, might be more sensitive to change.
+Again, such correlations are not without exception and will require experimental scrutiny.
 
@@
 
-\paragraph{Dynamics Model:} The next component in our architecture is the Dynamics Model which (optionally) incorporates information about the query \textit{and} the dynamic graph under consideration.
-
-The first feature of the Dynamics Model aims to capture how many triples change for a predicate in a time interval, with the idea that -- as in previous works~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12,DehghanzadehPKUHD14} -- predicates capture rich information about the dynamics of a dataset, where the results of queries with dynamic predicates will be more sensitive to change; furthermore, given that in most datasets the number of predicates is relatively low (in the thousands or tens of thousand), such statistics are quite lightweight to maintain and compute.
-
+\paragraph{Data Features (predicate dynamics):} The next component in our architecture can extract features that capture information about the query \textit{and} the dynamic graph under consideration.
+The first such feature we consider captures how many triples change for a predicate in a time interval, with the idea that -- as in previous works~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12,DehghanzadehPKUHD14} -- predicates capture rich information about the dynamics of a dataset, where the results of queries with dynamic predicates will be more sensitive to change; furthermore, given that in most datasets the number of predicates is relatively low (in the thousands or tens of thousand), such statistics are quite lightweight to maintain and compute.
 Formally, given two RDF graphs $G_i$ and $G_j$, we denote by $G_i \oplus G_j$ the set of triples $(G_i \cup G_j) - (G_i \cap G_j)$ where ``$-$'' denotes set difference; in other words, $G_i \oplus G_j$ denotes the triples in one graph or the other but not both (noting that $G_i \oplus G_j = G_j \oplus G_i$).
 Next, given an RDF graph $G$ and an IRI $p$, let $\#(G,p) \da |\{(x,y,z) \in G : y = p \}|$ denote the number of triples in $G$ with predicate $p$.
-Next, given a dynamic RDF graph $\mathcal{G} \da ( G_1, \ldots, G_n )$ and a predicate $p$, we denote by $\Delta(\mathcal{G},p) \da \Sigma_{i=1}^{n-1} \#(G_i \oplus G_{i+1},p)$ the sum of the number of triples with the predicate $p$ that changed between pairs of consecutive versions.
+Finally, given a dynamic RDF graph $\mathcal{G} \da ( G_1, \ldots, G_n )$ and a predicate $p$, we denote by $\Delta(\mathcal{G},p) \da \Sigma_{i=1}^{n-1} \#(G_i \oplus G_{i+1},p)$ the sum of the number of triples with the predicate $p$ that changed between pairs of consecutive versions.
 
@@
 Given a query $Q$ with a set of predicates $\{ p_i, ..., p_n \}$, we may then consider a variety of aggregate functions over $\Delta(\mathcal{G},p_1), ..., \Delta(\mathcal{G},p_n)$, such as $\mathrm{max}$, $\mathrm{mean}$, etc., to compute a final numeric feature for the query, representing a summary of the level of dynamicity of the predicates it contains.
+
+\paragraph{Data Features (result dynamics):} The second data feature we capture indicates the past versions of the dynamic graph in which the query results $Q$ changed (what Knuth et al.
+call the ``clairvoyant model''~\cite{KnuthHS16}).
+\ah{How precisely is this represented?} While this feature offers rich information for prediction, given a query that has not previously been seen, it may be costly to compute, potentially requiring the execution of the query over each version, which in turn requires maintaining indexes over a variety of past versions.
 
@@
 With 23 versions, we must hold out at least one version for testing, leaving a maximum interval of 22 past versions to use for predictions.
-However, the more versions we add to interval of past versions considered, the fewer examples we can generate; for example, if we consider 22 versions for training, we will only have one ground truth label for each query.
+However, the more past versions we consider for a prediction, the fewer examples we can generate; for example, if we consider 22 versions for training, we will only have one ground truth label for each query.
 In the end, we opted to consider the 3 previous versions ($G_{i-3}, G_{i-2}, G_{i-1}$) to predict changes for the subsequent version ($G_i$); this allows us to predict from $G_4$ to $G_{23}$ inclusive, providing 20 examples per query and 4420 examples in total.
@@
 \paragraph{Results:} Table~\ref{tab:classifier} shows the $F_1$-measure for the predictions made; for reference, we include a baseline that randomly guesses yes/no.
-The results show that all configurations clearly outperform the random baseline; however, with the highest result being an $F_1$-measure of 0.706, we can also see that predicting changes in query results is indeed a challenging problem.
+With the highest result being an $F_1$-measure of 0.706, the results show that predicting changes in query results is indeed a challenging problem; on the other hand, all configurations clearly outperform the random baseline.
 Looking in further detail, as can be expected, prediction performs best when the \textsc{r} feature -- with knowledge of changes in the historical value of results of the input query -- is available.
@@
 In this paper we propose methods to predict whether or not (OSC) and/or when (TTL) the results of an input query will change in the future.
-Such methods have applications for synchronization, caching, and more besides.
-More specifically, we propose a framework based on classifiers and regression that accept features extracted from the queries, from past versions of the data, or from the combination of both.
+More specifically, we propose a framework based on classifiers and regression that accept features extracted from the query, from past versions of the data, and/or from the combination of both.
 Considering this framework, we propose that there is a conceptual trade-off between the cost of computing features and their value for prediction: features extracted from queries alone are the most efficient to extract, not requiring historical data, but are quite coarse-grained for prediction; on the other hand, features based on changes in results in previous versions for an (unseen) input query are often the most costly to acquire, but offer fine-grained information for prediction; finally, features based on the dynamics of predicates used in the data offer a balance between the two, allowing to summarize historical data into succinct statistics, while offering more fine-grained information than static query features and less fine-grained information than historical results.
 
-Our goal then is to explore this trade-off experimentally.
+Our goal then was to explore this trade-off experimentally.
 For this, we define some desirable characteristics for the queries and data used in the experimental setting, such as availability, size, dynamics, diversity, etc.
@@
 Analysing these data and queries, we conclude that they meet the desired characteristics, offering a large-scale, diverse, real-world setting; more importantly, we find that our queries offer an ideal distribution of queries whose results always change, queries whose results never change, and (more importantly) queries whose results change intermittently and for which prediction is thus a non-trivial task.
-We make this gold standard available at \ah{...} -- including queries, results and features -- so that it can be reused in further research.
+We make this gold standard available at \ah{...} -- including queries, results and features -- so that the results presented here can be reproduced and so that it can be reused in further research.
 

### 2019-04-10 kgswc19.tex
revision 09/04

@@
 %
-\author{Alberto Moya \and
+\author{Alberto Moya\orcidID{0000-0002-7003-5087} \and 
 Aidan Hogan\orcidID{0000-0001-9482-1982}}
@@
 
-Some works \ah{Which?
-Cite at the end of the sentence} have proposed to predict dynamics based on the historical evolution of entities, but following such an approach for SPARQL queries is expensive because (1) a SPARQL query may involve potentially many entities; and (2) it is necessary to have the previous complete versions of data to analyse the entities relevant to a query.
-Recent works have explored the dynamics of Linked Data with the intention of finding patterns that allow for characterizing, recognizing, and predicting changes based on analysis of the domains, predicates, and schema~\cite{UmbrichHHPD10,KaferAUOH13,NishiokaS16,NishiokaS17,GonzalezH18}.
+Some works have studied the data dynamics based on the historical evolution of entities~\cite{KaferAUOH13,DividinoGSG14,NishiokaS15}, but following such an approach for SPARQL queries is expensive because (1) a SPARQL query may involve potentially many entities; and (2) it is necessary to have the previous complete versions of data to analyse the entities relevant to a query.
+Many works have explored the dynamics of Linked Data with the intention of finding patterns that allow for characterizing, recognizing, and predicting changes based on analysis of the domains, predicates, and schema~\cite{UmbrichHHPD10,KaferAUOH13,NishiokaS16,NishiokaS17,GonzalezH18}.
 Among these works, we can find hybrid approaches that have been developed to return fresher query results with faster response times by decomposing a query into dynamic sub-queries executed remotely, and static sub-queries executed over local caches~\cite{UmbrichKHP12}, but such an approach only focuses on estimating the dynamics of individual triple patterns, and does not consider, for example, query operators.
@@
 (2) We propose a number of features to instantiate this architecture based on analysis of the query, analysis of the dynamics of predicates in the data, and analysis of historical changes in the query results.
-(3) We create a gold standard for these tasks based on 23 weekly versions of Wikidata and a set of 221 real-world SPARQL queries.
+(3) We create a gold standard for these tasks based on 23 consecutive versions of Wikidata and a set of 221 real-world SPARQL queries.
 (4) We use this gold standard to compare the predictions obtained using different types of features and classifiers.
@@
 Our work thus complements existing works aiming for weak consistency, but (i) generalizes the problem, proposing a framework that can incorporate statistics on predicate dynamics~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12,DehghanzadehPKUHD14} and historical changes in query results~\cite{KnuthHS16}, as proposed in prior works, as well as novel types of features, (ii) introduces new features based on query operators and statistics; (iii) creates a novel gold standard based on Wikidata and presents comparative results that indicate the relative predictive power inherent in different types of features.
-Furthermore, to the best of our knowledge, no prior work uses Machine Learning models to predict changes in SPARQL query results.
-\ah{Is this last claim true?
-We should try to be as sure as possible.}
 
@@
 Our problem is thus concerned with predicting if/when the results of a query will change, rather than predicting how  -- or to what extent -- they will change.
-
-\ah{Is this how we define quality under bag semantics, without \texttt{DISTINCT}? Do we consider equality of unique results, or multiplicity of results?}
 
@@
 \paragraph{Query Features:} Our initial set of features are based on analysis of the input SPARQL query.
-Figure~\ref{fig:query} shows the list of evaluated features, as well as -- for the purposes of illustration -- their values for an accompanying example query.
-
-The first four features capture statistics about the query, indicating loosely its size and complexity.
-One may consider that the higher these four values are, in general, the more dynamic we can expect the query to be since there are more ``opportunities'' for the query results to be affected by change.
+Figure~\ref{fig:query} shows only a sample of the evaluated features, as well as -- for the purposes of illustration -- their values for an accompanying example query.
+
+The first block of features capture statistics about the query, indicating loosely its size and complexity.
+One may consider that the higher these values are, in general, the more dynamic we can expect the query to be since there are more ``opportunities'' for the query results to be affected by change.
 In some cases, however, the hypothesized correlation is not direct since, for example, adding more triple patterns may serve to narrow the query down and focus it on a static part of the graph (e.g., when looking for the movies of directors, adding a triple pattern to restrict the results to directors who have died).
@@
 
-The next six features indicate the query operators used; to reduce the dimensionality of the problem, we capture only a subset of popularly-used features and group related features into one dimension: in the case of \textit{Recursive path}, we group queries that contain as expression of the form $e$\texttt{*} or $e$\texttt{+}.
-On the other hand, in \textit{Negation}, we group non-monotonic features that allow for modeling difference (\texttt{MINUS}, \texttt{NOT EXISTS}, \texttt{OPTIONAL}, \texttt{!BOUND}).
+The second block indicate the query operators used; we capture the presence or absence of query operators and solution modifiers from SPARQL 1.1.
+In the third block, we group related features into one dimension: in the case of \textit{Recursive path}, we group queries that contain as expression of the form $e$\texttt{*} or $e$\texttt{+}.
+On the other hand, in \textit{Negation}, we group non-monotonic features that allow for modeling difference (\texttt{MINUS}, \texttt{NOT EXISTS}, \texttt{!BOUND}).
 These features -- though course-grained -- are straightforward to extract from a query, and offer valuable insights into how the query may behave in a dynamic query; for example, the \textit{Negation} feature captures information about the (non-)monotonicity of the query, while we suppose that \textit{Recursive paths}, which may traverse an arbitrary number of triples in the graph, might be more sensitive to change.
 Again, such correlations are not without exception and will require experimental scrutiny.
-
-\ah{Is OPTIONAL not captured? Why were these features captured? Why not BIND or OFFSET or GROUP BY?}
 
@@
 	\hfill
-	\resizebox{0.48\textwidth}{!}{
-	\begin{tabular}{rlr}    \toprule
-		1  & \ntextnumero of triple patterns        & 7			\\
-		2  & \ntextnumero of variables      & 3			\\
-		3  & \ntextnumero of projected variables & 1			\\
-		4  & \ntextnumero of predicates     & 6			\\ \midrule
-		5  & \texttt{FILTER}            	& \tickYes	\\
-		6  & \texttt{LIMIT}             	& \tickYes	\\
-		7  & \texttt{UNION}             	& \tickYes	\\
-		8  & \textit{Sub-query}         	& \tickNo	\\
-		9  & \textit{Recursive path}		& \tickYes	\\
-		10 & \textit{Negation}         		& \tickYes	\\ \midrule
-		11 & \textit{Predicates}			& $\{$ \texttt{:instance\_of}, ... $\}$     \\    \bottomrule
+	\resizebox{0.45\textwidth}{!}{
+	\begin{tabular}{lr}    \toprule
+		\ntextnumero of triple patterns        & 7			\\
+		\ntextnumero of variables      & 3			\\
+		\ntextnumero of projected variables & 1			\\
+		\ntextnumero of predicates     & 6			\\ \midrule
+		\texttt{FILTER}            	& \tickYes	\\
+		\texttt{LIMIT}             	& \tickYes	\\
+		\texttt{UNION}             	& \tickYes	\\
+		\texttt{GROUPBY}     & \tickNo			\\ \midrule 
+		\textit{Sub-query}         	& \tickNo	\\
+		\textit{Recursive path}		& \tickYes	\\
+		\textit{Negation}         		& \tickYes	\\ \midrule
+		\textit{Predicates}			& $\{$ \texttt{:instance\_of}, ... $\}$     \\    \bottomrule
 	\end{tabular}	
 	}
-\caption{Example query and the features extracted from it \label{fig:query}}	
+\caption{Example query and a sample of the features extracted from it \label{fig:query}}	
 \end{figure}
@@
 Next, given an RDF graph $G$ and an IRI $p$, let $\#(G,p) \da |\{(x,y,z) \in G : y = p \}|$ denote the number of triples in $G$ with predicate $p$.
-Finally, given a dynamic RDF graph $\mathcal{G} \da ( G_1, \ldots, G_n )$ and a predicate $p$, we denote by $\Delta(\mathcal{G},p) \da \Sigma_{i=1}^{n-1} \#(G_i \oplus G_{i+1},p)$ the sum of the number of triples with the predicate $p$ that changed between pairs of consecutive versions.
-
-\ah{We don't need to define the interval in the latter function since we could project an interval from $\mathcal{G}$ when needed (more general and readable I think; e.g., $\Delta(\mathcal{G}[i,j],p)$).
-Also I don't get why you divide by Aggregated Predicate Multiplicity; doesn't this just scale all predicates by the same constant for an interval?
-I think it would perhaps make more sense to normalize by $\#(G_i \cup G_{i+1},p)$, or even just go down the route of Jaccard Similarity: $\Delta(\mathcal{G},p) \da \Sigma_{i=1}^{n-1}\frac{\#(G_i \cap G_{i+1},p)}{\#(G_i \cup G_{i+1},p)}$}
+Finally, given a dynamic RDF graph $\mathcal{G} \da ( G_1, \ldots, G_n )$ and a predicate $p$, we denote by $\Delta(\mathcal{G},p) \da \Sigma_{i=1}^{n-1}\frac{\#(G_i \oplus G_{i+1},p)}{\#(G_i \cup G_{i+1},p)}$ the normalized sum of the number of triples with the predicate $p$ that changed between pairs of consecutive versions.
 
@@
 In Figure~\ref{fig:dgc} we show the pairwise changes between each version: $G_1 \oplus G_2$ and $G_3 \oplus G_4$.
-For a predicate $p$, the value $\Delta(\mathcal{G},p)$ is then the sum of triples with predicate $p$ in the graphs $G_1 \oplus G_2$ and $G_2 \oplus G_3$.
-Looking at Figure~\ref{fig:dgc}, for example, \texttt{:name} does not appear ($\Delta(\mathcal{G},\texttt{:name})=0$), \texttt{:children} appears twice in $G_1 \oplus G_2$ and twice in $G_2 \oplus G_3$ ($\Delta(\mathcal{G},\texttt{:children})=4$), and so forth.
+For a predicate $p$, the value $\Delta(\mathcal{G},p)$ is then the sum of triples with predicate $p$ in the graphs $G_1 \oplus G_2$ and $G_2 \oplus G_3$ divided by the number of triples with predicate $p$ in the graphs $G_1 \cup G_2$ and $G_2 \cup G_3$.
+Looking at Figure~\ref{fig:dgc}, for example, \texttt{:name} does not appear ($\Delta(\mathcal{G},\texttt{:name})=0$), \texttt{:children} appears twice in $G_1 \oplus G_2$ and $G_1 \cup G_2$, as well as, twice in $G_2 \oplus G_3$ and $G_2 \cup G_3$ ($\Delta(\mathcal{G},\texttt{:children})=4/4=1$), and so forth.
 \end{example}
@@
 
-\paragraph{Data Features (result dynamics):} The second data feature we capture indicates the past versions of the dynamic graph in which the query results $Q$ changed (what Knuth et al.
-call the ``clairvoyant model''~\cite{KnuthHS16}).
-\ah{How precisely is this represented?} While this feature offers rich information for prediction, given a query that has not previously been seen, it may be costly to compute, potentially requiring the execution of the query over each version, which in turn requires maintaining indexes over a variety of past versions.
+\paragraph{Data Features (result dynamics):} The second data feature we capture indicates how many times the query results $Q$ have changed over the past versions of the dynamic graph.
+While this feature offers rich information for prediction, given a query that has not previously been seen, it is costly to compute, since it requires the execution of the query over each past version within an interval, which in turn requires maintaining indexes over a variety of past versions.
 
@@
 
-\paragraph{RDF Data:} We use 23 Wikidata snapshots from 18/04/2017 to 27/09/2017, which are captured weekly in the \emph{truthy version} that contains triples (without qualifiers) that have the best non-deprecated rank for a given property.
+\paragraph{RDF Data:} We use 23 Wikidata snapshots from 18/04/2017 to 27/09/2017, which are captured almost weekly($\pm$ 1 day) in the \emph{truthy version} that contains triples (without qualifiers) that have the best non-deprecated rank for a given property.
 The first version has 1,102,242,331 triples and 3,276 unique predicates, while the final version has 1,924,967,162 triples (+74\%) and 3,660 unique predicates (+11\%).
-Figure~\ref{fig:wikidata}, on the left, shows the growth in triples as the weeks progress, while on the right, shows the numbers of triples added and removed week-to-week.
-We see that although many more triples are added than removed, there are some triples removed each week.
-We further note some peaks in triples added in some weeks, which may be due to bulk imports of data.
-The gap in week 12 is due to the fact that we were not able to obtain the data for that week (we thus exclude the triples added/removed from weeks 11 to 12, and 12 to 13).
+Figure~\ref{fig:wikidata}, on the left, shows the growth in triples as the time progress, while on the right, shows the numbers of triples added and removed version-to-version.
+We see that although many more triples are added than removed, there are some triples removed each version.
+We further note some peaks in triples added in some versions, which may be due to bulk imports of data.
+Between the version 11 and 12 we have almost 2 weeks because we were not able to obtain the data for that week, but this cause the third highest peak.
 The dynamic graph thus considers a total of 32.3 billion triples across 23 versions.
-
-\ah{Regarding: ``The first version has 1,102,242,331 triples, 57,499,191 unique IRIs, 3,276 unique predicates, 933,588,860 literals, and 7,449 blank nodes, while the final version has 1,924,967,162 triples (+74\%), 80,637,164 unique IRIs (+40\%), 3,660 unique predicates (+11\%), 1,676,930,476 literals (+79\%) and 8,909 blank nodes (+19\%).'' I assume that the literals and blank nodes are not unique?
-Also the IRIs are unique per position?
-(In any case, such data is not 100\% needed; what's more important are the unique triples and unique predicates.)}
-
-\ah{Would it be possible to get a table of the top-10 most and least dynamic predicates by number of triples added/removed over the weeks, and also the number of triples with each such predicate in the first and last week?}
 
@@
 		ymin=0,
-		xtick={1,2,...,24},
+		xtick={1,2,...,23},
 		xticklabels={},
-		extra x ticks={1,24},
+		extra x ticks={1,23},
 		legend pos=north west,
@@
 		coordinates {
-			(1,1102242331)(2,1138975810)(3,1141303957)(4,1170283653)(5,1176093821)(6,1184177503)(7,1190456714)(8,1211813217)(9,1239145151)(10,1271353260)(11,1293099057)
-		};
-
-		\addplot[
-		color=blue,
-		mark=diamond,
-		]
-		coordinates {
-			(13,1348622441)(14,1418460320)(15,1432143355)(16,1470298276)(17,1506140622)(18,1542278452)(19,1593334993)(20,1613832004)(21,1684774955)(22,1771601730)(23,1843086909)(24,1924967162)
+			(1,1102242331)(2,1138975810)(3,1141303957)(4,1170283653)(5,1176093821)(6,1184177503)(7,1190456714)(8,1211813217)(9,1239145151)(10,1271353260)(11,1293099057)(12,1348622441)(13,1418460320)(14,1432143355)(15,1470298276)(16,1506140622)(17,1542278452)(18,1593334993)(19,1613832004)(20,1684774955)(21,1771601730)(22,1843086909)(23,1924967162)
 		};
@@
 		\addplot 
-		coordinates {(1,40137188)(2,4563382)(3,29302968)(4,8906402)(5,11455047)(6,11435682)(7,27229516)(8,26968577)(9,38952400)(10,20283316)(11,0)(12,0)(13,70073220)(14,17685830)(15,45405782)(16,41199256)(17,42633484)(18,56683899)(19,25309217)(20,25030143)(21,22548371)(22,23330356)(23,37009794)};
+		coordinates {(1,40137188)(2,4563382)(3,29302968)(4,8906402)(5,11455047)(6,11435682)(7,27229516)(8,26968577)(9,38952400)(10,20283316)(11,49001368)(12,70073220)(13,17685830)(14,45405782)(15,41199256)(16,42633484)(17,56683899)(18,25309217)(19,25030143)(20,22548371)(21,23330356)(22,37009794)};
 		\addplot 
-		coordinates {(1,3403743)(2,2235268)(3,1116111)(4,3096270)(5,3371448)(6,5156519)(7,5873153)(8,1336452)(9,6744349)(10,999601)(11,0)(12,0)(13,1123311)(14,4002828)(15,7250907)(16,5356971)(17,6495717)(18,5627376)(19,4812243)(20,790590)(21,681118)(22,1194899)(23,998105)};
+		coordinates {(1,3403743)(2,2235268)(3,1116111)(4,3096270)(5,3371448)(6,5156519)(7,5873153)(8,1336452)(9,6744349)(10,999601)(11,786939)(12,1123311)(13,4002828)(14,7250907)(15,5356971)(16,6495717)(17,5627376)(18,4812243)(19,790590)(20,681118)(21,1194899)(22,998105)};
 		\legend{Added, Removed}
@@
 \caption{Evolution of total triples in Wikidata (left) and number of triples added to and removed from the $n$\textsuperscript{th} version for the $n+1$\textsuperscript{th} version (right) \label{fig:wikidata}}
+\end{figure}
+
+Although we have seen that the number of unique predicates did not change much over all versions, we are interested to see which predicates are indicative of dynamic statements.
+Table~\ref{tab:predicates} shows the ten most dynamic predicates according to the number statements involving that predicate (left) and the ratio of added (+) and deleted (−) statements divided by the total number of statements for that predicate across all snapshots; for the second, we only include predicates that appear in all snapshots and appear in ≥ 1, 000 statements overall.
+Note that the largest number of changes involves the predicate scheme:description, however, its dynamic value is low due to the number of statements with it in the versions.
+
+\begin{figure}[!t]
+	\caption{Top-10 dynamic predicates according with total (left) and proportional (right) changes}
+	\label{tab:predicates}
+	\resizebox{0.45\textwidth}{!}{
+		\begin{tabular}{rlrr}    \toprule
+			\ntextnumero        & Predicate & Total	& Dyn 	\\  \midrule
+			1 & schema:description  & 464567354  	& 0.04   \\
+			2 & schema:dateModified  & 99330382  	& 0.14   \\
+			3 & schema:version  & 99330164  	& 0.14  \\
+			4 & schema:name  & 44487864  	& 0.01   \\
+			5 & rdfs:label  & 44487864  	& 0.01   \\
+			6 & skos:prefLabel  & 44487864  	& 0.01   \\
+			7 & wdt:P2093(author name)  & 30671685  	& 0.09   \\
+			8 & rdf:type  & 21808957  	& 0.02   \\
+			9 & wdt:P31(instance of)  & 12483471  	& 0.02   \\
+			10 & schema:about  & 10773295  	& 0.02  \\    \bottomrule
+		\end{tabular}	
+	}
+	\hfill
+	\resizebox{0.5\textwidth}{!}{
+		\begin{tabular}{rlrr}    \toprule
+			\ntextnumero        & Predicate & Total	& Dyn 	\\  \midrule
+			1 & owl:complementOf & 156862 & 0.99	\\
+			2 & owl:onProperty & 156862 & 0.99	\\
+			3 & owl:someValuesFrom & 86674 & 0.71	\\
+			4 & wdt:P2462(member of the deme) & 3563 & 0.62	\\
+			5 & wdt:P3383(film poster) & 4388 & 0.58	\\
+			6 & wdt:P2331(Cycling Archives ID) & 4229 & 0.33	\\
+			7 & wdt:P1112() & 4366 & 0.20	\\
+			8 & wdt:P505(general manager) & 4959 & 0.18	\\
+			9 & schema:dateModified & 99330382 & 0.14	\\
+			10 & schema:version & 99330164 & 0.14	\\    \bottomrule
+		\end{tabular}	
+	}	
 \end{figure}
@@
 
-Finally, we look at how the results for these queries change over the 23 weeks.
-Figure~\ref{graph:changeV} shows how many queries had some results change between two consecutive versions, where we see that approximately half of the queries see changes in their results each week.
-But are these always the same queries that change every week?
-Figure~\ref{graph:changeV} shows the number of weeks in which some result changed for each query; the queries are ordered by the number of changes in their results, where we can see that the results of 44 queries (19.90\%) change each time, 18 queries (8\%) never change, and more generally, we note a quite uniform distribution of queries in between.
+Finally, we look at how the results for these queries change over the 23 versions.
+Figure~\ref{graph:changeV} shows how many queries had some results change between two consecutive versions, where we see that approximately half of the queries see changes in their results each version.
+But are these always the same queries that change every time?
+Figure~\ref{graph:changeV} shows the number of versions in which some result changed for each query; the queries are ordered by the number of changes in their results, where we can see that the results of 44 queries (19.90\%) change each time, 18 queries (8\%) never change, and more generally, we note a quite uniform distribution of queries in between.
 More generally speaking, we conclude that our query set has a good balance of queries whose results never change, queries whose results always change, and queries whose results sometimes change.
@@
 		coordinates {
-			(1,115)(2,112)(3,118)(4,116)(5,121)(6,111)(7,127)(8,109)(9,118)(10,115)
-		};
-
-		\addplot[
-		color=blue,
-		mark=diamond,
-		]
-		coordinates {
-			(12,105)(13,113)(14,118)(15,117)(16,111)(17,118)(18,112)(19,116)(20,106)(21,106)(22,129)
+			(1,115)(2,112)(3,118)(4,116)(5,121)(6,111)(7,127)(8,109)(9,118)(10,115)(11,129)(12,105)(13,113)(14,118)(15,117)(16,111)(17,118)(18,112)(19,116)(20,106)(21,106)(22,129)
 		};
@@
 We remark that no feature explicitly indicates which specific version numbers are given and which are to be predicted.
-(\ah{Discuss the gap: \textit{We evaluated the quality of our proposal by forming a set of classifiers to predict the change of 221 queries in 15 weeks, which is equivalent to 3315 objects to classify, described in terms of our proposed features.}}) \ah{Could we do experiments with different intervals?}
+(\ah{Discuss the gap: \textit{We evaluated the quality of our proposal by forming a set of classifiers to predict the change of 221 queries in 15 versions, which is equivalent to 3315 objects to classify, described in terms of our proposed features.}}) \ah{Could we do experiments with different intervals?}
 

### 2019-04-10 kgswc19.tex
first pass 10/04

@@
 
-\paragraph{Queries:} In order to achieve a set of SPARQL queries that are answerable over Wikidata and with which we could run experiments, we took the user-contributed example queries from the Wikidata Query Service\footnote{https://www.wikidata.org/wiki/Wikidata:SPARQL\_query\_service/queries/examples}, consisting of 388 SPARQL 1.1 \texttt{SELECT} queries of varying degrees of complexity, touching upon various domains of data, and with a mix of varying query operators.\footnote{Very recently, a large log of millions of SPARQL queries have been released for Wikidata~\cite{MalyshevKGGB18}; these queries were not available when conducting our experiments, but would be interesting to analyze as part of future work.}
-
-Certain queries had to be removed from the original set of 388: some had syntax errors; others asked for information external to Wikidata (using \texttt{SERVICE}) or asked for qualifiers that are not present in the truthy version.
-We also eliminated a handful of queries that returned bindings to blank nodes to facilitate comparison between the results.
+\paragraph{Queries:} In order to achieve a set of SPARQL queries that are answerable over Wikidata and with which we could run experiments, we took the user-contributed example queries from the Wikidata Query Service\footnote{https://www.wikidata.org/wiki/Wikidata:SPARQL\_query\_service/queries/examples}, consisting of 389 SPARQL 1.1 \texttt{SELECT} queries of varying degrees of complexity, touching upon various domains of data, and with a mix of varying query operators.\footnote{Very recently, a large log of millions of SPARQL queries have been released for Wikidata~\cite{MalyshevKGGB18}; these queries were not available when conducting our experiments, but would be interesting to analyze as part of future work.}
+
+96 queries had to be removed from the original set of 389 because they asked for information external to Wikidata (using \texttt{SERVICE}) or asked for qualifiers that are not present in the truthy version.
+We also eliminated 12 queries that returned bindings to blank nodes to facilitate comparison between the results.
+Furthermore, some queries featured non-deterministic elements -- such as \texttt{LIMIT}/\texttt{OFFSET} without ordering, \texttt{SAMPLE} or temporal functions -- such as \texttt{TODAY()} -- that may lead to changes in results not related to changes in the data; we remove 23 queries with \texttt{SAMPLE} and temporal functions, and add an \texttt{ORDER BY} clause to all queries to ensure determinism with \texttt{LIMIT}/\texttt{OFFSET} and more generally to facilitate quick comparison of results.
+We also eliminated 37 queries with 0 results in the executions over all the snapshots.
+As a result of this process of filtering, we end up with a total of 221 deterministic queries.
+
 Furthermore, we needed to transform certain queries for the purposes of our experiments.
-Some queries used features particular to the Wikidata Query Service, such as using a custom label \texttt{SERVICE} call to specify language preferences; we converted these queries to standard SPARQL using \texttt{COALESCE} to capture preferences.
-Furthermore, some queries featured non-deterministic elements -- such as \texttt{LIMIT}/\texttt{OFFSET} without ordering, \texttt{SAMPLE} or temporal functions -- such as \texttt{TODAY()} -- that may lead to changes in results not related to changes in the data; we remove queries with \texttt{SAMPLE} and temporal functions, and add an \texttt{ORDER BY} clause to all queries to ensure determinism with \texttt{LIMIT}/\texttt{OFFSET} and more generally to facilitate quick comparison of results.
+163 queries used features particular to the Wikidata Query Service, such as using a custom label \texttt{SERVICE} call to specify language preferences; we converted these queries to standard SPARQL using \texttt{COALESCE} to capture preferences.
 Figure~\ref{fig:transform} shows an example of a Wikidata query with the \texttt{LABEL} service rewritten and a deterministic ordering added.
-As a result of this process of filtering and transformation, we end up with a total of 221 deterministic queries.
 
@@
 
-In Table~\ref{tab:Pattern}, we provide statistics on the distribution of different types of triple patterns in the 221 queries; of key importance is that 90.35\% of the triple patterns have a constant in the predicate position, meaning that they are compatible with a model based on the dynamics of predicates.
-We also see that 83 triple patterns (7.93\%) feature a path expression, where Table~\ref{tab:Path} provides the distribution of these expressions (remarking that multiple patterns can be used in one expression); of note here is that recursion is commonly found (74.70\% use $e\texttt{*}$ while 9.64\% use $e\texttt{+}$), and that we find no negated property paths.
+In Table~\ref{tab:Pattern}, we provide statistics on the distribution of different types of triple patterns in the 221 queries before the transformation; of key importance is that 86.20\% of the triple patterns have a constant in the predicate position, meaning that they are compatible with a model based on the dynamics of predicates.
+We also see that 83 triple patterns (11.34\%) feature a path expression, where Table~\ref{tab:Path} provides the distribution of these expressions (remarking that multiple patterns can be used in one expression); of note here is that recursion is commonly found (74.70\% use $e\texttt{*}$ while 9.64\% use $e\texttt{+}$), and that we find no negated property paths.
 
@@
 			\textbf{Pattern} & \textbf{\ntextnumero} & \textbf{\%} \\    \midrule
-			V C V   & 751      & 71.73\%    \\
-			V C C   & 189      & 18.05\%    \\
-			V V V   & 9        & 0.86\%     \\
-			C C V   & 6        & 0.57\%     \\
-			V V C   & 5        & 0.48\%     \\
-			C V V   & 3        & 0.29\%     \\
-			C V C   & 1        & 0.10\%      \\
+			V C V   & 435      & 59.43\%    \\
+			V C C   & 190      & 25.96\%    \\
+			V V V   & 9        & 1.23\%     \\
+			C C V   & 6        & 0.82\%     \\
+			V V C   & 5        & 0.68\%     \\
+			C V V   & 3        & 0.41\%     \\
+			C V C   & 1        & 0.14\%     \\
 			C C C   & 0        & 0.00\%     \\    \midrule
-			V P C   & 69       & 6.59\%     \\
-			V P V   & 13       & 1.24\%     \\
-			C P V   & 1        & 0.10\%      \\
+			V P C   & 69       & 9.43\%     \\
+			V P V   & 13       & 1.78\%     \\
+			C P V   & 1        & 0.14\%     \\
 			C P C   & 0        & 0.00\%     \\    \bottomrule
@@
 But are these always the same queries that change every time?
-Figure~\ref{graph:changeV} shows the number of versions in which some result changed for each query; the queries are ordered by the number of changes in their results, where we can see that the results of 44 queries (19.90\%) change each time, 18 queries (8\%) never change, and more generally, we note a quite uniform distribution of queries in between.
+Figure~\ref{graph:changeV} shows the number of versions in which some result changed for each query; the queries are ordered by the number of changes in their results, where we can see that the results of 44 queries (19.90\%) change each time, 17 queries (7.69\%) never change, and more generally, we note a quite uniform distribution of queries in between.
 More generally speaking, we conclude that our query set has a good balance of queries whose results never change, queries whose results always change, and queries whose results sometimes change.
@@
 		coordinates {
-			(1,21)(2,21)(3,21)(4,21)(5,21)(6,21)(7,21)(8,21)(9,21)(10,21)(11,21)(12,21)(13,21)(14,21)(15,21)(16,21)(17,21)(18,21)(19,21)(20,21)(21,21)(22,21)(23,21)(24,21)(25,21)(26,21)(27,21)(28,21)(29,21)(30,21)(31,21)(32,21)(33,21)(34,21)(35,21)(36,21)(37,21)(38,21)(39,21)(40,21)(41,21)(42,21)(43,21)(44,21)(45,20)(46,20)(47,20)(48,20)(49,20)(50,20)(51,20)(52,19)(53,19)(54,19)(55,19)(56,19)(57,19)(58,19)(59,19)(60,18)(61,18)(62,18)(63,18)(64,18)(65,18)(66,18)(67,17)(68,17)(69,17)(70,17)(71,17)(72,17)(73,17)(74,16)(75,16)(76,16)(77,16)(78,16)(79,16)(80,15)(81,15)(82,15)(83,15)(84,14)(85,14)(86,14)(87,14)(88,14)(89,14)(90,14)(91,14)(92,14)(93,13)(94,13)(95,13)(96,13)(97,13)(98,13)(99,12)(100,12)(101,12)(102,12)(103,12)(104,12)(105,12)(106,12)(107,12)(108,12)(109,11)(110,11)(111,11)(112,11)(113,10)(114,10)(115,10)(116,10)(117,10)(118,9)(119,9)(120,9)(121,9)(122,9)(123,9)(124,9)(125,9)(126,8)(127,8)(128,8)(129,8)(130,8)(131,8)(132,8)(133,8)(134,8)(135,8)(136,7)(137,7)(138,7)(139,7)(140,7)(141,7)(142,7)(143,6)(144,6)(145,6)(146,6)(147,6)(148,6)(149,6)(150,6)(151,5)(152,5)(153,5)(154,4)(155,4)(156,4)(157,4)(158,4)(159,4)(160,4)(161,4)(162,4)(163,4)(164,4)(165,4)(166,3)(167,3)(168,3)(169,3)(170,3)(171,3)(172,2)(173,2)(174,2)(175,2)(176,2)(177,2)(178,2)(179,2)(180,2)(181,2)(182,2)(183,2)(184,2)(185,2)(186,2)(187,2)(188,1)(189,1)(190,1)(191,1)(192,1)(193,1)(194,1)(195,1)(196,1)(197,1)(198,1)(199,1)(200,1)(201,1)(202,1)(203,1)(204,0)(205,0)(206,0)(207,0)(208,0)(209,0)(210,0)(211,0)(212,0)(213,0)(214,0)(215,0)(216,0)(217,0)(218,0)(219,0)(220,0)(221,0)
+			(1,22)(2,22)(3,22)(4,22)(5,22)(6,22)(7,22)(8,22)(9,22)(10,22)(11,22)(12,22)(13,22)(14,22)(15,22)(16,22)(17,22)(18,22)(19,22)(20,22)(21,22)(22,22)(23,22)(24,22)(25,22)(26,22)(27,22)(28,22)(29,22)(30,22)(31,22)(32,22)(33,22)(34,22)(35,22)(36,22)(37,22)(38,22)(39,22)(40,22)(41,22)(42,22)(43,22)(44,22)(45,21)(46,21)(47,21)(48,21)(49,21)(50,21)(51,21)(52,20)(53,20)(54,20)(55,20)(56,20)(57,20)(58,20)(59,20)(60,19)(61,19)(62,19)(63,19)(64,19)(65,19)(66,18)(67,18)(68,18)(69,18)(70,18)(71,18)(72,18)(73,18)(74,17)(75,17)(76,17)(77,17)(78,16)(79,16)(80,16)(81,16)(82,16)(83,16)(84,15)(85,15)(86,15)(87,15)(88,15)(89,15)(90,15)(91,15)(92,14)(93,14)(94,14)(95,14)(96,13)(97,13)(98,13)(99,13)(100,13)(101,13)(102,13)(103,13)(104,12)(105,12)(106,12)(107,12)(108,12)(109,12)(110,11)(111,11)(112,11)(113,11)(114,11)(115,11)(116,11)(117,11)(118,10)(119,10)(120,10)(121,10)(122,10)(123,10)(124,10)(125,9)(126,9)(127,9)(128,9)(129,9)(130,9)(131,8)(132,8)(133,8)(134,8)(135,8)(136,7)(137,7)(138,7)(139,7)(140,7)(141,7)(142,7)(143,7)(144,6)(145,6)(146,6)(147,6)(148,6)(149,6)(150,6)(151,6)(152,5)(153,5)(154,5)(155,5)(156,5)(157,5)(158,4)(159,4)(160,4)(161,4)(162,4)(163,4)(164,4)(165,4)(166,4)(167,4)(168,4)(169,3)(170,3)(171,3)(172,3)(173,3)(174,3)(175,2)(176,2)(177,2)(178,2)(179,2)(180,2)(181,2)(182,2)(183,2)(184,2)(185,2)(186,2)(187,2)(188,2)(189,2)(190,1)(191,1)(192,1)(193,1)(194,1)(195,1)(196,1)(197,1)(198,1)(199,1)(200,1)(201,1)(202,1)(203,1)(204,1)(205,0)(206,0)(207,0)(208,0)(209,0)(210,0)(211,0)(212,0)(213,0)(214,0)(215,0)(216,0)(217,0)(218,0)(219,0)(220,0)(221,0)
 		};
@@
 Our experiments aim to ascertain the (relative) quality of predictions that can be made in the context of One Shot Change (OSC) and Time To Live (TTL) based on query features, statistics of predicate dynamics, as well as information about the historical evolution of queries.
-\ah{Update with research questions later}
-
-\paragraph{Setting:} Given that our tasks are prediction based, we must thus build a training set and a test set.
-With 23 versions, we must hold out at least one version for testing, leaving a maximum interval of 22 past versions to use for predictions.
-However, the more past versions we consider for a prediction, the fewer examples we can generate; for example, if we consider 22 versions for training, we will only have one ground truth label for each query.
-In the end, we opted to consider the 3 previous versions ($G_{i-3}, G_{i-2}, G_{i-1}$) to predict changes for the subsequent version ($G_i$); this allows us to predict from $G_4$ to $G_{23}$ inclusive, providing 20 examples per query and 4420 examples in total.
+
+\paragraph{Setting:} To build our final datasets, given that our data features are temporal based, we must thus define a past interval to consider.
+With 23 versions, we must hold out at least one version to label, leaving a maximum interval of 22 past versions to use for compute our data features.
+However, the more past versions we consider for a prediction, the fewer examples we can generate; for example, if we consider 22 versions for compute, we will only have one ground truth label for each query.
+In the end, we opted to consider intervals of 3, 5, 9 and 17 previous versions, for example, ($G_{i-3}, G_{i-2}, G_{i-1}$) to predict changes in queries results for the subsequent version ($G_i$); this allows us to predict from $G_4$ to $G_{23}$ inclusive, providing 20 examples per query and 4420 examples in the largest dataset and 1326 examples in the smallest dataset.
 We remark that no feature explicitly indicates which specific version numbers are given and which are to be predicted.
-(\ah{Discuss the gap: \textit{We evaluated the quality of our proposal by forming a set of classifiers to predict the change of 221 queries in 15 versions, which is equivalent to 3315 objects to classify, described in terms of our proposed features.}}) \ah{Could we do experiments with different intervals?}
 
 In the case of predicate dynamics, as discussed in Section~\ref{sec:approach}, there are potentially many predicates per query, each with its own value for $\Delta(\mathcal{G},p)$, where to reduce (and fix) dimensionality, we may apply an aggregate function to choose the min, max or mean value over all predicates.
-In preliminary experiments, we found the mean value to offer the best results, followed by the max value; hence in what follows we adopt the mean value in our experiments.
-
-We tested the Pearson correlation ($\rho$) between the features available in our data.
-We found only one pair of highly correlated variables: \textbf{\textnumero\ of triple patterns} and \textbf{\textnumero\ of variables} ($\rho = 0.943$).
-\ah{Not sure about this part: what does it add?
-What does ``highly correlated mean?'' Also it's surprising that, e.g., \textbf{\textnumero of predicates} is not also highly correlated here.}
+\ah{In preliminary experiments, we found the mean value to offer the best results, followed by the max value; hence in what follows we adopt the mean value in our experiments.}
 
@@
 We split the data to use 80\% for training and 20\% for tests.
-To avoid overfitting, we use cross-validation in the training data.
-We consider three types of features: static query features (\textsc{q}): predicate dynamics (\textsc{p}) and information on historical results (\texttt{r}).
-\ah{We should be more explicit on precisely what \textsc{r} is?} We may also consider combinations of these features: \textsc{qp}, \textsc{qs}, \textsc{pr} and \textsc{qpr}.
-In our experiments, we compare \textsc{qp} with \textsc{r}.
-\ah{Is this correct?
-Or would it be \textsc{qr}?} \ah{Why only 2?
-Could we not compare all seven?}
+To avoid overfitting, we use stratified cross-validation in the training data.
+We consider three types of features: static query features (\textsc{q}): predicate dynamics (\textsc{p}) considering information of $\Delta(\mathcal{G},p)$ and the cardinality of predicates and information on historical results (\texttt{r}) considering both, the number of past changes in the fixed interval and the cardinalities of those changes divided by the sizes of query results.
+We may also consider combinations of these features: \textsc{qp}, \textsc{qr}, \textsc{pr} and \textsc{qpr}.
 
@@
 To train the model we consider the changes observed in the monitored interval.
-For each update of the results of a query, we have created an example of training with the previously presented features.
-The target TTL value that we learn is the time interval during which the results of the query remain fresh from the current update time.
-\ah{This paragraph is a little vague/unclear to me.}
+For each update of the results of a query, we have created a training example with the previously presented features.
+The target TTL value that we learn is the time interval during which the results of the query remain fresh from the previous update time.
+As an example, suppose that the query q change three times ($G_{3}, G_{7}, G_{9}$) over all versions, we consider two examples from $G_{3}$ and $G_{7}$ with target TTL value 4 and 2 respectively.
 
@@
 \begin{table*}[t]
-	\centering
-	\caption{$F_1$-measure for tested classifiers on OSC}
-	\label{tab:classifier}
-	\begin{tabular}{lrr}    \toprule	
-		\textbf{Classifier}                   & \textsc{r} & \textsc{qp} \\    \midrule
-		\textit{Random Baseline}				 & 0.497869 & 0.486923 \\\midrule
-		Decision Trees                & 0.695591 & \underline{0.651462} \\
-		Linear SVM                   & \underline{0.705755} & 0.593846 \\
-		Naive Bayes                  & 0.661239 & 0.587692 \\ 
-		Nearest Neighbors            & 0.701587 & 0.635923 \\   \bottomrule
+	\begin{tabular}{lrrrrrrr}    \toprule
+		\textbf{Classifier}        & \textsc{Q}         & \textsc{P}         & \textsc{R}         & \textsc{QP}        & \textsc{QR}        & \textsc{PR}        & \textsc{QPR}       \\    \midrule
+		\textit{Random Baseline}   & 0.5002262 & 0.5126697 & 0.5027149 & 0.511086  & 0.5031674 & 0.5063348 & 0.4979638 \\
+		Decision Trees    & 0.4708145 & 0.5233032 & 0.7821267 & 0.4857466 & 0.5671946 & 0.6803167 & 0.6047511 \\
+		Naive Bayes       & 0.5169683 & 0.4647059 & 0.7638009 & 0.5242081 & 0.5420814 & 0.7588235 & 0.5337104 \\
+		Nearest Neighbors & 0.4846154 & 0.5106335 & 0.7819005 & 0.5056561 & 0.6776018 & 0.7760181 & 0.6832579 \\
+		Linear SVM        & 0.5914027 & 0.5108597 & 0.7638009 & 0.5914027 & 0.7751131 & 0.7647059 & 0.7746606 \\    \midrule
+		\textbf{Better}            & 0.5914027 & 0.5233032 & 0.7821267 & 0.5914027 & 0.7751131 & 0.7760181 & 0.7746606 \\    \bottomrule
+		\textit{Random Baseline}   & 0.5010055 & 0.5005028 & 0.4959779 & 0.4987431 & 0.4974862 & 0.4856712 & 0.4816491 \\
+		Decision Trees    & 0.4497235 & 0.5294118 & 0.782554  & 0.4954751 & 0.5723982 & 0.688537  & 0.6289593 \\
+		Naive Bayes       & 0.5160885 & 0.4705882 & 0.7946204 & 0.5160885 & 0.5510307 & 0.7787833 & 0.5334339 \\
+		Nearest Neighbors & 0.5125691 & 0.5296631 & 0.7868276 & 0.5115636 & 0.6792358 & 0.7875817 & 0.6726998 \\
+		Linear SVM        & 0.5957768 & 0.5075415 & 0.7946204 & 0.5849673 & 0.7895928 & 0.7885872 & 0.7880845 \\    \midrule
+		\textbf{Better}            & 0.5957768 & 0.5296631 & 0.7946204 & 0.5849673 & 0.7895928 & 0.7885872 & 0.7880845 \\    \bottomrule
+		\textit{Random Baseline}   & 0.4961215 & 0.5087266 & 0.5054945 & 0.487395  & 0.512605  & 0.4938591 & 0.5090498 \\
+		Decision Trees    & 0.448287  & 0.54234   & 0.7918552 & 0.4974144 & 0.6263736 & 0.7117001 & 0.6764706 \\
+		Naive Bayes       & 0.5258565 & 0.4605688 & 0.7983193 & 0.5235941 & 0.5688429 & 0.7685844 & 0.5585003 \\
+		Nearest Neighbors & 0.5206852 & 0.5336134 & 0.789916  & 0.5310278 & 0.6706529 & 0.7886231 & 0.6780866 \\
+		Linear SVM        & 0.5727214 & 0.5       & 0.7983193 & 0.5688429 & 0.7957337 & 0.7950873 & 0.7947641 \\    \midrule
+		\textbf{Better}          & 0.5727214 & 0.54234   & 0.7983193 & 0.5688429 & 0.7957337 & 0.7950873 & 0.7947641 \\    \bottomrule
+		\textit{Random Baseline}   & 0.4766214 & 0.4962293 & 0.5067873 & 0.5030166 & 0.4788839 & 0.484917  & 0.5030166 \\
+		Decision Trees    & 0.4969834 & 0.4819005 & 0.8205128 & 0.5346908 & 0.7292609 & 0.7503771 & 0.7277526 \\
+		Naive Bayes       & 0.5090498 & 0.4834087 & 0.8280543 & 0.5037707 & 0.5791855 & 0.7978884 & 0.5874811 \\
+		Nearest Neighbors & 0.5580694 & 0.4871795 & 0.8288084 & 0.5671192 & 0.739819  & 0.8220211 & 0.7239819 \\
+		Linear SVM        & 0.5656109 & 0.5098039 & 0.8288084 & 0.5791855 & 0.8273002 & 0.8288084 & 0.8295626 \\    \midrule
+		\textbf{Better}            & 0.5656109 & 0.5098039 & 0.8288084 & 0.5791855 & 0.8273002 & 0.8288084 & 0.8295626 \\     \bottomrule
 	\end{tabular}
 \end{table*}
+
 \begin{comment}
@@
 
-\ah{Some reference can be shortened to just expand the acronym, no more, for example; \textit{Advances in Information Retrieval -- 34th European Conference on IR Research, ECIR, 2012}; can become \textit{European Conference on IR Research, ECIR, 2012}.
-Addresses and dates are not necessary (keep the year of course).
-The workshop name is enough; no need to also give the conference.
-Keeping the publisher for conference is good, but maybe IEEE Computer Society can become IEEE.}
-
 \bibliographystyle{splncs04}

### 2019-04-11 kgswc19.tex
second

@@
     citecolor={dark-green}, urlcolor={dark-blue},
-    pdftitle={Estimating the dynamics of SPARQL query results over Wikidata},    % title
+    pdftitle={Estimating the dynamics of SPARQL query results},    % title
     pdfauthor={Alberto Moya, Aidan Hogan},     % author
@@
 %
-\title{Estimating the dynamics of\\SPARQL query results over Wikidata\thanks{This work was supported by CONICYT PFCHA, Doctorado Becas Chile/2017 -- 72180000, the Millennium Institute for Foundational Research on Data (IMFD) and FONDECYT Grant No.
+\title{Estimating the dynamics of SPARQL query results\thanks{This work was supported by CONICYT PFCHA, Doctorado Becas Chile/2017 -- 72180000, the Millennium Institute for Foundational Research on Data (IMFD) and FONDECYT Grant No.
 1181896.}}
@@
 In the case of predicate dynamics, as discussed in Section~\ref{sec:approach}, there are potentially many predicates per query, each with its own value for $\Delta(\mathcal{G},p)$, where to reduce (and fix) dimensionality, we may apply an aggregate function to choose the min, max or mean value over all predicates.
-\ah{In preliminary experiments, we found the mean value to offer the best results, followed by the max value; hence in what follows we adopt the mean value in our experiments.}
+In preliminary experiments, we found the mean value to offer the best results, followed by the max value; hence in what follows we adopt the mean value in our experiments.
 
@@
 		\textbf{Classifier}        & \textsc{Q}         & \textsc{P}         & \textsc{R}         & \textsc{QP}        & \textsc{QR}        & \textsc{PR}        & \textsc{QPR}       \\    \midrule
-		\textit{Random Baseline}   & 0.5002262 & 0.5126697 & 0.5027149 & 0.511086  & 0.5031674 & 0.5063348 & 0.4979638 \\
-		Decision Trees    & 0.4708145 & 0.5233032 & 0.7821267 & 0.4857466 & 0.5671946 & 0.6803167 & 0.6047511 \\
-		Naive Bayes       & 0.5169683 & 0.4647059 & 0.7638009 & 0.5242081 & 0.5420814 & 0.7588235 & 0.5337104 \\
-		Nearest Neighbors & 0.4846154 & 0.5106335 & 0.7819005 & 0.5056561 & 0.6776018 & 0.7760181 & 0.6832579 \\
-		Linear SVM        & 0.5914027 & 0.5108597 & 0.7638009 & 0.5914027 & 0.7751131 & 0.7647059 & 0.7746606 \\    \midrule
-		\textbf{Better}            & 0.5914027 & 0.5233032 & 0.7821267 & 0.5914027 & 0.7751131 & 0.7760181 & 0.7746606 \\    \bottomrule
-		\textit{Random Baseline}   & 0.5010055 & 0.5005028 & 0.4959779 & 0.4987431 & 0.4974862 & 0.4856712 & 0.4816491 \\
-		Decision Trees    & 0.4497235 & 0.5294118 & 0.782554  & 0.4954751 & 0.5723982 & 0.688537  & 0.6289593 \\
-		Naive Bayes       & 0.5160885 & 0.4705882 & 0.7946204 & 0.5160885 & 0.5510307 & 0.7787833 & 0.5334339 \\
-		Nearest Neighbors & 0.5125691 & 0.5296631 & 0.7868276 & 0.5115636 & 0.6792358 & 0.7875817 & 0.6726998 \\
-		Linear SVM        & 0.5957768 & 0.5075415 & 0.7946204 & 0.5849673 & 0.7895928 & 0.7885872 & 0.7880845 \\    \midrule
-		\textbf{Better}            & 0.5957768 & 0.5296631 & 0.7946204 & 0.5849673 & 0.7895928 & 0.7885872 & 0.7880845 \\    \bottomrule
-		\textit{Random Baseline}   & 0.4961215 & 0.5087266 & 0.5054945 & 0.487395  & 0.512605  & 0.4938591 & 0.5090498 \\
-		Decision Trees    & 0.448287  & 0.54234   & 0.7918552 & 0.4974144 & 0.6263736 & 0.7117001 & 0.6764706 \\
-		Naive Bayes       & 0.5258565 & 0.4605688 & 0.7983193 & 0.5235941 & 0.5688429 & 0.7685844 & 0.5585003 \\
-		Nearest Neighbors & 0.5206852 & 0.5336134 & 0.789916  & 0.5310278 & 0.6706529 & 0.7886231 & 0.6780866 \\
-		Linear SVM        & 0.5727214 & 0.5       & 0.7983193 & 0.5688429 & 0.7957337 & 0.7950873 & 0.7947641 \\    \midrule
-		\textbf{Better}          & 0.5727214 & 0.54234   & 0.7983193 & 0.5688429 & 0.7957337 & 0.7950873 & 0.7947641 \\    \bottomrule
-		\textit{Random Baseline}   & 0.4766214 & 0.4962293 & 0.5067873 & 0.5030166 & 0.4788839 & 0.484917  & 0.5030166 \\
-		Decision Trees    & 0.4969834 & 0.4819005 & 0.8205128 & 0.5346908 & 0.7292609 & 0.7503771 & 0.7277526 \\
-		Naive Bayes       & 0.5090498 & 0.4834087 & 0.8280543 & 0.5037707 & 0.5791855 & 0.7978884 & 0.5874811 \\
-		Nearest Neighbors & 0.5580694 & 0.4871795 & 0.8288084 & 0.5671192 & 0.739819  & 0.8220211 & 0.7239819 \\
-		Linear SVM        & 0.5656109 & 0.5098039 & 0.8288084 & 0.5791855 & 0.8273002 & 0.8288084 & 0.8295626 \\    \midrule
-		\textbf{Better}            & 0.5656109 & 0.5098039 & 0.8288084 & 0.5791855 & 0.8273002 & 0.8288084 & 0.8295626 \\     \bottomrule
+		\textit{Random Baseline}   & 0.490  & 0.486 & 0.500   & 0.502 & 0.486 & 0.495 & 0.491 \\
+		Decision Trees    & 0.471 & 0.513 & \underline{0.782} & 0.497 & 0.572 & 0.705 & 0.605 \\
+		Naive Bayes       & 0.514 & \underline{0.523} & 0.764 & 0.514 & 0.540 & 0.763 & 0.540  \\
+		Nearest Neighbors & 0.505 & 0.521 & 0.778 & 0.509 & 0.679 & \underline{0.781} & 0.680  \\
+		Linear SVM        & \underline{0.596} & 0.518 & 0.764 & \underline{0.593} & \underline{0.775} & 0.767 & \underline{0.774} \\    \midrule
+		\textbf{Better}   & 0.596 & 0.523 & 0.782 & 0.593 & 0.775 & 0.781 & 0.774 \\    \bottomrule
+		\textit{Random Baseline}   & 0.487 & 0.497 & 0.500 & 0.512 & 0.502 & 0.515 & 0.501 \\
+		Decision Trees    & 0.482 & 0.515 & 0.782 & 0.494 & 0.580  & 0.714 & 0.606 \\
+		Naive Bayes       & 0.514 & 0.514 & 0.782 & 0.513 & 0.561 & \underline{0.797} & 0.554 \\
+		Nearest Neighbors & 0.509 & \underline{0.535} & \underline{0.797} & 0.525 & 0.676 & 0.791 & 0.674 \\
+		Linear SVM        & \underline{0.600} & 0.516 & 0.785 & \underline{0.590} & \underline{0.788} & 0.790  & \underline{0.788} \\    \midrule
+		\textbf{Better}   & 0.600 & 0.535 & 0.797 & 0.590 & 0.788 & 0.797 & 0.788 \\    \bottomrule
+		\textit{Random Baseline}   & 0.507 & 0.498 & 0.487 & 0.506 & 0.486 & 0.497 & 0.495 \\
+		Decision Trees    & 0.469 & 0.514 & 0.794 & 0.457 & 0.656 & 0.731 & 0.642 \\
+		Naive Bayes       & 0.517 & 0.519 & \underline{0.797} & 0.516 & 0.577 & 0.791 & 0.576 \\
+		Nearest Neighbors & 0.514 & \underline{0.529} & 0.787 & 0.518 & 0.654 & \underline{0.800} & 0.65  \\
+		Linear SVM        & \underline{0.578} & 0.512 & \underline{0.797} & \underline{0.571} & \underline{0.796} & \underline{0.800} & \underline{0.796} \\    \midrule
+		\textbf{Better}   & 0.578 & 0.529 & 0.797 & 0.571 & 0.796 & 0.800 & 0.796 \\    \bottomrule
+		\textit{Random Baseline}   & 0.514 & 0.515 & 0.502 & 0.493 & 0.499 & 0.499 & 0.502 \\
+		Decision Trees    & 0.512 & 0.517 & \underline{0.827} & 0.493 & 0.725 & 0.716 & 0.704 \\
+		Naive Bayes       & 0.515 & \underline{0.519} & \underline{0.827} & 0.508 & 0.599 & 0.810  & 0.594 \\
+		Nearest Neighbors & 0.549 & 0.481 & \underline{0.827} & 0.555 & 0.744 & 0.821 & 0.732 \\
+		Linear SVM        & \underline{0.554} & \underline{0.519} & \underline{0.827} & \underline{0.560} & \underline{0.819} & \underline{0.831} & \underline{0.815} \\    \midrule
+		\textbf{Better}   & 0.554 & 0.519 & 0.827 & 0.560 & 0.819 & 0.831 & 0.815 \\    \bottomrule
 	\end{tabular}
 \end{table*}
+
+\begin{figure}[t]
+	%\begin{minipage}[b]{0.30\linewidth}
+	\centering
+	\begin{tikzpicture}[thick,font=\scriptsize]
+	\begin{axis}[
+	%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
+	xlabel={F1},
+	width=5cm, 
+	height=5cm,
+	ymin=0,
+	xtick={1,2,...,4},
+	xticklabels={},
+	extra x ticks={1,23},
+	legend pos=north west,
+	ymajorgrids=true,
+	grid style=dashed,
+	]
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.596)(2,0.600)(3,0.578)(4,0.554)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.523)(2,0.535)(3,0.529)(4,0.519)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.782)(2,0.797)(3,0.797)(4,0.827)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.593)(2,0.590)(3,0.571)(4,0.560)
+	};
+
+		\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.775)(2,0.788)(3,0.796)(4,0.819)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.781)(2,0.797)(3,0.800)(4,0.831)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.774)(2,0.788)(3,0.796)(4,0.815)
+	};
+
+	%\legend{# entities}
+	%\legend{# predicates}
+	%\legend{# literals}
+	%\legend{# blank nodes}
+
+	\end{axis}
+	\end{tikzpicture}
+	%		\caption{Evolution of the number of triples on Wikidata.}
+	%		\label{graph:datasize}
+	%	\end{minipage}
+	\hspace{0.1cm}
+	%	\begin{minipage}[b]{0.30\linewidth}
+	%		\centering
+	\begin{tikzpicture}[thick,font=\scriptsize]
+	\begin{axis}[
+	%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
+	xlabel={Precision},
+	width=5cm, 
+	height=5cm,
+	ymin=0,
+	xtick={1,2,...,23},
+	xticklabels={},
+	extra x ticks={1,23},
+	legend pos=north west,
+	ymajorgrids=true,
+	grid style=dashed,
+	]
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.49)(2,0.600)(3,0.578)(4,0.554)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.486)(2,0.535)(3,0.529)(4,0.519)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.5)(2,0.797)(3,0.797)(4,0.827)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.502)(2,0.590)(3,0.571)(4,0.560)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.486)(2,0.788)(3,0.796)(4,0.819)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.495)(2,0.797)(3,0.800)(4,0.831)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.491)(2,0.788)(3,0.796)(4,0.815)
+	};
+
+	%\legend{# entities}
+	%\legend{# predicates}
+	%\legend{# literals}
+	%\legend{# blank nodes}
+
+	\end{axis}
+	\end{tikzpicture}
+	%		\caption{Evolution of the number of triples on Wikidata.}
+	%		\label{graph:datasize}
+	%	\end{minipage}
+	\hspace{0.1cm}
+	%	\begin{minipage}[b]{0.30\linewidth}
+	%		\centering
+	\begin{tikzpicture}[thick,font=\scriptsize]
+	\begin{axis}[
+	%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
+	xlabel={Recall},
+	width=5cm, 
+	height=5cm,
+	ymin=0,
+	xtick={1,2,...,23},
+	xticklabels={},
+	extra x ticks={1,23},
+	legend pos=north west,
+	ymajorgrids=true,
+	grid style=dashed,
+	]
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.596)(2,0.600)(3,0.578)(4,0.554)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.523)(2,0.535)(3,0.529)(4,0.519)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.782)(2,0.797)(3,0.797)(4,0.827)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.593)(2,0.590)(3,0.571)(4,0.560)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.775)(2,0.788)(3,0.796)(4,0.819)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.781)(2,0.797)(3,0.800)(4,0.831)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.774)(2,0.788)(3,0.796)(4,0.815)
+	};
+
+	%\legend{# entities}
+	%\legend{# predicates}
+	%\legend{# literals}
+	%\legend{# blank nodes}
+
+	\end{axis}
+	\end{tikzpicture}
+	\caption{Evolution of total triples in Wikidata (left) and number of triples added to and removed from the $n$\textsuperscript{th} version for the $n+1$\textsuperscript{th} version (right) \label{fig:result}}
+\end{figure}
 

### 2019-04-11 kgswc19.tex
Tabla y graficas terminadas

@@
 
+% Please add the following required packages to your document preamble:
+% \usepackage[normalem]{ulem}
+% \useunder{\uline}{\ul}{}
 \begin{table*}[t]
@@
 		\textbf{Classifier}        & \textsc{Q}         & \textsc{P}         & \textsc{R}         & \textsc{QP}        & \textsc{QR}        & \textsc{PR}        & \textsc{QPR}       \\    \midrule
-		\textit{Random Baseline}   & 0.490  & 0.486 & 0.500   & 0.502 & 0.486 & 0.495 & 0.491 \\
-		Decision Trees    & 0.471 & 0.513 & \underline{0.782} & 0.497 & 0.572 & 0.705 & 0.605 \\
-		Naive Bayes       & 0.514 & \underline{0.523} & 0.764 & 0.514 & 0.540 & 0.763 & 0.540  \\
-		Nearest Neighbors & 0.505 & 0.521 & 0.778 & 0.509 & 0.679 & \underline{0.781} & 0.680  \\
-		Linear SVM        & \underline{0.596} & 0.518 & 0.764 & \underline{0.593} & \underline{0.775} & 0.767 & \underline{0.774} \\    \midrule
-		\textbf{Better}   & 0.596 & 0.523 & 0.782 & 0.593 & 0.775 & 0.781 & 0.774 \\    \bottomrule
-		\textit{Random Baseline}   & 0.487 & 0.497 & 0.500 & 0.512 & 0.502 & 0.515 & 0.501 \\
-		Decision Trees    & 0.482 & 0.515 & 0.782 & 0.494 & 0.580  & 0.714 & 0.606 \\
-		Naive Bayes       & 0.514 & 0.514 & 0.782 & 0.513 & 0.561 & \underline{0.797} & 0.554 \\
-		Nearest Neighbors & 0.509 & \underline{0.535} & \underline{0.797} & 0.525 & 0.676 & 0.791 & 0.674 \\
-		Linear SVM        & \underline{0.600} & 0.516 & 0.785 & \underline{0.590} & \underline{0.788} & 0.790  & \underline{0.788} \\    \midrule
-		\textbf{Better}   & 0.600 & 0.535 & 0.797 & 0.590 & 0.788 & 0.797 & 0.788 \\    \bottomrule
-		\textit{Random Baseline}   & 0.507 & 0.498 & 0.487 & 0.506 & 0.486 & 0.497 & 0.495 \\
-		Decision Trees    & 0.469 & 0.514 & 0.794 & 0.457 & 0.656 & 0.731 & 0.642 \\
-		Naive Bayes       & 0.517 & 0.519 & \underline{0.797} & 0.516 & 0.577 & 0.791 & 0.576 \\
-		Nearest Neighbors & 0.514 & \underline{0.529} & 0.787 & 0.518 & 0.654 & \underline{0.800} & 0.65  \\
-		Linear SVM        & \underline{0.578} & 0.512 & \underline{0.797} & \underline{0.571} & \underline{0.796} & \underline{0.800} & \underline{0.796} \\    \midrule
-		\textbf{Better}   & 0.578 & 0.529 & 0.797 & 0.571 & 0.796 & 0.800 & 0.796 \\    \bottomrule
-		\textit{Random Baseline}   & 0.514 & 0.515 & 0.502 & 0.493 & 0.499 & 0.499 & 0.502 \\
-		Decision Trees    & 0.512 & 0.517 & \underline{0.827} & 0.493 & 0.725 & 0.716 & 0.704 \\
-		Naive Bayes       & 0.515 & \underline{0.519} & \underline{0.827} & 0.508 & 0.599 & 0.810  & 0.594 \\
-		Nearest Neighbors & 0.549 & 0.481 & \underline{0.827} & 0.555 & 0.744 & 0.821 & 0.732 \\
-		Linear SVM        & \underline{0.554} & \underline{0.519} & \underline{0.827} & \underline{0.560} & \underline{0.819} & \underline{0.831} & \underline{0.815} \\    \midrule
-		\textbf{Better}   & 0.554 & 0.519 & 0.827 & 0.560 & 0.819 & 0.831 & 0.815 \\    \bottomrule
+		\textit{Random Baseline} & 0.499          & 0.489          & 0.502          & 0.489          & 0.495          & 0.495          & 0.516          \\
+		Decision Trees           & 0.497          & \underline{0.519}    & \underline{0.781}    & 0.497          & 0.569          & 0.706          & 0.6            \\
+		Naive Bayes              & 0.432          & 0.419          & 0.758          & 0.432          & 0.478          & 0.759          & 0.479          \\
+		Nearest Neighbors        & 0.505          & 0.517          & 0.776          & 0.509          & 0.679          & \underline{0.781}    & 0.68           \\
+		Linear SVM               & \underline{0.596}    & 0.372          & 0.758          & \underline{0.593}    & \underline{0.775}    & 0.762          & \underline{0.774}    \\    \midrule
+		\textbf{Better}          & \textbf{0.596} & \textbf{0.519} & \textbf{0.781} & \textbf{0.593} & \textbf{0.775} & \textbf{0.781} & \textbf{0.774} \\    \bottomrule
+		\textit{Random Baseline} & 0.48           & 0.496          & 0.499          & 0.505          & 0.495          & 0.497          & 0.499          \\
+		Decision Trees           & 0.485          & 0.514          & 0.782          & 0.494          & 0.581          & 0.711          & 0.603          \\
+		Naive Bayes              & 0.431          & 0.486          & 0.782          & 0.432          & 0.508          & \underline{0.797}    & 0.5            \\
+		Nearest Neighbors        & 0.517          & \underline{0.535}    & \underline{0.795}    & 0.525          & 0.676          & 0.79           & 0.673          \\
+		Linear SVM               & \underline{0.6}      & 0.384          & 0.784          & \underline{0.59}     & \underline{0.787}    & 0.789          & \underline{0.787}    \\    \midrule
+		\textbf{Better}          & \textbf{0.6}   & \textbf{0.535} & \textbf{0.795} & \textbf{0.59}  & \textbf{0.787} & \textbf{0.797} & \textbf{0.787} \\    \bottomrule
+		\textit{Random Baseline} & 0.497          & 0.505          & 0.493          & 0.483          & 0.497          & 0.511          & 0.515          \\
+		Decision Trees           & 0.456          & 0.517          & 0.793          & 0.48           & 0.668          & 0.735          & 0.651          \\
+		Naive Bayes              & 0.434          & 0.461          & \underline{0.797}    & 0.435          & 0.531          & 0.79           & 0.531          \\
+		Nearest Neighbors        & 0.514          & \underline{0.528}    & 0.787          & 0.517          & 0.653          & \underline{0.8}      & 0.648          \\
+		Linear SVM               & \underline{0.578}    & 0.397          & \underline{0.797}    & \underline{0.571}    & \underline{0.795}    & \underline{0.8}      & \underline{0.796}    \\    \midrule
+		\textbf{Better}          & \textbf{0.578} & \textbf{0.528} & \textbf{0.797} & \textbf{0.571} & \textbf{0.795} & \textbf{0.8}   & \textbf{0.796} \\    \bottomrule
+		\textit{Random Baseline} & 0.486          & 0.494          & 0.482          & 0.512          & 0.524          & 0.511          & 0.49           \\
+		Decision Trees           & 0.511          & \underline{0.511}    & \underline{0.827}    & 0.469          & 0.712          & 0.712          & 0.71           \\
+		Naive Bayes              & 0.416          & 0.465          & 0.826          & 0.414          & 0.555          & 0.81           & 0.554          \\
+		Nearest Neighbors        & 0.548          & 0.48           & 0.826          & 0.554          & 0.744          & 0.82           & 0.731          \\
+		Linear SVM               & \underline{0.554}    & 0.438          & 0.826          & \underline{0.56}     & \underline{0.819}    & \underline{0.831}    & \underline{0.815}    \\    \midrule
+		\textbf{Better}          & \textbf{0.554} & \textbf{0.511} & \textbf{0.827} & \textbf{0.56}  & \textbf{0.819} & \textbf{0.831} & \textbf{0.815}	\\    \bottomrule
 	\end{tabular}
@@
 	coordinates {
-		(1,0.596)(2,0.600)(3,0.578)(4,0.554)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.523)(2,0.535)(3,0.529)(4,0.519)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.782)(2,0.797)(3,0.797)(4,0.827)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.593)(2,0.590)(3,0.571)(4,0.560)
-	};
-
-		\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.775)(2,0.788)(3,0.796)(4,0.819)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.781)(2,0.797)(3,0.800)(4,0.831)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.774)(2,0.788)(3,0.796)(4,0.815)
+		(1,0.596)(2,0.6)(3,0.578)(4,0.554)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.519)(2,0.535)(3,0.528)(4,0.511)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.781)(2,0.795)(3,0.797)(4,0.827)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.593)(2,0.59)(3,0.571)(4,0.56)
+	};
+
+		\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.775)(2,0.787)(3,0.795)(4,0.819)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.781)(2,0.797)(3,0.8)(4,0.831)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.774)(2,0.787)(3,0.796)(4,0.815)
 	};
@@
 	coordinates {
-		(1,0.49)(2,0.600)(3,0.578)(4,0.554)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.486)(2,0.535)(3,0.529)(4,0.519)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.5)(2,0.797)(3,0.797)(4,0.827)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.502)(2,0.590)(3,0.571)(4,0.560)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.486)(2,0.788)(3,0.796)(4,0.819)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.495)(2,0.797)(3,0.800)(4,0.831)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.491)(2,0.788)(3,0.796)(4,0.815)
+		(1,0.598)(2,0.601)(3,0.601)(4,0.619)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.52)(2,0.535)(3,0.528)(4,0.512)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.796)(2,0.798)(3,0.798)(4,0.828)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.596)(2,0.591)(3,0.589)(4,0.579)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.775)(2,0.787)(3,0.795)(4,0.819)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.782)(2,0.797)(3,0.8)(4,0.831)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.774)(2,0.787)(3,0.796)(4,0.815)
 	};
@@
 	coordinates {
-		(1,0.596)(2,0.600)(3,0.578)(4,0.554)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.523)(2,0.535)(3,0.529)(4,0.519)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.782)(2,0.797)(3,0.797)(4,0.827)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.593)(2,0.590)(3,0.571)(4,0.560)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.775)(2,0.788)(3,0.796)(4,0.819)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.781)(2,0.797)(3,0.800)(4,0.831)
+		(1,0.598)(2,0.601)(3,0.579)(4,0.554)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.52)(2,0.535)(3,0.528)(4,0.512)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.787)(2,0.794)(3,0.798)(4,0.828)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.595)(2,0.591)(3,0.572)(4,0.561)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.775)(2,0.787)(3,0.796)(4,0.819)
+	};
+
+	\addplot[
+	color=blue,
+	mark=diamond,
+	]
+	coordinates {
+		(1,0.783)(2,0.797)(3,0.8)(4,0.831)
 	};

### 2019-04-11 kgswc19.tex
an update

@@
     pdftitle={Estimating the dynamics of SPARQL query results},    % title
-    pdfauthor={Alberto Moya, Aidan Hogan},     % author
-    pdfsubject={KGSWC 2019},   % subject of the document
+%    pdfauthor={Alberto Moya, Aidan Hogan},     % author
+    pdfsubject={ISWC 2019},   % subject of the document
     pdfkeywords={SPARQL;} {Linked Data;} {Dynamics;} {Wikidata;}, % list of keywords
@@
 %
-\title{Estimating the dynamics of SPARQL query results\thanks{This work was supported by CONICYT PFCHA, Doctorado Becas Chile/2017 -- 72180000, the Millennium Institute for Foundational Research on Data (IMFD) and FONDECYT Grant No.
-1181896.}}
+\title{Estimating the dynamics of SPARQL query results
+%\thanks{This work was supported by CONICYT PFCHA, Doctorado Becas Chile/2017 -- 72180000, the Millennium Institute for Foundational Research on Data (IMFD) and FONDECYT Grant No. 1181896.}
+}
 %
@@
 %
-\author{Alberto Moya\orcidID{0000-0002-7003-5087} \and 
-Aidan Hogan\orcidID{0000-0001-9482-1982}}
-%
-\authorrunning{A. Moya and A. Hogan}
+%\author{Alberto Moya\orcidID{0000-0002-7003-5087} \and 
+%Aidan Hogan\orcidID{0000-0001-9482-1982}}
+%
+%\authorrunning{A. Moya and A. Hogan}
 % First names are abbreviated in the running head.
@@
 %
-\institute{IMFD; DCC, University of Chile\\
-\email{\{amoya,ahogan\}@dcc.uhile.cl}}
+%\institute{IMFD; DCC, University of Chile\\
+%\email{\{amoya,ahogan\}@dcc.uhile.cl}}
 %
@@
 
-\paragraph{Results:} Table~\ref{tab:classifier} shows the $F_1$-measure for the predictions made; for reference, we include a baseline that randomly guesses yes/no.
-With the highest result being an $F_1$-measure of 0.706, the results show that predicting changes in query results is indeed a challenging problem; on the other hand, all configurations clearly outperform the random baseline.
-Looking in further detail, as can be expected, prediction performs best when the \textsc{r} feature -- with knowledge of changes in the historical value of results of the input query -- is available.
-However, the \textsc{qp} features -- which do not assume the availability of such information -- are often competitive: more specifically, the best result for \textsc{r} is 0.706 with Linear SVM, while the best results for \textsc{qp} is 0.651 with Decision Trees.
-
-Finally, to predict the TTL of the results of the query -- outputting the number of versions for which the query results can be expected to remain the same -- we employed techniques of multiple linear regression using the \textsc{sklearn} and \textsc{tensorflow} libraries.
-The latter allows us to see how the periods of estimation and the convergence analysis will behave.
-We were able to predict the TTL (in relation to the number of invariant weeks) with an average square error of 0.43 and a variance score of 0.01.
-These results allow to predict future changes in query results with good accuracy using linear regression techniques.
-\ah{Could do with a comparison here; the 0.43 result is quite abstract.}
+\paragraph{Results:} Table~\ref{tab:classifier} shows the $F_1$ -measure for the predictions made considering the union of different combinations of the sets of features (Q, P, R) and varying the window sizes (3, 5, 9 , 17); for reference, we include a baseline that randomly guesses yes / no.
+We can notice the best results were obtained using the characteristics of R reaching values ​​of 0.83 of f1.
+The results show that predicting changes in query results is indeed a challenging problem.
+Looking in further detail, as can be expected, prediction performs best when the \ textsc {r} feature - with knowledge of changes in the historical value of results of the input query - is available.
+And, the features from sets \textsc {q} and  \textsc {p} - which do not assume the availability of such information - obtain less F1-score: more specifically, the best result using \ textsc {pr} is 0.83 with Linear SVM, window size 17, while the best results without \ textsc {r} is 0.60 with Linear SVM, window size 5.
+
+Furthermore, we show in Figure~\ref{fig:result} that effectively the size of the window influences the quality of the results when considering r, being better as the window increases.
+However, these results are not affected when we consider only predicates in q or p.
+Also, we show, on the other hand, that all configurations outperform the random baseline.
 
@@
 \begin{table*}[t]
+	\caption{$F_1$-measure for tested classifiers considering different sets of features (Q, P, R) and varying the the window sizes (2,5,9,17)}
+	\label{tab:classifier}
 	\begin{tabular}{lrrrrrrr}    \toprule
@@
 	\end{tikzpicture}
-	\caption{Evolution of total triples in Wikidata (left) and number of triples added to and removed from the $n$\textsuperscript{th} version for the $n+1$\textsuperscript{th} version (right) \label{fig:result}}
+	\caption{Best results for each feature sets in term of $F_1$-measure (left), Precision score (center) and Recall (right) with different windows sizes 
+		\label{fig:result}}
 \end{figure}
+
+Finally, to predict the TTL of the results of the query -- outputting the number of versions for which the query results can be expected to remain the same -- we employed techniques of multiple linear regression using the \textsc{sklearn}.
+The latter allows us to see how the periods of estimation and the convergence analysis will behave.
+We were able to predict the TTL (in relation to the number of invariant weeks) with an average square error of 0.43 weeks and a variance score of 0.01.
+These results allow to predict future changes in query results with good accuracy using linear regression techniques.
 
@@
 
-In this paper we propose methods to predict whether or not (OSC) and/or when (TTL) the results of an input query will change in the future.
-More specifically, we propose a framework based on classifiers and regression that accept features extracted from the query, from past versions of the data, and/or from the combination of both.
-Considering this framework, we propose that there is a conceptual trade-off between the cost of computing features and their value for prediction: features extracted from queries alone are the most efficient to extract, not requiring historical data, but are quite coarse-grained for prediction; on the other hand, features based on changes in results in previous versions for an (unseen) input query are often the most costly to acquire, but offer fine-grained information for prediction; finally, features based on the dynamics of predicates used in the data offer a balance between the two, allowing to summarize historical data into succinct statistics, while offering more fine-grained information than static query features and less fine-grained information than historical results.
+In this paper we evaluate methods to predict whether or not (OSC) and/or when (TTL) the results of an input query will change in the future.
+More specifically, we evaluate a framework based on classifiers and regression that accept features extracted from the query, from past versions of the data, and/or from the combination of both.
+Considering this framework, we show that there is a conceptual trade-off between the cost of computing features and their value for prediction: features extracted from queries alone are the most efficient to extract, not requiring historical data, but are quite coarse-grained for prediction; on the other hand, features based on changes in results in previous versions for an (unseen) input query are often the most costly to acquire, but offer fine-grained information for prediction; finally, features based on the dynamics of predicates used in the data offer a balance between the two, allowing to summarize historical data into succinct statistics, while offering more fine-grained information than static query features and less fine-grained information than historical results.
 
@@
 For OSC, we evaluate the success rates of a variety of classifiers based on the $F_1$-measure of the predictions.
-Our results show that although features based on historical changes to query results perform best ($F_1 = 0.706$), considering static query features and predicate dynamics alone is competitive ($F_1 =0.651$) when historical results are not available or are prohibitively costly to compute.
+Our results show that the features based on historical changes to query results perform best ($F_1 = 0.83$), considering static query features and predicate dynamics alone is less competitive ($F_1 =0.60$) when historical results are not available or are prohibitively costly to compute.
 

### 2019-04-11 kgswc19.tex
...

@@
 \begin{table*}[t]
-	\caption{$F_1$-measure for tested classifiers considering different sets of features (Q, P, R) and varying the the window sizes (2,5,9,17)}
+	\centering
+	\caption{$F_1$-measure for tested classifiers considering different sets of features (Q, P, R) and varying the the window sizes (3,5,9,17)}
 	\label{tab:classifier}
-	\begin{tabular}{lrrrrrrr}    \toprule
-		\textbf{Classifier}        & \textsc{Q}         & \textsc{P}         & \textsc{R}         & \textsc{QP}        & \textsc{QR}        & \textsc{PR}        & \textsc{QPR}       \\    \midrule
-		\textit{Random Baseline} & 0.499          & 0.489          & 0.502          & 0.489          & 0.495          & 0.495          & 0.516          \\
-		Decision Trees           & 0.497          & \underline{0.519}    & \underline{0.781}    & 0.497          & 0.569          & 0.706          & 0.6            \\
-		Naive Bayes              & 0.432          & 0.419          & 0.758          & 0.432          & 0.478          & 0.759          & 0.479          \\
-		Nearest Neighbors        & 0.505          & 0.517          & 0.776          & 0.509          & 0.679          & \underline{0.781}    & 0.68           \\
-		Linear SVM               & \underline{0.596}    & 0.372          & 0.758          & \underline{0.593}    & \underline{0.775}    & 0.762          & \underline{0.774}    \\    \midrule
-		\textbf{Better}          & \textbf{0.596} & \textbf{0.519} & \textbf{0.781} & \textbf{0.593} & \textbf{0.775} & \textbf{0.781} & \textbf{0.774} \\    \bottomrule
-		\textit{Random Baseline} & 0.48           & 0.496          & 0.499          & 0.505          & 0.495          & 0.497          & 0.499          \\
-		Decision Trees           & 0.485          & 0.514          & 0.782          & 0.494          & 0.581          & 0.711          & 0.603          \\
-		Naive Bayes              & 0.431          & 0.486          & 0.782          & 0.432          & 0.508          & \underline{0.797}    & 0.5            \\
-		Nearest Neighbors        & 0.517          & \underline{0.535}    & \underline{0.795}    & 0.525          & 0.676          & 0.79           & 0.673          \\
-		Linear SVM               & \underline{0.6}      & 0.384          & 0.784          & \underline{0.59}     & \underline{0.787}    & 0.789          & \underline{0.787}    \\    \midrule
-		\textbf{Better}          & \textbf{0.6}   & \textbf{0.535} & \textbf{0.795} & \textbf{0.59}  & \textbf{0.787} & \textbf{0.797} & \textbf{0.787} \\    \bottomrule
-		\textit{Random Baseline} & 0.497          & 0.505          & 0.493          & 0.483          & 0.497          & 0.511          & 0.515          \\
-		Decision Trees           & 0.456          & 0.517          & 0.793          & 0.48           & 0.668          & 0.735          & 0.651          \\
-		Naive Bayes              & 0.434          & 0.461          & \underline{0.797}    & 0.435          & 0.531          & 0.79           & 0.531          \\
-		Nearest Neighbors        & 0.514          & \underline{0.528}    & 0.787          & 0.517          & 0.653          & \underline{0.8}      & 0.648          \\
-		Linear SVM               & \underline{0.578}    & 0.397          & \underline{0.797}    & \underline{0.571}    & \underline{0.795}    & \underline{0.8}      & \underline{0.796}    \\    \midrule
-		\textbf{Better}          & \textbf{0.578} & \textbf{0.528} & \textbf{0.797} & \textbf{0.571} & \textbf{0.795} & \textbf{0.8}   & \textbf{0.796} \\    \bottomrule
-		\textit{Random Baseline} & 0.486          & 0.494          & 0.482          & 0.512          & 0.524          & 0.511          & 0.49           \\
-		Decision Trees           & 0.511          & \underline{0.511}    & \underline{0.827}    & 0.469          & 0.712          & 0.712          & 0.71           \\
-		Naive Bayes              & 0.416          & 0.465          & 0.826          & 0.414          & 0.555          & 0.81           & 0.554          \\
-		Nearest Neighbors        & 0.548          & 0.48           & 0.826          & 0.554          & 0.744          & 0.82           & 0.731          \\
-		Linear SVM               & \underline{0.554}    & 0.438          & 0.826          & \underline{0.56}     & \underline{0.819}    & \underline{0.831}    & \underline{0.815}    \\    \midrule
-		\textbf{Better}          & \textbf{0.554} & \textbf{0.511} & \textbf{0.827} & \textbf{0.56}  & \textbf{0.819} & \textbf{0.831} & \textbf{0.815}	\\    \bottomrule
+	\begin{tabular}{l@{\hskip 0.1in}l@{\hskip 0.1in}r@{\hskip 0.1in}r@{\hskip 0.1in}r@{\hskip 0.1in}r@{\hskip 0.1in}r@{\hskip 0.1in}r@{\hskip 0.1in}r}    \toprule		  
+		& \textbf{Classifier}        & \textsc{Q}         & \textsc{P}         & \textsc{R}         & \textsc{QP}        & \textsc{QR}        & \textsc{PR}        & \textsc{QPR}       \\    \midrule
+	w=3 	& \textit{Random Baseline} & 0.499          & 0.489          & 0.502          & 0.489          & 0.495          & 0.495          & 0.516          \\
+		& Decision Trees           & 0.497          & \underline{0.519}    & \underline{0.781}    & 0.497          & 0.569          & 0.706          & 0.6            \\
+		& Naive Bayes              & 0.432          & 0.419          & 0.758          & 0.432          & 0.478          & 0.759          & 0.479          \\
+		& Nearest Neighbors        & 0.505          & 0.517          & 0.776          & 0.509          & 0.679          & \underline{0.781}    & 0.68           \\
+		& Linear SVM               & \underline{0.596}    & 0.372          & 0.758          & \underline{0.593}    & \underline{0.775}    & 0.762          & \underline{0.774}    \\    \midrule
+		& \textbf{Better}          & \textbf{0.596} & \textbf{0.519} & \textbf{0.781} & \textbf{0.593} & \textbf{0.775} & \textbf{0.781} & \textbf{0.774} \\    \bottomrule
+	w=5	& \textit{Random Baseline} & 0.48           & 0.496          & 0.499          & 0.505          & 0.495          & 0.497          & 0.499          \\
+		& Decision Trees           & 0.485          & 0.514          & 0.782          & 0.494          & 0.581          & 0.711          & 0.603          \\
+		& Naive Bayes              & 0.431          & 0.486          & 0.782          & 0.432          & 0.508          & \underline{0.797}    & 0.5            \\
+		& Nearest Neighbors        & 0.517          & \underline{0.535}    & \underline{0.795}    & 0.525          & 0.676          & 0.79           & 0.673          \\
+		& Linear SVM               & \underline{0.6}      & 0.384          & 0.784          & \underline{0.59}     & \underline{0.787}    & 0.789          & \underline{0.787}    \\    \midrule
+		& \textbf{Better}          & \textbf{0.6}   & \textbf{0.535} & \textbf{0.795} & \textbf{0.59}  & \textbf{0.787} & \textbf{0.797} & \textbf{0.787} \\    \bottomrule
+	w=9	& \textit{Random Baseline} & 0.497          & 0.505          & 0.493          & 0.483          & 0.497          & 0.511          & 0.515          \\
+		& Decision Trees           & 0.456          & 0.517          & 0.793          & 0.48           & 0.668          & 0.735          & 0.651          \\
+		& Naive Bayes              & 0.434          & 0.461          & \underline{0.797}    & 0.435          & 0.531          & 0.79           & 0.531          \\
+		& Nearest Neighbors        & 0.514          & \underline{0.528}    & 0.787          & 0.517          & 0.653          & \underline{0.8}      & 0.648          \\
+		& Linear SVM   				& \underline{0.578}    & 0.397          & \underline{0.797}    & \underline{0.571}    & \underline{0.795}    & \underline{0.8}      & \underline{0.796}    \\     \midrule
+		& \textbf{Better}          & \textbf{0.578} & \textbf{0.528} & \textbf{0.797} & \textbf{0.571} & \textbf{0.795} & \textbf{0.8}   & \textbf{0.796} \\    \bottomrule
+	w=17	& \textit{Random Baseline} & 0.486          & 0.494          & 0.482          & 0.512          & 0.524          & 0.511          & 0.49           \\
+		& Decision Trees           & 0.511          & \underline{0.511}    & \underline{0.827}    & 0.469          & 0.712          & 0.712          & 0.71           \\
+		& Naive Bayes              & 0.416          & 0.465          & 0.826          & 0.414          & 0.555          & 0.81           & 0.554          \\
+		& Nearest Neighbors        & 0.548          & 0.48           & 0.826          & 0.554          & 0.744          & 0.82           & 0.731          \\
+		& Linear SVM               & \underline{0.554}    & 0.438          & 0.826          & \underline{0.56}     & \underline{0.819}    & \underline{0.831}    & \underline{0.815}    \\    \midrule
+		& \textbf{Better}          & \textbf{0.554} & \textbf{0.511} & \textbf{0.827} & \textbf{0.56}  & \textbf{0.819} & \textbf{0.831} & \textbf{0.815}	\\    \bottomrule
 	\end{tabular}

### 2019-04-11 kgswc19.tex
update

@@
 We address the problem of estimating when the results of an input SPARQL query results over dynamic RDF datasets will change.
-We propose a framework that extracts features from the query and/or from past versions of the target dataset.
+We evaluate a framework that extracts features from the query and/or from past versions of the target dataset.
 The resulting features are then used to predict (1) whether or not the results for a query will change in the near future using classifiers, and (2) when they might change (if at all) using linear regression.
-To evaluate our proposal, we create a gold standard based on 23 versions of Wikidata and a curated collection of 221 SPARQL queries.
-Our results show that the quality of predictions using query features and lightweight statistics of the dynamics of individual predicates is competitive with that obtained using (more costly to derive) knowledge of the complete historical changes in the query results.
+To evaluate, we create a gold standard based on 23 versions of Wikidata and a curated collection of 221 SPARQL queries.
+Our results show that the quality of predictions using query features and lightweight statistics of the dynamics of individual predicates is less competitive with that obtained using (more costly to derive) knowledge of the complete historical changes in the query results.
 
@@
 In this work, we address the problem of estimating if and/or when SPARQL query results will change in a future version of a dynamic RDF dataset.
-We propose a general architecture based on Machine Learning that extracts static features from a query, as well as dynamic features from the query and dataset.
+We evaluate a general architecture based on Machine Learning that extracts static features from a query, as well as dynamic features from the query and dataset.
 These features are fed into a classifier to predict whether or not the query results will change in a fixed point in the future, or, more ambitiously, can be fed into a regression model to predict when the query results are likely to change.
-With respect to the features used, we propose that there is a trade-off between those that are coarse-grained but easy to compute (e.g., static query features) versus those that are fine-grained but more costly to compute (e.g., historical changes in query results), with further features falling in the middle (e.g., high-level statistics about the dynamics of triples with particular predicates).
+With respect to the features used, we show that there is a trade-off between those that are coarse-grained but easy to compute (e.g., static query features) versus those that are fine-grained but more costly to compute (e.g., historical changes in query results), with further features falling in the middle (e.g., high-level statistics about the dynamics of triples with particular predicates).
 Per this trade-off, which features to use for predicting the dynamics of query results may then depend on the particular application.
@@
 In order to better understand this conceptual trade-off -- and more generally, to evaluate the quality of predictions made by our framework -- we create a novel gold standard based on 23 weekly versions of the Wikidata knowledge graph~\cite{VrandecicK14} and 221 user-generated queries; we show this gold standard to have a variety of desirable features, including (most importantly) a balance of queries whose results never change, always change, as well as non-trivial queries whose results intermittently change.
-Using this gold standard, our experiments show that while the most accurate predictions are based on historical changes to the input query's results, predictions based only on query features and predicate dynamics remain competitive (while being much more efficient to maintain/extract).
-
-\paragraph{Contributions:} (1) We propose a general architecture for predicting when/if the results of a SPARQL query will change in a future version of a dynamic RDF dataset.
-(2) We propose a number of features to instantiate this architecture based on analysis of the query, analysis of the dynamics of predicates in the data, and analysis of historical changes in the query results.
+Using this gold standard, our experiments show that while the most accurate predictions are based on historical changes to the input query's results, predictions based only on query features and predicate dynamics remain less competitive (while being much more efficient to maintain/extract).
+
+\paragraph{Contributions:} (1) We evaluate a general architecture for predicting when/if the results of a SPARQL query will change in a future version of a dynamic RDF dataset.
+(2) We evaluate a number of features to instantiate this architecture based on analysis of the query, analysis of the dynamics of predicates in the data, and analysis of historical changes in the query results.
 (3) We create a gold standard for these tasks based on 23 consecutive versions of Wikidata and a set of 221 real-world SPARQL queries.
@@
 Section~\ref{sec:preliminar} provides a formal statement of the problem.
-Section~\ref{sec:approach} describes our proposed architecture and features.
+Section~\ref{sec:approach} describes the evaluated architecture and features.
 In Section~\ref{sec:data}, we describe the creation of our gold standard based on Wikidata.
@@
 \paragraph{Novelty:} Given a query and a dynamic RDF dataset, we aim to predict (1) whether or not the query's results will have changed in a fixed point in the near future and/or (2) when the results will change.
-Our work thus complements existing works aiming for weak consistency, but (i) generalizes the problem, proposing a framework that can incorporate statistics on predicate dynamics~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12,DehghanzadehPKUHD14} and historical changes in query results~\cite{KnuthHS16}, as proposed in prior works, as well as novel types of features, (ii) introduces new features based on query operators and statistics; (iii) creates a novel gold standard based on Wikidata and presents comparative results that indicate the relative predictive power inherent in different types of features.
+Our work thus complements existing works aiming for weak consistency, but (i) generalizes the problem, evaluating a framework that can incorporate statistics on predicate dynamics~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12,DehghanzadehPKUHD14} and historical changes in query results~\cite{KnuthHS16}, as proposed in prior works, as well as novel types of features, (ii) introduces new features based on query operators and statistics; (iii) creates a novel gold standard based on Wikidata and presents comparative results that indicate the relative predictive power inherent in different types of features.
 

### 2019-04-11 kgswc19.tex
update

@@
 
-\paragraph{Architecture:} In Figure~\ref{fig:schema}, we provide an overview of our proposed architecture for making predictions with respect to the OSC and TTL tasks.
+\paragraph{Architecture:} In Figure~\ref{fig:schema}, we provide an overview of a general architecture for making predictions with respect to the OSC and TTL tasks.
 A SPARQL query $Q$ and a dynamic RDF graph $\mathcal{G}$ are given as input.
@@
 
-In order to evaluate the effectiveness of our proposal for predicting changes in SPARQL query results, we require a dynamic RDF graph, with access to various historical versions; preferably this graph contains real-world, large-scale, diverse data, and with sufficient changes between versions to provide both positive and negative examples of queries whose results change.
+In order to evaluate the effectiveness for predicting changes in SPARQL query results, we require a dynamic RDF graph, with access to various historical versions; preferably this graph contains real-world, large-scale, diverse data, and with sufficient changes between versions to provide both positive and negative examples of queries whose results change.
 Furthermore, we require a set of SPARQL queries that can be answered against this dynamic graph; preferably these queries again should be diverse, representative of real-world user queries, of a variety of shapes and sizes, using a variety of query operators, and with a mix of both dynamic and static results over the dynamic graph.
@@
 Although we have seen that the number of unique predicates did not change much over all versions, we are interested to see which predicates are indicative of dynamic statements.
-Table~\ref{tab:predicates} shows the ten most dynamic predicates according to the number statements involving that predicate (left) and the ratio of added (+) and deleted (−) statements divided by the total number of statements for that predicate across all snapshots; for the second, we only include predicates that appear in all snapshots and appear in ≥ 1, 000 statements overall.
+Table~\ref{tab:predicates} shows the ten most dynamic predicates according to the number of added and deleted statements involving that predicate (Total) and the ratio of added and deleted statements divided by the total number of statements for that predicate across all snapshots (Dyn); for the second, we only include predicates that appear in all snapshots and appear in ≥ 1, 000 statements overall.
 Note that the largest number of changes involves the predicate scheme:description, however, its dynamic value is low due to the number of statements with it in the versions.
@@
 To avoid overfitting, we use stratified cross-validation in the training data.
-We consider three types of features: static query features (\textsc{q}): predicate dynamics (\textsc{p}) considering information of $\Delta(\mathcal{G},p)$ and the cardinality of predicates and information on historical results (\texttt{r}) considering both, the number of past changes in the fixed interval and the cardinalities of those changes divided by the sizes of query results.
+We consider three types of features: static query features (\textsc{q}): predicate dynamics (\textsc{p}) considering information of $\Delta(\mathcal{G},p)$ and the cardinality of predicates and information on historical results (\textsc{r}) considering the number of past changes in the fixed interval.
 We may also consider combinations of these features: \textsc{qp}, \textsc{qr}, \textsc{pr} and \textsc{qpr}.

### 2019-04-11 kgswc19.tex
update

@@
 
-\paragraph{Results:} Table~\ref{tab:classifier} shows the $F_1$ -measure for the predictions made considering the union of different combinations of the sets of features (Q, P, R) and varying the window sizes (3, 5, 9 , 17); for reference, we include a baseline that randomly guesses yes / no.
-We can notice the best results were obtained using the characteristics of R reaching values ​​of 0.83 of f1.
+\paragraph{Results:} Table~\ref{tab:classifier} shows the $F_1$-score for the predictions made considering the union of different combinations of the sets of features (Q, P, R) and varying the window sizes (3, 5, 9, 17); for reference, we include a baseline that randomly guesses yes/no.
+We can note the best results were obtained using the features from \textsc{r} reaching values ​​of 0.83 of $F_1$.
 The results show that predicting changes in query results is indeed a challenging problem.
-Looking in further detail, as can be expected, prediction performs best when the \ textsc {r} feature - with knowledge of changes in the historical value of results of the input query - is available.
-And, the features from sets \textsc {q} and  \textsc {p} - which do not assume the availability of such information - obtain less F1-score: more specifically, the best result using \ textsc {pr} is 0.83 with Linear SVM, window size 17, while the best results without \ textsc {r} is 0.60 with Linear SVM, window size 5.
-
-Furthermore, we show in Figure~\ref{fig:result} that effectively the size of the window influences the quality of the results when considering r, being better as the window increases.
-However, these results are not affected when we consider only predicates in q or p.
-Also, we show, on the other hand, that all configurations outperform the random baseline.
+Looking in further detail, as can be expected, prediction performs best when the \textsc{r} feature - with knowledge of changes in the historical value of results of the input query - is available.
+And, the features from sets \textsc{q} and  \textsc{p} - which do not assume the availability of such information - obtain less $F_1$-score: more specifically, the best result was using \textsc{pr} (0.83) with Linear SVM and window size 17, while the best results without \textsc{r} was using \textsc{Q} (0.60) with Linear SVM and window size 5.
+
+Furthermore, we show in Figure~\ref{fig:result} that effectively the size of the window influences the quality of the results when considering \textsc{r}, being better as the window increases.
+However, these results are not affected when we consider only predicates in \textsc{q} or  \textsc{p}.
+Also, we show, that all configurations outperform the random baseline.
 
@@
 	\centering
-	\caption{$F_1$-measure for tested classifiers considering different sets of features (Q, P, R) and varying the the window sizes (3,5,9,17)}
+	\caption{$F_1$-score for tested classifiers considering different sets of features (Q, P, R) and window sizes (3,5,9,17)}
 	\label{tab:classifier}
@@
 	\centering
-	\begin{tikzpicture}[thick,font=\scriptsize]
+	\begin{tikzpicture}[thick,font=\scriptsize,scale=0.95]
 	\begin{axis}[
@@
 	ymin=0,
-	xtick={1,2,...,4},
+	xtick={3,5,9,17},
 	xticklabels={},
-	extra x ticks={1,23},
+	extra x ticks={3,5,9,17},
 	legend pos=north west,
@@
 	\addplot[
+	color=red,
+	mark=diamond,
+	]
+	coordinates {
+		(3,0.516)(5,0.505)(9,0.515)(17,0.524)
+	};
+
+	\addplot[
+	color=brown,
+	mark=diamond,
+	]
+	coordinates {
+		(3,0.596)(5,0.6)(9,0.578)(17,0.554)
+	};
+
+	\addplot[
+	color=green,
+	mark=diamond,
+	]
+	coordinates {
+		(3,0.519)(5,0.535)(9,0.528)(17,0.511)
+	};
+
+	\addplot[
 	color=blue,
@@
 	coordinates {
-		(1,0.596)(2,0.6)(3,0.578)(4,0.554)
+		(3,0.781)(5,0.795)(9,0.797)(17,0.827)
+	};
+
+	\addplot[
+	color=black,
+	mark=diamond,
+	]
+	coordinates {
+		(3,0.593)(5,0.59)(9,0.571)(17,0.56)
 	};
@@
 	coordinates {
-		(1,0.519)(2,0.535)(3,0.528)(4,0.511)
+		(3,0.775)(5,0.787)(9,0.795)(17,0.819)
 	};
@@
 	coordinates {
-		(1,0.781)(2,0.795)(3,0.797)(4,0.827)
+		(3,0.781)(5,0.797)(9,0.8)(17,0.831)
 	};
@@
 	coordinates {
-		(1,0.593)(2,0.59)(3,0.571)(4,0.56)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.775)(2,0.787)(3,0.795)(4,0.819)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.781)(2,0.797)(3,0.8)(4,0.831)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.774)(2,0.787)(3,0.796)(4,0.815)
+		(3,0.774)(5,0.787)(9,0.796)(17,0.815)
 	};
@@
 	%		\centering
-	\begin{tikzpicture}[thick,font=\scriptsize]
+	\begin{tikzpicture}[thick,font=\scriptsize,scale=0.95]
 	\begin{axis}[
@@
 	ymin=0,
-	xtick={1,2,...,23},
+	xtick={3,5,9,17},
 	xticklabels={},
-	extra x ticks={1,23},
+	extra x ticks={3,5,9,17},
 	legend pos=north west,
@@
 	\addplot[
+	color=red,
+	mark=diamond,
+	]
+	coordinates {
+		(3,0.516)(5,0.505)(9,0.515)(17,0.524)
+	};
+
+	\addplot[
+	color=brown,
+	mark=diamond,
+	]
+	coordinates {
+		(3,0.598)(5,0.601)(9,0.601)(17,0.619)
+	};
+
+	\addplot[
+	color=green,
+	mark=diamond,
+	]
+	coordinates {
+		(3,0.52)(5,0.535)(9,0.528)(17,0.512)
+	};
+
+	\addplot[
 	color=blue,
@@
 	coordinates {
-		(1,0.598)(2,0.601)(3,0.601)(4,0.619)
+		(3,0.796)(5,0.798)(9,0.798)(17,0.828)
+	};
+
+	\addplot[
+	color=black,
+	mark=diamond,
+	]
+	coordinates {
+		(3,0.596)(5,0.591)(9,0.589)(17,0.579)
 	};
@@
 	coordinates {
-		(1,0.52)(2,0.535)(3,0.528)(4,0.512)
+		(3,0.775)(5,0.787)(9,0.795)(17,0.819)
 	};
@@
 	coordinates {
-		(1,0.796)(2,0.798)(3,0.798)(4,0.828)
+		(3,0.782)(5,0.797)(9,0.8)(17,0.831)
 	};
@@
 	coordinates {
-		(1,0.596)(2,0.591)(3,0.589)(4,0.579)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.775)(2,0.787)(3,0.795)(4,0.819)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.782)(2,0.797)(3,0.8)(4,0.831)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.774)(2,0.787)(3,0.796)(4,0.815)
+		(3,0.774)(5,0.787)(3,0.796)(4,0.815)
 	};
@@
 	\addplot[
+	color=red,
+	mark=diamond,
+	]
+	coordinates {
+		(3,0.516)(5,0.505)(9,0.515)(17,0.524)
+	};
+
+	\addplot[
+	color=brown,
+	mark=diamond,
+	]
+	coordinates {
+		(3,0.598)(5,0.601)(9,0.579)(17,0.554)
+	};
+
+	\addplot[
+	color=green,
+	mark=diamond,
+	]
+	coordinates {
+		(3,0.52)(5,0.535)(9,0.528)(17,0.512)
+	};
+
+	\addplot[
 	color=blue,
@@
 	coordinates {
-		(1,0.598)(2,0.601)(3,0.579)(4,0.554)
+		(3,0.787)(5,0.794)(9,0.798)(17,0.828)
+	};
+
+	\addplot[
+	color=black,
+	mark=diamond,
+	]
+	coordinates {
+		(3,0.595)(5,0.591)(9,0.572)(17,0.561)
 	};
@@
 	coordinates {
-		(1,0.52)(2,0.535)(3,0.528)(4,0.512)
+		(3,0.775)(5,0.787)(9,0.796)(17,0.819)
 	};
@@
 	coordinates {
-		(1,0.787)(2,0.794)(3,0.798)(4,0.828)
+		(3,0.783)(5,0.797)(9,0.8)(17,0.831)
 	};
@@
 	coordinates {
-		(1,0.595)(2,0.591)(3,0.572)(4,0.561)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.775)(2,0.787)(3,0.796)(4,0.819)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.783)(2,0.797)(3,0.8)(4,0.831)
-	};
-
-	\addplot[
-	color=blue,
-	mark=diamond,
-	]
-	coordinates {
-		(1,0.774)(2,0.788)(3,0.796)(4,0.815)
+		(3,0.774)(5,0.788)(9,0.796)(17,0.815)
 	};
@@
 
-Finally, to predict the TTL of the results of the query -- outputting the number of versions for which the query results can be expected to remain the same -- we employed techniques of multiple linear regression using the \textsc{sklearn}.
+Finally, to predict the TTL of the results of the query -- outputting the number of versions for which the query results can be expected to remain the same -- we employed techniques of multiple linear regression using the \textsc{sklearn} library.
 The latter allows us to see how the periods of estimation and the convergence analysis will behave.
@@
 Concerning the prediction tasks themselves, there are a variety of other features and approaches that could be applied; one promising direction may be, for example, to apply a more fine-grained analysis of the query, considering for example estimates of the selectivity of particular triples patterns, and combining this information with statistics on the dynamics of the data and the query operators used; our current gold standard could be used to validate such developments.
-Finally, it would be interesting to investigate the effectiveness of these techniques in practice, developing caching systems, synchronization schedules, and other applications, based on the proposed predictions.
+Finally, it would be interesting to investigate the effectiveness of these techniques in practice, developing caching systems, synchronization schedules, and other applications, based on the evaluated predictions.
 

### 2019-04-11 kgswc19.tex
update

@@
 
-One topic we do wish to highlight however -- as it relates to the behavior of a query over a dynamic RDF graph -- is that of the \emph{monotonicity} of SPARQL queries~\cite{ArenasP11,ArenasU17}.
+One topic we do wish to highlight, however -- as it relates to the behavior of a query over a dynamic RDF graph -- is that of the \emph{monotonicity} of SPARQL queries~\cite{ArenasP11,ArenasU17}.
 We say that a \texttt{SELECT} query $Q$ is monotone if and only if $G_1 \subseteq G_2$ implies that $Q(G_1) \subseteq Q(G_2)$ for RDF graphs $G_1$ and $G_2$; intuitively, as data are extended, the results of a monotone query can only be extended.
@@
 
-The second block indicate the query operators used; we capture the presence or absence of query operators and solution modifiers from SPARQL 1.1.
+The second block indicates the query operators used; we capture the presence or absence of query operators and solution modifiers from SPARQL 1.1.
 In the third block, we group related features into one dimension: in the case of \textit{Recursive path}, we group queries that contain as expression of the form $e$\texttt{*} or $e$\texttt{+}.
@@
 We further note some peaks in triples added in some versions, which may be due to bulk imports of data.
-Between the version 11 and 12 we have almost 2 weeks because we were not able to obtain the data for that week, but this cause the third highest peak.
+Between version 11 and 12 we have almost 2 weeks because we were not able to obtain the data for that week, but this cause the third highest peak.
 The dynamic graph thus considers a total of 32.3 billion triples across 23 versions.
@@
 
-\paragraph{Setting:} To build our final datasets, given that our data features are temporal based, we must thus define a past interval to consider.
-With 23 versions, we must hold out at least one version to label, leaving a maximum interval of 22 past versions to use for compute our data features.
-However, the more past versions we consider for a prediction, the fewer examples we can generate; for example, if we consider 22 versions for compute, we will only have one ground truth label for each query.
+\paragraph{Setting:} To build our final datasets, given that our data features are temporally based, we must thus define a past interval to consider.
+With 23 versions, we must hold out at least one version to label, leaving a maximum interval of 22 past versions to use to compute our data features.
+However, the more past versions we consider for a prediction, the fewer examples we can generate; for example, if we consider 22 versions to compute, we will only have one ground truth label for each query.
 In the end, we opted to consider intervals of 3, 5, 9 and 17 previous versions, for example, ($G_{i-3}, G_{i-2}, G_{i-1}$) to predict changes in queries results for the subsequent version ($G_i$); this allows us to predict from $G_4$ to $G_{23}$ inclusive, providing 20 examples per query and 4420 examples in the largest dataset and 1326 examples in the smallest dataset.
@@
 
-For OSC prediction, we selected four classifiers for testing: Decision Trees, Linear SVM, Naive Bayes and Nearest Neighbors.
+For OSC prediction, we selected four classifiers for testing: Decision Trees, Linear SVM, Naive Bayes, and Nearest Neighbors.
 We split the data to use 80\% for training and 20\% for tests.
@@
 	coordinates {
-		(3,0.774)(5,0.787)(3,0.796)(4,0.815)
+		(3,0.774)(5,0.787)(9,0.796)(17,0.815)
 	};
@@
 Finally, to predict the TTL of the results of the query -- outputting the number of versions for which the query results can be expected to remain the same -- we employed techniques of multiple linear regression using the \textsc{sklearn} library.
-The latter allows us to see how the periods of estimation and the convergence analysis will behave.
+The latter allow us to see how the periods of estimation and the convergence analysis will behave.
 We were able to predict the TTL (in relation to the number of invariant weeks) with an average square error of 0.43 weeks and a variance score of 0.01.
-These results allow to predict future changes in query results with good accuracy using linear regression techniques.
+These results allow us to predict future changes in query results with good accuracy using linear regression techniques.
 
@@
 Analysing these data and queries, we conclude that they meet the desired characteristics, offering a large-scale, diverse, real-world setting; more importantly, we find that our queries offer an ideal distribution of queries whose results always change, queries whose results never change, and (more importantly) queries whose results change intermittently and for which prediction is thus a non-trivial task.
-We make this gold standard available at \ah{...} -- including queries, results and features -- so that the results presented here can be reproduced and so that it can be reused in further research.
+We make this gold standard available at \ah{...} -- including queries, results, and features -- so that the results presented here can be reproduced and so that it can be reused in further research.
 

### 2019-04-11 kgswc19.tex
update

@@
 	\resizebox{0.45\textwidth}{!}{
-		\begin{tabular}{rlrr}    \toprule
+		\begin{tabular}{r@{\hskip 0.1in}lr@{\hskip 0.1in}r}    \toprule
 			\ntextnumero        & Predicate & Total	& Dyn 	\\  \midrule
@@
 	\resizebox{0.5\textwidth}{!}{
-		\begin{tabular}{rlrr}    \toprule
+		\begin{tabular}{r@{\hskip 0.1in}lr@{\hskip 0.1in}r}    \toprule
 			\ntextnumero        & Predicate & Total	& Dyn 	\\  \midrule
@@
 		\caption{Distribution of triple patterns (C: Constant, V: Variable, P: Path). \label{tab:Pattern}}
-		\begin{tabular}{lrr}    \toprule
+		\begin{tabular}{l@{\hskip 0.2in}r@{\hskip 0.2in}r}    \toprule
 			\textbf{Pattern} & \textbf{\ntextnumero} & \textbf{\%} \\    \midrule
@@
 		\caption{Distribution of paths\label{tab:Path}}
-		\begin{tabular}{lrr}    \toprule
+		\begin{tabular}{l@{\hskip 0.2in}r@{\hskip 0.2in}r}    \toprule
 			\textbf{Pattern}      & \textbf{\ntextnumero} & \textbf{\%} \\    \midrule
@@
 
-Furthermore, we show in Figure~\ref{fig:result} that effectively the size of the window influences the quality of the results when considering \textsc{r}, being better as the window increases.
-However, these results are not affected when we consider only predicates in \textsc{q} or  \textsc{p}.
-Also, we show, that all configurations outperform the random baseline.
+Furthermore, we show in Figure~\ref{fig:result} that effectively the size of the window influences the quality of the results when considering \textsc{r} (blue), being better as the window increases.
+However, these results are not affected when we consider only predicates in \textsc{q} (brown), \textsc{p} (green)  or \textsc{qp} (black).
+Also, we show, that all configurations outperform the random baseline (red).
 

### 2019-04-11 kgswc19.tex

@@
 
+\makeatletter \renewcommand\subsubsection{\@startsection{subsubsection}{4}{\z@}%
+	{1ex \@plus0ex \@minus0ex}% 
+	{-1ex}% 
+	{\normalfont\normalsize\bfseries}} 
+\makeatother
+
 \usepackage{textcomp}
@@
 \hypersetup{
-    colorlinks, linkcolor={dark-red},
+    colorlinks, 
+    linkcolor={dark-red},
     citecolor={dark-green}, urlcolor={dark-blue},
@@
 %
-\title{Estimating the dynamics of SPARQL query results
+\title{Estimating the dynamics of \\SPARQL query results
 %\thanks{This work was supported by CONICYT PFCHA, Doctorado Becas Chile/2017 -- 72180000, the Millennium Institute for Foundational Research on Data (IMFD) and FONDECYT Grant No. 1181896.}
@@
 
-Some works have studied the data dynamics based on the historical evolution of entities~\cite{KaferAUOH13,DividinoGSG14,NishiokaS15}, but following such an approach for SPARQL queries is expensive because (1) a SPARQL query may involve potentially many entities; and (2) it is necessary to have the previous complete versions of data to analyse the entities relevant to a query.
+Some works study data dynamics based on the historical evolution of entities~\cite{KaferAUOH13,DividinoGSG14,NishiokaS15}, but following such an approach for SPARQL queries is expensive because (1) a SPARQL query may involve potentially many entities; and (2) it is necessary to have the previous complete versions of data to analyse the entities relevant to a query.
 Many works have explored the dynamics of Linked Data with the intention of finding patterns that allow for characterizing, recognizing, and predicting changes based on analysis of the domains, predicates, and schema~\cite{UmbrichHHPD10,KaferAUOH13,NishiokaS16,NishiokaS17,GonzalezH18}.
-Among these works, we can find hybrid approaches that have been developed to return fresher query results with faster response times by decomposing a query into dynamic sub-queries executed remotely, and static sub-queries executed over local caches~\cite{UmbrichKHP12}, but such an approach only focuses on estimating the dynamics of individual triple patterns, and does not consider, for example, query operators.
+Among these works are hybrid approaches developed to return fresher query results with faster response times by decomposing a query into dynamic sub-queries executed remotely, and static sub-queries executed over local caches~\cite{UmbrichKHP12}, but such an approach only focuses on estimating the dynamics of individual triple patterns, and does not consider, for example, query operators.
 More generally, coping with changes in remote data still presents a major challenge for applications leveraging dynamic Linked Data.
@@
 
-\paragraph{Paper structure:} In Section~\ref{sec:related}, we review related works.
+%\paragraph{Paper structure:} In Section~\ref{sec:related}, we review related works.
 Section~\ref{sec:preliminar} provides a formal statement of the problem.
@@
 \paragraph{Query Features:} Our initial set of features are based on analysis of the input SPARQL query.
-Figure~\ref{fig:query} shows only a sample of the evaluated features, as well as -- for the purposes of illustration -- their values for an accompanying example query.
+Figure~\ref{fig:query} shows only a sample of evaluated features, as well as -- for the purposes of illustration -- their values for an example query.
 
@@
 Although we have seen that the number of unique predicates did not change much over all versions, we are interested to see which predicates are indicative of dynamic statements.
-Table~\ref{tab:predicates} shows the ten most dynamic predicates according to the number of added and deleted statements involving that predicate (Total) and the ratio of added and deleted statements divided by the total number of statements for that predicate across all snapshots (Dyn); for the second, we only include predicates that appear in all snapshots and appear in ≥ 1, 000 statements overall.
+Table~\ref{tab:predicates} shows the ten most dynamic predicates according to the number of added and deleted statements involving that predicate (Total) and the ratio of added and deleted statements divided by the total number of statements for that predicate across all snapshots (Dyn); for the second, we only include predicates that appear in all snapshots and appear in $\geq$ 1,000 statements overall.
 Note that the largest number of changes involves the predicate scheme:description, however, its dynamic value is low due to the number of statements with it in the versions.
@@
 
-96 queries had to be removed from the original set of 389 because they asked for information external to Wikidata (using \texttt{SERVICE}) or asked for qualifiers that are not present in the truthy version.
+A total of 96 queries had to be removed from the original set of 389 because they asked for information external to Wikidata (using \texttt{SERVICE}) or asked for qualifiers that are not present in the truthy version.
 We also eliminated 12 queries that returned bindings to blank nodes to facilitate comparison between the results.
 Furthermore, some queries featured non-deterministic elements -- such as \texttt{LIMIT}/\texttt{OFFSET} without ordering, \texttt{SAMPLE} or temporal functions -- such as \texttt{TODAY()} -- that may lead to changes in results not related to changes in the data; we remove 23 queries with \texttt{SAMPLE} and temporal functions, and add an \texttt{ORDER BY} clause to all queries to ensure determinism with \texttt{LIMIT}/\texttt{OFFSET} and more generally to facilitate quick comparison of results.
-We also eliminated 37 queries with 0 results in the executions over all the snapshots.
+We also eliminated 37 queries with 0 results forr all the snapshots.
 As a result of this process of filtering, we end up with a total of 221 deterministic queries.
 
-Furthermore, we needed to transform certain queries for the purposes of our experiments.
+%Furthermore, we needed to transform certain queries for the purposes of our experiments.
 163 queries used features particular to the Wikidata Query Service, such as using a custom label \texttt{SERVICE} call to specify language preferences; we converted these queries to standard SPARQL using \texttt{COALESCE} to capture preferences.
 Figure~\ref{fig:transform} shows an example of a Wikidata query with the \texttt{LABEL} service rewritten and a deterministic ordering added.
-
-\begin{figure}[!t]
-\centering
-{
-	\begin{minipage}{0.8\textwidth}
-	\begin{lstlisting}[style=sparqld]
-#Cats
-SELECT ?item ?itemLabel
-WHERE {
- ?item wdt:P31 wd:Q146 .
- SERVICE wikibase:label { bd:serviceParam wikibase:language "en,es" }
-}
-	\end{lstlisting}
-	\end{minipage}	
-}
-
-{
-\begin{minipage}{0.8\textwidth}
-	\begin{lstlisting}[style=sparqld]
-#Cats
-SELECT ?item ?itemLabel
-WHERE {
-  ?item wdt:P31 wd:Q146 .
-  OPTIONAL { ?item rdfs:label ?en . FILTER(lang(?en)="en") }
-  OPTIONAL { ?item rdfs:label ?es . FILTER(lang(?es)="es") }
-  BIND(str(coalesce(?en, ?es, strafter(str(?item),
-    "http://www.wikidata.org/entity/"))) AS ?itemLabel )
-}
-ORDER BY ?item ?itemLabel
-	\end{lstlisting}
-\end{minipage}	
-}
-\caption{Query transformation before (above) and after (below) \label{fig:transform}}
-\end{figure}
+%
+%\begin{figure}[!t]
+%\centering
+%{
+%	\begin{minipage}{0.8\textwidth}
+%	\begin{lstlisting}[style=sparqld]
+%#Cats
+%SELECT ?item ?itemLabel
+%WHERE {
+% ?item wdt:P31 wd:Q146 .
+% SERVICE wikibase:label { bd:serviceParam wikibase:language "en,es" }
+%}
+%	\end{lstlisting}
+%	\end{minipage}	
+%}
+%
+%{
+%\begin{minipage}{0.8\textwidth}
+%	\begin{lstlisting}[style=sparqld]
+%#Cats
+%SELECT ?item ?itemLabel
+%WHERE {
+%  ?item wdt:P31 wd:Q146 .
+%  OPTIONAL { ?item rdfs:label ?en . FILTER(lang(?en)="en") }
+%  OPTIONAL { ?item rdfs:label ?es . FILTER(lang(?es)="es") }
+%  BIND(str(coalesce(?en, ?es, strafter(str(?item),
+%    "http://www.wikidata.org/entity/"))) AS ?itemLabel )
+%}
+%ORDER BY ?item ?itemLabel
+%	\end{lstlisting}
+%\end{minipage}	
+%}
+%\caption{Query transformation before (above) and after (below) \label{fig:transform}}
+%\end{figure}
 
@@
 
-\paragraph{Setting:} To build our final datasets, given that our data features are temporally based, we must thus define a past interval to consider.
+\paragraph{Setting:} To build our final datasets, we must define a past interval to consider.
 With 23 versions, we must hold out at least one version to label, leaving a maximum interval of 22 past versions to use to compute our data features.
-However, the more past versions we consider for a prediction, the fewer examples we can generate; for example, if we consider 22 versions to compute, we will only have one ground truth label for each query.
-In the end, we opted to consider intervals of 3, 5, 9 and 17 previous versions, for example, ($G_{i-3}, G_{i-2}, G_{i-1}$) to predict changes in queries results for the subsequent version ($G_i$); this allows us to predict from $G_4$ to $G_{23}$ inclusive, providing 20 examples per query and 4420 examples in the largest dataset and 1326 examples in the smallest dataset.
-We remark that no feature explicitly indicates which specific version numbers are given and which are to be predicted.
+However, the more past versions we consider for a prediction, the fewer examples we can generate; for example, if we consider 22 versions, we will only have one ground truth label for each query.
+In the end, we opted to consider intervals of 3, 5, 9 and 17 previous versions, for example, ($G_{i-3}, G_{i-2}, G_{i-1}$) to predict changes in queries results for the subsequent version ($G_i$); this allows us to predict from $G_4$ to $G_{23}$ inclusive, providing 20 examples per query, 4,420 examples in the largest dataset and 1,326 examples in the smallest dataset.
+No feature explicitly indicates which specific version numbers are given and which are to be predicted.
 
@@
 The target TTL value that we learn is the time interval during which the results of the query remain fresh from the previous update time.
-As an example, suppose that the query q change three times ($G_{3}, G_{7}, G_{9}$) over all versions, we consider two examples from $G_{3}$ and $G_{7}$ with target TTL value 4 and 2 respectively.
+As an example, suppose that the query $Q$ change three times ($G_{3}, G_{7}, G_{9}$) over all versions; we consider two examples from $G_{3}$ and $G_{7}$ with target TTL value 4 and 2 respectively.
 
 \paragraph{Results:} Table~\ref{tab:classifier} shows the $F_1$-score for the predictions made considering the union of different combinations of the sets of features (Q, P, R) and varying the window sizes (3, 5, 9, 17); for reference, we include a baseline that randomly guesses yes/no.
-We can note the best results were obtained using the features from \textsc{r} reaching values ​​of 0.83 of $F_1$.
+We can note the best results were obtained using the features from \textsc{r} reaching values of 0.83 $F_1$.
 The results show that predicting changes in query results is indeed a challenging problem.
-Looking in further detail, as can be expected, prediction performs best when the \textsc{r} feature - with knowledge of changes in the historical value of results of the input query - is available.
-And, the features from sets \textsc{q} and  \textsc{p} - which do not assume the availability of such information - obtain less $F_1$-score: more specifically, the best result was using \textsc{pr} (0.83) with Linear SVM and window size 17, while the best results without \textsc{r} was using \textsc{Q} (0.60) with Linear SVM and window size 5.
-
-Furthermore, we show in Figure~\ref{fig:result} that effectively the size of the window influences the quality of the results when considering \textsc{r} (blue), being better as the window increases.
+Looking in further detail, as can be expected, prediction performs best when the \textsc{r} feature -- with knowledge of changes in the historical value of results of the input query -- is available.
+The features from sets \textsc{q} and  \textsc{p} -- which do not assume the availability of such information -- obtain a lower $F_1$-score: more specifically, the best result was using \textsc{pr} (0.83) with Linear SVM and window size 17, while the best results without \textsc{r} was using \textsc{Q} (0.60) with Linear SVM and window size 5.
+
+Furthermore, we show in Figure~\ref{fig:result} that the size of the window influences the quality of the results when considering \textsc{r} (blue), being better as the window increases.
 However, these results are not affected when we consider only predicates in \textsc{q} (brown), \textsc{p} (green)  or \textsc{qp} (black).
@@
 		& Linear SVM               & \underline{0.596}    & 0.372          & 0.758          & \underline{0.593}    & \underline{0.775}    & 0.762          & \underline{0.774}    \\    \midrule
-		& \textbf{Better}          & \textbf{0.596} & \textbf{0.519} & \textbf{0.781} & \textbf{0.593} & \textbf{0.775} & \textbf{0.781} & \textbf{0.774} \\    \bottomrule
+		& \textbf{Better}          & \textbf{0.596} & \textbf{0.519} & \textbf{0.781} & \textbf{0.593} & \textbf{0.775} & \textbf{0.781} & \textbf{0.774} \\    \midrule
 	w=5	& \textit{Random Baseline} & 0.48           & 0.496          & 0.499          & 0.505          & 0.495          & 0.497          & 0.499          \\
@@
 		& Linear SVM               & \underline{0.6}      & 0.384          & 0.784          & \underline{0.59}     & \underline{0.787}    & 0.789          & \underline{0.787}    \\    \midrule
-		& \textbf{Better}          & \textbf{0.6}   & \textbf{0.535} & \textbf{0.795} & \textbf{0.59}  & \textbf{0.787} & \textbf{0.797} & \textbf{0.787} \\    \bottomrule
+		& \textbf{Better}          & \textbf{0.6}   & \textbf{0.535} & \textbf{0.795} & \textbf{0.59}  & \textbf{0.787} & \textbf{0.797} & \textbf{0.787} \\    \midrule
 	w=9	& \textit{Random Baseline} & 0.497          & 0.505          & 0.493          & 0.483          & 0.497          & 0.511          & 0.515          \\
@@
 		& Linear SVM   				& \underline{0.578}    & 0.397          & \underline{0.797}    & \underline{0.571}    & \underline{0.795}    & \underline{0.8}      & \underline{0.796}    \\     \midrule
-		& \textbf{Better}          & \textbf{0.578} & \textbf{0.528} & \textbf{0.797} & \textbf{0.571} & \textbf{0.795} & \textbf{0.8}   & \textbf{0.796} \\    \bottomrule
+		& \textbf{Better}          & \textbf{0.578} & \textbf{0.528} & \textbf{0.797} & \textbf{0.571} & \textbf{0.795} & \textbf{0.8}   & \textbf{0.796} \\    \midrule
 	w=17	& \textit{Random Baseline} & 0.486          & 0.494          & 0.482          & 0.512          & 0.524          & 0.511          & 0.49           \\
@@
 	width=5cm, 
-	height=5cm,
+	height=4cm,
 	ymin=0,
@@
 
-Our goal then was to explore this trade-off experimentally.
+We thus explore this trade-off experimentally.
 For this, we define some desirable characteristics for the queries and data used in the experimental setting, such as availability, size, dynamics, diversity, etc.
@@
 Analysing these data and queries, we conclude that they meet the desired characteristics, offering a large-scale, diverse, real-world setting; more importantly, we find that our queries offer an ideal distribution of queries whose results always change, queries whose results never change, and (more importantly) queries whose results change intermittently and for which prediction is thus a non-trivial task.
-We make this gold standard available at \ah{...} -- including queries, results, and features -- so that the results presented here can be reproduced and so that it can be reused in further research.
+We will make this gold standard available after double-blind review -- including queries, results, and features -- so that the results presented here can be reproduced and so that it can be reused in further research.
 

### 2019-05-29 rebuttal/rebuttal.txt
rebuttal

Given the scores as they stand, it looks like the paper is leaning towards a reject. While we disagree with the verdict, we are admittedly somewhat biased, and the reviewers' comments are very constructive. We hence take this opportunity to thank the reviewers for their time and thoughts, and to express some high level perspectives regarding the work and the paper.

# Reviewer 1:

The reviewer notes that the paper lacks technical depth, and that applying machine learning techniques seems straightforward. We accept this point of view but we saw/see the contribution as going beyond the solution proposed. We submitted the paper as we felt that the novelty was in the problem itself: establishing the conceptual framework for it, putting together the benchmark, identifying the trade-offs, giving initial results. We felt (perhaps optimistically) that what we had done required more thought and (conceptual/physical) effort than what would be needed to publish another algorithm for an incremental improvement on an existing problem with bibliographies, benchmarks, datasets, problem statements, etc., already available. It might even be considered in its own way as more novel: if we define novelty exclusively in terms of novel *algorithms*, we perhaps run the risk of disincentivising the study of new problems, rather incentivising incremental work that beats the state-of-the-art.

Maybe ISWC could use a track for work on new problems, or maybe the work might just be better suited to a workshop (and then we can work on new algorithms for a conference submission).

# Reviewer 2:

We very much appreciate the detailed comments. There are plenty of great suggestions for us to explore (though perhaps taken together they go beyond the scope of a conference paper?). Regarding the novel features related to the query, we were confident these would be more useful than they were. Intuitively we figured that a query with more triple patterns and more variables had higher probability to "hit" something that changed; of course static query features would not suffice to beat features based on historical results, but they would be a lot cheaper to manage/compute than historical features. Unfortunately the query features did not turn out to be very useful. Correlation analysis surprisingly showed no relation between the size of the query (number of variables, number of triple patterns, etc.) and how dynamic its results were. We still don't know why this is, but possibly it's because adding more triple patterns might actually help focus the query (e.g., { (?s a :Person) } will be more dynamic than { (?s a :Person), (?s :name "Bill Gates") }); in any case, our initial hypothesis was mostly rejected. As the deadline approached, the paper was a bit stuck between what we wrote before the results, and what we wrote after them. A more appropriate narrative would have been a negative result (but these can be hard to sell).

On a side note, we of course plan to make it up to the authors of [11] for using their data without full credit (though really we just used their dumps; most work was in the queries, deltas, feature extraction, etc.).

# Reviewer 3:

We appreciate that the reviewer recognises the (months of) work that went into the benchmark; we were very happy after all that work when we saw that the queries happened to be almost perfectly balanced for the static/dynamic classes. The reviewer is correct that the model is probably not very generalisable, at least with predicate features; indeed the model will have to be trained for a specific dataset (though maybe transfer learning could be applied; or alternatively a more generalisable model might be just to use the 1/0 feature vector of past changes). Regarding the cost of annotation, we disagree (or maybe misunderstand) the part that states "your approach will be as expensive": static query features and predicate features are much cheaper to work with. (Maybe we misunderstood.) We accept the point about the straightforward approach (as mentioned by reviewer 1).

# Reviewer 4:

We had not considered the resource track; maybe in retrospect that might have been a better option. We acknowledge that more queries and datasets would be better, though the ones we have already required a lot of work and we felt they would be sufficient to establish initial conclusions for the problem (then again, maybe our query features will work better for other datasets). As before, we accept the reviewer's opinion that the technical contribution is not that high; actually we agree a lot with (W2). We promise that we had planned on sharing the code after double-blind review and accept the point about supplementary material.

### 2019-07-06 main.tex
Renaming file for clarity

--- 2019-04-11 kgswc19.tex
@@
 \usepackage{hyperref}
-\definecolor{dark-blue}{rgb}{0.0,0.0,0.2}
-\definecolor{dark-green}{rgb}{0.0,0.2,0.0}
-\definecolor{dark-red}{rgb}{0.2,0.0,0.0}
 \hypersetup{
     colorlinks, 
-    linkcolor={dark-red},
-    citecolor={dark-green}, urlcolor={dark-blue},
+    linkcolor={black},
+    citecolor={black}, urlcolor={black},
     pdftitle={Estimating the dynamics of SPARQL query results},    % title
-%    pdfauthor={Alberto Moya, Aidan Hogan},     % author
+    pdfauthor={Alberto Moya, Aidan Hogan},     % author
     pdfsubject={ISWC 2019},   % subject of the document
@@
 %
-\title{Estimating the dynamics of \\SPARQL query results
+\title{Estimating the dynamics of SPARQL query results using binary classification
 %\thanks{This work was supported by CONICYT PFCHA, Doctorado Becas Chile/2017 -- 72180000, the Millennium Institute for Foundational Research on Data (IMFD) and FONDECYT Grant No. 1181896.}
@@
 %
-%\titlerunning{Abbreviated paper title}
+\titlerunning{Estimating the dynamics of SPARQL query results}
 % If the paper title is too long for the running head, you can set
@@
 %
-%\author{Alberto Moya\orcidID{0000-0002-7003-5087} \and 
-%Aidan Hogan\orcidID{0000-0001-9482-1982}}
+\author{Alberto Moya \and Aidan Hogan}
 %
@@
 %
-%\institute{IMFD; DCC, University of Chile\\
-%\email{\{amoya,ahogan\}@dcc.uhile.cl}}
+\institute{IMFD; DCC, University of Chile\\
+\email{\{amoya,ahogan\}@dcc.uhile.cl}}
 %
@@
 \begin{abstract}
-We address the problem of estimating when the results of an input SPARQL query results over dynamic RDF datasets will change.
-We evaluate a framework that extracts features from the query and/or from past versions of the target dataset.
-The resulting features are then used to predict (1) whether or not the results for a query will change in the near future using classifiers, and (2) when they might change (if at all) using linear regression.
-To evaluate, we create a gold standard based on 23 versions of Wikidata and a curated collection of 221 SPARQL queries.
-Our results show that the quality of predictions using query features and lightweight statistics of the dynamics of individual predicates is less competitive with that obtained using (more costly to derive) knowledge of the complete historical changes in the query results.
+We address the problem of estimating when the results of an input SPARQL query over dynamic RDF datasets will change.
+We evaluate a framework that extracts features from the query and/or from past versions of the target dataset and inputs them into binary classifiers to predict whether or not the results for a query will change at a fixed point in the near future.
+For this evaluation, we create a gold standard based on 23 versions of Wikidata and a curated collection of 221 SPARQL queries.
+Our results show that the quality of predictions possible using (only) features based on the query structure and lightweight statistics of the predicate dynamics -- though capable of beating a random baseline -- are not competitive with results obtained using (more costly to derive) knowledge of the complete historical changes in the query results.
 
@@
 As a result, without further information as to the dynamics of a particular SPARQL query, applications face the choice of either performing frequent query executions that may be redundant and repeatedly return the same results (aiming for stronger consistency at the cost of more frequent updates), or performing infrequent query executions that may lead to stale data being persisted in the application when the underlying sources change (accepting weaker consistency to improve efficiency/scalability).
-Given the costs on clients and servers of repetitive requests served over the Web  and the potential efficiency gains offered by local caches, several weak consistency approaches have been proposed that try to keep the local data of the applications updated at lower cost by predicting changes~\cite{KnuthHS16,DividinoGS15,NishiokaS17}.
+Given the costs for clients and servers of repetitive requests served over the Web  and the potential efficiency gains offered by local caches, several weak consistency approaches have been proposed that try to keep the local data of the applications updated at lower cost by predicting changes~\cite{KnuthHS16,DividinoGS15,NishiokaS17}.
 
@@
 
-In this work, we address the problem of estimating if and/or when SPARQL query results will change in a future version of a dynamic RDF dataset.
-We evaluate a general architecture based on Machine Learning that extracts static features from a query, as well as dynamic features from the query and dataset.
-These features are fed into a classifier to predict whether or not the query results will change in a fixed point in the future, or, more ambitiously, can be fed into a regression model to predict when the query results are likely to change.
-With respect to the features used, we show that there is a trade-off between those that are coarse-grained but easy to compute (e.g., static query features) versus those that are fine-grained but more costly to compute (e.g., historical changes in query results), with further features falling in the middle (e.g., high-level statistics about the dynamics of triples with particular predicates).
+Given a SPARQL query and a dynamic RDF graph (consisting of multiple historical versions), in this work, we address the problem of predicting whether or not the query's results will change in the next version of the RDF graph.
+To the best of our knowledge, this is the first work to address this problem.
+Along these lines, we evaluate a general architecture based on Machine Learning that extracts static features from a query, as well as dynamic features from the query and dataset.
+These features are fed into a binary classifier to predict whether or not the query results will change in a fixed point in the future, or, more ambitiously, can be fed into a regression model to predict when the query results are likely to change.
+With respect to the features used, we show that there is a trade-off between those that are easy to compute but offer less accurate predictions (e.g., static query features) versus those that are more costly to compute but offer more accurate predictions (e.g., historical changes in query results).
 Per this trade-off, which features to use for predicting the dynamics of query results may then depend on the particular application.
@@
 In order to better understand this conceptual trade-off -- and more generally, to evaluate the quality of predictions made by our framework -- we create a novel gold standard based on 23 weekly versions of the Wikidata knowledge graph~\cite{VrandecicK14} and 221 user-generated queries; we show this gold standard to have a variety of desirable features, including (most importantly) a balance of queries whose results never change, always change, as well as non-trivial queries whose results intermittently change.
-Using this gold standard, our experiments show that while the most accurate predictions are based on historical changes to the input query's results, predictions based only on query features and predicate dynamics remain less competitive (while being much more efficient to maintain/extract).
+Using this gold standard, our experiments show that although features based on static characteristics of the query and statistics of changes in the data for individual predicates are more efficient to compute and maintain, they do not offer the same prediction quality as features based on knowledge of historical changes of the input query's results.
 
@@
 A variety of works have addressed the issue of modeling and consuming dynamic Linked Data from a broad range of perspectives~\cite{PassantM10,TrampFEA10,UmbrichHHPD10,UmbrichKL10,KaferAUOH13,DividinoSGG13,DividinoKG14,Kjernsmo15,NishiokaS16,GonzalezH18}.
-One of the major challenges considered is that of keeping cached copies of remote dynamic data -- cached for reasons of efficiency and scalability -- up-to-date on the consumer side, which we refer to as the synchronization problem.
+One of the major challenges considered is that of keeping cached copies of remote dynamic data -- cached for reasons of efficiency and scalability -- up-to-date on the consumer side, which we refer to as the \textit{synchronization problem}.
 
@@
 
-On the other hand, a variety of works have looked at building models of remote data that can help to predict which data are most dynamic, and which are most static, indicating which subsets of the data may need be refreshed from the remote source more often~\cite{PassantM10,TrampFEA10,UmbrichHHPD10,UmbrichKL10,KaferAUOH13,DividinoSGG13,NishiokaS16,GonzalezH18}; such works consider changes in RDF datasets at differing levels of granularity, including documents~\cite{KaferAUOH13}, domains~\cite{KaferAUOH13,NishiokaS16,NishiokaS17}, predicates~\cite{KaferAUOH13,NishiokaS17}, characteristic sets\footnote{A characteristic set is the set of predicate terms used to describe a given subject~\cite{NeumannM11}.}~\cite{NishiokaS16,GonzalezH18}, etc.
+Conversely, a variety of works have looked at building models of remote data that can help to predict which data are most dynamic, and which are most static, indicating which subsets of the data may need be refreshed from the remote source more often~\cite{PassantM10,TrampFEA10,UmbrichHHPD10,UmbrichKL10,KaferAUOH13,DividinoSGG13,NishiokaS16,GonzalezH18}; such works consider changes in RDF datasets at differing levels of granularity, including documents~\cite{KaferAUOH13}, domains~\cite{KaferAUOH13,NishiokaS16,NishiokaS17}, predicates~\cite{KaferAUOH13,NishiokaS17}, characteristic sets\footnote{A characteristic set is the set of predicate terms used to describe a given subject~\cite{NeumannM11}.}~\cite{NishiokaS16,GonzalezH18}, etc.
 Features at different levels of granularity can be fed into different predictive models based on Poisson Processes~\cite{UmbrichHHPD10}, Markov Chains~\cite{UmbrichMP15}, Empirical Distributions~\cite{NeumaierU16}, Machine Learning classification and regression~\cite{NishiokaS17,GonzalezH18}, as well as a variety of other heuristics~\cite{AliciAOCU12,UmbrichMP15,KnuthHS16} and metrics~\cite{DividinoGS15,KnuthHS16,AkhtarAL17}.
 Such approaches obviate the need for a subscription/notification mechanism.
-On the other hand, to ensure strong consistency in the presence of highly dynamic data, consumers may need to conservatively send a great many refresh requests to the server, which may be even more costly than a subscription/notification mechanism; hence such approaches are better suited for scenarios where weak consistency is more acceptable.
+However, to ensure strong consistency in the presence of highly dynamic data, consumers may need to conservatively send a great many refresh requests to the server, which may be even more costly than a subscription/notification mechanism; hence such approaches are better suited for scenarios where weak consistency is more acceptable.
 
@@
 Specifically regarding the dynamics of SPARQL query results, Passant and Mendes~\cite{PassantM10} proposed \textsc{sparqlPuSH} as a notification framework based  on \textsc{PubSubHubBub} (recently standardized as \textsc{WebSub}~\cite{webSub18}) aiming for strong consistency.
-On the other hand, aiming for weak consistency, Umbrich et al.~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12} proposed various methods to obtain knowledge about dynamics for different query patterns, mainly based on predicates.
+Rather aiming for weak consistency, Umbrich et al.~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12} proposed various methods to obtain knowledge about dynamics for different query patterns, mainly based on predicates.
 Dehghanzadeh et al.~\cite{DehghanzadehPKUHD14} later proposed a method to estimate the freshness using cardinality estimation techniques based on predicates and characteristic sets.
@@
 
-\paragraph{Novelty:} Given a query and a dynamic RDF dataset, we aim to predict (1) whether or not the query's results will have changed in a fixed point in the near future and/or (2) when the results will change.
-Our work thus complements existing works aiming for weak consistency, but (i) generalizes the problem, evaluating a framework that can incorporate statistics on predicate dynamics~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12,DehghanzadehPKUHD14} and historical changes in query results~\cite{KnuthHS16}, as proposed in prior works, as well as novel types of features, (ii) introduces new features based on query operators and statistics; (iii) creates a novel gold standard based on Wikidata and presents comparative results that indicate the relative predictive power inherent in different types of features.
+\paragraph{Novelty:} Given a query and a dynamic RDF dataset, we aim to predict whether or not the query's results will have changed in a fixed point in the near future.
+Our work thus complements existing works aiming for weak consistency, but (i) generalizes the problem, evaluating a framework that can incorporate statistics on predicate dynamics~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12,DehghanzadehPKUHD14} and historical changes in query results~\cite{KnuthHS16} (as proposed in prior works for addressing related but yet distinct problems relating to dynamic data), as well as novel types of features, (ii) introduces new features based on query operators and statistics; (iii) creates a novel gold standard based on Wikidata and presents comparative results that indicate the relative predictive power inherent in different types of features.
 
@@
 
-We consider a dynamic RDF graph to be a sequence of $n$ versions of an RDF graph denoted $\mathcal{G} \da ( G_1, \ldots, G_n )$; in practice, we assume these versions to have regular intervals (e.g., hourly, daily, weekly, etc.).
+We consider a \textit{dynamic RDF graph} to be a sequence of $n$ discrete versions of an RDF graph denoted $\mathcal{G} \da ( G_1, \ldots, G_n )$; in practice, we assume these versions to have regular intervals (e.g., hourly, daily, weekly, etc.).
 Further given a SPARQL (1.1) \texttt{SELECT} query $Q$, we would like to have knowledge of the dynamics of its results.
-The first such problem we consider -- which we call One Shot Change (OSC) -- accepts $\mathcal{G}$, $Q$ and a positive integer $k$ as input and outputs a boolean value predicting whether or not the results will change from $G_n$ to some future version $G_{n+k}$ (i.e., $Q(G_n) \neq Q(G_{n+k})$?).
-The second (more difficult) problem -- which we call Time-To-Live (TTL) -- accepts only $\mathcal{G}$ and $Q$ and outputs a positive integer $k$ as a prediction for the lowest such value where $Q(G_n) \neq Q(G_{n+k})$; in the case of bag semantics, we define ``$\neq$'' in terms of bag inequality, meaning that two bags of query results are different if the multiplicity of any result differs.
-Our problem is thus concerned with predicting if/when the results of a query will change, rather than predicting how  -- or to what extent -- they will change.
+The specific problem we consider in this work -- which we call One Shot Change (OSC) -- accepts $\mathcal{G}$, $Q$ and a positive integer $k$ as input and outputs a boolean value predicting whether or not the results will change from $G_n$ to some future version $G_{n+k}$ (i.e., $Q(G_n) \neq Q(G_{n+k})$?).
+In the case of bag semantics, we define ``$\neq$'' in terms of bag inequality, meaning that two bags of query results are different if the multiplicity of any result differs.
+We thus currently focus on predicting if the results of a query will change, rather than predicting when, or to what extent, they will change; these latter problems are left for future work.
 
@@
 
-\paragraph{Architecture:} In Figure~\ref{fig:schema}, we provide an overview of a general architecture for making predictions with respect to the OSC and TTL tasks.
+\paragraph{Architecture:} In Figure~\ref{fig:schema}, we provide an overview of a general architecture for making predictions with respect to the OSC prediction task.
 A SPARQL query $Q$ and a dynamic RDF graph $\mathcal{G}$ are given as input.
-The first component parses the query and generates static features which include statistics about the number of elements in the query (e.g., number of triple patterns, variables, etc.) as well as the presence of particular query operators (e.g., \texttt{UNION}, \texttt{FILTER}, recursive path expressions, etc.).
-The second component may take as input some features of the query, as well as information from the dynamic graph, and produce a set of further features; this component may compute, for example, statistics about the dynamicity of predicates used in the query, or statistics about changes in the query results over past versions.
-The final features are passed to a classifier in the case of the OSC problem, and to a regression model in the case of the TTL problem, to generate the final output prediction.
-We will now describe in more detail the query and data features considered in this work.
+The \textsf{Query Feature Extractor} extracts static features from the query, which include statistics about the number of elements in the query (e.g., number of triple patterns, variables, etc.) as well as the presence of particular query operators (e.g., \texttt{UNION}, \texttt{FILTER}, recursive path expressions, etc.).
+The \textsf{Predicate Feature Extractor} takes as input the predicates from the query and statistics about those predicates from the dynamic graph to produce a set of aggregated numerical features about the dynamicity of predicates used in the query.
+The \textsf{Result Feature Extractor} takes as input the results of the query over past versions of the dynamic graph from which it produces further features.
+All such features are passed to a (pre-trained) binary classifier to generate the final OSC prediction.
+We will now describe in more detail the query, data and result features considered in this work.
 
@@
 		font=\footnotesize,
-		minimum width=1.3em,
+		minimum width=2.1em,
 		minimum height=1.2em
@@
 
-\node[c,right=\hgap of q] (qfe) {\tb{Query \\ Feature \\ Extractor}}
+\node[c,right=1.7\hgap of q] (qfe) {\tb{Query \\ Feature \\ Extractor}}
   edge[arrin] (q);
 
-\node[f,right=1.5\hgap of qfe] (f4) {\tiny ...};
-\node[f,anchor=south] (f3) at (f4.north) {\tiny ...};
+\node[f,right=1.5\hgap of qfe,anchor=north] (f3) {\tiny ...};
 \node[f,anchor=south] (f2) at (f3.north) {\tiny ...};
 \node[f,anchor=south] (f1) at (f2.north) {$f_1$};
-\node[f,anchor=north] (f5) at (f4.south) {\tiny ...};
-\node[f,anchor=north] (f6) at (f5.south) {\tiny ...};
-\node[f,anchor=north] (fn) at (f6.south) {$f_n$};
+\node[f,anchor=north] (f4) at (f3.south) {$f_i$};
+%\node[f,anchor=north] (f5) at (f4.south) {$f_i$};
 
@@
 \draw [arrout] (qfe.east) -- + (0.3,0) |- (f4);
-\draw [arrout] (qfe.east) -- + (0.3,0) |- (f5);
-\draw [arrout] (qfe.east) -- + (0.3,0) |- (f6);
-\draw [arrout] (qfe.east) -- + (0.3,0) |- (fn);
-
-\node[c,right=\hgap of fn] (dfe) {\tb{Data\\Feature\\Extractor}}
-  edge[arrin] (fn);
-
-\draw[arrout] (f6.east) -- (dfe.west|-f6.east);
-
-\node[i,anchor=south] (g) at (q.mid|-dfe.south) {$\mathcal{G}$}
+
+\node[f,anchor=north] (f5) at (f4.south) {$f_{i+1}$};
+\node[f,anchor=north] (f6) at (f5.south) {\tiny ...};
+\node[f,anchor=north] (f7) at (f6.south) {\tiny ...};
+\node[f,anchor=north] (f8) at (f7.south) {$f_j$};
+
+\node[c] (dfe) at (qfe|-f6.south) {\tb{Predicate\\Feature\\Extractor}};
+
+\draw [arrout] (q.south) |- (dfe.west |-f5.south);
+
+\node[i,anchor=mid,left=\hgap of dfe] (g) {$\mathcal{G}$}
   edge[arrout] (dfe.west|-g.east);
 
-\node[f,right=\hgap of dfe] (dfk) {$f'_k$}
-  edge[arrin] (dfe);    
-\node[f,anchor=south] (df6) at (dfk.north) {\tiny ...}
-  edge[arrin] (dfe.east|-df6);  
-\node[f,anchor=south] (df5) at (df6.north) {$f_j$}
-  edge[arrin] (f5);    
-\node[f,anchor=south] (df4) at (df5.north) {\tiny ...}
-  edge[arrin] (f4);  
-\node[f,anchor=south] (df3) at (df4.north) {\tiny ...}
-  edge[arrin] (f3);      
-\node[f,anchor=south] (df2) at (df3.north) {\tiny ...}
-  edge[arrin] (f2);   
-\node[f,anchor=south] (df1) at (df2.north) {$f_1$}
-  edge[arrin] (f1);
-
-\node[c,right=1.5\hgap of df2,minimum width=6.5em] (c) {Classification\hsp};
-
-\draw [arrout] (df1.east) -- + (0.3,0) |- (c);
-\draw [arrout] (df2.east) -- + (0.3,0) |- (c);
-\draw [arrout] (df3.east) -- + (0.3,0) |- (c);
-\draw [arrout] (df4.east) -- + (0.3,0) |- (c);
-\draw [arrout] (df5.east) -- + (0.3,0) |- (c);
-\draw [arrout] (df6.east) -- + (0.3,0) |- (c);
-\draw [arrout] (dfk.east) -- + (0.3,0) |- (c);
-
-\node[c,right=1.5\hgap of df6,minimum width=6.5em] (r) {Regression\hsp};
-
-\draw [arrout] (df1.east) -- + (0.3,0) |- (r);
-\draw [arrout] (df2.east) -- + (0.3,0) |- (r);
-\draw [arrout] (df3.east) -- + (0.3,0) |- (r);
-\draw [arrout] (df4.east) -- + (0.3,0) |- (r);
-\draw [arrout] (df5.east) -- + (0.3,0) |- (r);
-\draw [arrout] (df6.east) -- + (0.3,0) |- (r);
-\draw [arrout] (dfk.east) -- + (0.3,0) |- (r);
+\draw [arrout] (dfe.east) -- + (0.3,0) |- (f5);  
+\draw [arrout] (dfe.east) -- + (0.3,0) |- (f6);
+\draw [arrout] (dfe.east) -- + (0.3,0) |- (f7);
+\draw [arrout] (dfe.east) -- + (0.3,0) |- (f8);
+
+\node[f,anchor=north] (f9) at (f8.south) {$f_{j+1}$};
+\node[f,anchor=north] (f10) at (f9.south) {\tiny ...};
+\node[f,anchor=north] (f11) at (f10.south) {\tiny ...};
+\node[f,anchor=north] (f12) at (f11.south) {$f_k$};
+
+\node[c] (rfe) at (qfe|-f10.south) {\tb{Result\\Feature\\Extractor}};
+
+%\draw [arrout] (g.south) |- (rfe);  
+%\draw [arrout] (q.south) |- (rfe);
+
+\draw [arrout] (rfe.east) -- + (0.3,0) |- (f9);  
+\draw [arrout] (rfe.east) -- + (0.3,0) |- (f10);
+\draw [arrout] (rfe.east) -- + (0.3,0) |- (f11);
+\draw [arrout] (rfe.east) -- + (0.3,0) |- (f12);
+
+\node[i,anchor=mid,left=\hgap of rfe] (qg) {$\mathcal{Q}(\mathcal{G})$}
+edge[arrout] (rfe.west|-qg.east);
+
+\draw [arrout] (q.south) -- (q.south |-qg.north);
+\draw [arrout] (g.south) -- (g.south |-qg.north);
+
+\node[c,right=1.5\hgap of f6.south,minimum width=6.5em] (c) {Classification\hsp};
+
+\draw [arrout] (f1.east) -- + (0.3,0) |- (c);
+\draw [arrout] (f2.east) -- + (0.3,0) |- (c);
+\draw [arrout] (f3.east) -- + (0.3,0) |- (c);
+\draw [arrout] (f4.east) -- + (0.3,0) |- (c);
+\draw [arrout] (f5.east) -- + (0.3,0) |- (c);
+\draw [arrout] (f6.east) -- + (0.3,0) |- (c);
+\draw [arrout] (f7.east) -- + (0.3,0) |- (c);
+\draw [arrout] (f8.east) -- + (0.3,0) |- (c);
+\draw [arrout] (f9.east) -- + (0.3,0) |- (c);
+\draw [arrout] (f10.east) -- + (0.3,0) |- (c);
+\draw [arrout] (f11.east) -- + (0.3,0) |- (c);
+\draw [arrout] (f12.east) -- + (0.3,0) |- (c);
 
@@
 
-\node[o,right=\hgap of r] (ttl) {\tb{TTL \\ Prediction}}
-  edge[arrin] (r);
-
 \end{tikzpicture}
@@
 
-\paragraph{Query Features:} Our initial set of features are based on analysis of the input SPARQL query.
+\paragraph{Query Features:} Our initial set of features is based on static analysis of the input SPARQL query.
 Figure~\ref{fig:query} shows only a sample of evaluated features, as well as -- for the purposes of illustration -- their values for an example query.
 
-The first block of features capture statistics about the query, indicating loosely its size and complexity.
-One may consider that the higher these values are, in general, the more dynamic we can expect the query to be since there are more ``opportunities'' for the query results to be affected by change.
-In some cases, however, the hypothesized correlation is not direct since, for example, adding more triple patterns may serve to narrow the query down and focus it on a static part of the graph (e.g., when looking for the movies of directors, adding a triple pattern to restrict the results to directors who have died).
+The first group of features captures statistics about the query, indicating loosely its size and complexity.
+One might consider that the higher these values are, in general, the more dynamic we can expect the query to be since there are more ``opportunities'' for the query results to be affected by change.
+In some cases, however, the hypothesized correlation is not direct since, for example, adding more triple patterns may serve to narrow the query down and focus it on a static part of the graph (e.g., when looking for the movies of directors, adding a triple pattern to restrict the results to directors who have died may reduce the likelihood of changes in the results in a future version).
 Hence it will be of interest to see how these features affect the predictions made experimentally.
 
-The second block indicates the query operators used; we capture the presence or absence of query operators and solution modifiers from SPARQL 1.1.
-In the third block, we group related features into one dimension: in the case of \textit{Recursive path}, we group queries that contain as expression of the form $e$\texttt{*} or $e$\texttt{+}.
+The second group indicates the query operators used; we capture the presence or absence of query operators and solution modifiers from SPARQL 1.1.
+
+In the third group gathers related features into one dimension: in the case of \textit{Recursive path}, we group queries with path expressions of the form $e$\texttt{*} or $e$\texttt{+}.
 On the other hand, in \textit{Negation}, we group non-monotonic features that allow for modeling difference (\texttt{MINUS}, \texttt{NOT EXISTS}, \texttt{!BOUND}).
 These features -- though course-grained -- are straightforward to extract from a query, and offer valuable insights into how the query may behave in a dynamic query; for example, the \textit{Negation} feature captures information about the (non-)monotonicity of the query, while we suppose that \textit{Recursive paths}, which may traverse an arbitrary number of triples in the graph, might be more sensitive to change.
-Again, such correlations are not without exception and will require experimental scrutiny.
-
-The final feature is the set of all predicates used in a triple pattern, including (where applicable) the IRIs used in a path expression; this is rather a meta-feature that serves as input for the next part of the proposed model.
+Again, such correlations are not without exception and will require empirical study.
 
@@
 
-\paragraph{Data Features (predicate dynamics):} The next component in our architecture can extract features that capture information about the query \textit{and} the dynamic graph under consideration.
-The first such feature we consider captures how many triples change for a predicate in a time interval, with the idea that -- as in previous works~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12,DehghanzadehPKUHD14} -- predicates capture rich information about the dynamics of a dataset, where the results of queries with dynamic predicates will be more sensitive to change; furthermore, given that in most datasets the number of predicates is relatively low (in the thousands or tens of thousand), such statistics are quite lightweight to maintain and compute.
+\paragraph{Predicate Features:} The next component extracts features that capture information about the dynamics of predicates used in the query (without \textit{evaluating} the query on the dynamic graph).
+This component captures how often and how many triples change for a predicate in a time interval, with the idea that -- as in previous works~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12,DehghanzadehPKUHD14} -- predicates capture rich information about the dynamics of a dataset, where the results of queries with dynamic predicates will be more sensitive to change.
+Furthermore, given that the number of predicates is often \textit{relatively} low (in the thousands or tens of thousand) for most datasets, such statistics are quite lightweight to maintain and compute.
 Formally, given two RDF graphs $G_i$ and $G_j$, we denote by $G_i \oplus G_j$ the set of triples $(G_i \cup G_j) - (G_i \cap G_j)$ where ``$-$'' denotes set difference; in other words, $G_i \oplus G_j$ denotes the triples in one graph or the other but not both (noting that $G_i \oplus G_j = G_j \oplus G_i$).
@@
  :children 1 , 2 ;
- :played :2014WC .
+ :played :2018WC .
 	\end{lstlisting}
@@
 Consider the example dynamic RDF graph $\mathcal{G}$ in Figure~\ref{fig:dg} (based on real data from Wikidata, with IRIs modified for the purposes of readability).
-In Figure~\ref{fig:dgc} we show the pairwise changes between each version: $G_1 \oplus G_2$ and $G_3 \oplus G_4$.
+In Figure~\ref{fig:dgc} we show the pairwise changes between each version: $G_1 \oplus G_2$ and $G_2 \oplus G_3$.
 For a predicate $p$, the value $\Delta(\mathcal{G},p)$ is then the sum of triples with predicate $p$ in the graphs $G_1 \oplus G_2$ and $G_2 \oplus G_3$ divided by the number of triples with predicate $p$ in the graphs $G_1 \cup G_2$ and $G_2 \cup G_3$.
@@
 
-\paragraph{Data Features (result dynamics):} The second data feature we capture indicates how many times the query results $Q$ have changed over the past versions of the dynamic graph.
-While this feature offers rich information for prediction, given a query that has not previously been seen, it is costly to compute, since it requires the execution of the query over each past version within an interval, which in turn requires maintaining indexes over a variety of past versions.
+\paragraph{Result features:} The third type of feature we capture indicates how many times the query results $Q$ have changed over the past versions of the dynamic graph.
+While this feature offers rich information for prediction, given a query that has not previously been seen, it is costly to compute, since it requires the evaluation of the query over each past version within an interval, which in turn requires maintaining indexes over the full data for a variety of past versions.
 
@@
 
-In order to evaluate the effectiveness for predicting changes in SPARQL query results, we require a dynamic RDF graph, with access to various historical versions; preferably this graph contains real-world, large-scale, diverse data, and with sufficient changes between versions to provide both positive and negative examples of queries whose results change.
+In order to evaluate the effectiveness for predicting changes in SPARQL query results, we require a dynamic RDF graph, with access to various historical versions; preferably this graph contains real-world, large-scale, diverse RDF data, and with sufficient changes between versions to provide both positive and negative examples of queries whose results change.
 Furthermore, we require a set of SPARQL queries that can be answered against this dynamic graph; preferably these queries again should be diverse, representative of real-world user queries, of a variety of shapes and sizes, using a variety of query operators, and with a mix of both dynamic and static results over the dynamic graph.
@@
 
-\paragraph{RDF Data:} We use 23 Wikidata snapshots from 18/04/2017 to 27/09/2017, which are captured almost weekly($\pm$ 1 day) in the \emph{truthy version} that contains triples (without qualifiers) that have the best non-deprecated rank for a given property.
+\paragraph{RDF Data:} We use 23 Wikidata snapshots from 18/04/2017 to 27/09/2017, which are captured (roughly) weekly in the \emph{truthy version} that contains triples (without qualifiers) that have the best non-deprecated rank for a given property; this data corpus was previously collected and used by Gonzalez and Hogan~\cite{GonzalezH18}.
 The first version has 1,102,242,331 triples and 3,276 unique predicates, while the final version has 1,924,967,162 triples (+74\%) and 3,660 unique predicates (+11\%).
@@
 We further note some peaks in triples added in some versions, which may be due to bulk imports of data.
-Between version 11 and 12 we have almost 2 weeks because we were not able to obtain the data for that week, but this cause the third highest peak.
-The dynamic graph thus considers a total of 32.3 billion triples across 23 versions.
+Between version 11 and 12 we have an almost 2 week gap because we were not able to obtain the data for that week, but this causes only the third highest peak.
+The dynamic graph considers a total of 32.3 billion triples across 23 versions.
 
@@
 
-Although we have seen that the number of unique predicates did not change much over all versions, we are interested to see which predicates are indicative of dynamic statements.
-Table~\ref{tab:predicates} shows the ten most dynamic predicates according to the number of added and deleted statements involving that predicate (Total) and the ratio of added and deleted statements divided by the total number of statements for that predicate across all snapshots (Dyn); for the second, we only include predicates that appear in all snapshots and appear in $\geq$ 1,000 statements overall.
-Note that the largest number of changes involves the predicate scheme:description, however, its dynamic value is low due to the number of statements with it in the versions.
-
-\begin{figure}[!t]
-	\caption{Top-10 dynamic predicates according with total (left) and proportional (right) changes}
+Table~\ref{tab:predicates} shows the ten most dynamic predicates according to the number of added and deleted statements involving that predicate (Total) and the ratio of added and deleted statements divided by the total number of statements for that predicate across all snapshots (Dyn); for the latter, we only include predicates that appear in all snapshots and appear in $\geq$ 1,000 statements overall.
+Though the largest number of changes involves the predicate \texttt{schema:description}, its dynamic value is low due to it being a common predicate.
+On the other hand, the OWL properties have high dynamicity ratios due to taking blank node values; we currently do not consider isomorphism of graphs due to blank nodes.
+
+\begin{table}[!t]
+	\caption{Top-10 dynamic predicates per total (left) and proportional (right) changes}
 	\label{tab:predicates}
@@
 			6 & skos:prefLabel  & 44487864  	& 0.01   \\
-			7 & wdt:P2093(author name)  & 30671685  	& 0.09   \\
+			7 & wdt:P2093 (author name)  & 30671685  	& 0.09   \\
 			8 & rdf:type  & 21808957  	& 0.02   \\
-			9 & wdt:P31(instance of)  & 12483471  	& 0.02   \\
+			9 & wdt:P31 (instance of)  & 12483471  	& 0.02   \\
 			10 & schema:about  & 10773295  	& 0.02  \\    \bottomrule
@@
 			3 & owl:someValuesFrom & 86674 & 0.71	\\
-			4 & wdt:P2462(member of the deme) & 3563 & 0.62	\\
-			5 & wdt:P3383(film poster) & 4388 & 0.58	\\
-			6 & wdt:P2331(Cycling Archives ID) & 4229 & 0.33	\\
-			7 & wdt:P1112() & 4366 & 0.20	\\
-			8 & wdt:P505(general manager) & 4959 & 0.18	\\
+			4 & wdt:P2462 (member of the deme) & 3563 & 0.62	\\
+			5 & wdt:P3383 (film poster) & 4388 & 0.58	\\
+			6 & wdt:P2331 (Cycling Archives ID) & 4229 & 0.33	\\
+			7 & wdt:P1112 (\textit{deleted}) & 4366 & 0.20	\\
+			8 & wdt:P505 (general manager) & 4959 & 0.18	\\
 			9 & schema:dateModified & 99330382 & 0.14	\\
@@
 	}	
-\end{figure}
+\end{table}
 
@@
 
-\paragraph{Queries:} In order to achieve a set of SPARQL queries that are answerable over Wikidata and with which we could run experiments, we took the user-contributed example queries from the Wikidata Query Service\footnote{https://www.wikidata.org/wiki/Wikidata:SPARQL\_query\_service/queries/examples}, consisting of 389 SPARQL 1.1 \texttt{SELECT} queries of varying degrees of complexity, touching upon various domains of data, and with a mix of varying query operators.\footnote{Very recently, a large log of millions of SPARQL queries have been released for Wikidata~\cite{MalyshevKGGB18}; these queries were not available when conducting our experiments, but would be interesting to analyze as part of future work.}
+\paragraph{Queries:} The corpus collected by Gonzalez and Hogan~\cite{GonzalezH18} does not offer queries.
+In order to achieve a set of SPARQL queries that are answerable over Wikidata and with which we could run experiments, we took the user-contributed example queries from the Wikidata Query Service\footnote{\url{https://www.wikidata.org/wiki/Wikidata:SPARQL_query_service/queries/examples}}, consisting of 389 SPARQL 1.1 \texttt{SELECT} queries of varying degrees of complexity, touching upon various domains of data, and with a mix of varying query operators.
+%\footnote{Very recently, a large log of millions of SPARQL queries have been released for Wikidata~\cite{MalyshevKGGB18}; these queries were not available when conducting our experiments, but would be interesting to analyze as part of future work.}
 
@@
 Furthermore, some queries featured non-deterministic elements -- such as \texttt{LIMIT}/\texttt{OFFSET} without ordering, \texttt{SAMPLE} or temporal functions -- such as \texttt{TODAY()} -- that may lead to changes in results not related to changes in the data; we remove 23 queries with \texttt{SAMPLE} and temporal functions, and add an \texttt{ORDER BY} clause to all queries to ensure determinism with \texttt{LIMIT}/\texttt{OFFSET} and more generally to facilitate quick comparison of results.
-We also eliminated 37 queries with 0 results forr all the snapshots.
-As a result of this process of filtering, we end up with a total of 221 deterministic queries.
+We also eliminated 37 queries with empty results for all snapshots.
+As a result of this process of filtering, we end up with a total of 221 non-empty deterministic queries.
 
@@
 Finally, we look at how the results for these queries change over the 23 versions.
-Figure~\ref{graph:changeV} shows how many queries had some results change between two consecutive versions, where we see that approximately half of the queries see changes in their results each version.
+Figure~\ref{graph:changeV} shows how many queries had some results change between two consecutive versions, where we see that approximately half of the result sets change in each version.
 But are these always the same queries that change every time?
@@
 
-Our experiments aim to ascertain the (relative) quality of predictions that can be made in the context of One Shot Change (OSC) and Time To Live (TTL) based on query features, statistics of predicate dynamics, as well as information about the historical evolution of queries.
+Our experiments evaluate the (relative) quality of One Shot Change (OSC) predictions that can be made based on query features, predicate features, result features, as well as combinations thereof, using a selection of binary classifiers.
 
 \paragraph{Setting:} To build our final datasets, we must define a past interval to consider.
-With 23 versions, we must hold out at least one version to label, leaving a maximum interval of 22 past versions to use to compute our data features.
+With 23 versions, we must hold out at least one version to predict, leaving a maximum interval of 22 past versions to use for computing our features.
 However, the more past versions we consider for a prediction, the fewer examples we can generate; for example, if we consider 22 versions, we will only have one ground truth label for each query.
-In the end, we opted to consider intervals of 3, 5, 9 and 17 previous versions, for example, ($G_{i-3}, G_{i-2}, G_{i-1}$) to predict changes in queries results for the subsequent version ($G_i$); this allows us to predict from $G_4$ to $G_{23}$ inclusive, providing 20 examples per query, 4,420 examples in the largest dataset and 1,326 examples in the smallest dataset.
-No feature explicitly indicates which specific version numbers are given and which are to be predicted.
+In the end, we opted to consider intervals of 3, 5, 9 and 17 previous versions, for example, in the case of an interval of three versions, we use ($G_{i-3}, G_{i-2}, G_{i-1}$) to predict changes in queries results for the subsequent version ($G_i$), which allows us to predict OSC from $G_4$ to $G_{23}$ inclusive, providing 20 labeled examples per query.
+We thus have 4,420 labeled examples for intervals of 3, ranging down to 1,326 examples for intervals of 17.
+No feature explicitly indicates the versions given or the version to be predicted.
 
@@
 
-For OSC prediction, we selected four classifiers for testing: Decision Trees, Linear SVM, Naive Bayes, and Nearest Neighbors.
+For OSC prediction, we test four binary classifiers: Decision Trees, Linear SVM, Naive Bayes, and Nearest Neighbors.
 We split the data to use 80\% for training and 20\% for tests.
-To avoid overfitting, we use stratified cross-validation in the training data.
-We consider three types of features: static query features (\textsc{q}): predicate dynamics (\textsc{p}) considering information of $\Delta(\mathcal{G},p)$ and the cardinality of predicates and information on historical results (\textsc{r}) considering the number of past changes in the fixed interval.
-We may also consider combinations of these features: \textsc{qp}, \textsc{qr}, \textsc{pr} and \textsc{qpr}.
+To avoid overfitting, we use stratified 5-fold cross-validation.
+We consider three types of features, as previously described -- query features (\textsc{q}), predicate features (\textsc{p}) and result features (\textsc{r}) -- considering historical data (only) from the fixed interval.
+We also consider combinations of these features: \textsc{qp}, \textsc{qr}, \textsc{pr} and \textsc{qpr}.
 
@@
 
-In addition, we train a Machine Learning model to determine the TTL of a query's results.
+%In addition, we train a Machine Learning model to determine the TTL of a query's results.
 To train the model we consider the changes observed in the monitored interval.
@@
 
-\paragraph{Results:} Table~\ref{tab:classifier} shows the $F_1$-score for the predictions made considering the union of different combinations of the sets of features (Q, P, R) and varying the window sizes (3, 5, 9, 17); for reference, we include a baseline that randomly guesses yes/no.
-We can note the best results were obtained using the features from \textsc{r} reaching values of 0.83 $F_1$.
-The results show that predicting changes in query results is indeed a challenging problem.
-Looking in further detail, as can be expected, prediction performs best when the \textsc{r} feature -- with knowledge of changes in the historical value of results of the input query -- is available.
-The features from sets \textsc{q} and  \textsc{p} -- which do not assume the availability of such information -- obtain a lower $F_1$-score: more specifically, the best result was using \textsc{pr} (0.83) with Linear SVM and window size 17, while the best results without \textsc{r} was using \textsc{Q} (0.60) with Linear SVM and window size 5.
-
-Furthermore, we show in Figure~\ref{fig:result} that the size of the window influences the quality of the results when considering \textsc{r} (blue), being better as the window increases.
-However, these results are not affected when we consider only predicates in \textsc{q} (brown), \textsc{p} (green)  or \textsc{qp} (black).
-Also, we show, that all configurations outperform the random baseline (red).
+\paragraph{Results:} Table~\ref{tab:classifier} shows the $F_1$-score for the predictions made considering the union of different combinations of the sets of features (\textsc{q}, \textsc{p}, \textsc{r}) and varying the window sizes (3, 5, 9, 17); for reference, we include a baseline that randomly guesses \texttt{yes}/\texttt{no}.
+Given the number of configurations presented, we only include $F_1$ scores for reasons of space; however, we remark that the Precision and Recall scores were in general quite balanced.
+The best results were obtained using the features from \textsc{r}, where, as can be expected, prediction performs best when knowledge of changes in the historical results of the input query is available.
+The features from sets \textsc{q} and \textsc{p} -- which do not assume the availability of such information -- obtain a significantly lower $F_1$-score; the best result including \textsc{r} was with \textsc{pr} ($F_1 = 0.831$) using Linear SVM and window size 17, while the best results without \textsc{r} was with \textsc{q} ($F_1 = 0.60$) using Linear SVM and window size 5.
+
+Furthermore, in Figure~\ref{fig:result}, we show that the size of the window influences the quality of the results when considering \textsc{r} features (blue), being better as the window increases.
+However, there is no clear trend in the results for increasing window sizes when considering models without using \textsc{r} features (red).
+(As aforementioned, we can see that Precision and Recall results are comparable.)
 
@@
 	\centering
-	\caption{$F_1$-score for tested classifiers considering different sets of features (Q, P, R) and window sizes (3,5,9,17)}
+	\caption{$F_1$-score for tested classifiers considering different sets of features (\textsc{q}, \textsc{p}, \textsc{r}) and window sizes (3,5,9,17)}
 	\label{tab:classifier}
 	\begin{tabular}{l@{\hskip 0.1in}l@{\hskip 0.1in}r@{\hskip 0.1in}r@{\hskip 0.1in}r@{\hskip 0.1in}r@{\hskip 0.1in}r@{\hskip 0.1in}r@{\hskip 0.1in}r}    \toprule		  
-		& \textbf{Classifier}        & \textsc{Q}         & \textsc{P}         & \textsc{R}         & \textsc{QP}        & \textsc{QR}        & \textsc{PR}        & \textsc{QPR}       \\    \midrule
-	w=3 	& \textit{Random Baseline} & 0.499          & 0.489          & 0.502          & 0.489          & 0.495          & 0.495          & 0.516          \\
-		& Decision Trees           & 0.497          & \underline{0.519}    & \underline{0.781}    & 0.497          & 0.569          & 0.706          & 0.6            \\
+		& \textbf{Classifier}        & \textsc{q}         & \textsc{p}         & \textsc{r}         & \textsc{qp}        & \textsc{qr}        & \textsc{pr}        & \textsc{qpr}       \\    \midrule
+	$w=3$ 	& \textit{Random Baseline} & 0.499          & 0.489          & 0.502          & 0.489          & 0.495          & 0.495          & 0.516          \\
+		& Decision Trees           & 0.497          & \underline{0.519}    & \underline{0.781}    & 0.497          & 0.569          & 0.706          & 0.600            \\
 		& Naive Bayes              & 0.432          & 0.419          & 0.758          & 0.432          & 0.478          & 0.759          & 0.479          \\
-		& Nearest Neighbors        & 0.505          & 0.517          & 0.776          & 0.509          & 0.679          & \underline{0.781}    & 0.68           \\
+		& Nearest Neighbors        & 0.505          & 0.517          & 0.776          & 0.509          & 0.679          & \underline{0.781}    & 0.680           \\
 		& Linear SVM               & \underline{0.596}    & 0.372          & 0.758          & \underline{0.593}    & \underline{0.775}    & 0.762          & \underline{0.774}    \\    \midrule
-		& \textbf{Better}          & \textbf{0.596} & \textbf{0.519} & \textbf{0.781} & \textbf{0.593} & \textbf{0.775} & \textbf{0.781} & \textbf{0.774} \\    \midrule
-	w=5	& \textit{Random Baseline} & 0.48           & 0.496          & 0.499          & 0.505          & 0.495          & 0.497          & 0.499          \\
+		& \textbf{Best}          & \textbf{0.596} & \textbf{0.519} & \textbf{0.781} & \textbf{0.593} & \textbf{0.775} & \textbf{0.781} & \textbf{0.774} \\    \midrule
+	$w=5$	& \textit{Random Baseline} & 0.480           & 0.496          & 0.499          & 0.505          & 0.495          & 0.497          & 0.499          \\
 		& Decision Trees           & 0.485          & 0.514          & 0.782          & 0.494          & 0.581          & 0.711          & 0.603          \\
-		& Naive Bayes              & 0.431          & 0.486          & 0.782          & 0.432          & 0.508          & \underline{0.797}    & 0.5            \\
-		& Nearest Neighbors        & 0.517          & \underline{0.535}    & \underline{0.795}    & 0.525          & 0.676          & 0.79           & 0.673          \\
-		& Linear SVM               & \underline{0.6}      & 0.384          & 0.784          & \underline{0.59}     & \underline{0.787}    & 0.789          & \underline{0.787}    \\    \midrule
-		& \textbf{Better}          & \textbf{0.6}   & \textbf{0.535} & \textbf{0.795} & \textbf{0.59}  & \textbf{0.787} & \textbf{0.797} & \textbf{0.787} \\    \midrule
-	w=9	& \textit{Random Baseline} & 0.497          & 0.505          & 0.493          & 0.483          & 0.497          & 0.511          & 0.515          \\
+		& Naive Bayes              & 0.431          & 0.486          & 0.782          & 0.432          & 0.508          & \underline{0.797}    & 0.500            \\
+		& Nearest Neighbors        & 0.517          & \underline{0.535}    & \underline{0.795}    & 0.525          & 0.676          & 0.790           & 0.673          \\
+		& Linear SVM               & \underline{0.600}      & 0.384          & 0.784          & \underline{0.590}     & \underline{0.787}    & 0.789          & \underline{0.787}    \\    \midrule
+		& \textbf{Best}          & \textbf{0.600}   & \textbf{0.535} & \textbf{0.795} & \textbf{0.590}  & \textbf{0.787} & \textbf{0.797} & \textbf{0.787} \\    \midrule
+	$w=9$	& \textit{Random Baseline} & 0.497          & 0.505          & 0.493          & 0.483          & 0.497          & 0.511          & 0.515          \\
 		& Decision Trees           & 0.456          & 0.517          & 0.793          & 0.48           & 0.668          & 0.735          & 0.651          \\
-		& Naive Bayes              & 0.434          & 0.461          & \underline{0.797}    & 0.435          & 0.531          & 0.79           & 0.531          \\
-		& Nearest Neighbors        & 0.514          & \underline{0.528}    & 0.787          & 0.517          & 0.653          & \underline{0.8}      & 0.648          \\
-		& Linear SVM   				& \underline{0.578}    & 0.397          & \underline{0.797}    & \underline{0.571}    & \underline{0.795}    & \underline{0.8}      & \underline{0.796}    \\     \midrule
-		& \textbf{Better}          & \textbf{0.578} & \textbf{0.528} & \textbf{0.797} & \textbf{0.571} & \textbf{0.795} & \textbf{0.8}   & \textbf{0.796} \\    \midrule
-	w=17	& \textit{Random Baseline} & 0.486          & 0.494          & 0.482          & 0.512          & 0.524          & 0.511          & 0.49           \\
+		& Naive Bayes              & 0.434          & 0.461          & \underline{0.797}    & 0.435          & 0.531          & 0.790           & 0.531          \\
+		& Nearest Neighbors        & 0.514          & \underline{0.528}    & 0.787          & 0.517          & 0.653          & \underline{0.800}      & 0.648          \\
+		& Linear SVM   				& \underline{0.578}    & 0.397          & \underline{0.797}    & \underline{0.571}    & \underline{0.795}    & \underline{0.800}      & \underline{0.796}    \\     \midrule
+		& \textbf{Best}          & \textbf{0.578} & \textbf{0.528} & \textbf{0.797} & \textbf{0.571} & \textbf{0.795} & \textbf{0.800}   & \textbf{0.796} \\    \midrule
+	$w=17$	& \textit{Random Baseline} & 0.486          & 0.494          & 0.482          & 0.512          & 0.524          & 0.511          & 0.490           \\
 		& Decision Trees           & 0.511          & \underline{0.511}    & \underline{0.827}    & 0.469          & 0.712          & 0.712          & 0.71           \\
-		& Naive Bayes              & 0.416          & 0.465          & 0.826          & 0.414          & 0.555          & 0.81           & 0.554          \\
-		& Nearest Neighbors        & 0.548          & 0.48           & 0.826          & 0.554          & 0.744          & 0.82           & 0.731          \\
-		& Linear SVM               & \underline{0.554}    & 0.438          & 0.826          & \underline{0.56}     & \underline{0.819}    & \underline{0.831}    & \underline{0.815}    \\    \midrule
-		& \textbf{Better}          & \textbf{0.554} & \textbf{0.511} & \textbf{0.827} & \textbf{0.56}  & \textbf{0.819} & \textbf{0.831} & \textbf{0.815}	\\    \bottomrule
+		& Naive Bayes              & 0.416          & 0.465          & 0.826          & 0.414          & 0.555          & 0.810           & 0.554          \\
+		& Nearest Neighbors        & 0.548          & 0.480           & 0.826          & 0.554          & 0.744          & 0.820           & 0.731          \\
+		& Linear SVM               & \underline{0.554}    & 0.438          & 0.826          & \underline{0.560}     & \underline{0.819}    & \underline{0.831}    & \underline{0.815}    \\    \midrule
+		& \textbf{Best}          & \textbf{0.554} & \textbf{0.511} & \textbf{0.827} & \textbf{0.560}  & \textbf{0.819} & \textbf{0.831} & \textbf{0.815}	\\    \bottomrule
 	\end{tabular}
@@
 	ymin=0,
+	ymax=1,
 	xtick={3,5,9,17},
@@
 	\addplot[
-	color=brown,
+	color=red,
 	mark=diamond,
@@
 	\addplot[
-	color=green,
+	color=red,
 	mark=diamond,
@@
 	\addplot[
-	color=black,
+	color=red,
 	mark=diamond,
@@
 	\addplot[
-	color=blue,
+	color=red,
 	mark=diamond,
@@
 	\end{tikzpicture}
-	\caption{Best results for each feature sets in term of $F_1$-measure (left), Precision score (center) and Recall (right) with different windows sizes 
+	\caption{Best results for combination of features in term of $F_1$-measure (left), Precision score (center) and Recall (right) with increasing windows sizes; blue indicates results using features from \textsc{r} while red indicates results not using features from \textsc{r}
 		\label{fig:result}}
@@
 
-Finally, to predict the TTL of the results of the query -- outputting the number of versions for which the query results can be expected to remain the same -- we employed techniques of multiple linear regression using the \textsc{sklearn} library.
+%Finally, to predict the TTL of the results of the query -- outputting the number of versions for which the query results can be expected to remain the same -- we employed techniques of multiple linear regression using the \textsc{sklearn} library.
 The latter allow us to see how the periods of estimation and the convergence analysis will behave.
@@
 
-\section{Conclusion and Future Work}
+\section{Summary, Conclusions and Future Work}
 \label{sec:conclusion}
 
-In this paper we evaluate methods to predict whether or not (OSC) and/or when (TTL) the results of an input query will change in the future.
-More specifically, we evaluate a framework based on classifiers and regression that accept features extracted from the query, from past versions of the data, and/or from the combination of both.
-Considering this framework, we show that there is a conceptual trade-off between the cost of computing features and their value for prediction: features extracted from queries alone are the most efficient to extract, not requiring historical data, but are quite coarse-grained for prediction; on the other hand, features based on changes in results in previous versions for an (unseen) input query are often the most costly to acquire, but offer fine-grained information for prediction; finally, features based on the dynamics of predicates used in the data offer a balance between the two, allowing to summarize historical data into succinct statistics, while offering more fine-grained information than static query features and less fine-grained information than historical results.
+In this paper we evaluate methods to predict whether or not (OSC) the results of an input query will change in a fixed point in the near future.
+More specifically, we evaluate a framework based on binary classifiers that accept features extracted from the query, from past versions of the data, and/or from the combination of both.
+Considering this framework, we hypothesize that there is a conceptual trade-off between the cost of computing features and their value for prediction: features extracted from queries alone are the most efficient to extract, not requiring historical data, but are quite coarse-grained for prediction; on the other hand, features based on changes in results over previous versions for an (unseen) input query are often the most costly to acquire, but offer fine-grained information for prediction; finally, features based on the dynamics of predicates offer a balance between the two, allowing to summarize historical data into succinct statistics, thus offering more fine-grained information than static query features but more coarse-grained information than historical results.
 
@@
 Settling on Wikidata, we collect 23 versions and curate 221 user-generated queries from which we compute results, statistics and ultimately features that form the gold standard for the tasks.
-Analysing these data and queries, we conclude that they meet the desired characteristics, offering a large-scale, diverse, real-world setting; more importantly, we find that our queries offer an ideal distribution of queries whose results always change, queries whose results never change, and (more importantly) queries whose results change intermittently and for which prediction is thus a non-trivial task.
-We will make this gold standard available after double-blind review -- including queries, results, and features -- so that the results presented here can be reproduced and so that it can be reused in further research.
-
-Finally, we use this gold standard dataset to evaluate the trade-off identified between different types of features for predictions in the OSC and TTL tasks.
+Analysing these data and queries, we conclude that they meet the desired characteristics, offering a large-scale, diverse, real-world setting; more importantly, we find that our queries offer an ideal distribution of queries whose results always change, queries whose results never change, and (more importantly) queries whose results change intermittently and for which prediction is thus non-trivial.
+
+Finally, we use this gold standard dataset to evaluate the trade-off identified between different types of features for OSC predictions.
 For OSC, we evaluate the success rates of a variety of classifiers based on the $F_1$-measure of the predictions.
-Our results show that the features based on historical changes to query results perform best ($F_1 = 0.83$), considering static query features and predicate dynamics alone is less competitive ($F_1 =0.60$) when historical results are not available or are prohibitively costly to compute.
-
-In terms of future work, gold standards based on datasets other than Wikidata could be pursued.
-Also, our gold standard based on Wikidata could be extended to consider more versions spanning a longer period of time and/or more queries; as previously mentioned, a very large query dataset for Wikidata was recently published~\cite{MalyshevKGGB18} from which it should be possible to extend the Wikidata gold standard (the challenge being to keep a balance of static and dynamic queries with diverse features).
+Our results show that the features based on historical changes to query results perform best (reaching $F_1 = 0.831$ in the best configuration), whereas considering static query features and predicate dynamics alone is less competitive (reaching $F_1 =0.600$ in the best configuration) when historical results are not available or are prohibitively costly to compute.
+Comparing query and predicate features, in fact, query features performed better, where predicate features alone only barely outperformed a random baseline.
+
+As a conclusion, our results show that having knowledge of the historical changes of the results of a query is important for improving the quality of OSC predictions using binary classifiers.
+However, in many scenarios, such knowledge is often not available or may not be practical to compute.
+Considering a real-world caching use-case, for example, historical changes in results may be maintained for queries that are frequently repeated with relatively low-cost, but for a previously unseen query, computing results for past versions at runtime would incur a prohibitive cost.
+Hence, we identify an open research question: is it possible to estimate the dynamics of query results without relying on (costly) knowledge about historical changes of query results while staying competitive with the quality of prediction possible when such knowledge is available?
+
+In terms of other future work, gold standards based on datasets other than Wikidata could be pursued.
+Also, our gold standard based on Wikidata could be extended to consider more versions spanning a longer period of time and/or more queries; a very large query dataset for Wikidata was recently published~\cite{MalyshevKGGB18} from which it should be possible to extend the Wikidata gold standard (the challenge being to keep a balance of static and dynamic queries with diverse features).
 Concerning the prediction tasks themselves, there are a variety of other features and approaches that could be applied; one promising direction may be, for example, to apply a more fine-grained analysis of the query, considering for example estimates of the selectivity of particular triples patterns, and combining this information with statistics on the dynamics of the data and the query operators used; our current gold standard could be used to validate such developments.
+Admittedly, there may also be better ways to perform such predictions without relying on binary classifiers, but rather using more analytical methods.
 Finally, it would be interesting to investigate the effectiveness of these techniques in practice, developing caching systems, synchronization schedules, and other applications, based on the evaluated predictions.
+
+%\paragraph{Acknowledgements} This work was supported by the Millennium Institute for Foundational Research on Data (IMFD) and by Fondecyt Grant No. 1181896.
 

### 2019-08-05 main.tex
camera-ready updates following this comments: - In the introduction it is written: “(aiming for stronger consistency at the cost of more frequent updates),”, I would say “aiming for stronger consistency at the cost of more frequent querying”
- In Figure 8, for F1, Precision and Recall several red lines to
indicate results not using features from R. Why we have several lines?
Do they refer to different classification methods? Please clarify this.

- I regret that the experimentations are not reproducible. Code, data,
queries are not available. I believe the gold standard databases (RDF
data, qu

@@
 Since datasets change over time, long-running applications that cache and repeatedly use query results obtained from an external SPARQL endpoint may resubmit the queries regularly to ensure up-to-dateness.
-As a result, without further information as to the dynamics of a particular SPARQL query, applications face the choice of either performing frequent query executions that may be redundant and repeatedly return the same results (aiming for stronger consistency at the cost of more frequent updates), or performing infrequent query executions that may lead to stale data being persisted in the application when the underlying sources change (accepting weaker consistency to improve efficiency/scalability).
+As a result, without further information as to the dynamics of a particular SPARQL query, applications face the choice of either performing frequent query executions that may be redundant and repeatedly return the same results (aiming for stronger consistency at the cost of more frequent querying), or performing infrequent query executions that may lead to stale data being persisted in the application when the underlying sources change (accepting weaker consistency to improve efficiency/scalability).
 Given the costs for clients and servers of repetitive requests served over the Web  and the potential efficiency gains offered by local caches, several weak consistency approaches have been proposed that try to keep the local data of the applications updated at lower cost by predicting changes~\cite{KnuthHS16,DividinoGS15,NishiokaS17}.
@@
 	\end{tikzpicture}
-	\caption{Best results for combination of features in term of $F_1$-measure (left), Precision score (center) and Recall (right) with increasing windows sizes; blue indicates results using features from \textsc{r} while red indicates results not using features from \textsc{r}
+	\caption{Best results for combination of features in term of $F_1$-measure (left), Precision score (center) and Recall (right) with increasing windows sizes; blue indicates results using features from \textsc{r} (\textsc{r}, \textsc{qr}, \textsc{pr} and \textsc{qpr}) while red indicates results not using features from \textsc{r} (\textsc{q}, \textsc{p}, \textsc{qp} and the Baselines)
 		\label{fig:result}}
@@
 Analysing these data and queries, we conclude that they meet the desired characteristics, offering a large-scale, diverse, real-world setting; more importantly, we find that our queries offer an ideal distribution of queries whose results always change, queries whose results never change, and (more importantly) queries whose results change intermittently and for which prediction is thus non-trivial.
+We make this gold standard available at \url{https://users.dcc.uchile.cl/~amoya/quweda2019/}.
 

### 2019-08-05 main.tex
clarifying this: On page 12, a random baseline is used as a reference. Was the yes/no random guess adjusted to the apriori probability of results to change?

@@
 
-\paragraph{Results:} Table~\ref{tab:classifier} shows the $F_1$-score for the predictions made considering the union of different combinations of the sets of features (\textsc{q}, \textsc{p}, \textsc{r}) and varying the window sizes (3, 5, 9, 17); for reference, we include a baseline that randomly guesses \texttt{yes}/\texttt{no}.
+\paragraph{Results:} Table~\ref{tab:classifier} shows the $F_1$-score for the predictions made considering the union of different combinations of the sets of features (\textsc{q}, \textsc{p}, \textsc{r}) and varying the window sizes (3, 5, 9, 17); for reference, we include a baseline that randomly guesses \texttt{yes}/\texttt{no} generating random predictions by respecting the training set’s class distribution.
 Given the number of configurations presented, we only include $F_1$ scores for reasons of space; however, we remark that the Precision and Recall scores were in general quite balanced.

### 2019-08-06 main.tex
Camera ready draft

--- 2019-08-05 main.tex
@@
 To the best of our knowledge, this is the first work to address this problem.
-Along these lines, we evaluate a general architecture based on Machine Learning that extracts static features from a query, as well as dynamic features from the query and dataset.
+Along these lines, we evaluate a general architecture based on Machine Learning that extracts static features from a query, as well as features from the query and dynamic dataset.
 These features are fed into a binary classifier to predict whether or not the query results will change in a fixed point in the future, or, more ambitiously, can be fed into a regression model to predict when the query results are likely to change.
@@
 
-A variety of works have addressed the issue of modeling and consuming dynamic Linked Data from a broad range of perspectives~\cite{PassantM10,TrampFEA10,UmbrichHHPD10,UmbrichKL10,KaferAUOH13,DividinoSGG13,DividinoKG14,Kjernsmo15,NishiokaS16,GonzalezH18}.
+A variety of works have addressed the issue of modeling and consuming dynamic Linked Data from a broad range of perspectives~\cite{PassantM10,TrampFEA10,UmbrichHHPD10,UmbrichKL10,KaferAUOH13,DividinoSGG13,IbanezSMC14,DividinoKG14,Kjernsmo15,NishiokaS16,GonzalezH18}.
 One of the major challenges considered is that of keeping cached copies of remote dynamic data -- cached for reasons of efficiency and scalability -- up-to-date on the consumer side, which we refer to as the \textit{synchronization problem}.
 
-Some works have addressed the synchronization problem on the publisher side, proposing notification mechanisms that keep registered consumers informed about relevant changes to the data~\cite{MissierACDG07,PassantM10,TrampFEA10,MaderMS14,webSub18}; although such approaches may facilitate strong consistency -- meaning that consumers are kept up-to-date with the remote data on the publisher side -- they centralize the burden  of synchronization on the publisher, potentially leading to scalability issues.
+Some works have addressed the synchronization problem on the publisher side, proposing notification mechanisms that keep registered consumers informed about relevant changes to the data~\cite{MissierACDG07,PassantM10,TrampFEA10,MaderMS14,IbanezSMC14,webSub18}; although such approaches may facilitate strong consistency -- meaning that consumers are kept up-to-date with the remote data on the publisher side -- they centralize the burden  of synchronization on the publisher, potentially leading to scalability issues.
 
@@
 The semantics of a SPARQL \texttt{SELECT} query $Q$ is defined in terms of its \emph{evaluation} over an RDF graph $G$, denoted $Q(G)$, giving a set of partial mappings from projected variables to the set $\Iris \cup \Lits \cup \Bnodes$; we refer to P{\'{e}}rez et al.~\cite{PerezAG09} for definitions.
-(According to the standard, a SPARQL query is evaluated over a set of \emph{named graphs}; however, since the Wikidata query service indexes data in one RDF graph, to keep discussion concise, we consider RDF graphs, though our methods can be generalized in a straightforward manner to the named graphs setting.)
+(A SPARQL query is in fact evaluated over a set of \emph{named graphs}; though we consider RDF graphs here, our methods also generalize to the named graphs setting.)
 
@@
 \centering
-\resizebox{0.9\textwidth}{!}{
+\resizebox{0.8\textwidth}{!}{
 \begin{tikzpicture}
@@
 
-\node[c,right=1.5\hgap of f6.south,minimum width=6.5em] (c) {Classification\hsp};
+\node[c,right=2\hgap of f6.south,minimum width=6.5em] (c) {Classification\hsp};
 
@@
 In some cases, however, the hypothesized correlation is not direct since, for example, adding more triple patterns may serve to narrow the query down and focus it on a static part of the graph (e.g., when looking for the movies of directors, adding a triple pattern to restrict the results to directors who have died may reduce the likelihood of changes in the results in a future version).
-Hence it will be of interest to see how these features affect the predictions made experimentally.
+Hence it will be of interest to see, experimentally, how these features affect the predictions.
 
@@
 In the third group gathers related features into one dimension: in the case of \textit{Recursive path}, we group queries with path expressions of the form $e$\texttt{*} or $e$\texttt{+}.
-On the other hand, in \textit{Negation}, we group non-monotonic features that allow for modeling difference (\texttt{MINUS}, \texttt{NOT EXISTS}, \texttt{!BOUND}).
+On the other hand, in \textit{Negation}, we group non-monotonic features that allow for modeling difference (\texttt{MINUS}, \texttt{NOT EXISTS}, \texttt{!BOUND}\footnote{We recall that \texttt{!BOUND} can be combined with \texttt{OPTIONAL} to express negation.}.
 These features -- though course-grained -- are straightforward to extract from a query, and offer valuable insights into how the query may behave in a dynamic query; for example, the \textit{Negation} feature captures information about the (non-)monotonicity of the query, while we suppose that \textit{Recursive paths}, which may traverse an arbitrary number of triples in the graph, might be more sensitive to change.
@@
 This component captures how often and how many triples change for a predicate in a time interval, with the idea that -- as in previous works~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12,DehghanzadehPKUHD14} -- predicates capture rich information about the dynamics of a dataset, where the results of queries with dynamic predicates will be more sensitive to change.
-Furthermore, given that the number of predicates is often \textit{relatively} low (in the thousands or tens of thousand) for most datasets, such statistics are quite lightweight to maintain and compute.
+Assuming that the number of predicates is \textit{relatively} low, such statistics can be easier to maintain.
 Formally, given two RDF graphs $G_i$ and $G_j$, we denote by $G_i \oplus G_j$ the set of triples $(G_i \cup G_j) - (G_i \cap G_j)$ where ``$-$'' denotes set difference; in other words, $G_i \oplus G_j$ denotes the triples in one graph or the other but not both (noting that $G_i \oplus G_j = G_j \oplus G_i$).
@@
 
-\paragraph{Results:} Table~\ref{tab:classifier} shows the $F_1$-score for the predictions made considering the union of different combinations of the sets of features (\textsc{q}, \textsc{p}, \textsc{r}) and varying the window sizes (3, 5, 9, 17); for reference, we include a baseline that randomly guesses \texttt{yes}/\texttt{no}.
+\paragraph{Results:} Table~\ref{tab:classifier} shows the $F_1$-score for the predictions made considering the union of different combinations of the sets of features (\textsc{q}, \textsc{p}, \textsc{r}) and varying the window sizes (3, 5, 9, 17); for reference, we include a baseline that randomly guesses \texttt{yes}/\texttt{no} respecting the observed class distribution.
 Given the number of configurations presented, we only include $F_1$ scores for reasons of space; however, we remark that the Precision and Recall scores were in general quite balanced.
@@
 
-In this paper we evaluate methods to predict whether or not (OSC) the results of an input query will change in a fixed point in the near future.
+In this paper we evaluate methods to predict whether or not (OSC) the results of an input query will change at a fixed point in the near future.
 More specifically, we evaluate a framework based on binary classifiers that accept features extracted from the query, from past versions of the data, and/or from the combination of both.
@@
 
-We thus explore this trade-off experimentally.
-For this, we define some desirable characteristics for the queries and data used in the experimental setting, such as availability, size, dynamics, diversity, etc.
-Settling on Wikidata, we collect 23 versions and curate 221 user-generated queries from which we compute results, statistics and ultimately features that form the gold standard for the tasks.
-Analysing these data and queries, we conclude that they meet the desired characteristics, offering a large-scale, diverse, real-world setting; more importantly, we find that our queries offer an ideal distribution of queries whose results always change, queries whose results never change, and (more importantly) queries whose results change intermittently and for which prediction is thus non-trivial.
-We make this gold standard available at \url{https://users.dcc.uchile.cl/~amoya/quweda2019/}.
-
-Finally, we use this gold standard dataset to evaluate the trade-off identified between different types of features for OSC predictions.
-For OSC, we evaluate the success rates of a variety of classifiers based on the $F_1$-measure of the predictions.
+We thus explore this trade-off experimentally using 23 versions and 221 user-contributed queries for Wikidata to form a gold standard dataset.
+We use this gold standard dataset to evaluate the trade-off identified between different types of features for OSC predictions.
 Our results show that the features based on historical changes to query results perform best (reaching $F_1 = 0.831$ in the best configuration), whereas considering static query features and predicate dynamics alone is less competitive (reaching $F_1 =0.600$ in the best configuration) when historical results are not available or are prohibitively costly to compute.
@@
 
-As a conclusion, our results show that having knowledge of the historical changes of the results of a query is important for improving the quality of OSC predictions using binary classifiers.
+In conclusion, our results show that having knowledge of the historical changes of the results of a query is important for improving the quality of OSC predictions using binary classifiers.
 However, in many scenarios, such knowledge is often not available or may not be practical to compute.
@@
 
-In terms of other future work, gold standards based on datasets other than Wikidata could be pursued.
-Also, our gold standard based on Wikidata could be extended to consider more versions spanning a longer period of time and/or more queries; a very large query dataset for Wikidata was recently published~\cite{MalyshevKGGB18} from which it should be possible to extend the Wikidata gold standard (the challenge being to keep a balance of static and dynamic queries with diverse features).
-Concerning the prediction tasks themselves, there are a variety of other features and approaches that could be applied; one promising direction may be, for example, to apply a more fine-grained analysis of the query, considering for example estimates of the selectivity of particular triples patterns, and combining this information with statistics on the dynamics of the data and the query operators used; our current gold standard could be used to validate such developments.
-Admittedly, there may also be better ways to perform such predictions without relying on binary classifiers, but rather using more analytical methods.
+Regarding future work, our gold standard based on Wikidata could be extended to consider more versions spanning a longer period of time and/or more queries; a very large query dataset for Wikidata was recently published~\cite{MalyshevKGGB18}, which may be of use here.
+Building gold standards based on other datasets would also help to diversify the evaluation process.
+Concerning the prediction tasks themselves, one promising direction may be to apply a more fine-grained analysis of the query, considering (for example) the selectivity of particular triples patterns as well as their dynamicity.
+There may also be better ways to perform such predictions without relying on binary classifiers, but rather using more analytical methods.
 Finally, it would be interesting to investigate the effectiveness of these techniques in practice, developing caching systems, synchronization schedules, and other applications, based on the evaluated predictions.
 
-%\paragraph{Acknowledgements} This work was supported by the Millennium Institute for Foundational Research on Data (IMFD) and by Fondecyt Grant No. 1181896.
+\paragraph{Material} We make supplementary material (queries, data, results, etc.) available at \url{https://users.dcc.uchile.cl/~amoya/quweda2019/}.
+
+\paragraph{Acknowledgements} This work was supported by the Millennium Institute for Foundational Research on Data (IMFD) and by Fondecyt Grant No. 1181896.
 

### 2019-08-06 main.tex
hyperref fixes

@@
     citecolor={black}, urlcolor={black},
-    pdftitle={Estimating the dynamics of SPARQL query results},    % title
+    pdftitle={Estimating the dynamics of SPARQL query results using binary classification},    % title
     pdfauthor={Alberto Moya, Aidan Hogan},     % author
-    pdfsubject={ISWC 2019},   % subject of the document
+    pdfsubject={QuWeDa @ ISWC 2019},   % subject of the document
     pdfkeywords={SPARQL;} {Linked Data;} {Dynamics;} {Wikidata;}, % list of keywords

### 2019-08-06 main.tex
Author name and Acknowledgements updated

--- 2019-08-06 main.tex
@@
     citecolor={black}, urlcolor={black},
-    pdftitle={Estimating the dynamics of SPARQL query results},    % title
-    pdfauthor={Alberto Moya, Aidan Hogan},     % author
-    pdfsubject={ISWC 2019},   % subject of the document
+    pdftitle={Estimating the dynamics of SPARQL query results using binary classification},    % title
+    pdfauthor={Alberto Moya Loustaunau, Aidan Hogan},     % author
+    pdfsubject={QuWeDa @ ISWC 2019},   % subject of the document
     pdfkeywords={SPARQL;} {Linked Data;} {Dynamics;} {Wikidata;}, % list of keywords
@@
 %
-\author{Alberto Moya \and Aidan Hogan}
+\author{Alberto Moya Loustaunau\and Aidan Hogan}
 %
@@
 
-\paragraph{Acknowledgements} This work was supported by the Millennium Institute for Foundational Research on Data (IMFD) and by Fondecyt Grant No. 1181896.
+\paragraph{Acknowledgements} This work was supported by the Millennium Institute for Foundational Research on Data (IMFD), by Fondecyt Grant No.
+1181896 and by CONICYT PFCHA/Doctorado Nacional/2017 - 21171070.
 

### 2019-11-13 main.tex
Updating title case

@@
     citecolor={black}, urlcolor={black},
-    pdftitle={Estimating the dynamics of SPARQL query results using binary classification},    % title
+    pdftitle={Estimating the Dynamics of SPARQL Query Results Using Binary Classification},    % title
     pdfauthor={Alberto Moya Loustaunau, Aidan Hogan},     % author
@@
 %
-\title{Estimating the dynamics of SPARQL query results using binary classification
+\title{Estimating the Dynamics of SPARQL Query Results Using Binary Classification
 %\thanks{This work was supported by CONICYT PFCHA, Doctorado Becas Chile/2017 -- 72180000, the Millennium Institute for Foundational Research on Data (IMFD) and FONDECYT Grant No. 1181896.}
@@
 %
-\titlerunning{Estimating the dynamics of SPARQL query results}
+\titlerunning{Estimating the Dynamics of SPARQL Query Results}
 % If the paper title is too long for the running head, you can set
